### 2024-03-12 draft/rectangulotopes.tex
start paper

\documentclass{amsart}

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
\usepackage[bottom]{footmisc}
\usepackage{tikz}
%\usepackage{tkz-graph}
%\usepackage{tikz-qtree}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles, cd}
\usepackage[external]{forest}
%\tikzexternalize
\graphicspath{{figures/}{figures/nodes/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}
\renewcommand{\topfraction}{1} % possibility to have one page of pictures
\renewcommand{\bottomfraction}{1} % possibility to have one page of pictures
\usepackage[noabbrev,capitalise]{cleveref}
\usepackage[export]{adjustbox}
\usepackage{ulem}\normalem

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% theorems
\newtheorem{theorem}{Theorem}%[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem*{theorem*}{Theorem}%[section]

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{question}[theorem]{Question}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{assumption}[theorem]{Assumption}
\crefname{notation}{Notation}{Notations}
\crefname{problem}{Problem}{Problems}

% math special letters
\newcommand{\R}{\mathbb{R}} % reals
\newcommand{\N}{\mathbb{N}} % naturals
\newcommand{\Z}{\mathbb{Z}} % integers
\newcommand{\C}{\mathbb{C}} % complex
\newcommand{\I}{\mathbb{I}} % set of integers
\newcommand{\HH}{\mathbb{H}} % hyperplane
\newcommand{\K}{\mathbb{K}} % field
\newcommand{\fA}{\mathfrak{A}} % alternating group
\newcommand{\fB}{\mathfrak{S}^\textsc{b}} % signed symmetric group
\newcommand{\cA}{\mathcal{A}} % algebra
\newcommand{\cC}{\mathcal{C}} % collection
\newcommand{\cS}{\mathcal{S}} % ground set
\newcommand{\uR}{\underline{R}} % underline set
\newcommand{\uS}{\underline{S}} % underline set
\newcommand{\uT}{\underline{T}} % underline set
\newcommand{\oS}{\overline{S}} % overline set
\newcommand{\ucS}{\underline{\cS}} % underline ground set
\renewcommand{\c}[1]{\mathcal{#1}} % caligraphic letters
\renewcommand{\b}[1]{{\boldsymbol{#1}}} % bold letters
\newcommand{\bb}[1]{\mathbb{#1}} % bb letters
\newcommand{\f}[1]{\mathfrak{#1}} % frak letters
\newcommand{\h}{\widehat} % hat letters

% math commands
\newcommand{\set}[2]{\left\{ #1 \;\middle|\; #2 \right\}} % set notation
\newcommand{\bigset}[2]{\big\{ #1 \;\big|\; #2 \big\}} % big set notation
\newcommand{\Bigset}[2]{\Big\{ #1 \;\Big|\; #2 \Big\}} % Big set notation
\newcommand{\setangle}[2]{\left\langle #1 \;\middle|\; #2 \right\rangle} % set notation
\newcommand{\ssm}{\smallsetminus} % small set minus
\newcommand{\dotprod}[2]{\left\langle \, #1 \; \middle| \; #2 \, \right\rangle} % dot product
\newcommand{\symdif}{\,\triangle\,} % symmetric difference
\newcommand{\one}{\b{1}} % the all one vector
\newcommand{\eqdef}{\mbox{\,\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}}\ensuremath{=}\,}} % :=
\newcommand{\defeq}{\mbox{~\ensuremath{=}\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}} }} % =:
\newcommand{\simplex}{\b{\triangle}} % simplex
\renewcommand{\implies}{\Rightarrow} % imply sign
\newcommand{\transpose}[1]{{#1}^t} % transpose matrix

% operators
\DeclareMathOperator{\conv}{conv} % convex hull
\DeclareMathOperator{\vect}{vect} % linear span
\DeclareMathOperator{\cone}{cone} % cone hull
\DeclareMathOperator{\inv}{inv} % inversions
\DeclareMathOperator{\ninv}{ninv} % inversions
\DeclareMathOperator{\Ima}{Im} % image
\DeclareMathOperator{\Vol}{Vol} % (mixed) volume

% others
\newcommand{\ie}{\textit{i.e.}~} % id est
\newcommand{\eg}{\textit{e.g.}~} % exempli gratia
\newcommand{\Eg}{\textit{E.g.}~} % exempli gratia
\newcommand{\apriori}{\textit{a priori}} % a priori
\newcommand{\viceversa}{\textit{vice versa}} % vice versa
\newcommand{\versus}{\textit{vs.}~} % versus
\newcommand{\aka}{\textit{a.k.a.}~} % also known as
\newcommand{\perse}{\textit{per se}} % per se
\newcommand{\ordinal}{\textsuperscript{th}} % th for ordinals
\newcommand{\ordinalst}{\textsuperscript{st}} % st for ordinals
\definecolor{darkblue}{rgb}{0,0,0.7} % darkblue color
\definecolor{green}{RGB}{57,181,74} % darkblue color
\definecolor{violet}{RGB}{147,39,143} % darkblue color
\newcommand{\darkblue}{\color{darkblue}} % darkblue command
\newcommand{\defn}[1]{\textsl{\darkblue #1}} % emphasis of a definition
\newcommand{\para}[1]{\smallskip\noindent\uline{#1.}} % paragraph
\renewcommand{\topfraction}{1} % possibility to have one page of pictures
\renewcommand{\bottomfraction}{1} % possibility to have one page of pictures
%\renewcommand\labelitemi{$\diamond$} % redefine itemize default symbol

% marginal comments
\usepackage{todonotes}
\newcommand{\vincent}[1]{\todo[color=blue!30]{\rm #1 \\ \hfill --- V.}}
\newcommand{\Vincent}[1]{\todo[inline, color=blue!30]{\rm #1 \\ \hfill --- V.}}
\newcommand{\jean}[2][]{\todo[size=\scriptsize, color=orange!30,#1]{\rm #2 \\ \hfill --- J.}}
\newcommand{\Jean}[2][]{\todo[inline, size=\scriptsize, color=orange!30,#1]{\rm #2 \\ \hfill --- J.}}

% lattices
\newcommand{\meet}{\wedge} % meet
\newcommand{\join}{\vee} % join
\newcommand{\bigMeet}{\bigwedge} % meet
\newcommand{\bigJoin}{\bigvee} % join
\newcommandx{\projDown}[1][1={}]{\smash{\pi_\downarrow^{#1}}} % down projection map
\newcommandx{\projUp}[1][1={}]{\smash{\pi^\uparrow_{#1}}} % up projection map
\newcommand{\con}{\mathrm{con}} % congruence

% geometry
\newcommandx{\Fan}[1][1=D]{\mathcal{F}_{#1}} % fan
\newcommand{\polytope}[1]{\mathds{#1}} % font polytope

% specific rectangulotope
\newcommand{\Asso}{\polytope{A}\mathsf{sso}} % associahedron
\newcommand{\WRP}{\polytope{WRP}} % weak rectangulation polytope
\newcommand{\SRP}{\polytope{SRP}} % strong rectangulation polytope
\newcommand{\horizontalPattern}{\smash{\raisebox{-.15cm}{\includegraphics[scale=.5]{horizontalPattern}}}} % horizontal pattern
\newcommand{\verticalPattern}{\smash{\raisebox{-.25cm}{\includegraphics[scale=.5]{verticalPattern}}}} % vertical pattern

% formating the table of contents
\setcounter{tocdepth}{4}
\makeatletter
\def\l@part{\@tocline{1}{8pt}{0pc}{}{}}
\def\l@section{\@tocline{1}{4pt}{0pc}{}{}}
\makeatother
\let\oldtocpart=\tocpart
\renewcommand{\tocpart}[2]{\sc\large\oldtocpart{#1}{#2}}
\let\oldtocsection=\tocsection
\renewcommand{\tocsection}[2]{\bf\oldtocsection{#1}{#2}}
\let\oldtocsubsubsection=\tocsubsubsection
\renewcommand{\tocsubsubsection}[2]{\quad\oldtocsubsubsection{#1}{#2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Rectangulation polytopes}

\thanks{VP was partially supported by the Spanish grant PID2022-137283NB-C21 of MCIN/AEI/10.13039/501100011033 / FEDER, UE, by Departament de Recerca i Universitats de la Generalitat de Catalunya (2021 SGR 00697), by the French grant CHARMS (ANR-19-CE40-0017), and by the French--Austrian projects PAGCAP (ANR-21-CE48-0020 \& FWF I 5788).}

\author{Jean Cardinal}
\address{Université Libre de Bruxelles}
\email{jean.cardinal@ulb.be}
\urladdr{\url{https://jean.cardinal.web.ulb.be}}

\author{Vincent Pilaud}
\address{Universitat de Barcelona}
\email{vincent.pilaud@ub.edu}
\urladdr{\url{https://www.ub.edu/comb/vincentpilaud/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{abstract}
\end{abstract}

\maketitle

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{Rectangulations}
\label{sec:rectangulations}

%%%%%%%%%%%%

\subsection{Combinatorics of (generic) rectangulations}
\label{subsec:rectangulations}

\Vincent{
Note that there seem to be a canonical integer drawing of~$R$ described as follows.
For vertical segment~$v$ of~$R$, consider the path~$p$ passing through~$v$ with maximal slope.
The abscissa of~$v$ is then the number of rectangle on the left of~$p$ plus the number of \horizontalPattern{} patterns along~$p$ before~$v$ minus the number of \horizontalPattern{} patterns along~$p$ after~$v$.
The ordinate of an horizontal segment~$h$ of~$R$ is defined symmetrically.
}

%%%%%%%%%%%%

\subsection{Source and target trees of a rectangulation}
\label{subsec:sourceTargetTrees}

\begin{figure}
	\centerline{\includegraphics[scale=.7]{strongRectangulation} \qquad \includegraphics[scale=.7]{strongRectangulationTrees}}
	\caption{A strong rectangulation (left) and its pair of intertwining twin trees (right). The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$. Example from \cite{AsinowskiCardinalFelsnerFusy}.}
\end{figure}

Fix a rectangulation~$R$ with $n$ rectangles, and consider the directed graph~$D(R)$ whose vertex set is the set of all vertices of~$R$, and whose edges are obtained as follows.
For each rectangle~$r$ of~$R$, we include in~$D(R)$ four edges joining the vertices of~$r$, where the horizontal edges are oriented from left to right if they are horizontal, and the vertical edges are oriented from bottom to top if they are vertical.
Hence, the bottom left corner of~$r$ is its source~$s(r)$, while its top right corner is its target~$t(r)$.
Note that since~$R$ is generic, the sources~$\set{s(r)}{r \in [n]}$ and the targets~$\set{t(r)}{r \in [n]}$ are disjoint.
The \defn{source tree}~$S(R)$ (resp.~\defn{target tree}~$T(R)$) is the subgraph of~$D(R)$ induced by the sources~$\set{s(r)}{r \in [n]}$ (resp.~by the targets~$\set{t(r)}{r \in [n]}$).
See \cref{fig:sourceTargetTrees}.

\begin{lemma}
The source tree~$S(R)$ (resp.~target tree~$T(R)$) is a binary tree, rooted at the bottom left (resp.~top right) corner of~$R$, and oriented from (resp.~towards) its root.
\end{lemma}

\begin{proof}
By symmetry, we only prove the statement for the source tree~$S(R)$.
As the triangulation is generic, each source~$s(r)$ distinct from the bottom left corner of~$R$ is either on the top edge of a rectangle~$r'$ or on the right edge of a rectangle~$r'$ (but not both).
This shows that~$r$ distinct from the bottom left rectangle of~$R$ has a unique parent~$r'$ in~$S(R)$, so that~$S(R)$ is indeed a tree on~$[n]$.
This tree is binary as each node has at most one vertical child and one horizontal child, and no other children.
\end{proof}

We complete each node of~$S(R)$ and~$T(R)$ with a vertical (resp.~horizontal) leaf if it has no vertical (resp.~horizontal) child.
See \cref{fig:sourceTargetTrees}.
Recall that the \defn{inorder labeling} of a binary tree~$T$ is the labeling of the nodes of~$T$ such that the label of each node~$t$ of~$T$ is larger than all labels in the left subtree of~$t$ and larger than all labels in the right subtree of~$t$.

\begin{lemma}
For any rectangle~$r$ of~$R$, the inorder label of the source~$s(r)$ in~$S(R)$ coincides with the inorder label of the target~$t(r)$ in~$T(R)$.
\end{lemma}

\begin{proof}
\end{proof}

This enables to unambiguously label the rectangles of~$R$ by the inorder 
The resulting labeling coincides with the NW--SE labeling of~\cite{AsinowskiCardinalFelsnerFusy}.
We prefer our presentation with~$S(R)$ and~$T(R)$ here as we will actually use these two binary trees.

\begin{lemma}
The source tree~$S(R)$ and target tree~$T(R)$ are \defn{twin binary trees}, meaning that they satisfy the following equivalent conditions:
\begin{enumerate}[(i)]
\item $S$ and~$T$ admit a common linear extension,
\item for any~$i \in [n-1]$, the node~$i$ is in the left subtree of the node~$i+1$ in~$S$ if and only if the node~$i+1$ is in the right subtree of the node~$i$ in~$R$,
\item for any~$i \in [n-1]$, the $(i+1)$-st leaf of~$S$ is a left leaf if and only if the $(i+1)$-st leaf of~$R$ is a right leaf.
\end{enumerate}
\end{lemma}

\begin{proof}
\end{proof}

%%%%%%%%%%%%

\subsection{Rectangulation insertion of permutations}
\label{subsec:insertion}

\Vincent{If my canonical drawing indeed makes sense, I would love to describe the insertion map directly on rectangles.}

\Vincent{Describe the strong poset here?}

The insertion fiber of a rectangulation~$R$ precisely consists in the linear extensions of the strong poset.

%%%%%%%%%%%%

\subsection{Diagonal rectangulations}
\label{subsec:diagonalRectangulations}

\begin{figure}
	\centerline{\includegraphics[scale=.7]{weakRectangulation} \qquad \includegraphics[scale=.7]{weakRectangulationTrees}}
	\caption{A weak rectangulation (left) and its pair of separated twin trees (right). The corresponding cotwisted Baxter permutation is $[7,14,15,16,8,11,13,10,5,6,1,4,9,2,3,12]$ and its inverse is $[11, 14, 15, 12, 9, 10, 1, 5, 13, 8, 6, 16, 7, 2, 3, 4]$.}
\end{figure}

We call \defn{antidiagonal} the segment joining the top left corner to the bottom right corner.
A rectangulation~$R$ is \defn{diagonal} (or \defn{weak}) if it satisfies the following equivalent conditions:
\begin{enumerate}[(i)]
\item all rectangles of~$R$ meet the antidiagonal,
\item the source~$s(r)$ (resp.~target~$t(r)$) of each rectangle~$r$ of~$R$ is below (resp.~above) the antidiagonal,
\item the source tree~$S(R)$ (resp.~target tree~$T(R)$) remains below (resp.~above) the antidiagonal,
\item the source tree~$S(R)$ and the target tree~$T(R)$ do not overlap along a segment of~$R$,
\item $R$ avoids the patterns \horizontalPattern{} and \verticalPattern{},
\item the $2$-clumped permutation~$\projDown(R)$ avoids the patterns~$2413$ and~$3412$,
\item the co-$2$-clumped permutation~$\projDown(R)$ avoids the patterns~$2143$ and~$3142$,
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Quotientopes}
\label{sec:quotientopes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Weak rectangulation polytope}
\label{sec:weakRectangulationPolytope}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Strong rectangulation polytope}
\label{sec:weakRectangulationPolytope}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\clearpage
\addtocontents{toc}{ \vspace{.1cm} }
\bibliographystyle{alpha}
\bibliography{rectangulotopes}
\label{sec:biblio}

\end{document}

### 2024-03-12 draft/rectangulotopes.tex
formulas diagonal insertion

@@
 
+%For a permutation~$\sigma$ of~$[n]$ and~$i \in [n]$, we define
+%\begin{alignat*}{2}
+%pv(i, \sigma) & \eqdef p(i) \cap v(i, \sigma), &
+%pV(i, \sigma) & \eqdef p(i) \cap V(i, \sigma), \\
+%Pv(i, \sigma) & \eqdef P(i) \cap v(i, \sigma) & \qquad\text{and}\qquad
+%PV(i, \sigma) & \eqdef P(i) \cap V(i, \sigma).
+%\end{alignat*}
+%where~$p(i) \eqdef [i-1]$ while~$P(i) \eqdef [n] \ssm [i]$, and~$v(i, \sigma) \eqdef \sigma([i-1])$ while $V(i, \sigma) \eqdef \sigma([n] \ssm [i])$.
+
 %%%%%%%%%%%%
@@
 	\caption{A weak rectangulation (left) and its pair of separated twin trees (right).
-The corresponding cotwisted Baxter permutation is $[7,14,15,16,8,11,13,10,5,6,1,4,9,2,3,12]$ and its inverse is $[11, 14, 15, 12, 9, 10, 1, 5, 13, 8, 6, 16, 7, 2, 3, 4]$.}
+The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.}
 \end{figure}
 
-We call \defn{antidiagonal} the segment joining the top left corner to the bottom right corner.
+We call \defn{diagonal} the segment joining the top left corner to the bottom right corner.
 A rectangulation~$R$ is \defn{diagonal} (or \defn{weak}) if it satisfies the following equivalent conditions:
 \begin{enumerate}[(i)]
-\item all rectangles of~$R$ meet the antidiagonal,
-\item the source~$s(r)$ (resp.~target~$t(r)$) of each rectangle~$r$ of~$R$ is below (resp.~above) the antidiagonal,
-\item the source tree~$S(R)$ (resp.~target tree~$T(R)$) remains below (resp.~above) the antidiagonal,
+\item all rectangles of~$R$ meet the diagonal,
+\item the source~$s(r)$ (resp.~target~$t(r)$) of each rectangle~$r$ of~$R$ is below (resp.~above) the diagonal,
+\item the source tree~$S(R)$ (resp.~target tree~$T(R)$) remains below (resp.~above) the diagonal,
 \item the source tree~$S(R)$ and the target tree~$T(R)$ do not overlap along a segment of~$R$,
@@
 \end{enumerate}
+
+\begin{proposition}
+The weak rectangulation insertion of any permutation~$\sigma$ of~$[n]$ is given by the rectangles~$[w\ell(i, \sigma), wr(i, \sigma)-1] \times [n-wd(i, \sigma)+1, n-wu(i, \sigma)]$ for~$i \in [n]$ where
+\begin{align*}
+w\ell(i, \sigma) & \eqdef \max \; \{0\} \cup \set{j}{j < i \text{ and } \sigma^{-1}(j) < \sigma^{-1}(i)} \\
+wr(i, \sigma) & \eqdef \min \; \{n+1\} \cup \set{j}{i < j \text{ and } \sigma^{-1}(i) < \sigma^{-1}(j)} \\
+wd(i, \sigma) & \eqdef \min \; \{n+1\} \cup \set{j}{i < j \text{ and } \sigma^{-1}(j) < \sigma^{-1}(i)} \\
+\text{and} \qquad
+wu(i, \sigma) & \eqdef \max \; \{0\} \cup \set{j}{j < i \text{ and } \sigma^{-1}(i) < \sigma^{-1}(j)}.
+\end{align*}
+\end{proposition}
 

### 2024-03-20 draft/rectangulotopes.tex
essai
comment
programmation + files rectangulations
tentative plan and statements

@@
-\documentclass{amsart}
+\documentclass[12pt]{amsart}
 
 \usepackage[T1]{fontenc}
-\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont}
+\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont, verbatim}
 \usepackage{marginnote}
@@
 \newcommand{\h}{\widehat} % hat letters
+
+% hello
 
@@
 \newcommand{\Asso}{\polytope{A}\mathsf{sso}} % associahedron
-\newcommand{\WRP}{\polytope{WRP}} % weak rectangulation polytope
-\newcommand{\SRP}{\polytope{SRP}} % strong rectangulation polytope
+\newcommand{\WRP}{\polytope{WR}} % weak rectangulation polytope
+\newcommand{\SRP}{\polytope{SR}} % strong rectangulation polytope
 \newcommand{\horizontalPattern}{\smash{\raisebox{-.15cm}{\includegraphics[scale=.5]{horizontalPattern}}}} % horizontal pattern
@@
 \author{Jean Cardinal}
-\address{Université Libre de Bruxelles}
+\address{Université libre de Bruxelles (ULB)}
 \email{jean.cardinal@ulb.be}
@@
 \begin{abstract}
+  Rectangulations are decompositions of a square into finitely many axis-aligned rectangles.
+  We give simple realizations of $(n-1)$-dimensional polytopes associated with rectangulations composed of $n$ rectangles.
+They correspond to quotientopes of natural congruence relations on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
+We give  elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedra.
 \end{abstract}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Polytopes of permutations, triangulations, and rectangulations}
 
-\newpage
-\section{Rectangulations}
-\label{sec:rectangulations}
+\begin{itemize}
+\item Permutohedra
+\item Associahedra
+\item Generalized permutohedra
+\item Quotientopes
+\item Quotientopes of rectangulations, natural successors to associahedra; Definitions
+\item Flip graphs on rectangulations
+\item Loday popularized a simple realization of the associahedron
+\item We give a Loday-style description of two classes of rectangulation polytopes
+\item This intro states the main results; proofs are in the following sections.
+\end{itemize}
 
-%%%%%%%%%%%%
+\subsection{permutohedra}
 
-\subsection{Combinatorics of (generic) rectangulations}
-\label{subsec:rectangulations}
+\begin{theorem}
+  Let $\polytope{P} (n)$ be the convex hull of the points
+  $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, 
+  for all permutations $\pi\in\f{S}_n$.
+  Then $\polytope{P} (n)$ is a realization of the $(n-1)$-dimensional permutohedron.
+  \end{theorem}
 
-\Vincent{
-Note that there seem to be a canonical integer drawing of~$R$ described as follows.
-For vertical segment~$v$ of~$R$, consider the path~$p$ passing through~$v$ with maximal slope.
-The abscissa of~$v$ is then the number of rectangle on the left of~$p$ plus the number of \horizontalPattern{} patterns along~$p$ before~$v$ minus the number of \horizontalPattern{} patterns along~$p$ after~$v$.
-The ordinate of an horizontal segment~$h$ of~$R$ is defined symmetrically.
-}
+\subsection{Loday's associahedra}
 
-%%%%%%%%%%%%
+In 2004, Loday published the following elegant realization of the ubiquitous associahedron~\cite{}.
+
+\begin{theorem}
+  Let $\Asso (n)$ be the convex hull of the points
+  $\sum_{i\in [n]} w_i\cdot \mathbf{e}_i$, where
+  \[
+  w_i := \ell(T, i)\cdot r(T,i),
+  \]
+  for all binary trees $T$ on $n$ internal nodes, where $\ell(T,i)$ and $r(T,i)$ denote respectively the number of leaves in the left and right subtrees of $i$ in $T$.
+  Then $\Asso (n)$ is a realization of the $(n-1)$-dimensional associahedron.
+  \end{theorem}
+
+\subsection{Weak and strong rectangulations}
+
+\begin{itemize}
+\item Terminology: Rectangles, vertices, segments
+\item Wall slides
+\item Two equivalence classes: weak equivalence preserving rectangle-segment adjacencies, and strong equivalence preserving rectangle-rectangle adjacencies.
+\item Weak equivalence classes have a \emph{diagonal} representative.
+\end{itemize}
 
@@
 
+\jean{Here only give the definitions allowing to state the results.}
+
 \begin{figure}
 	\centerline{\includegraphics[scale=.7]{strongRectangulation} \qquad \includegraphics[scale=.7]{strongRectangulationTrees}}
-	\caption{A strong rectangulation (left) and its pair of intertwining twin trees (right).
-The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
-Example from \cite{AsinowskiCardinalFelsnerFusy}.}
+	\caption{\label{fig:sourceTargetTrees}A strong rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{AsinowskiCardinalFelsnerFusy}.}
+        % The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
 \end{figure}
@@
 Fix a rectangulation~$R$ with $n$ rectangles, and consider the directed graph~$D(R)$ whose vertex set is the set of all vertices of~$R$, and whose edges are obtained as follows.
-For each rectangle~$r$ of~$R$, we include in~$D(R)$ four edges joining the vertices of~$r$, where the horizontal edges are oriented from left to right if they are horizontal, and the vertical edges are oriented from bottom to top if they are vertical.
+For each rectangle~$r$ of~$R$, we include in~$D(R)$ four edges joining the vertices of~$r$, where the horizontal edges are oriented from left to right, and the vertical edges are oriented from bottom to top.
 Hence, the bottom left corner of~$r$ is its source~$s(r)$, while its top right corner is its target~$t(r)$.
@@
 
+\begin{comment}
 \begin{proof}
@@
 \end{proof}
+\end{comment}
 
-We complete each node of~$S(R)$ and~$T(R)$ with a vertical (resp.~horizontal) leaf if it has no vertical (resp.~horizontal) child.
-See \cref{fig:sourceTargetTrees}.
+We complete each node of~$S(R)$ and~$T(R)$ with a vertical (resp.~horizontal) leaf if it has no vertical (resp.~horizontal) child, see \cref{fig:sourceTargetTrees}.
 Recall that the \defn{inorder labeling} of a binary tree~$T$ is the labeling of the nodes of~$T$ such that the label of each node~$t$ of~$T$ is larger than all labels in the left subtree of~$t$ and larger than all labels in the right subtree of~$t$.
@@
 
-\begin{proof}
-\end{proof}
-
 This enables to unambiguously label the rectangles of~$R$ by the inorder 
 The resulting labeling coincides with the NW--SE labeling of~\cite{AsinowskiCardinalFelsnerFusy}.
-We prefer our presentation with~$S(R)$ and~$T(R)$ here as we will actually use these two binary trees.
+
+\subsection{Weak rectangulation polytopes}
+
+We are now ready to give our first main result on the realizations of weak rectangulation polytopes.
+
+\begin{theorem}
+  Let $\WRP (n)$ be the convex hull of the points
+  $\sum_{i\in [n]} w_i\cdot \mathbf{e}_i$,
+  where
+  \[
+  \begin{split}
+    w_i := \quad & h(T(R), i)\cdot v(T(R),i) \\
+           \quad & - h(S(R), i)\cdot v(S(R),i),
+  \end{split}
+  \]
+  for all weak rectangulations $R$ on $n$ rectangles, where $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
+  Then $\WRP (n)$ is a realization of the $(n-1)$-dimensional weak rectangulation polytope.
+\end{theorem}
+
+\subsection{Strong rectangulation polytopes}
+
+Our second main result on the realizations of the strong rectangulation polytopes does not require any additional definition. It is, however, slightly more involved.
+
+\begin{theorem}
+  Let $\SRP (n)$ be the convex hull of the points
+  \[
+  \sum_{i,j\in [n]} w_{i,j}\cdot (\mathbf{e}_i - \mathbf{e}_j),
+  \]
+  with
+  \[
+  \begin{split}
+  w_{i,j} := \quad & h(T(R), i) \cdot cv (T(R), S(R), i, j)\cdot h(S(R), j) \\
+            \quad & - v(S(R), j) \cdot ch (T(R), S(R), i, j)\cdot v(T(R), i)
+  \end{split}
+  \]
+  \jean{Write the correct formula here, and sum only on the suitable pairs $i,j$.}
+  for all strong rectangulations $R$ on $n$ rectangles, where:
+  \begin{itemize}
+  \item $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$,
+  \item $ch (T,S,i,j)$ and $cv (T,S,i,j)$ denote respectively the number of \emph{common leaves} to the horizontal (vertical) subtree of $i$ in $T$ and the horizontal (vertical) subtree of $j$ in $S$
+  \end{itemize}
+Then $\SRP (n)$ is a realization of the $(n-1)$-dimensional strong rectangulation polytope.  
+\end{theorem}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Quotientopes}
+\label{sec:quotientopes}
+
+\subsection{Quotients of the weak Bruhat order on permutations}
+
+\subsection{Arc diagrams}
+
+\jean{State the main results that we will use. I just give sloppy statements for now.}
+
+\begin{theorem}
+  Congruences on the weak Bruhat order are one-to-one with arc ideals.
+\end{theorem}
+
+\subsection{Shard polytopes}
+
+\begin{theorem}
+  The quotientope of a congruence $\equiv$ on the weak Bruhat order is realized by the Minkowski sum of the shard polytopes corresponding to the arcs in the arc ideal defining $\equiv$.
+\end{theorem}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Weak rectangulation polytopes}
+\label{sec:weakRectangulationPolytopes}
+
+%%%%%%%%%%%%
+\subsection{The Baxter congruence}
+
+Define the weak poset $P_w(R)$ of a rectangulation $R$ as the intersection of two partial orders defined by the twin binary trees
+
+\begin{theorem}
+  The following subsets of $\f{S}_n$ 
+  \[
+  \{\mathcal{L}(P_w(R)) : R \text{ is a weak rectangulation} \}
+  \]
+  are equivalence classes of a congruence relation on the weak Bruhat order on $\f{S}_n$.
+  The minimal elements of each equivalence classes are twisted Baxter permutations, and the maximal
+  elements are co-twisted Baxter permutations.
+\end{theorem}
+
+\begin{figure}
+	\centerline{\includegraphics[scale=.7]{weakRectangulation} \qquad \includegraphics[scale=.7]{weakRectangulationTrees}}
+	\caption{A diagonal rectangulation (left) and its pair of separated twin trees (right).
+The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.}
+\end{figure}
+
+We call this congruence the \emph{Baxter congruence}.
+The weak rectangulation polytope $\WRP(n)$ is the quotientope of the Baxter congruence of the weak Bruhat order on permutations.
+
+\subsection{Loday and anti-Loday arcs}
+
+The Baxter congruence is defined by forbidding arcs that cross the horizontal line.
+Therefore, the quotientope is defined by a Minkowski sum of the shard polytopes defined by two families of arcs: the Loday and the anti-Loday arcs.
+
+\subsection{Loday-type coordinates for weak rectangulation polytopes}
+
+We now prove our first main result.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Strong rectangulation polytopes}
+\label{sec:strongRectangulationPolytopes}
+
+%%%%%%%%%%%%
+\subsection{The strong rectangulation congruence}
+
+Define the strong poset $P_s(R)$ of a rectangulation $R$ using the intertwined binary trees.
+
+\begin{theorem}
+  The following subsets of $\f{S}_n$ 
+  \[
+  \{\mathcal{L}(P_s(R)) : R \text{ is a strong rectangulation} \}
+  \]
+  are equivalence classes of a congruence relation on the weak Bruhat order on $\f{S}_n$.
+  The minimal elements of each equivalence classes are 2-clumped permutations, and the maximal
+  elements are co-2-clumped permutations.
+\end{theorem}
+
+The strong rectangulation polytope $\SRP(n)$ is the quotientope of the congruence of the weak Bruhat order on permutations that is defined by the 2-clumped permutations.
+
+\jean{Does the congruence have a little name?}
+
+\subsection{Yin and yang arcs}
+
+The congruence is defined by forbidding the arcs that cross the horizontal line more than once.
+The quotientope is therefore the Minkowski sum of the shard polytopes corresponding to four families of arcs:
+\begin{itemize}
+\item the Loday and anti-Loday arcs, like in the weak case,
+\item the Yin and yang arcs.
+\end{itemize}
+
+\subsection{Shard polytopes of yin and yang arcs}
+
+While the shard polytopes of Loday and anti-Loday arcs are simplices, the shard polytopes of the yin and yang arcs are pyramids.
+
+\subsection{Loday-type coordinates for strong rectangulation polytopes}
+
+Proof of our second main result:
+\begin{itemize}
+\item Identify what are the vertices of each yin summands that contribute to a given rectangulation,
+\item group the contributions for each pair of rectangles $i,j$,
+\item prove that the count is the triple product of numbers of leaves of the trees,
+\item argue about what pairs $i,j$ are to be considered,
+\item argue symmetrically about yang arcs and conclude.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\clearpage
+\addtocontents{toc}{ \vspace{.1cm} }
+\bibliographystyle{alpha}
+\bibliography{rectangulotopes}
+\label{sec:biblio}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%
+\newpage
+\appendix
+\section{Material}
+
+Properties/results we may need along the way.
+
+\subsection{Twin binary trees.}
 
@@
 \item for any~$i \in [n-1]$, the node~$i$ is in the left subtree of the node~$i+1$ in~$S$ if and only if the node~$i+1$ is in the right subtree of the node~$i$ in~$R$,
-\item for any~$i \in [n-1]$, the $(i+1)$-st leaf of~$S$ is a left leaf if and only if the $(i+1)$-st leaf of~$R$ is a right leaf.
+\item for any~$i \in [n-1]$, the $(i+1)$th leaf of~$S$ is a left leaf if and only if the $(i+1)$th leaf of~$R$ is a right leaf.
 \end{enumerate}
@@
 
-\begin{proof}
-\end{proof}
-
-%%%%%%%%%%%%
-
-\subsection{Rectangulation insertion of permutations}
-\label{subsec:insertion}
-
-\Vincent{If my canonical drawing indeed makes sense, I would love to describe the insertion map directly on rectangles.}
-
-\Vincent{Describe the strong poset here?}
-
-The insertion fiber of a rectangulation~$R$ precisely consists in the linear extensions of the strong poset.
-
-%For a permutation~$\sigma$ of~$[n]$ and~$i \in [n]$, we define
-%\begin{alignat*}{2}
-%pv(i, \sigma) & \eqdef p(i) \cap v(i, \sigma), &
-%pV(i, \sigma) & \eqdef p(i) \cap V(i, \sigma), \\
-%Pv(i, \sigma) & \eqdef P(i) \cap v(i, \sigma) & \qquad\text{and}\qquad
-%PV(i, \sigma) & \eqdef P(i) \cap V(i, \sigma).
-%\end{alignat*}
-%where~$p(i) \eqdef [i-1]$ while~$P(i) \eqdef [n] \ssm [i]$, and~$v(i, \sigma) \eqdef \sigma([i-1])$ while $V(i, \sigma) \eqdef \sigma([n] \ssm [i])$.
-
-%%%%%%%%%%%%
-
-\subsection{Diagonal rectangulations}
-\label{subsec:diagonalRectangulations}
-
-\begin{figure}
-	\centerline{\includegraphics[scale=.7]{weakRectangulation} \qquad \includegraphics[scale=.7]{weakRectangulationTrees}}
-	\caption{A weak rectangulation (left) and its pair of separated twin trees (right).
-The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.}
-\end{figure}
+\subsection{Properties of diagonal rectangulations.}
 
 We call \defn{diagonal} the segment joining the top left corner to the bottom right corner.
-A rectangulation~$R$ is \defn{diagonal} (or \defn{weak}) if it satisfies the following equivalent conditions:
+A rectangulation~$R$ is \defn{diagonal} if it satisfies the following equivalent conditions:
 \begin{enumerate}[(i)]
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\Vincent{
+Note that there seem to be a canonical integer drawing of~$R$ described as follows.
+For vertical segment~$v$ of~$R$, consider the path~$p$ passing through~$v$ with maximal slope.
+The abscissa of~$v$ is then the number of rectangle on the left of~$p$ plus the number of \horizontalPattern{} patterns along~$p$ before~$v$ minus the number of \horizontalPattern{} patterns along~$p$ after~$v$.
+The ordinate of an horizontal segment~$h$ of~$R$ is defined symmetrically.
+}
 
-\section{Quotientopes}
-\label{sec:quotientopes}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Weak rectangulation polytope}
-\label{sec:weakRectangulationPolytope}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Strong rectangulation polytope}
-\label{sec:weakRectangulationPolytope}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\clearpage
-\addtocontents{toc}{ \vspace{.1cm} }
-\bibliographystyle{alpha}
-\bibliography{rectangulotopes}
-\label{sec:biblio}
+\begin{comment}
+\newpage
+\noindent
+\input{figures/rectangulations2.tikz}
+\vspace{1cm}
+\noindent
+\input{figures/rectangulations3.tikz}
+\vspace{1cm}
+\noindent
+\input{figures/rectangulations4.tikz}
+\newpage
+\noindent
+\input{figures/rectangulations5.tikz}
+\newpage
+\noindent
+\input{figures/rectangulations6.tikz}
+\clearpage
+\end{comment}
 

### 2024-03-20 draft/rectangulotopes.tex
additional statements

@@
-\documentclass[12pt]{amsart}
+\documentclass{amsart}
 
@@
 
+% shard polytopes
+\newcommand{\SP}{\polytope{SP}}
+
+% Loday and anti-Loday accents (?)
+\newcommand{\loday}[1]{\overset{\frown}{#1}}
+\newcommand{\antiloday}[1]{\overset{\smile}{#1}}
+% yin and yang accents (?)
+\newcommand{\yang}[1]{\overset{\backsim}{#1}}
+\newcommand{\yin}[1]{\overset{\sim}{#1}}
+
 % formating the table of contents
@@
   Rectangulations are decompositions of a square into finitely many axis-aligned rectangles.
-  We give simple realizations of $(n-1)$-dimensional polytopes associated with rectangulations composed of $n$ rectangles.
-They correspond to quotientopes of natural congruence relations on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
+  We describe realizations of $(n-1)$-dimensional polytopes associated with rectangulations composed of $n$ rectangles.
+They are defined as quotientopes of natural congruence relations on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
 We give  elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedra.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
 
@@
 \begin{itemize}
-\item Terminology: Rectangles, vertices, segments
+\item Terminology: rectangles, vertices, segments, generic rectangulation
 \item Wall slides
-\item Two equivalence classes: weak equivalence preserving rectangle-segment adjacencies, and strong equivalence preserving rectangle-rectangle adjacencies.
-\item Weak equivalence classes have a \emph{diagonal} representative.
+\item Two equivalence classes: \defn{weak equivalence} preserving rectangle-segment adjacencies, and \defn{strong equivalence} preserving rectangle-rectangle adjacencies.
+\item Weak equivalence classes have a \defn{diagonal} representative.
 \end{itemize}
@@
   Let $\WRP (n)$ be the convex hull of the points
-  $\sum_{i\in [n]} w_i\cdot \mathbf{e}_i$,
+  $\sum_{i\in [n]} (\loday{w}_i - \antiloday{w}_i)\cdot \mathbf{e}_i$,
   where
@@
   \begin{split}
-    w_i := \quad & h(T(R), i)\cdot v(T(R),i) \\
-           \quad & - h(S(R), i)\cdot v(S(R),i),
+    \loday{w}_i := & h(T(R), i)\cdot v(T(R),i) \\
+    \antiloday{w}_i := & h(S(R), i)\cdot v(S(R),i),
   \end{split}
@@
 
-Our second main result on the realizations of the strong rectangulation polytopes does not require any additional definition. It is, however, slightly more involved.
+Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if they are both either horizontal or vertical, and the edges to their parents lie on the same segment of $R$.
+We also denote by $i|j$ and $\frac{\ j\ }i$ the situations where the rectangle labeled $i$ touches the rectangle labeled $j$ from the left and from below in $R$, respectively.
 
@@
   \[
-  \sum_{i,j\in [n]} w_{i,j}\cdot (\mathbf{e}_i - \mathbf{e}_j),
+  \sum_{i,j\in [n], i< j} (\yin{w}_{i,j} - \yang{w}_{i,j})\cdot (\mathbf{e}_i - \mathbf{e}_j),
   \]
@@
   \[
-  \begin{split}
-  w_{i,j} := \quad & h(T(R), i) \cdot cv (T(R), S(R), i, j)\cdot h(S(R), j) \\
-            \quad & - v(S(R), j) \cdot ch (T(R), S(R), i, j)\cdot v(T(R), i)
-  \end{split}
+  \yin{w}_{i,j} := 
+  \begin{cases}
+    h(T(R), i) \cdot cv (T(R), S(R), i, j)\cdot h(S(R), j) & \text{ if neither } i|j \text{ nor } \frac{\ j\ }i \\
+    0 & \text{ otherwise }
+  \end{cases}
   \]
-  \jean{Write the correct formula here, and sum only on the suitable pairs $i,j$.}
+  and
+    \[
+  \yang{w}_{i,j} := 
+  \begin{cases}
+    v(S(R), j) \cdot ch (T(R), S(R), i, j)\cdot v(T(R), i) & \text{ if neither } j|i \text{ nor } \frac{\ i\ }j \\
+    0 & \text{ otherwise }
+  \end{cases}
+  \]
+
+  \jean{Double check the formula.}
   for all strong rectangulations $R$ on $n$ rectangles, where:
@@
   \item $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$,
-  \item $ch (T,S,i,j)$ and $cv (T,S,i,j)$ denote respectively the number of \emph{common leaves} to the horizontal (vertical) subtree of $i$ in $T$ and the horizontal (vertical) subtree of $j$ in $S$
+  \item $ch (T,S,i,j)$ and $cv (T,S,i,j)$ denote respectively the number of common leaves of the horizontal (vertical) subtree of $i$ in $T$ and the horizontal (vertical) subtree of $j$ in $S$
   \end{itemize}
@@
 \end{theorem}
+
+\Jean{Is there a more concise way to write this down/compute? Here we do not take advantage of the fact that at most one of the two factors $yin{w}_{i,j}$ and $\yang{w}_{i,j}$ is nonzero.}
 
@@
 \subsection{Arc diagrams}
-
-\jean{State the main results that we will use. I just give sloppy statements for now.}
 
@@
 
-Define the weak poset $P_w(R)$ of a rectangulation $R$ as the intersection of two partial orders defined by the twin binary trees
+Define the weak poset $P_w(R)$ of a rectangulation $R$ as the union of two partial orders defined by the twin binary trees
 
 \begin{theorem}
-  The following subsets of $\f{S}_n$ 
-  \[
-  \{\mathcal{L}(P_w(R)) : R \text{ is a weak rectangulation} \}
-  \]
+  The subsets $\{\mathcal{L}(P_w(R)) : R \text{ is a rectangulation} \}$ of $\f{S}_n$ 
   are equivalence classes of a congruence relation on the weak Bruhat order on $\f{S}_n$.
@@
 We call this congruence the \emph{Baxter congruence}.
-The weak rectangulation polytope $\WRP(n)$ is the quotientope of the Baxter congruence of the weak Bruhat order on permutations.
+The weak rectangulation polytope $\WRP(n)$ is the quotientope of the Baxter congruence of the weak Bruhat order on $\f{S}_n$.
 
@@
 
-Define the strong poset $P_s(R)$ of a rectangulation $R$ using the intertwined binary trees.
+Define the strong poset $P_s(R)$ of a rectangulation $R$. Can we use the intertwined binary trees?
 
 \begin{theorem}
-  The following subsets of $\f{S}_n$ 
-  \[
-  \{\mathcal{L}(P_s(R)) : R \text{ is a strong rectangulation} \}
-  \]
+  The subsets $\{\mathcal{L}(P_s(R)) : R \text{ is a rectangulation} \}$ of $\f{S}_n$ 
   are equivalence classes of a congruence relation on the weak Bruhat order on $\f{S}_n$.
@@
 
-The strong rectangulation polytope $\SRP(n)$ is the quotientope of the congruence of the weak Bruhat order on permutations that is defined by the 2-clumped permutations.
+The strong rectangulation polytope $\SRP(n)$ is the quotientope of the congruence of the weak Bruhat order on $\f{S}_n$ that is defined by the 2-clumped permutations.
 
-\jean{Does the congruence have a little name?}
+\jean{Does the congruence have a name?}
 
@@
 \item the Loday and anti-Loday arcs, like in the weak case,
-\item the Yin and yang arcs.
+\item the yin and yang arcs.
 \end{itemize}
@@
 \begin{itemize}
-\item Identify what are the vertices of each yin summands that contribute to a given rectangulation,
-\item group the contributions for each pair of rectangles $i,j$,
+\item Identify what are the yin summands contributing to a given pair $\mathbf{e}_i - \mathbf{e}_j$,
 \item prove that the count is the triple product of numbers of leaves of the trees,
@@
 \end{itemize}
+
+\begin{lemma}
+  \label{lem:yinminmax}
+  Let $\pi\in\f{S}_n$ and consider a yin arc $A$ composed of two intervals $I$ and $J$.
+  Let $i := \arg\max_{k\in I} \pi^{-1}(k)$ and $j := \arg\min_{k\in J} \pi^{-1}(k)$.
+  The vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
+  is $\mathbf{e}_i-\mathbf{e}_j$ if $\pi^{-1}(i)>\pi^{-1}(j)$, and $0$ otherwise.
+\end{lemma}
+
+\begin{lemma}
+  Let $R$ be a rectangulation on $n$ rectangles and $p(R)$ the vertex of $\SRP(n)$ associated with $R$.
+  Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $A$ such that $\mathbf{e}_i-\mathbf{e}_j$
+  is the extremal vertex of $\SP(A)$ contributing to $p(R)$ is
+  \[
+  \yin{w}_{i,j} = 
+  \begin{cases}
+    h(T(R), i) \cdot cv (T(R), S(R), i, j)\cdot h(S(R), j) & \text{ if neither } i|j \text{ nor } \frac{\ j\ }i \\
+    0 & \text{ otherwise }
+  \end{cases}
+  \]
+\end{lemma}
+\begin{proof}
+  Let $\pi$ be the 2-clumped permutation corresponding to $R$, hence the minimal element of $\mathcal{L}(P_s(R))$ in the weak Bruhat order.
+  From Lemma~\ref{lem:yinminmax}, the set of yin arcs $A$ such that $\SP(A)$ contributes to $\mathbf{e}_i-\mathbf{e}_j$
+  are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that:
+  \begin{itemize}
+  \item $i = \arg\max_{k\in I} \pi^{-1}(k)$,
+  \item $j = \arg\min_{k\in J} \pi^{-1}(k)$, and
+  \item $i,j$ is an inversion in $\pi$, hence $\pi^{-1}(i)>\pi^{-1}(j)$.
+  \end{itemize}
+  We observe that provided the third condition holds, the number of choices of the three endpoints defining such a pair of intervals $I$ and $J$ is the product of the
+  number of leaves of the subtree of $T(R)$ rooted at $i$, the number of common leaves to the vertical subtrees of $T(R)$ and $S(R)$ rooted
+  at $i$ and $j$, respectively, and the number of leaves of the subtree of $S(R)$ rooted at $j$.
+
+  It remains to observe that the only case where this product is nonzero and $i,j$ is not an inversion in $\pi$ is when either $i|j$ or $\frac{\ j\ }i$.
+\end{proof}
 

### 2024-03-20 draft/rectangulotopes.tex
additional statements

@@
 
-Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if they are both either horizontal or vertical, and the edges to their parents lie on the same segment of $R$.
-We also denote by $i|j$ and $\frac{\ j\ }i$ the situations where the rectangle labeled $i$ touches the rectangle labeled $j$ from the left and from below in $R$, respectively.
+Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$.
+We denote by $i|j$ and $\frac{\ j\ }i$ the situations where the rectangle labeled $i$ touches the rectangle labeled $j$ from the left and from below in $R$, respectively.
 

### 2024-03-20 draft/rectangulotopes.tex
additional statements

@@
 By symmetry, we only prove the statement for the source tree~$S(R)$.
-As the triangulation is generic, each source~$s(r)$ distinct from the bottom left corner of~$R$ is either on the top edge of a rectangle~$r'$ or on the right edge of a rectangle~$r'$ (but not both).
+As the rectangulation is generic, each source~$s(r)$ distinct from the bottom left corner of~$R$ is either on the top edge of a rectangle~$r'$ or on the right edge of a rectangle~$r'$ (but not both).
 This shows that~$r$ distinct from the bottom left rectangle of~$R$ has a unique parent~$r'$ in~$S(R)$, so that~$S(R)$ is indeed a tree on~$[n]$.
@@
 The resulting labeling coincides with the NW--SE labeling of~\cite{AsinowskiCardinalFelsnerFusy}.
+From now on, the labels of the rectangles in a rectangulations are the inorder labels, and the two trees $T(R)$ and $S(R)$ are defined on the vertex set $[n]$ accordingly.
+For each $i\in [n]$, we refer to its two subtrees in $T(R)$ as the \defn{horizontal} and \defn{vertical subtrees}, and similarly for $S(R)$.
 
@@
   \begin{cases}
-    h(T(R), i) \cdot cv (T(R), S(R), i, j)\cdot h(S(R), j) & \text{ if neither } i|j \text{ nor } \frac{\ j\ }i \\
-    0 & \text{ otherwise }
+    h(T(R), i) \cdot cv (T(R), S(R), i, j)\cdot h(S(R), j) & \text{ if neither } i|j \text{ nor } \frac{\ j\ }i, \\
+    0 & \text{ otherwise.}
   \end{cases}
@@
   \begin{cases}
-    v(S(R), j) \cdot ch (T(R), S(R), i, j)\cdot v(T(R), i) & \text{ if neither } j|i \text{ nor } \frac{\ i\ }j \\
-    0 & \text{ otherwise }
+    v(S(R), j) \cdot ch (T(R), S(R), i, j)\cdot v(T(R), i) & \text{ if neither } j|i \text{ nor } \frac{\ i\ }j, \\
+    0 & \text{ otherwise.}
   \end{cases}
@@
 
-Define the weak poset $P_w(R)$ of a rectangulation $R$ as the union of two partial orders defined by the twin binary trees
+Given a rectangulation $R$, the two trees $T(R)$ and $S(R)$, where the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top,
+define two partial orders on $[n]$.
+We define the \defn{weak poset} $P_w(R)=([n],\prec_w)$ of a rectangulation $R$ as the transitive closure of the union of the two partial orders defined by the intertwining binary trees $T(R)$ and $S(R)$.
 
 \begin{theorem}
-  The subsets $\{\mathcal{L}(P_w(R)) : R \text{ is a rectangulation} \}$ of $\f{S}_n$ 
+  The subsets $\mathcal{L}(P_w(R))$ of $\f{S}_n$, for all rectangulations $R$ on $n$ rectangles, 
   are equivalence classes of a congruence relation on the weak Bruhat order on $\f{S}_n$.
@@
 
-We call this congruence the \emph{Baxter congruence}.
+We call this congruence the \defn{Baxter congruence}.
 The weak rectangulation polytope $\WRP(n)$ is the quotientope of the Baxter congruence of the weak Bruhat order on $\f{S}_n$.
@@
 The Baxter congruence is defined by forbidding arcs that cross the horizontal line.
-Therefore, the quotientope is defined by a Minkowski sum of the shard polytopes defined by two families of arcs: the Loday and the anti-Loday arcs.
+Therefore, the quotientope is defined by a Minkowski sum of the shard polytopes defined by two families of arcs: the \defn{Loday arcs} and the \defn{anti-Loday arcs}, that go respectively above and below all elements of an interval $I$ of $[n]$.
+
+\begin{lemma}
+  Given a Loday arc $A$ defined by an interval $I$, the Loday shard polytope $\SP(A)$ is defined as the standard simplex $\conv \{ \mathbf{e}_i : i\in I \}$.
+\end{lemma}
+
+\begin{lemma}
+  Given an anti-Loday arc $A$ defined by an interval $I$, the anti-Loday shard polytope $\SP(A)$ is defined as the negative standard simplex $\conv \{ - \mathbf{e}_i : i\in I \}$.
+\end{lemma}
+
+\begin{lemma}
+  The weak rectangulation polytope $\WRP(n)$ is realized by the Minkowski sum of all Loday and anti-Loday shard polytopes. In particular, it is the Minkowski sum of two opposite Loday associahedra.
+\end{lemma}
 
@@
 
-We now prove our first main result.
+\begin{lemma}
+  Let $R$ be a rectangulation on $n$ rectangles and $p(R)$ the vertex of $\WRP(n)$ associated with $R$.
+  Given $i\in [n]$, the number of Loday arcs $A$ such that $\mathbf{e}_i$ is the extremal vertex of $\SP(A)$ contributing to $p(R)$ is
+  \[
+  \loday{w}_i =  h(T(R), i)\cdot v(T(R),i),
+  \]
+   where $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
+\end{lemma}
 
@@
 
-Define the strong poset $P_s(R)$ of a rectangulation $R$. Can we use the intertwined binary trees?
+We define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of a rectangulation $R$.
+For a poset $P$ on $[n]$, we define $\mathcal{L}(P)$ as the set of permutations of $\f{S}_n$ corresponding to linear extensions of $P$.
 
 \begin{theorem}
-  The subsets $\{\mathcal{L}(P_s(R)) : R \text{ is a rectangulation} \}$ of $\f{S}_n$ 
+  The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ on $n$ rectangles, 
   are equivalence classes of a congruence relation on the weak Bruhat order on $\f{S}_n$.
@@
 
-The congruence is defined by forbidding the arcs that cross the horizontal line more than once.
-The quotientope is therefore the Minkowski sum of the shard polytopes corresponding to four families of arcs:
-\begin{itemize}
-\item the Loday and anti-Loday arcs, like in the weak case,
-\item the yin and yang arcs.
-\end{itemize}
+The strong rectangulation congruence is defined by forbidding the arcs that cross the horizontal line more than once.
+We classify the remaining arcs into two families, depending on whether the arcs cross from above or below.
 
-\subsection{Shard polytopes of yin and yang arcs}
+A \defn{yin arc} is defined by two consecutive nonempty intervals $I$ and $J$ of $[n]$ such that the arc goes above all elements of $I$ and
+below all elements of $J$. Note that the Loday arc are obtained when $|J|=1$ and the anti-Loday arc are obtained with $|I|=1$.
+A \defn{yang arc} is defined similarly, except that the arc goes below the elements of $I$ and above the elements of $J$. Note that Loday and anti-Loday arcs are also yang arcs.
 
-While the shard polytopes of Loday and anti-Loday arcs are simplices, the shard polytopes of the yin and yang arcs are pyramids.
+\begin{lemma}
+  Given a yin arc $A=(I,J)$, the shard polytope $\SP(A)$ is defined as the convex hull of the points
+  $\{\mathbf{0} \}\cup \{ \mathbf{e}_i - \mathbf{e}_j : (i,j)\in I\times J\}$.
+  %  \SP(A) = \conv (\{\mathbf{0} \}\cup \{ \mathbf{e}_i - \mathbf{e}_j : (i,j)\in I\times J\})
+\end{lemma}
+
+\begin{lemma}
+  Given a yang arc $A=(I,J)$, the shard polytope $\SP(A)$ is defined as the convex hull of the points
+  $\{\mathbf{0} \}\cup \{ \mathbf{e}_j - \mathbf{e}_i : (i,j)\in I\times J\}$.
+  %  \SP(A) = \conv (\{\mathbf{0} \}\cup \{ \mathbf{e}_j - \mathbf{e}_i : (i,j)\in I\times J\})
+\end{lemma}
+
+The following is a direct application of the main result of Padrol, Pilaud, and Ritter~\cite{}.
+
+\begin{lemma}
+  The strong rectangulation polytope $\SRP(n)$ is realized by the Minkowski sum of all shard polytopes of yin and yang arcs.
+\end{lemma}
+
+Note that since Loday and anti-Loday arcs are both yin and yang, each Loday and anti-Loday shard polytope is summed twice in the proposed realization.
+This does not change the combinatorial type of the polytope.
 
@@
 
-Proof of our second main result:
-\begin{itemize}
-\item Identify what are the yin summands contributing to a given pair $\mathbf{e}_i - \mathbf{e}_j$,
-\item prove that the count is the triple product of numbers of leaves of the trees,
-\item argue about what pairs $i,j$ are to be considered,
-\item argue symmetrically about yang arcs and conclude.
-\end{itemize}
+For every rectangulation $R$, we wish to give the coordinates of a point $p(R)\in\R^n$ such that
+$\SRP(n)$ is the convex hull of the points $p(R)$ for all rectangulations $R$.
+We first consider the contributions, to these coordinates, of the Minkowski summand $\SP(A)$ for all yin arcs $A$.
 
@@
   \label{lem:yinminmax}
-  Let $\pi\in\f{S}_n$ and consider a yin arc $A$ composed of two intervals $I$ and $J$.
+  Let $\pi\in\f{S}_n$ and consider a yin arc $A=(I,J)$.
   Let $i := \arg\max_{k\in I} \pi^{-1}(k)$ and $j := \arg\min_{k\in J} \pi^{-1}(k)$.
-  The vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
-  is $\mathbf{e}_i-\mathbf{e}_j$ if $\pi^{-1}(i)>\pi^{-1}(j)$, and $0$ otherwise.
+  The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
+  is $\mathbf{e}_i-\mathbf{e}_j$ if $\pi^{-1}(i)>\pi^{-1}(j)$, and $\mathbf{0}$ otherwise.
 \end{lemma}

### 2024-03-21 draft/rectangulotopes.tex
additional statements

@@
 \usepackage[T1]{fontenc}
-\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont, verbatim}
+\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont, verbatim, stackrel}
 \usepackage{marginnote}
@@
 
-% specific rectangulotope
+% polytopes
 \newcommand{\Asso}{\polytope{A}\mathsf{sso}} % associahedron
+% rectangulotopes
 \newcommand{\WRP}{\polytope{WR}} % weak rectangulation polytope
 \newcommand{\SRP}{\polytope{SR}} % strong rectangulation polytope
-\newcommand{\horizontalPattern}{\smash{\raisebox{-.15cm}{\includegraphics[scale=.5]{horizontalPattern}}}} % horizontal pattern
-\newcommand{\verticalPattern}{\smash{\raisebox{-.25cm}{\includegraphics[scale=.5]{verticalPattern}}}} % vertical pattern
-
 % shard polytopes
 \newcommand{\SP}{\polytope{SP}}
+
+% horizontal pattern
+\newcommand{\horizontalPattern}{\smash{\raisebox{-.15cm}{\includegraphics[scale=.5]{horizontalPattern}}}} 
+% vertical pattern
+\newcommand{\verticalPattern}{\smash{\raisebox{-.25cm}{\includegraphics[scale=.5]{verticalPattern}}}}
 
@@
 \newcommand{\yin}[1]{\overset{\sim}{#1}}
+% weak (Baxter) and strong equivalence on S_n
+\newcommand{\baxtereq}{\asymp}%{\stackrel{\frown}{\smile}}
+\newcommand{\strongeq}{\simeq}% % J: I would like to stack \backsim and \sim instead
 
@@
 
-\begin{itemize}
-\item Permutohedra
-\item Associahedra
-\item Generalized permutohedra
-\item Quotientopes
-\item Quotientopes of rectangulations, natural successors to associahedra; Definitions
-\item Flip graphs on rectangulations
-\item Loday popularized a simple realization of the associahedron
-\item We give a Loday-style description of two classes of rectangulation polytopes
-\item This intro states the main results; proofs are in the following sections.
-\end{itemize}
+The encoding of combinatorial objects in the form of a polyhedral structure is a recurrent theme in geometric and algebraic combinatorics, matroid and polymatroid theory, and combinatorial optimization.
 
-\subsection{permutohedra}
+A striking yet elementary example of this idea is the \defn{permutahedron} $\polytope{P} (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\pi\in\f{S}_n$.
+The facets of $\polytope{P} (n)$ are indexed by the nonempty and nontrivial subsets of $[n]$, while its skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
 
-\begin{theorem}
-  Let $\polytope{P} (n)$ be the convex hull of the points
-  $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, 
-  for all permutations $\pi\in\f{S}_n$.
-  Then $\polytope{P} (n)$ is a realization of the $(n-1)$-dimensional permutohedron.
-  \end{theorem}
+The \defn{associahedron} is an ubiquitous polytope, first defined by Tamari~\cite{}, then studied by Stasheff in a topological context~\cite{}, that has since then found applications in various areas of mathematics~\cite{}.
+It was first defined as a combinatorial object, before geometric realizations were shown to exist~\cite{}.
 
-\subsection{Loday's associahedra}
+The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
+In particular, it skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, whose structure has been the subject of many investigations, with applications in computer science~\cite{}.
 
-In 2004, Loday published the following elegant realization of the ubiquitous associahedron~\cite{}.
+In 2004, Loday published the following elegant realization of the associahedron~\cite{}.
 
@@
 
+\begin{itemize}
+\item \defn{Generalized (deformed) permutahedra}
+\item \defn{Quotientopes}
+\item Quotientopes of rectangulations, natural successors to associahedra; Definitions
+\item We give a Loday-style description of two classes of rectangulation polytopes
+\item This intro states the main results; proofs are in the following sections.
+\end{itemize}
+
 \subsection{Weak and strong rectangulations}
@@
 \begin{itemize}
-\item Terminology: rectangles, vertices, segments, generic rectangulation
-\item Wall slides
+\item Terminology: rectangles, \defn{vertices}, \defn{segments}, \defn{generic} rectangulation
+\item \defn{Wall slides}
 \item Two equivalence classes: \defn{weak equivalence} preserving rectangle-segment adjacencies, and \defn{strong equivalence} preserving rectangle-rectangle adjacencies.
 \item Weak equivalence classes have a \defn{diagonal} representative.
+\item \defn{Flip graphs} on rectangulations
 \end{itemize}
@@
 
-\Jean{Is there a more concise way to write this down/compute? Here we do not take advantage of the fact that at most one of the two factors $yin{w}_{i,j}$ and $\yang{w}_{i,j}$ is nonzero.}
+\Jean{Is there a more concise way to write this down/compute? Here we do not take advantage of the fact that at most one of the two factors $\yin{w}_{i,j}$ and $\yang{w}_{i,j}$ is nonzero.}
 
@@
 
+\jean{Clash of terminology: weak vs. Baxter?}
+
 Given a rectangulation $R$, the two trees $T(R)$ and $S(R)$, where the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top,
@@
 We define the \defn{weak poset} $P_w(R)=([n],\prec_w)$ of a rectangulation $R$ as the transitive closure of the union of the two partial orders defined by the intertwining binary trees $T(R)$ and $S(R)$.
+The weak poset is a two-dimensional lattice, whose minimum is the root of $S(R)$ and the maximum is the root of $T(R)$.
 
@@
   The subsets $\mathcal{L}(P_w(R))$ of $\f{S}_n$, for all rectangulations $R$ on $n$ rectangles, 
-  are equivalence classes of a congruence relation on the weak Bruhat order on $\f{S}_n$.
-  The minimal elements of each equivalence classes are twisted Baxter permutations, and the maximal
-  elements are co-twisted Baxter permutations.
+  are equivalence classes of a congruence relation $\baxtereq$ on the weak Bruhat order on $\f{S}_n$.
+  Furthermore:
+  \begin{itemize}
+  \item The congruence classes are one-to-one with weak rectangulations,  
+  \item the minimal elements of each congruence class are the twisted Baxter permutations,
+  \item the maximal elements are the co-twisted Baxter permutations,
+  \item the congruence classes each contain a single Baxter permutation, and are one-to-one with Baxter permutations.
+  \end{itemize}
 \end{theorem}
@@
 We call this congruence the \defn{Baxter congruence}.
-The weak rectangulation polytope $\WRP(n)$ is the quotientope of the Baxter congruence of the weak Bruhat order on $\f{S}_n$.
+The \defn{weak rectangulation polytope} $\WRP(n)$ is the quotientope of the Baxter congruence $\baxtereq$ on the weak Bruhat order on $\f{S}_n$.
 
@@
 
-The Baxter congruence is defined by forbidding arcs that cross the horizontal line.
-Therefore, the quotientope is defined by a Minkowski sum of the shard polytopes defined by two families of arcs: the \defn{Loday arcs} and the \defn{anti-Loday arcs}, that go respectively above and below all elements of an interval $I$ of $[n]$.
+\begin{lemma}
+The Baxter congruence $\baxtereq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that do not cross the horizontal line.
+\end{lemma}
+
+Therefore, the quotientope is realized by a Minkowski sum of the shard polytopes defined by two families of arcs: the \defn{Loday arcs} and the \defn{anti-Loday arcs}, that go respectively above and below all elements of an interval $I$ of $[n]$.
 
 \begin{lemma}
-  Given a Loday arc $A$ defined by an interval $I$, the Loday shard polytope $\SP(A)$ is defined as the standard simplex $\conv \{ \mathbf{e}_i : i\in I \}$.
+  \label{lem:lodaysp}
+  Given a Loday arc $A$ defined by an interval $I$, the Loday shard polytope $\SP(A)$ is a translate of the standard simplex $\conv \{ \mathbf{e}_i : i\in I \}$.
 \end{lemma}
@@
 \begin{lemma}
-  Given an anti-Loday arc $A$ defined by an interval $I$, the anti-Loday shard polytope $\SP(A)$ is defined as the negative standard simplex $\conv \{ - \mathbf{e}_i : i\in I \}$.
+  \label{lem:antilodaysp}
+  Given an anti-Loday arc $A$ defined by an interval $I$, the anti-Loday shard polytope $\SP(A)$ is a translate of the negative standard simplex $\conv \{ - \mathbf{e}_i : i\in I \}$.
 \end{lemma}
@@
 \subsection{Loday-type coordinates for weak rectangulation polytopes}
+
+For every rectangulation $R$, we wish to give the coordinates of a point $p(R)\in\R^n$ such that $\WRP(n)$ is the convex hull of the points $p(R)$ for all rectangulations $R$.
+Recall that in a Minkowski sum of polytopes, the vertex that is extremal in a generic direction is the sum of the vertices of the summands that are extremal in this direction.
+We therefore need to understand which vertices of the translates of the shard polytopes defined in Lemmas~\ref{lem:lodaysp} and~\ref{lem:antilodaysp} are extremal for a direction given by a permutation.
+
+\begin{lemma}
+  \label{lem:lodaymax}
+  Let $\pi\in\f{S}_n$ and consider a Loday arc $A$ defined by an interval $I$.
+  The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
+  is $\mathbf{e}_i$, where $i := \arg\max_{j\in I} \pi^{-1}(j)$.
+  Similarly, if $A$ is an anti-Loday arc defined by $I$, the unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
+  is $-\mathbf{e}_i$, where $i := \arg\min_{j\in I} \pi^{-1}(j)$.
+\end{lemma}
+
+Note that this extremal vertex should only depend on the Baxter congruence class of $\pi$.
+This, in passing, gives us an alternative characterization of the Baxter congruence:
+$\pi\baxtereq\sigma$ if and only if for any interval $I$ of $[n]$, $\arg\max_{j\in I} \pi^{-1}(j)=\arg\max_{j\in I} \sigma^{-1}(j)$, and
+$\arg\min_{j\in I} \pi^{-1}(j)=\arg\min_{j\in I} \sigma^{-1}(j)$.
 
@@
   The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ on $n$ rectangles, 
-  are equivalence classes of a congruence relation on the weak Bruhat order on $\f{S}_n$.
-  The minimal elements of each equivalence classes are 2-clumped permutations, and the maximal
-  elements are co-2-clumped permutations.
+  are equivalence classes of a congruence relation $\strongeq$ on the weak Bruhat order on $\f{S}_n$.
+    Furthermore:
+  \begin{itemize}
+  \item The congruence classes are one-to-one with strong rectangulations,  
+  \item the minimal elements of each congruence class are the 2-clumped permutations,
+  \item the maximal elements are the co-2-clumped permutations.
+  \end{itemize}
 \end{theorem}
 
-The strong rectangulation polytope $\SRP(n)$ is the quotientope of the congruence of the weak Bruhat order on $\f{S}_n$ that is defined by the 2-clumped permutations.
+We call the congruence $\strongeq$ the \defn{strong rectangulation congruence}.
+The \defn{strong rectangulation polytope} $\SRP(n)$ is the quotientope of the strong rectangulation congruence $\strongeq$.
 
@@
 
-The strong rectangulation congruence is defined by forbidding the arcs that cross the horizontal line more than once.
-We classify the remaining arcs into two families, depending on whether the arcs cross from above or below.
+\begin{lemma}
+  The strong rectangulation congruence $\strongeq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that cross the horizontal line at most once.
+\end{lemma}
 
-A \defn{yin arc} is defined by two consecutive nonempty intervals $I$ and $J$ of $[n]$ such that the arc goes above all elements of $I$ and
-below all elements of $J$. Note that the Loday arc are obtained when $|J|=1$ and the anti-Loday arc are obtained with $|I|=1$.
-A \defn{yang arc} is defined similarly, except that the arc goes below the elements of $I$ and above the elements of $J$. Note that Loday and anti-Loday arcs are also yang arcs.
+We classify the arcs into two families, depending on whether the arcs cross from above or below.
+A \defn{yin arc} is defined by two consecutive nonempty intervals $I$ and $J$ of $[n]$ such that the arc goes above all elements of $I$ and below all elements of $J$.
+A \defn{yang arc} is defined similarly, except that the arc goes below the elements of $I$ and above the elements of $J$.
 
 \begin{lemma}
-  Given a yin arc $A=(I,J)$, the shard polytope $\SP(A)$ is defined as the convex hull of the points
+  \label{lem:yinsp}
+  Given a yin arc $A=(I,J)$, the yin shard polytope $\SP(A)$ is defined as the convex hull of the points
   $\{\mathbf{0} \}\cup \{ \mathbf{e}_i - \mathbf{e}_j : (i,j)\in I\times J\}$.
-  %  \SP(A) = \conv (\{\mathbf{0} \}\cup \{ \mathbf{e}_i - \mathbf{e}_j : (i,j)\in I\times J\})
 \end{lemma}
@@
 \begin{lemma}
-  Given a yang arc $A=(I,J)$, the shard polytope $\SP(A)$ is defined as the convex hull of the points
+  \label{lem:yangsp}
+  Given a yang arc $A=(I,J)$, the yang shard polytope $\SP(A)$ is a translate of the convex hull of the points
   $\{\mathbf{0} \}\cup \{ \mathbf{e}_j - \mathbf{e}_i : (i,j)\in I\times J\}$.
-  %  \SP(A) = \conv (\{\mathbf{0} \}\cup \{ \mathbf{e}_j - \mathbf{e}_i : (i,j)\in I\times J\})
 \end{lemma}
+
+Note that yin shard polytopes are translates of the Loday shard polytopes when $|J|=1$, and of the anti-Loday shard polytopes when $|I|=1$.
+similarly, Loday and anti-Loday shard polytopes are translates of some yang shard polytopes.
 
@@
 \begin{lemma}
+  \label{lem:strongsum}
   The strong rectangulation polytope $\SRP(n)$ is realized by the Minkowski sum of all shard polytopes of yin and yang arcs.
@@
 
-Note that since Loday and anti-Loday arcs are both yin and yang, each Loday and anti-Loday shard polytope is summed twice in the proposed realization.
-This does not change the combinatorial type of the polytope.
+Since the combinatorial type of a Minkowski sum is invariant to translation of the summands, we can use the polytopes defined in Lemmas~\ref{lem:yinsp} and~\ref{lem:yangsp} as a definition of the yin and yang shard polytopes.
+Also note that since Loday and anti-Loday shard polytopes are, up to a translation, both yin and yang, each Loday and anti-Loday shard polytope is summed twice in the proposed realization.
+This, again, does not change the combinatorial type of the polytope.
+
+Lemma~\ref{lem:strongsum} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.
 
@@
 
-For every rectangulation $R$, we wish to give the coordinates of a point $p(R)\in\R^n$ such that
-$\SRP(n)$ is the convex hull of the points $p(R)$ for all rectangulations $R$.
+For every rectangulation $R$, we wish to give the coordinates of a point $p(R)\in\R^n$ such that $\SRP(n)$ is the convex hull of the points $p(R)$ for all rectangulations $R$.
 We first consider the contributions, to these coordinates, of the Minkowski summand $\SP(A)$ for all yin arcs $A$.
+Recall that in a Minkowski sum of polytopes, the vertex that is extremal in a generic direction is the sum of the vertices of the summands that are extremal in this direction.
+We therefore need to understand which vertices of the translates of the yin shard polytopes defined in Lemmas~\ref{lem:yinsp} are extremal for a direction given by a permutation.
 
@@
 \end{lemma}
+
+Note again that by definition, this extremal vertex should only depend on the strong congruence class of $\pi$.
 

### 2024-03-25 draft/rectangulotopes.tex
figures rectangulation lattices
updates

@@
 % rectangulotopes
-\newcommand{\WRP}{\polytope{WR}} % weak rectangulation polytope
-\newcommand{\SRP}{\polytope{SR}} % strong rectangulation polytope
+\newcommand{\WRP}{\polytope{WR}} % weak rectangulotope
+\newcommand{\SRP}{\polytope{SR}} % strong rectangulotope
 % shard polytopes
@@
 
-\title{Rectangulation polytopes}
+%\title{Rectangulation polytopes}
+\title{Rectangulotopes}
 
@@
 They are defined as quotientopes of natural congruence relations on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
-We give  elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedra.
+We give elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedron.
 \end{abstract}
@@
 
-A striking yet elementary example of this idea is the \defn{permutahedron} $\polytope{P} (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\pi\in\f{S}_n$.
+An elementary yet striking example of this idea is the \defn{permutahedron} $\polytope{P} (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\pi\in\f{S}_n$.
 The facets of $\polytope{P} (n)$ are indexed by the nonempty and nontrivial subsets of $[n]$, while its skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
 
-The \defn{associahedron} is an ubiquitous polytope, first defined by Tamari~\cite{}, then studied by Stasheff in a topological context~\cite{}, that has since then found applications in various areas of mathematics~\cite{}.
+The \defn{associahedron} is an ubiquitous polytope, first defined by Tamari~\cite{T51}, then studied by Stasheff in a topological context~\cite{S63}, that has since then found applications in various areas of mathematics, including topology~\cite{}, operads~\cite{}, cluster algebras~\cite{}, combinatorial Hopf algebras~\cite{}, physics~\cite{}, etc.
 It was first defined as a combinatorial object, before geometric realizations were shown to exist~\cite{}.
@@
 \begin{theorem}
+  \label{thm:loday}
   Let $\Asso (n)$ be the convex hull of the points
@@
 
-\begin{itemize}
-\item \defn{Generalized (deformed) permutahedra}
-\item \defn{Quotientopes}
-\item Quotientopes of rectangulations, natural successors to associahedra; Definitions
-\item We give a Loday-style description of two classes of rectangulation polytopes
-\item This intro states the main results; proofs are in the following sections.
-\end{itemize}
+A natural object associated with the associahedron is the \defn{Tamari lattice}, a lattice whose cover graph is isomorphic to the skeleton of the associahedron.
+The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} relation called the \defn{sylvester congruence}.
+Pilaud and Santos~\cite{}, answering a question of Reading~\cite{}, proved that with \emph{every} lattice congruence of the weak order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more generally whose normal fan is the fan defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
+They called such polytopes \defn{quotientopes}.
+
+The goal of this paper is to initiate the study of quotientopes of lattice congruences whose classes are in bijections with \defn{rectangulations}, defined as decompositions of a square into finitely many axis-aligned rectangles.
+
+Those lattice congruences are the \defn{Baxter congruence} on one hand, studied by Law and Reading~\cite{}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}), and the \defn{strong rectangulation congruence}, whose classes are in bijection with so-called \defn{strong rectangulations}.
+The latter is mentioned by Reading~\cite{} (seee for instance in Theorem 1.2, and Example 4.11), and studied more recently by Asinowski et al.~\cite{}.
+
+One of our main results is an elementary description of the vertex coordinates of a realization of both families of quotientopes, in the spirit of Theorem~\ref{thm:loday}.
+The quotientopes corresponding to the Baxter congruence will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
+similarly, the polytopes corresponding to the strong rectangulation congruence will be called the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
+
+In this introduction, we give a minimal set of definition to be able to completely state our results on the vertex description of the weak and strong rectangulotopes.
+Section~\ref{sec:quotientopes} is dedicated to the background on quotientopes and their realizations as Minkowski sums of \defn{shard polytopes}, due to Padrol, Pilaud, and Ritter~\cite{}.
+Section~\ref{sec:wr} details the case of weak rectangulations, while Section~\ref{sec:sr} deals with strong rectangulations.
 
@@
 \end{itemize}
+
+Law and Reading~\cite{} defined a flip graph on weak rectangulations.
+
+\begin{lemma}
+  The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$.
+\end{lemma}
+
+Meehan~\cite{} (See also Asinowski et al.~\cite{}) describe the flip graph on strong rectangulations.
+
+\begin{lemma}
+  The skeleton of $\SRP(n)$ is isomorphic to the flip graph of strong rectangulations of size $n$.
+\end{lemma}
 
@@
 
-\subsection{Weak rectangulation polytopes}
+\subsection{Weak rectangulotopes}
 
-We are now ready to give our first main result on the realizations of weak rectangulation polytopes.
+We are now ready to give our first main result on the realizations of weak rectangulotopes.
 
@@
   \begin{split}
-    \loday{w}_i := & h(T(R), i)\cdot v(T(R),i) \\
-    \antiloday{w}_i := & h(S(R), i)\cdot v(S(R),i),
+    \loday{w}_i := & h(T, i)\cdot v(T,i) \\
+    \antiloday{w}_i := & h(S, i)\cdot v(S,i),
   \end{split}
   \]
-  for all weak rectangulations $R$ on $n$ rectangles, where $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
-  Then $\WRP (n)$ is a realization of the $(n-1)$-dimensional weak rectangulation polytope.
+  for all weak rectangulations $R$ on $n$ rectangles, where $T:=T(R)$, $S:=S(R)$,
+  and $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
+  Then $\WRP (n)$ is a realization of the $(n-1)$-dimensional weak rectangulotope.
 \end{theorem}
 
-\subsection{Strong rectangulation polytopes}
+\subsection{Strong rectangulotopes}
 
-Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$.
-We denote by $i|j$ and $\frac{\ j\ }i$ the situations where the rectangle labeled $i$ touches the rectangle labeled $j$ from the left and from below in $R$, respectively.
+Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$. We denote by %$i|j$ and $\frac{\ j\ }i$
+$i\backslash j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$. %, respectively.
 
@@
   \[
-  \yin{w}_{i,j} := 
-  \begin{cases}
-    h(T(R), i) \cdot cv (T(R), S(R), i, j)\cdot h(S(R), j) & \text{ if neither } i|j \text{ nor } \frac{\ j\ }i, \\
-    0 & \text{ otherwise.}
-  \end{cases}
+  \begin{split}
+    \yin{w}_{i,j} := & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot [ \neg i\backslash j ] \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
+    \yang{w}_{i,j} := & v(S, j) \cdot ch (T, S, i, j)\cdot v(T, i) \cdot [ \neg j\backslash i ]
+  \end{split}
   \]
-  and
-    \[
-  \yang{w}_{i,j} := 
-  \begin{cases}
-    v(S(R), j) \cdot ch (T(R), S(R), i, j)\cdot v(T(R), i) & \text{ if neither } j|i \text{ nor } \frac{\ i\ }j, \\
-    0 & \text{ otherwise.}
-  \end{cases}
-  \]
+
+%  \begin{cases}
+%    h(T(R), i) \cdot cv (T, S, i, j)\cdot h(S, j) & \text{ if neither } i|j \text{ nor } \frac{\ j\ }i, \\
+%    0 & \text{ otherwise.}
+%  \end{cases}
+
+  %\begin{cases}
+  %  v(S, j) \cdot ch (T, S, i, j)\cdot v(T, i) & \text{ if neither } j|i \text{ nor } \frac{\ i\ }j, \\
+  %  0 & \text{ otherwise.}
+  %\end{cases}
 
@@
   \begin{itemize}
+  \item $T := T(R)$ and $S:=S(R)$,
   \item $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$,
@@
   \end{itemize}
-Then $\SRP (n)$ is a realization of the $(n-1)$-dimensional strong rectangulation polytope.  
+Then $\SRP (n)$ is a realization of the $(n-1)$-dimensional strong rectangulotope.  
 \end{theorem}
@@
 
-\section{Weak rectangulation polytopes}
-\label{sec:weakRectangulationPolytopes}
+\section{Weak rectangulotopes}
+\label{sec:wr}
 
@@
 We call this congruence the \defn{Baxter congruence}.
-The \defn{weak rectangulation polytope} $\WRP(n)$ is the quotientope of the Baxter congruence $\baxtereq$ on the weak Bruhat order on $\f{S}_n$.
+The \defn{weak rectangulotope} $\WRP(n)$ is the quotientope of the Baxter congruence $\baxtereq$ on the weak Bruhat order on $\f{S}_n$.
 
@@
 \begin{lemma}
-  The weak rectangulation polytope $\WRP(n)$ is realized by the Minkowski sum of all Loday and anti-Loday shard polytopes. In particular, it is the Minkowski sum of two opposite Loday associahedra.
+  The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all Loday and anti-Loday shard polytopes. In particular, it is the Minkowski sum of two opposite Loday associahedra.
 \end{lemma}
 
-\subsection{Loday-type coordinates for weak rectangulation polytopes}
+\subsection{Loday-type coordinates for weak rectangulotopes}
 
@@
 
-\section{Strong rectangulation polytopes}
-\label{sec:strongRectangulationPolytopes}
+\section{Strong rectangulotopes}
+\label{sec:sr}
 
@@
 We call the congruence $\strongeq$ the \defn{strong rectangulation congruence}.
-The \defn{strong rectangulation polytope} $\SRP(n)$ is the quotientope of the strong rectangulation congruence $\strongeq$.
+The \defn{strong rectangulotope} $\SRP(n)$ is the quotientope of the strong rectangulation congruence $\strongeq$.
 
@@
   \label{lem:strongsum}
-  The strong rectangulation polytope $\SRP(n)$ is realized by the Minkowski sum of all shard polytopes of yin and yang arcs.
+  The strong rectangulotope $\SRP(n)$ is realized by the Minkowski sum of all shard polytopes of yin and yang arcs.
 \end{lemma}
@@
 
-\subsection{Loday-type coordinates for strong rectangulation polytopes}
+\subsection{Loday-type coordinates for strong rectangulotopes}
 
@@
   \[
-  \yin{w}_{i,j} = 
-  \begin{cases}
-    h(T(R), i) \cdot cv (T(R), S(R), i, j)\cdot h(S(R), j) & \text{ if neither } i|j \text{ nor } \frac{\ j\ }i \\
-    0 & \text{ otherwise }
-  \end{cases}
+    h(T(R), i) \cdot cv (T(R), S(R), i, j)\cdot h(S(R), j) 
   \]
+  if neither $i|j$ nor $\frac{\ j\ }i$, and $0$ otherwise.
 \end{lemma}

### 2024-03-25 programmation/submodular.py
added submodular.py in programmation

@@ -1,0 +1,85 @@
+# submodular functions for rectangulotopes
+
+from itertools import *
+
+# simplex for a Loday arc I
+# 1 if S and I intersects, 0 otherwise
+def f_loday(S, I):
+    return int(bool(set(S) & set (I)))
+
+# anti-simplex for an anti-Loday arc I
+# -1 if I is a subset of S, 0 otherwise
+def f_antiloday(S, I):
+    return - int(bool(set(I).issubset(set(S))))
+
+# simplex + antisimplex
+# 1 if S and I intersects but I is not a subset of S, 0 otherwise
+def f_cuboc(S, I):
+    return int(bool(set(S) & set (I)) and not bool(set(I).issubset(set(S))))
+
+# weak rectangulation
+# number of I such that S and I intersects but I is not a subset of S
+def f_weakrect(S, n):
+    return sum([int(bool(set(S) & set (range(i, j+1))) and not bool(set(range(i, j+1)).issubset(set(S)))) for i in range(n-1) for j in range(i+1, n)])
+
+# powerset recipe
+def powerset(iterable):
+    "powerset([1,2,3]) --> () (1,) (2,) (3,) (1,2) (1,3) (2,3) (1,2,3)"
+    s = list(iterable)
+    return chain.from_iterable(combinations(s, r) for r in range(len(s)+1))
+
+# compute the polyhedron for a given submod function
+def submod_polytope(f, n):
+    all_ineq = []
+    for S in powerset(range(n)):
+        all_ineq.append((f(S),) + tuple(-1 if x in S else 0 for x in range(n)))
+    return Polyhedron(ieqs=all_ineq)
+
+# test
+n = 6
+g = lambda S : f_antiloday(S, range(n-2))
+print(submod_polytope(g, n))
+
+# associahedron as a Minkowski sum
+def assoc(n):
+    A = Polyhedron([(0,)*n])
+    for i in range(n-1):
+        for j in range(i+1, n):
+            I = range(i, j+1)
+            g_loday = lambda S : f_loday(S, range(i, j+1))
+            A = A + submod_polytope(g_loday, n)
+    return Polyhedron(A.vertices()) # remove rays
+
+# diagonal/Baxter/weak rectangulation polytope
+# either sum all lodays and antilodays, or the cubocs
+def wrp(n):
+    A = Polyhedron([(0,)*n])
+    for i in range(n-1):
+        for j in range(i+1, n):
[…]

### 2024-03-26 draft/rectangulotopes.tex
bibliography and introduction

@@
 \usepackage[T1]{fontenc}
-\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont, verbatim, stackrel}
+\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont, verbatim, stackrel, stmaryrd}
 \usepackage{marginnote}
@@
 They are defined as quotientopes of natural congruence relations on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
-We give elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedron.
+We give a thorough description of these polytopes, including elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedron.
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-
 \section{Introduction}
@@
 
-The encoding of combinatorial objects in the form of a polyhedral structure is a recurrent theme in geometric and algebraic combinatorics, matroid and polymatroid theory, and combinatorial optimization.
-
+The encoding of combinatorial objects in the form of a polyhedral structure is a recurrent theme in geometric and algebraic combinatorics.
 An elementary yet striking example of this idea is the \defn{permutahedron} $\polytope{P} (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\pi\in\f{S}_n$.
 The facets of $\polytope{P} (n)$ are indexed by the nonempty and nontrivial subsets of $[n]$, while its skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
+The many generalizations and extensions of the permutohedra gave rise to a flourishing theory of so-called \defn{generalized permutohedra}.
 
-The \defn{associahedron} is an ubiquitous polytope, first defined by Tamari~\cite{T51}, then studied by Stasheff in a topological context~\cite{S63}, that has since then found applications in various areas of mathematics, including topology~\cite{}, operads~\cite{}, cluster algebras~\cite{}, combinatorial Hopf algebras~\cite{}, physics~\cite{}, etc.
-It was first defined as a combinatorial object, before geometric realizations were shown to exist~\cite{}.
-
+Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined by Tamari~\cite{T51}, then studied by Stasheff in a topological context~\cite{S63}, that has since then found applications in various areas of mathematics, including topology, operads, cluster algebras, combinatorial Hopf algebras, and physics~\cite{MR4675114}.
+It was first defined as a combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894}.
 The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
-In particular, it skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, whose structure has been the subject of many investigations, with applications in computer science~\cite{}.
-
-In 2004, Loday published the following elegant realization of the associahedron~\cite{}.
+In particular, it skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.
+In 2004, Loday published the following elegant realization of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex given the corresponding binary tree.
 
@@
 
-A natural object associated with the associahedron is the \defn{Tamari lattice}, a lattice whose cover graph is isomorphic to the skeleton of the associahedron.
+A natural object associated with the associahedron is the \defn{Tamari lattice}, the cover graph of which is isomorphic to the skeleton of the associahedron.
 The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} relation called the \defn{sylvester congruence}.
-Pilaud and Santos~\cite{}, answering a question of Reading~\cite{}, proved that with \emph{every} lattice congruence of the weak order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more generally whose normal fan is the fan defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
-They called such polytopes \defn{quotientopes}.
+Pilaud and Santos~\cite{MR3964495}, answering a question of Reading~\cite{MR2142177}, proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more generally whose normal fan is the fan defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
+Those polytopes are generalized permutohedra that they called \defn{quotientopes}.
 
-The goal of this paper is to initiate the study of quotientopes of lattice congruences whose classes are in bijections with \defn{rectangulations}, defined as decompositions of a square into finitely many axis-aligned rectangles.
+Associahedra are therefore the quotientopes of a congruence of the weak Bruhat order whose classes are in bijection with triangulations.  
+The goal of this paper is to initiate the study of quotientopes of congruences of the weak Bruhat order whose classes are in bijection with (equivalence classes of) \defn{rectangulations}, defined as decompositions of a square into finitely many axis-aligned rectangles.
+The combinatorics of rectangulations has been studied for decades~\cite{MR2233287,MR2871762,MR2864445,MR2763051,MR3084577,MR3878132,MR4598046}, and the order-theoretic take on these structures has already proved useful.
+%, in particular in the context of electronic engineering applications~\cite{}.
 
-Those lattice congruences are the \defn{Baxter congruence} on one hand, studied by Law and Reading~\cite{}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}), and the \defn{strong rectangulation congruence}, whose classes are in bijection with so-called \defn{strong rectangulations}.
-The latter is mentioned by Reading~\cite{} (seee for instance in Theorem 1.2, and Example 4.11), and studied more recently by Asinowski et al.~\cite{}.
+The relevant lattice congruences are the \defn{Baxter congruence} on one hand, studied by Law and Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}) and \defn{Baxter permutations}, and the \defn{strong rectangulation congruence}, whose classes are in bijection with so-called \defn{strong rectangulations}~\cite{MR2864445}.
+The latter is mentioned by Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by Asinowski et al.~\cite{ACFF24}.
 
-One of our main results is an elementary description of the vertex coordinates of a realization of both families of quotientopes, in the spirit of Theorem~\ref{thm:loday}.
+One of our main results is an elementary description of the vertex coordinates of a realization of both families of quotientopes, in the spirit of Loday's realization in Theorem~\ref{thm:loday}.
 The quotientopes corresponding to the Baxter congruence will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
-similarly, the polytopes corresponding to the strong rectangulation congruence will be called the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
+Similarly, the polytopes corresponding to the strong rectangulation congruence will be called the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
+  The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
+Similarly, the skeleton of $\SRP(n)$ is isomorphic to the flip graph of strong rectangulations of size $n$ defined by Meehan~\cite{MR3697823} (See also Asinowski et al.~\cite{ACFF24}).
+Rectangulotopes enrich the family of natural quotientopes, alongside associahedra and permutreehedra~\cite{MR3856522}, and are instructive examples of application of the theory of lattice congruences to concrete combinatorial objects.
 
-In this introduction, we give a minimal set of definition to be able to completely state our results on the vertex description of the weak and strong rectangulotopes.
-Section~\ref{sec:quotientopes} is dedicated to the background on quotientopes and their realizations as Minkowski sums of \defn{shard polytopes}, due to Padrol, Pilaud, and Ritter~\cite{}.
-Section~\ref{sec:wr} details the case of weak rectangulations, while Section~\ref{sec:sr} deals with strong rectangulations.
+% jean:{Vincent, you may want to say more about permutreehedra, incl. the Loday coordinates.}
+
+We give the necessary definitions for stating our results on the vertex description of the weak and strong rectangulotopes.
 
@@
 
-Law and Reading~\cite{} defined a flip graph on weak rectangulations.
-
-\begin{lemma}
-  The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$.
-\end{lemma}
-
-Meehan~\cite{} (See also Asinowski et al.~\cite{}) describe the flip graph on strong rectangulations.
-
-\begin{lemma}
-  The skeleton of $\SRP(n)$ is isomorphic to the flip graph of strong rectangulations of size $n$.
-\end{lemma}
-
 \subsection{Source and target trees of a rectangulation}
@@
 
-\jean{Here only give the definitions allowing to state the results.}
+% jean:{Here only give the definitions allowing to state the results.}
 
 \begin{figure}
-	\centerline{\includegraphics[scale=.7]{strongRectangulation} \qquad \includegraphics[scale=.7]{strongRectangulationTrees}}
-	\caption{\label{fig:sourceTargetTrees}A strong rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{AsinowskiCardinalFelsnerFusy}.}
+	\centerline{\includegraphics[width=.5\textwidth]{strongRectangulation} \qquad \includegraphics[width=.5\textwidth]{strongRectangulationTrees}}
+	\caption{\label{fig:sourceTargetTrees}A strong rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{ACFF24}.}
         % The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
@@
 This enables to unambiguously label the rectangles of~$R$ by the inorder 
-The resulting labeling coincides with the NW--SE labeling of~\cite{AsinowskiCardinalFelsnerFusy}.
+The resulting labeling coincides with the NW--SE labeling of~\cite{ACFF24}.
 From now on, the labels of the rectangles in a rectangulations are the inorder labels, and the two trees $T(R)$ and $S(R)$ are defined on the vertex set $[n]$ accordingly.
@@
 
-We are now ready to give our first main result on the realizations of weak rectangulotopes.
+Our first result is a concise formula for the coordinates of the vertices of the  weak rectangulotopes $\WRP(n)$, that consists of applying Loday's formula on each of the source and target trees of the rectangulation.
+This materializes the result of Law and Reading that the weak rectangulotopes are Minkowski sums of two opposite associahedra~\cite{MR2871762}.
 
@@
 
+In order to exhibit a similar realization for the strong rectangulotopes, we need to count subsets of leaves that lie in some subtrees of both the source and target trees of a rectangulation.
 Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$. We denote by %$i|j$ and $\frac{\ j\ }i$
@@
   \begin{split}
-    \yin{w}_{i,j} := & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot [ \neg i\backslash j ] \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
-    \yang{w}_{i,j} := & v(S, j) \cdot ch (T, S, i, j)\cdot v(T, i) \cdot [ \neg j\backslash i ]
+    \yin{w}_{i,j} := & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot \llbracket \neg i\backslash j \rrbracket \\% \text { if not } i\backslash j, \text{and } 0\text{ otherwise.} \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
+    \yang{w}_{i,j} := & v(S, i) \cdot ch (S, T, i, j)\cdot v(T, j) \cdot \llbracket \neg j\backslash i \rrbracket
   \end{split}
@@
 
-  \jean{Double check the formula.}
+  % jean:{Double check the formula.}
   for all strong rectangulations $R$ on $n$ rectangles, where:
@@
 
-\Jean{Is there a more concise way to write this down/compute? Here we do not take advantage of the fact that at most one of the two factors $\yin{w}_{i,j}$ and $\yang{w}_{i,j}$ is nonzero.}
+% jean:{Is there better way to write this down/compute? Here we do not take advantage of the fact that at most one of the two factors $\yin{w}_{i,j}$ and $\yang{w}_{i,j}$ is nonzero.}
+
+\subsection{Plan of the paper}
+
+ Section~\ref{sec:quotientopes} is dedicated to the background on quotientopes and their realizations as Minkowski sums of \defn{shard polytopes}, due to Padrol, Pilaud, and Ritter~\cite{MR4584712}, which will be our main tool throughout.
+Section~\ref{sec:wr} details the case of weak rectangulations, while Section~\ref{sec:sr} deals with strong rectangulations.
+In both cases, in addition to the vertex coordinates, we give the description of the face lattices and the submodular functions defining the polytopes.
+
+\subsection*{Acknowledgments}
+
+This work was initiated at the Workshop on Combinatorics, Algorithms, and Geometry organized by Torsten M\"utze on March 4-8, 2024 in Dresden, Germany.
+The authors thank Torsten M\"utze and Namrata for the organization and the other participants of the workshop for the stimulating interactions.
 
@@
 
-\jean{Clash of terminology: weak vs. Baxter?}
+% jean:{Clash of terminology: weak vs. Baxter?}
 
@@
 define two partial orders on $[n]$.
-We define the \defn{weak poset} $P_w(R)=([n],\prec_w)$ of a rectangulation $R$ as the transitive closure of the union of the two partial orders defined by the intertwining binary trees $T(R)$ and $S(R)$.
-The weak poset is a two-dimensional lattice, whose minimum is the root of $S(R)$ and the maximum is the root of $T(R)$.
+We define the \defn{weak poset} $P_w(R)=([n],\prec_w)$ of a rectangulation $R$ as the transitive closure of the union of the two partial orders defined by the intertwining binary trees $T(R)$ and $S(R)$~\cite{MR4014603}.
+The weak poset is a two-dimensional lattice, the minimum of which is the root of $S(R)$ and the maximum is the root of $T(R)$.
+We define the set $\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
+\[
+\mathcal{L}(P_w(R)) := \{\pi\in\f{S}_n : i\prec_w j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
+\]
 
@@
 \begin{figure}
-	\centerline{\includegraphics[scale=.7]{weakRectangulation} \qquad \includegraphics[scale=.7]{weakRectangulationTrees}}
+	\centerline{\includegraphics[width=.5\textwidth]{weakRectangulation} \qquad \includegraphics[width=.5\textwidth]{weakRectangulationTrees}}
 	\caption{A diagonal rectangulation (left) and its pair of separated twin trees (right).
@@
 \end{lemma}
+
+% jean: From this, do we directly have a description of the face lattice of $\WRP(n)$?}
+
+\subsection{Shard polytopes of Loday and anti-Loday arcs}
 
@@
   The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all Loday and anti-Loday shard polytopes. In particular, it is the Minkowski sum of two opposite Loday associahedra.
+\end{lemma}
+
+\subsection{Submodular functions of weak rectangulotopes}
+
+\begin{lemma}
+  The weak rectangulotope $\WRP(n)$ is realized as
+  \[
+  \{x\in\R^n : \sum_{i\in S} x_i \leq f(S) \text { and } \sum_{i\in [n]} x_i = f([n])\},
+  \]
+  where $f:2^{[n]}\mapsto \R$ is the submodular function that maps $S$ to the number of intervals
+  $I$ of $[n]$ such that $I\cap S\not=\emptyset$ and $I\not\subseteq S$. 
 \end{lemma}
@@
 This, in passing, gives us an alternative characterization of the Baxter congruence:
-$\pi\baxtereq\sigma$ if and only if for any interval $I$ of $[n]$, $\arg\max_{j\in I} \pi^{-1}(j)=\arg\max_{j\in I} \sigma^{-1}(j)$, and
+$\pi\baxtereq\sigma$ if and only if for any interval $I$ of $[n]$,
+%the maximum and the minimum of $\pi^{-1}$ and $\sigma^{-1}$ on $I$ have the same index.
+$\arg\max_{j\in I} \pi^{-1}(j)=\arg\max_{j\in I} \sigma^{-1}(j)$, and
 $\arg\min_{j\in I} \pi^{-1}(j)=\arg\min_{j\in I} \sigma^{-1}(j)$.
@@
 
-We define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of a rectangulation $R$.
-For a poset $P$ on $[n]$, we define $\mathcal{L}(P)$ as the set of permutations of $\f{S}_n$ corresponding to linear extensions of $P$.
+We define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of a rectangulation $R$, and $\mathcal{L}(P_s)$ as the set of permutations of $\f{S}_n$ corresponding to linear extensions of $P_s$:
+\[
+\mathcal{L}(P_s(R)) := \{\pi\in\f{S}_n : i\prec_s j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
+\]
 
@@
 
-\jean{Does the congruence have a name?}
+% jean:{Does the congruence have a name?}
 
@@
 
+% jean:{From this, do we directly have a description of the face lattice of $\SRP(n)$?}
+
 We classify the arcs into two families, depending on whether the arcs cross from above or below.
@@
 A \defn{yang arc} is defined similarly, except that the arc goes below the elements of $I$ and above the elements of $J$.
+
+\subsection{Shard polytopes of yin and yang arcs}
 
@@
 
-The following is a direct application of the main result of Padrol, Pilaud, and Ritter~\cite{}.
+The following is a direct application of the main result of Padrol, Pilaud, and Ritter~\cite{MR4584712}.
 
@@
 Lemma~\ref{lem:strongsum} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.
+
+\subsection{Submodular functions of strong rectangulotopes}
+
+% jean:{From this, do we directly have the submodular function defining $\SRP(n)$?}
 
@@
 \begin{lemma}
-  Let $R$ be a rectangulation on $n$ rectangles and $p(R)$ the vertex of $\SRP(n)$ associated with $R$.
+  Let $R$ be a rectangulation on $n$ rectangles, $([n],\prec_s)$ the associated strong order, and $p(R)$ the vertex of $\SRP(n)$ corresponding to $R$.
   Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $A$ such that $\mathbf{e}_i-\mathbf{e}_j$
@@
   \]
-  if neither $i|j$ nor $\frac{\ j\ }i$, and $0$ otherwise.
+  if $i\succ_s j$, and $0$ otherwise.
 \end{lemma}
@@
   Let $\pi$ be the 2-clumped permutation corresponding to $R$, hence the minimal element of $\mathcal{L}(P_s(R))$ in the weak Bruhat order.
-  From Lemma~\ref{lem:yinminmax}, the set of yin arcs $A$ such that $\SP(A)$ contributes to $\mathbf{e}_i-\mathbf{e}_j$
-  are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that:
-  \begin{itemize}
-  \item $i = \arg\max_{k\in I} \pi^{-1}(k)$,
-  \item $j = \arg\min_{k\in J} \pi^{-1}(k)$, and
-  \item $i,j$ is an inversion in $\pi$, hence $\pi^{-1}(i)>\pi^{-1}(j)$.
-  \end{itemize}
-  We observe that provided the third condition holds, the number of choices of the three endpoints defining such a pair of intervals $I$ and $J$ is the product of the
+  From Lemma~\ref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc $A$ whose shard polytope $\SP(A)$ contributes to $\mathbf{e}_i-\mathbf{e}_j$ only if
+  $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.
+  Provided this holds, the set of yin arcs $A$ such that $\SP(A)$ contributes to $\mathbf{e}_i-\mathbf{e}_j$ are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that $i$ is the index of the maximum of $\pi^{-1}$ in $I$, and $j$ is the index of the maximum of $\pi^{-1}$ in $J$.
+  %$i = \arg\max_{k\in I} \pi^{-1}(k)$ and $j = \arg\min_{k\in J} \pi^{-1}(k)$.
+  We claim that the number of choices of the three endpoints defining such a pair of intervals $I$ and $J$ is the product of the
   number of leaves of the subtree of $T(R)$ rooted at $i$, the number of common leaves to the vertical subtrees of $T(R)$ and $S(R)$ rooted
   at $i$ and $j$, respectively, and the number of leaves of the subtree of $S(R)$ rooted at $j$.
+\end{proof}
 
-  It remains to observe that the only case where this product is nonzero and $i,j$ is not an inversion in $\pi$ is when either $i|j$ or $\frac{\ j\ }i$.
-\end{proof}
+It remains to observe that the only case where this product is nonzero and $i,j$ is not an inversion in $\pi$ is when $i\backslash j$.
+
+\begin{lemma}
+  Let $R$ be a rectangulation on $n$ rectangles and $([n],\prec_s)$ the associated strong order.
+  Given a pair $i,j\in [n]$ such that $i<j$ and $i\prec_s j$, we have
+  \[
+    cv (T(R), S(R), i, j) > 0 \implies i\backslash j.
+  \]
+\end{lemma}
 
@@
 
+\begin{figure}
+	\centerline{\includegraphics[scale=.9]{weakRectangulationLattice}}
+	\caption{The weak rectangulation lattice.}
+	\label{fig:weakRectangulationLattice}
+\end{figure}
+
+\begin{figure}
+	\centerline{\includegraphics[scale=.9]{strongRectangulationLattice}}
+	\caption{The strong rectangulation lattice.}
+	\label{fig:strongRectangulationLattice}
+\end{figure}
+
 \end{document}

### 2024-04-02 draft/rectangulotopes.tex
minor changes

@@
 \newcommand{\one}{\b{1}} % the all one vector
-\newcommand{\eqdef}{\mbox{\,\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}}\ensuremath{=}\,}} % :=
+\newcommand{\eqdef}{\mbox{\,\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}}\ensuremath{=}\,}} % \eqdef
 \newcommand{\defeq}{\mbox{~\ensuremath{=}\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}} }} % =:
@@
 % polytopes
-\newcommand{\Asso}{\polytope{A}\mathsf{sso}} % associahedron
+\newcommand{\Perm}{\polytope{P}} % permutahedron
+\newcommand{\Asso}{\polytope{A}} % associahedron
 % rectangulotopes
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Introduction}
@@
 The encoding of combinatorial objects in the form of a polyhedral structure is a recurrent theme in geometric and algebraic combinatorics.
-An elementary yet striking example of this idea is the \defn{permutahedron} $\polytope{P} (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\pi\in\f{S}_n$.
-The facets of $\polytope{P} (n)$ are indexed by the nonempty and nontrivial subsets of $[n]$, while its skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
-The many generalizations and extensions of the permutohedra gave rise to a flourishing theory of so-called \defn{generalized permutohedra}.
+An elementary yet striking example of this idea is the \defn{permutahedron} $\Perm (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\pi\in\f{S}_n$.
+Its facets are indexed by the proper nonempty subsets of $[n]$, while its skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
+The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra} (also known as \defn{generalized permutahedra}).
 
@@
 It was first defined as a combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894}.
-The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
+The face lattice of the associahedron~$\Asso (n)$ is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
 In particular, it skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.
@@
   \[
-  w_i := \ell(T, i)\cdot r(T,i),
+  w_i \eqdef \ell(T, i)\cdot r(T,i),
   \]
@@
 Pilaud and Santos~\cite{MR3964495}, answering a question of Reading~\cite{MR2142177}, proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more generally whose normal fan is the fan defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
-Those polytopes are generalized permutohedra that they called \defn{quotientopes}.
+Those polytopes are generalized permutahedra that they called \defn{quotientopes}.
 
@@
 
-The relevant lattice congruences are the \defn{Baxter congruence} on one hand, studied by Law and Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}) and \defn{Baxter permutations}, and the \defn{strong rectangulation congruence}, whose classes are in bijection with so-called \defn{strong rectangulations}~\cite{MR2864445}.
+The relevant lattice congruences are the \defn{Baxter congruence} on one hand, studied by Law and Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}) and \defn{Baxter permutations}, and the \defn{strong rectangulation congruence}, whose classes are in bijection with \defn{strong rectangulations}~\cite{MR2864445}.
 The latter is mentioned by Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by Asinowski et al.~\cite{ACFF24}.
@@
 The quotientopes corresponding to the Baxter congruence will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
-Similarly, the polytopes corresponding to the strong rectangulation congruence will be called the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
 The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
-Similarly, the skeleton of $\SRP(n)$ is isomorphic to the flip graph of strong rectangulations of size $n$ defined by Meehan~\cite{MR3697823} (See also Asinowski et al.~\cite{ACFF24}).
+Similarly, the quotientopes corresponding to the strong rectangulation congruence will be called the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
+The skeleton of $\SRP(n)$ is isomorphic to the flip graph of strong rectangulations of size $n$ defined by Meehan~\cite{MR3697823} (see also~\cite{ACFF24}).
+
 Rectangulotopes enrich the family of natural quotientopes, alongside associahedra and permutreehedra~\cite{MR3856522}, and are instructive examples of application of the theory of lattice congruences to concrete combinatorial objects.
@@
 
+\subsection{Weak and strong rectangulations}
+
 We give the necessary definitions for stating our results on the vertex description of the weak and strong rectangulotopes.
-
-\subsection{Weak and strong rectangulations}
 
@@
 	\centerline{\includegraphics[width=.5\textwidth]{strongRectangulation} \qquad \includegraphics[width=.5\textwidth]{strongRectangulationTrees}}
-	\caption{\label{fig:sourceTargetTrees}A strong rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{ACFF24}.}
+	\caption{A strong rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{ACFF24}.}
         % The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
+        \label{fig:strongRectangulation}
 \end{figure}
@@
 The \defn{source tree}~$S(R)$ (resp.~\defn{target tree}~$T(R)$) is the subgraph of~$D(R)$ induced by the sources~$\set{s(r)}{r \in [n]}$ (resp.~by the targets~$\set{t(r)}{r \in [n]}$).
-See \cref{fig:sourceTargetTrees}.
+See \cref{fig:strongRectangulation}.
 
@@
 
-We complete each node of~$S(R)$ and~$T(R)$ with a vertical (resp.~horizontal) leaf if it has no vertical (resp.~horizontal) child, see \cref{fig:sourceTargetTrees}.
+We complete each node of~$S(R)$ and~$T(R)$ with a vertical (resp.~horizontal) leaf if it has no vertical (resp.~horizontal) child, see \cref{fig:strongRectangulation}.
 Recall that the \defn{inorder labeling} of a binary tree~$T$ is the labeling of the nodes of~$T$ such that the label of each node~$t$ of~$T$ is larger than all labels in the left subtree of~$t$ and larger than all labels in the right subtree of~$t$.
@@
   \begin{split}
-    \loday{w}_i := & h(T, i)\cdot v(T,i) \\
-    \antiloday{w}_i := & h(S, i)\cdot v(S,i),
+    \loday{w}_i \eqdef & h(T, i)\cdot v(T,i) \\
+    \antiloday{w}_i \eqdef & h(S, i)\cdot v(S,i),
   \end{split}
   \]
-  for all weak rectangulations $R$ on $n$ rectangles, where $T:=T(R)$, $S:=S(R)$,
+  for all weak rectangulations $R$ on $n$ rectangles, where $T \eqdef T(R)$, $S \eqdef S(R)$,
   and $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
@@
 \end{theorem}
+
+\begin{example}
+The vertex of~$\WRP (16)$ corresponding to the weak rectangulation of \cref{fig:weakRectangulation} is
+\[
+(-3, 0, 26, 2, -9, 5, -69, -4, 17, 0, -8, 59, 2, -20, 0, 2).
+\]
+For instance, the $4$th coordinate is $h(T, 4) \cdot v(T, 4) - h(S, 4) \cdot v(S, 4) = 1 \cdot 5 - 1 \cdot 3 = 2$.
+\end{example}
 
@@
   \begin{split}
-    \yin{w}_{i,j} := & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot \llbracket \neg i\backslash j \rrbracket \\% \text { if not } i\backslash j, \text{and } 0\text{ otherwise.} \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
-    \yang{w}_{i,j} := & v(S, i) \cdot ch (S, T, i, j)\cdot v(T, j) \cdot \llbracket \neg j\backslash i \rrbracket
+    \yin{w}_{i,j} \eqdef & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot \llbracket \neg i\backslash j \rrbracket \\% \text { if not } i\backslash j, \text{and } 0\text{ otherwise.} \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
+    \yang{w}_{i,j} \eqdef & v(S, i) \cdot ch (S, T, i, j)\cdot v(T, j) \cdot \llbracket \neg j\backslash i \rrbracket
   \end{split}
@@
   \begin{itemize}
-  \item $T := T(R)$ and $S:=S(R)$,
+  \item $T \eqdef T(R)$ and $S \eqdef S(R)$,
   \item $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$,
@@
 % jean:{Is there better way to write this down/compute? Here we do not take advantage of the fact that at most one of the two factors $\yin{w}_{i,j}$ and $\yang{w}_{i,j}$ is nonzero.}
+
+\begin{example}
+The vertex of~$\WRP (16)$ corresponding to the weak rectangulation of \cref{fig:strongRectangulation} is
+\[
+???
+\]
+For instance, the $4$th coordinate is $???$.
+
+\end{example}
 
@@
 \[
-\mathcal{L}(P_w(R)) := \{\pi\in\f{S}_n : i\prec_w j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
+\mathcal{L}(P_w(R)) \eqdef \{\pi\in\f{S}_n : i\prec_w j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
 \]
@@
 	\centerline{\includegraphics[width=.5\textwidth]{weakRectangulation} \qquad \includegraphics[width=.5\textwidth]{weakRectangulationTrees}}
-	\caption{A diagonal rectangulation (left) and its pair of separated twin trees (right).
-The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.}
+	\caption{A diagonal rectangulation (left) and its pair of separated twin trees (right).}
+	% The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.
+	\label{fig:weakRectangulation}
 \end{figure}
@@
 
+\vincent{Say also the minkowski sum of all faces of the standard simplex corresponding to intervals, and all their opposite.}
+
 \subsection{Submodular functions of weak rectangulotopes}
@@
   The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
-  is $\mathbf{e}_i$, where $i := \arg\max_{j\in I} \pi^{-1}(j)$.
+  is $\mathbf{e}_i$, where $i \eqdef \arg\max_{j\in I} \pi^{-1}(j)$.
   Similarly, if $A$ is an anti-Loday arc defined by $I$, the unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
-  is $-\mathbf{e}_i$, where $i := \arg\min_{j\in I} \pi^{-1}(j)$.
+  is $-\mathbf{e}_i$, where $i \eqdef \arg\min_{j\in I} \pi^{-1}(j)$.
 \end{lemma}
@@
 \[
-\mathcal{L}(P_s(R)) := \{\pi\in\f{S}_n : i\prec_s j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
+\mathcal{L}(P_s(R)) \eqdef \{\pi\in\f{S}_n : i\prec_s j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
 \]
@@
   Let $\pi\in\f{S}_n$ and consider a yin arc $A=(I,J)$.
-  Let $i := \arg\max_{k\in I} \pi^{-1}(k)$ and $j := \arg\min_{k\in J} \pi^{-1}(k)$.
+  Let $i \eqdef \arg\max_{k\in I} \pi^{-1}(k)$ and $j \eqdef \arg\min_{k\in J} \pi^{-1}(k)$.
   The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$

### 2024-04-03 draft/rectangulotopes.tex
details

@@
   They are defined as quotientopes of natural congruence relations on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
-We give a thorough description of these polytopes, including elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedron.
+  We give a thorough description of geometric realizations of these polytopes, including elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedron.
 \end{abstract}
@@
 
-The relevant lattice congruences are the \defn{Baxter congruence} on one hand, studied by Law and Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}) and \defn{Baxter permutations}, and the \defn{strong rectangulation congruence}, whose classes are in bijection with \defn{strong rectangulations}~\cite{MR2864445}.
+The relevant lattice congruences are the \defn{weak rectangulation congruence} on one hand, studied by Law and Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}) and \defn{Baxter permutations}, and the \defn{strong rectangulation congruence}, whose classes are in bijection with \defn{strong rectangulations}~\cite{MR2864445}.
 The latter is mentioned by Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by Asinowski et al.~\cite{ACFF24}.
@@
 One of our main results is an elementary description of the vertex coordinates of a realization of both families of quotientopes, in the spirit of Loday's realization in Theorem~\ref{thm:loday}.
-The quotientopes corresponding to the Baxter congruence will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
+The quotientopes corresponding to the weak rectangulation congruence will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
 The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
@@
 %%%%%%%%%%%%
-\subsection{The Baxter congruence}
-
-% jean:{Clash of terminology: weak vs. Baxter?}
+\subsection{The weak rectangulation congruence}
 
@@
 
-We call this congruence the \defn{Baxter congruence}.
-The \defn{weak rectangulotope} $\WRP(n)$ is the quotientope of the Baxter congruence $\baxtereq$ on the weak Bruhat order on $\f{S}_n$.
+We call this congruence the \defn{weak rectangulation congruence}.
+The \defn{weak rectangulotope} $\WRP(n)$ is the quotientope of the weak rectangulation congruence $\baxtereq$ on the weak Bruhat order on $\f{S}_n$.
 
@@
 \begin{lemma}
-The Baxter congruence $\baxtereq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that do not cross the horizontal line.
+The weak rectangulation congruence $\baxtereq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that do not cross the horizontal line.
 \end{lemma}
@@
 \begin{lemma}
+  \label{lem:weakMinkowski}
   The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all Loday and anti-Loday shard polytopes. In particular, it is the Minkowski sum of two opposite Loday associahedra.
@@
 
+Weak rectangulotopes are deformed permutahedra, hence are defined by a \defn{submodular} set function $f:2^{[n]}\mapsto \R$.
+The description of the weak rectangulotope $\WRP(n)$ as a Minkowski sum in Lemma~\ref{lem:weakMinkowski} allows us to easily compute the submodular function defining it.
+
 \begin{lemma}
@@
   \]
-  where $f:2^{[n]}\mapsto \R$ is the submodular function that maps $S$ to the number of intervals
-  $I$ of $[n]$ such that $I\cap S\not=\emptyset$ and $I\not\subseteq S$. 
+  where $f:2^{[n]}\mapsto \R$ is the following submodular function:
+  \[
+  f(S) := |\{ I\text{ interval of } [n] : I\cap S\not=\emptyset \text{ and } I\cap ([n]\setminus S) \not= \emptyset \}|.
+  \]
+  %that maps $S$ to the number of intervals $I$ of $[n]$ such that $I\cap S\not=\emptyset$ and $I\not\subseteq S$. 
 \end{lemma}
+
+The proof relies on the following simple lemma stating that the submodular function of a Minkowski sum of deformed permutahedra is the sum of the submodular function of the summands.
+
+\begin{lemma}
+  \label{lem:submodsum}
+  Consider a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by the submodular functions $f_1,f_2,\ldots ,f_k$.
+Then the Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f:=\sum_i f_i$.
+\end{lemma}
+
+\begin{proof}
+  From Lemmas~\ref{lem:weakMinkowski} and~\ref{lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the Loday and anti-Loday shard polytope in Lemmas~\ref{lem:lodaysp} and~\ref{lem:antilodaysp}.
+For a Loday arc $A$ defined by an interval $I$ of $[n]$, we have
+  \[
+  \SP(A) = \{ x\in\R^n : \sum_{i\in S} x_i \leq \loday{f}_I(S) \text { and } \sum_{i\in [n]} x_i = \loday{f}_I([n]) \},
+  \]
+  where
+  \[
+  \loday{f}_I(S) := \llbracket S\cap I\not=\emptyset \rrbracket .
+  \]
+  Similarly, for an anti-Loday arc $A$ defined by an interval $I$, the shard polytope $\SP(A)$ is defined by the following submodular function:
+  \[
+  \antiloday{f}_I(S) := - \llbracket S\subseteq I \rrbracket .
+  \]
+  Their sum
+  \[
+  f(S) := \sum_{I\text{ interval of }[n]} \loday{f}_I(S) + \antiloday{f}_I(S)
+  \]
+  is therefore exactly the number of intervals of $I$ that intersect $S$ but are not contained in $S$, hence the number of intervals intersecting both $S$ and its complement.
+\end{proof}
 
@@
 
-Note that this extremal vertex should only depend on the Baxter congruence class of $\pi$.
-This, in passing, gives us an alternative characterization of the Baxter congruence:
+Note that this extremal vertex should only depend on the weak rectangulation congruence class of $\pi$.
+This, in passing, gives us an alternative characterization of the weak rectangulation congruence:
 $\pi\baxtereq\sigma$ if and only if for any interval $I$ of $[n]$,
@@
 
-% jean:{Does the congruence have a name?}
-
 \subsection{Yin and yang arcs}
@@
 \end{lemma}
-
-% jean:{From this, do we directly have a description of the face lattice of $\SRP(n)$?}
 
@@
 
-% jean:{From this, do we directly have the submodular function defining $\SRP(n)$?}
+\jean{Compute the submodular functions of each yin and yang shard polytopes and sum them.}
 

### 2024-04-03 draft/rectangulotopes.tex
minor changes statements theorems

--- 2024-04-02 draft/rectangulotopes.tex
@@
 % Loday and anti-Loday accents (?)
-\newcommand{\loday}[1]{\overset{\frown}{#1}}
-\newcommand{\antiloday}[1]{\overset{\smile}{#1}}
+\newcommand{\loday}[1]{\smash{\overset{\frown}{#1}}}
+\newcommand{\antiloday}[1]{\smash{\overset{\smile}{#1}}}
 % yin and yang accents (?)
-\newcommand{\yang}[1]{\overset{\backsim}{#1}}
-\newcommand{\yin}[1]{\overset{\sim}{#1}}
-% weak (Baxter) and strong equivalence on S_n
+\newcommand{\yang}[1]{\smash{\overset{\backsim}{#1}}}
+\newcommand{\yin}[1]{\smash{\overset{\sim}{#1}}}
+% weak and strong equivalence on S_n
 \newcommand{\baxtereq}{\asymp}%{\stackrel{\frown}{\smile}}
@@
 An elementary yet striking example of this idea is the \defn{permutahedron} $\Perm (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\pi\in\f{S}_n$.
-Its facets are indexed by the proper nonempty subsets of $[n]$, while its skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
-The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra} (also known as \defn{generalized permutahedra}).
+It is an $(n-1)$-dimensional polytope, whose facets are indexed by the proper nonempty subsets of $[n]$, and whose skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
+The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{generalized permutahedra}~\cite{Postnikov} (or \defn{deformed permutahedra}).
 
@@
 It was first defined as a combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894}.
-The face lattice of the associahedron~$\Asso (n)$ is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
+The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
 In particular, it skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.
@@
   \label{thm:loday}
-  Let $\Asso (n)$ be the convex hull of the points
-  $\sum_{i\in [n]} w_i\cdot \mathbf{e}_i$, where
+  The $(n-1)$-dimensional associahedron is realized by the convex hull~$\Asso (n)$ of the points
+  $\sum_{i\in [n]} w^T_i \cdot \mathbf{e}_i$ for all binary trees $T$ on $n$ internal nodes, with
   \[
-  w_i \eqdef \ell(T, i)\cdot r(T,i),
+  w^T_i \eqdef \ell^T_i\cdot r^T_i,
   \]
-  for all binary trees $T$ on $n$ internal nodes, where $\ell(T,i)$ and $r(T,i)$ denote respectively the number of leaves in the left and right subtrees of $i$ in $T$.
-  Then $\Asso (n)$ is a realization of the $(n-1)$-dimensional associahedron.
+  where $\ell^T_i$ and $r^T_i$ respectively denote the number of leaves in the left and right subtrees of $i$ in $T$.
 \end{theorem}
@@
 The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} relation called the \defn{sylvester congruence}.
-Pilaud and Santos~\cite{MR3964495}, answering a question of Reading~\cite{MR2142177}, proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more generally whose normal fan is the fan defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
+Pilaud and Santos~\cite{MR3964495}, answering a question of Reading~\cite{MR2142177}, proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more precisely whose normal fan is the \defn{quotient fan} defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
 Those polytopes are generalized permutahedra that they called \defn{quotientopes}.
+Associahedra are therefore the quotientopes of a congruence of the weak Bruhat order whose classes are in bijection with triangulations.
 
-Associahedra are therefore the quotientopes of a congruence of the weak Bruhat order whose classes are in bijection with triangulations.  
-The goal of this paper is to initiate the study of quotientopes of congruences of the weak Bruhat order whose classes are in bijection with (equivalence classes of) \defn{rectangulations}, defined as decompositions of a square into finitely many axis-aligned rectangles.
+The goal of this paper is to describe the quotientopes of congruences of the weak Bruhat order whose classes are in bijection with (equivalence classes of) \defn{rectangulations}, defined as decompositions of a square into finitely many axis-aligned rectangles.
 The combinatorics of rectangulations has been studied for decades~\cite{MR2233287,MR2871762,MR2864445,MR2763051,MR3084577,MR3878132,MR4598046}, and the order-theoretic take on these structures has already proved useful.
@@
 
-The relevant lattice congruences are the \defn{Baxter congruence} on one hand, studied by Law and Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}) and \defn{Baxter permutations}, and the \defn{strong rectangulation congruence}, whose classes are in bijection with \defn{strong rectangulations}~\cite{MR2864445}.
-The latter is mentioned by Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by Asinowski et al.~\cite{ACFF24}.
+The relevant lattice congruences are the \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) on one hand, studied by Law and Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}) and \defn{Baxter permutations}, and the \defn{strong rectangulation congruence}, whose classes are in bijection with \defn{strong rectangulations}~\cite{MR2864445}.
+The latter is mentioned by Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by Asinowski, Cardinal, Felsner, and Fusy~\cite{ACFF24}.
 
 One of our main results is an elementary description of the vertex coordinates of a realization of both families of quotientopes, in the spirit of Loday's realization in Theorem~\ref{thm:loday}.
-The quotientopes corresponding to the Baxter congruence will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
+The quotientopes corresponding to the weak rectangulation congruence will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
 The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
@@
 % jean:{Here only give the definitions allowing to state the results.}
+
+\begin{figure}
+	\centerline{\includegraphics[width=.5\textwidth]{weakRectangulation} \qquad \includegraphics[width=.5\textwidth]{weakRectangulationTrees}}
+	\caption{A weak rectangulation (left) and its pair of separated twin trees (right). Example from \cite{ACFF24}.}
+	% The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.
+	\label{fig:weakRectangulation}
+\end{figure}
 
@@
 \begin{theorem}
-  Let $\WRP (n)$ be the convex hull of the points
-  $\sum_{i\in [n]} (\loday{w}_i - \antiloday{w}_i)\cdot \mathbf{e}_i$,
-  where
+  The $(n-1)$-dimensional weak rectangulotope is realized by the convex hull $\WRP (n)$ of the points
+  $\sum_{i\in [n]} (\loday{w}^R_i - \antiloday{w}^R_i)\cdot \mathbf{e}_i$ for all weak rectangulations $R$ on $n$ rectangles,
+  with
   \[
-  \begin{split}
-    \loday{w}_i \eqdef & h(T, i)\cdot v(T,i) \\
-    \antiloday{w}_i \eqdef & h(S, i)\cdot v(S,i),
-  \end{split}
+%  \begin{split}
+%    \loday{w}^R_i \eqdef & h(T, i)\cdot v(T,i) \\
+%    \antiloday{w}^R_i \eqdef & h(S, i)\cdot v(S,i),
+%  \end{split}
+    \loday{w}^R_i \eqdef h^T_i\cdot v^T_i
+    \qquad\text{and}\qquad
+    \antiloday{w}^R_i \eqdef h^S_i\cdot v^S_i,
   \]
-  for all weak rectangulations $R$ on $n$ rectangles, where $T \eqdef T(R)$, $S \eqdef S(R)$,
-  and $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
-  Then $\WRP (n)$ is a realization of the $(n-1)$-dimensional weak rectangulotope.
+  where~$S$ and~$T$ are the source and target trees of the rectangulation~$R$, and $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$.
 \end{theorem}
@@
   \]
-For instance, the $4$th coordinate is $h(T, 4) \cdot v(T, 4) - h(S, 4) \cdot v(S, 4) = 1 \cdot 5 - 1 \cdot 3 = 2$.
+  For instance, the $4$th coordinate is $h^T_4 \cdot v^T_4 - h^S_4 \cdot v^S_4 = 1 \cdot 5 - 1 \cdot 3 = 2$.
 \end{example}
@@
 In order to exhibit a similar realization for the strong rectangulotopes, we need to count subsets of leaves that lie in some subtrees of both the source and target trees of a rectangulation.
-Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$. We denote by %$i|j$ and $\frac{\ j\ }i$
-$i\backslash j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$. %, respectively.
+Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$.
+We denote by %$i|j$ and $\frac{\ j\ }i$
+$i \, \backslash \, j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$. %, respectively.
 
 \begin{theorem}
-  Let $\SRP (n)$ be the convex hull of the points
+  The $(n-1)$-dimensional strong rectangulotope is realized by the convex hull~$\SRP (n)$ of the points
   \[
-  \sum_{i,j\in [n], i< j} (\yin{w}_{i,j} - \yang{w}_{i,j})\cdot (\mathbf{e}_i - \mathbf{e}_j),
+  \sum_{i,j\in [n], i< j} (\yin{w}^R_{i,j} - \yang{w}^R_{i,j})\cdot (\mathbf{e}_i - \mathbf{e}_j),
   \]
-  with
+   for all strong rectangulations $R$ on $n$ rectangles, with
   \[
-  \begin{split}
-    \yin{w}_{i,j} \eqdef & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot \llbracket \neg i\backslash j \rrbracket \\% \text { if not } i\backslash j, \text{and } 0\text{ otherwise.} \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
-    \yang{w}_{i,j} \eqdef & v(S, i) \cdot ch (S, T, i, j)\cdot v(T, j) \cdot \llbracket \neg j\backslash i \rrbracket
-  \end{split}
+%  \begin{split}
+%    \yin{w}_{i,j} \eqdef & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot \llbracket \neg i\backslash j \rrbracket \\% \text { if not } i\backslash j, \text{and } 0\text{ otherwise.} \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
+%    \yang{w}_{i,j} \eqdef & v(S, i) \cdot ch (S, T, i, j)\cdot v(T, j) \cdot \llbracket \neg j\backslash i \rrbracket,
+%  \end{split}
+    \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \neg i \, \backslash \, j \rrbracket
+    \qquad\text{and}\qquad
+    \yang{w}^R_{i,j} \eqdef v^S_i \cdot ch^{S,T}_{i,j}\cdot v§T_j \cdot \llbracket \neg j \, \backslash \, i \rrbracket,
   \]
-
-%  \begin{cases}
-%    h(T(R), i) \cdot cv (T, S, i, j)\cdot h(S, j) & \text{ if neither } i|j \text{ nor } \frac{\ j\ }i, \\
-%    0 & \text{ otherwise.}
-%  \end{cases}
-
-  %\begin{cases}
-  %  v(S, j) \cdot ch (T, S, i, j)\cdot v(T, i) & \text{ if neither } j|i \text{ nor } \frac{\ i\ }j, \\
-  %  0 & \text{ otherwise.}
-  %\end{cases}
-
-  % jean:{Double check the formula.}
-  for all strong rectangulations $R$ on $n$ rectangles, where:
+ where:
   \begin{itemize}
-  \item $T \eqdef T(R)$ and $S \eqdef S(R)$,
-  \item $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$,
-  \item $ch (T,S,i,j)$ and $cv (T,S,i,j)$ denote respectively the number of common leaves of the horizontal (vertical) subtree of $i$ in $T$ and the horizontal (vertical) subtree of $j$ in $S$
+  \item $S$ and $T$ are the source and target trees of the rectangulation~$R$,
+  \item $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
+  \item $ch^{T,S}_{i,j}$ (resp.~$cv^{T,S}_{i,j}$) denote the number of common leaves of the horizontal (resp.~vertical) subtree of $i$ in $T$ and the horizontal (resp.~vertical) subtree of $j$ in $S$.
   \end{itemize}
-Then $\SRP (n)$ is a realization of the $(n-1)$-dimensional strong rectangulotope.  
 \end{theorem}
@@
 This work was initiated at the Workshop on Combinatorics, Algorithms, and Geometry organized by Torsten M\"utze on March 4-8, 2024 in Dresden, Germany.
-The authors thank Torsten M\"utze and Namrata for the organization and the other participants of the workshop for the stimulating interactions.
+The authors thank Namrata and Torsten M\"utze for the organization and the other participants of the workshop for the stimulating interactions, notably on other combinatorial aspects of rectangulations.
 
@@
 %%%%%%%%%%%%
-\subsection{The Baxter congruence}
 
-% jean:{Clash of terminology: weak vs. Baxter?}
+\subsection{The weak rectangulation congruence}
 
@@
 
-\begin{figure}
-	\centerline{\includegraphics[width=.5\textwidth]{weakRectangulation} \qquad \includegraphics[width=.5\textwidth]{weakRectangulationTrees}}
-	\caption{A diagonal rectangulation (left) and its pair of separated twin trees (right).}
-	% The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.
-	\label{fig:weakRectangulation}
-\end{figure}
-
-We call this congruence the \defn{Baxter congruence}.
-The \defn{weak rectangulotope} $\WRP(n)$ is the quotientope of the Baxter congruence $\baxtereq$ on the weak Bruhat order on $\f{S}_n$.
+We call this congruence the \defn{weak rectangulation congruence} (it is sometimes referred to as the \defn{Baxter congruence}).
+The \defn{weak rectangulotope} $\WRP(n)$ is the quotientope of the weak rectangulation congruence $\baxtereq$ on the weak Bruhat order on $\f{S}_n$.
 
@@
 \begin{lemma}
-  The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all Loday and anti-Loday shard polytopes. In particular, it is the Minkowski sum of two opposite Loday associahedra.
+  The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all Loday and anti-Loday shard polytopes.
+  In particular, it is the Minkowski sum of two opposite Loday associahedra.
 \end{lemma}

### 2024-04-03 draft/rectangulotopes.tex

@@
   They are defined as quotientopes of natural congruence relations on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
-  We give a thorough description of these polytopes, including elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedron.
+  We give a thorough description of geometric realizations of these polytopes, including elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedron.
 \end{abstract}
@@
 \begin{lemma}
+  \label{lem:weakMinkowski}
   The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all Loday and anti-Loday shard polytopes.
@@
   \]
-  where $f:2^{[n]}\mapsto \R$ is the submodular function that maps $S$ to the number of intervals
-  $I$ of $[n]$ such that $I\cap S\not=\emptyset$ and $I\not\subseteq S$.
+  where $f:2^{[n]}\mapsto \R$ is the following submodular function:
+  \[
+  f(S) := |\{ I\text{ interval of } [n] : I\cap S\not=\emptyset \text{ and } I\cap ([n]\setminus S) \not= \emptyset \}|.
+  \]
+  %that maps $S$ to the number of intervals $I$ of $[n]$ such that $I\cap S\not=\emptyset$ and $I\not\subseteq S$. 
 \end{lemma}
+
+The proof relies on the following simple lemma stating that the submodular function of a Minkowski sum of deformed permutahedra is the sum of the submodular function of the summands.
+
+\begin{lemma}
+  \label{lem:submodsum}
+  Consider a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by the submodular functions $f_1,f_2,\ldots ,f_k$.
+Then the Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f:=\sum_i f_i$.
+\end{lemma}
+
+\begin{proof}
+  From Lemmas~\ref{lem:weakMinkowski} and~\ref{lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the Loday and anti-Loday shard polytope in Lemmas~\ref{lem:lodaysp} and~\ref{lem:antilodaysp}.
+For a Loday arc $A$ defined by an interval $I$ of $[n]$, we have
+  \[
+  \SP(A) = \{ x\in\R^n : \sum_{i\in S} x_i \leq \loday{f}_I(S) \text { and } \sum_{i\in [n]} x_i = \loday{f}_I([n]) \},
+  \]
+  where
+  \[
+  \loday{f}_I(S) := \llbracket S\cap I\not=\emptyset \rrbracket .
+  \]
+  Similarly, for an anti-Loday arc $A$ defined by an interval $I$, the shard polytope $\SP(A)$ is defined by the following submodular function:
+  \[
+  \antiloday{f}_I(S) := - \llbracket S\subseteq I \rrbracket .
+  \]
+  Their sum
+  \[
+  f(S) := \sum_{I\text{ interval of }[n]} \loday{f}_I(S) + \antiloday{f}_I(S)
+  \]
+  is therefore exactly the number of intervals of $I$ that intersect $S$ but are not contained in $S$, hence the number of intervals intersecting both $S$ and its complement.
+\end{proof}
 

### 2024-04-03 draft/rectangulotopes.tex
details

--- 2024-04-03 draft/rectangulotopes.tex
@@
 $i\backslash j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$. %, respectively.
+We also use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.
 
@@
   \item $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$,
-  \item $ch (T,S,i,j)$ and $cv (T,S,i,j)$ denote respectively the number of common leaves of the horizontal (vertical) subtree of $i$ in $T$ and the horizontal (vertical) subtree of $j$ in $S$
+  \item $ch (T,S,i,j)$ and $cv (T,S,i,j)$ denote respectively the number of common leaves of the horizontal (vertical) subtree of $i$ in $T$ and the horizontal (vertical) subtree of $j$ in $S$.
   \end{itemize}
@@
 Section~\ref{sec:wr} details the case of weak rectangulations, while Section~\ref{sec:sr} deals with strong rectangulations.
-In both cases, in addition to the vertex coordinates, we give the description of the face lattices and the submodular functions defining the polytopes.
+In both cases, in addition to the vertex coordinates, we also give the submodular functions defining the polytopes.
 
@@
   \]
+  (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and 0 otherwise.)
   Similarly, for an anti-Loday arc $A$ defined by an interval $I$, the shard polytope $\SP(A)$ is defined by the following submodular function:
   \[
-  \antiloday{f}_I(S) := - \llbracket S\subseteq I \rrbracket .
+  \antiloday{f}_I(S) := - \llbracket I\subseteq S \rrbracket .
   \]

### 2024-04-03 draft/rectangulotopes.tex

--- 2024-04-03 draft/rectangulotopes.tex
@@
 $i \, \backslash \, j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$. %, respectively.
+We also use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.
 

### 2024-04-03 draft/rectangulotopes.tex
abstract

@@
   They are defined as quotientopes of natural congruence relations on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
-  We give a thorough description of geometric realizations of these polytopes, including elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedron.
+  We give simple geometric realizations of these polytopes, including elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedron.
 \end{abstract}

### 2024-04-03 draft/rectangulotopes.tex
sect 3.4

@@
 
-Weak rectangulotopes are deformed permutahedra, hence are defined by a \defn{submodular} set function $f:2^{[n]}\mapsto \R$.
+Recall that every deformed permutahedra can be defined as
+  \[
+  \{x\in\R^n : \sum_{i\in S} x_i \leq f(S)\ \forall S\subseteq [n], S\not=\emptyset, \text { and } \sum_{i\in [n]} x_i = f([n])\},
+  \]
+where $f$ is a \defn{submodular} set function $f:2^{[n]}\mapsto \R$.
 The description of the weak rectangulotope $\WRP(n)$ as a Minkowski sum in Lemma~\ref{lem:weakMinkowski} allows us to easily compute the submodular function defining it.
@@
 \begin{lemma}
-  The weak rectangulotope $\WRP(n)$ is realized as
-  \[
-  \{x\in\R^n : \sum_{i\in S} x_i \leq f(S) \text { and } \sum_{i\in [n]} x_i = f([n])\},
-  \]
-  where $f:2^{[n]}\mapsto \R$ is the following submodular function:
+  The weak rectangulotope $\WRP(n)$ is realized by the following submodular function:
   \[
@@
   From Lemmas~\ref{lem:weakMinkowski} and~\ref{lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the Loday and anti-Loday shard polytope in Lemmas~\ref{lem:lodaysp} and~\ref{lem:antilodaysp}.
-For a Loday arc $A$ defined by an interval $I$ of $[n]$, we have
-  \[
-  \SP(A) = \{ x\in\R^n : \sum_{i\in S} x_i \leq \loday{f}_I(S) \text { and } \sum_{i\in [n]} x_i = \loday{f}_I([n]) \},
-  \]
-  where
+For a Loday arc $A$ defined by an interval $I$ of $[n]$, we have that its shard polytope $\SP(A)$ is realized by the submodular function
   \[

### 2024-04-03 draft/rectangulotopes.tex
quotientopes

--- 2024-04-03 draft/rectangulotopes.tex
@@
 $i \, \backslash \, j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$. %, respectively.
-We also use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.
 
@@
 %  \end{split}
-    \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \neg i \, \backslash \, j \rrbracket
+    \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \neg \, i \, \backslash \, j \rrbracket
     \qquad\text{and}\qquad
-    \yang{w}^R_{i,j} \eqdef v^S_i \cdot ch^{S,T}_{i,j}\cdot v§T_j \cdot \llbracket \neg j \, \backslash \, i \rrbracket,
+    \yang{w}^R_{i,j} \eqdef v^S_i \cdot ch^{S,T}_{i,j}\cdot v^T_j \cdot \llbracket \neg \, j \, \backslash \, i \rrbracket,
   \]
@@
 \begin{example}
-The vertex of~$\WRP (16)$ corresponding to the weak rectangulation of \cref{fig:strongRectangulation} is
+The vertex of~$\SRP (16)$ corresponding to the strong rectangulation of \cref{fig:strongRectangulation} is
 \[
@@
 Section~\ref{sec:wr} details the case of weak rectangulations, while Section~\ref{sec:sr} deals with strong rectangulations.
-In both cases, in addition to the vertex coordinates, we also give the submodular functions defining the polytopes.
+In both cases, in addition to the vertex coordinates, we give the description of the face lattices and the submodular functions defining the polytopes.
 
@@
 
-\subsection{Quotients of the weak Bruhat order on permutations}
+We now briefly recall some results on the weak order and its quotients, both from a lattice and geometric perspectives.
+We refer to \cite{Reading-arcDiagrams, Reading-Chapters, PadrolPilaudRitter} for details.
 
-\subsection{Arc diagrams}
+\subsection{Weak order and noncrossing arc diagrams}
 
-\begin{theorem}
-  Congruences on the weak Bruhat order are one-to-one with arc ideals.
-\end{theorem}
+Denote by~$\f{S}_n$ the set of permutations of~$[n]$.
+The \defn{inversion set} of~$\sigma \in \f{S}_n$ is~$\inv(\sigma) \eqdef \set{(\sigma_i, \sigma_j)}{1 \le i < j \le n \text{ and } \sigma_i > \sigma_j}$.
+The \defn{weak order} is the lattice on~$\f{S}_n$ defined by the inclusion of their inversion sets.
+Note that the cover relations in the weak order are given by the transposition of two adjacent letters.
+%The weak order is known to be a congruence uniform lattice~\cite{}.
+%In particular, it is join semidistributive, so that any permutation admits a canonical join representation.
 
-\subsection{Shard polytopes}
+An \defn{arc} on~$[n]$ is a quadruple~$(a, b, A, B)$ where~$1 \le a < b \le n$ and~$A \sqcup B$ forms a partition of the interval~${]a,b[} \eqdef \{a+1, \dots, b-1\}$.
+We represent an arc by an abscissa monotone curve wiggling around the horizontal axis, starting at point~$a$ and ending at point~$b$, and passing above the points of~$A$ and below the points of~$B$.
+A \defn{noncrossing arc diagram} on~$[n]$ is a collection of arcs on~$[n]$ where any two arcs do not cross in their interior and have distinct left endpoints and distinct right endpoints (but the right endpoint of an arc can be the right endpoint of another arc).
 
-\begin{theorem}
-  The quotientope of a congruence $\equiv$ on the weak Bruhat order is realized by the Minkowski sum of the shard polytopes corresponding to the arcs in the arc ideal defining $\equiv$.
-\end{theorem}
+In~\cite{Reading-arcDiagrams}, N.~Reading defined an elegant bijection between permutations of~$[n]$ and noncrossing arc diagrams on~$[n]$.
+It sends a permutation~$\sigma$ to the noncrossing arc diagram with an arc~$(\sigma_{j+1}, \sigma_j, A_j, B_j)$ for each descent~$j \in [n-1]$ of~$\sigma$ (\ie with~$\sigma_j > \sigma_{j+1}$), where \linebreak $A_j \eqdef \set{\sigma_i}{1 \le i < j \text{ and } \sigma_{j+1} < \sigma_i < \sigma_j}$ and~$B_j \eqdef \set{\sigma_k}{j+1 < k \le n \text{ and } \sigma_{j+1} < \sigma_k < \sigma_j}$.
+\linebreak
+This can also been visualized by representing the table~$(\sigma_j,j)$ of the permutation~$\sigma$, drawing the segments corresponding to the descents of~$\sigma$, and letting all points fall on the horizontal axis, allowing the segments to bend but not to cross each other nor to pass through a point.
+The single arcs correspond to permutations with a single descent, that is, to join irreducible permutations of the weak order.
+In general, the noncrossing arc diagram of a permutation~$\sigma$ actually encodes the canonical join representation of~$\sigma$ in the weak order (which was known to exist, as the weak order is join semidistributive).
+See~\cite{Reading-arcDiagrams} for details.
+
+\subsection{Quotients and arc ideals}
+
+A \defn{lattice congruence} of the weak order is an equivalence relation~$\equiv$ on~$\f{S}_n$ that respects the meet and join operations, \ie such that $x \equiv x'$ and~$y \equiv y'$ implies $x \meet y \, \equiv \, x' \meet y'$ and~$x \join y \, \equiv \, x' \join y'$.
+The \defn{lattice quotient}~$\f{S}_n/{\equiv}$ is the lattice on the congruence classes of~$\equiv$ where~$X \le Y$ if and only if there exist~$x \in X$ and~$y \in Y$ such that~$x \le y$, and~$X \meet Y$ (resp.~$X \join Y$) is the congruence class of~$x \meet y$ (resp.~$x \join y$) for any~$x \in X$~and~$y \in Y$.
+
+Respecting the meet and join operation is a strong condition that imposes additional structure on~$\equiv$.
+In fact, all the congruence classes of~$\equiv$ are intervals of the weak order, and the quotient~$\f{S}_n/{\equiv}$ is isomorphic (as a poset) to the subposet of the weak order induced by permutations which are minimal in their class.
+Moreover, the latter are precisely the permutations whose noncrossing arc diagrams only use arcs corresponding to join irreducible permutations which are minimal in their class -- we denote by~$\c{A}_\equiv$ this set of arcs.
+In other words, the classes of~$\equiv$ are in bijection with noncrossing arc diagrams using only arcs in~$\c{A}_\equiv$.
+This bijection actually translates the fact that canonical join representations behave properly under lattice quotients.
+
+Moreover, the sets of arcs of the form~$\c{A}_\equiv$ are easily described.
+An arc~$(a, b, A, B)$ is a \defn{subarc} of an arc~$(a', b', A', B')$ if~$a' \le a < b \le b$ and~$A \subseteq A'$ while~$B \subseteq B'$.
+An \defn{arc ideal} is a subset of arcs closed by subarcs.
+The map~${\equiv} \mapsto \c{A}_\equiv$ is a bijection between the lattice congruences of the weak order on~$\f{S}_n$ and the arc ideals.
+(In other words, the lattice of congruences of the weak order is distributive, and its poset of join irreducibles is isomorphic to the subarc order).
+
+\subsection{Quotient fans and shards}
+
+The \defn{braid arrangement} is the arrangement of hyperplanes~$\set{\b{x} \in \R^n}{x_a = x_b}$ for all~$1 \le a < b \le n$.
+It has a chamber for each permutation~$\sigma$ of~$[n]$, given by the set of vectors whose coordinates are ordered as~$\sigma$.
+
+The \defn{shard} of an arc~$(a, b, A, B)$ is the piece of the braid hyperplane~${x_a = x_b}$ defined by the inequalities~$x_{a'} < x_a$ for all~$a' \in A$ and~$x_b < x_{b'}$ for all~$b' \in B$.
+%$\Sigma(a, b, A, B) \eqdef \set{\b{x} \in \R^n}{x_a = x_b \text{ and } x_{a'} < x_a \text{ for all } a' \in A \text{ and }  x_b < x_{b'} \text{ for all } b' \in B}$.
+The \defn{quotient fan} of a lattice congruence~$\equiv$ of the weak order is the polyhedral fan~$\c{F}_\equiv$ where
+\begin{itemize}
+\item the maximal cones are obtained by glueing together the chambers of the braid arrangement corresponding to permutations in the same congruence class of~$\equiv$,
+\item the union of the codimension~$1$ cones is the union of the shards of the arcs of~$\c{A}_\equiv$.
+\end{itemize}
+(These two descriptions are equivalent.)
+By construction, the dual graph of the braid fan~$\c{F}_\equiv$ is isomorphic to the cover graph of the quotient~$\f{S}_n/{\equiv}$.
+
+\subsection{Quotientopes and shard polytopes}
+
+A \defn{quotientope} for a lattice congruence~$\equiv$ of the weak order is a polytope whose normal fan is the quotient fan~$\c{F}_\equiv$.
+In particular, the skeleton of a quotientope is isomorphic to the cover graph of the quotient~$\f{S}_n/{\equiv}$.
+The existence of quotientopes was first proved in \cite{PilaudSantos} and later better understood in~\cite{PadrolPilaudRitter} using Minkowski sums of shard polytopes.
+
+Consider an arc~$\alpha \eqdef (a, b, A, B)$ on~$[n]$.
+An \defn{$\alpha$-alternating matching} is a sequence~$a \le i_1 < j_1 < i_2 < j_2 < \dots < i_q < j_q \le b$ such that~$i_p \in \{a\} \cup A$ and~$j_p \in \{b\} \cup B$ for all~$p \in [q]$.
+Its \defn{characteristic vector} is~$\sum_{p \in [q]} \b{e}_{i_p} - \b{e}_{j_p}$.
+The \defn{shard polytope} of~$\alpha$ is the convex hull~$\SP(\alpha)$ of the characteristic vectors of all $\alpha$-alternating matchings.
+It was shown in~\cite{PadrolPilaudRitter} that any Minkowski sum of positive scalings of the shard polytopes~$\SP(\alpha)$ for all arcs~$\alpha$ in the arc ideal~$\c{A}_\equiv$ is a quotientope for~$\equiv$.
 
@@
   \]
-  (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and 0 otherwise.)
   Similarly, for an anti-Loday arc $A$ defined by an interval $I$, the shard polytope $\SP(A)$ is defined by the following submodular function:
   \[
-  \antiloday{f}_I(S) := - \llbracket I\subseteq S \rrbracket .
+  \antiloday{f}_I(S) := - \llbracket S\subseteq I \rrbracket .
   \]

### 2024-04-03 draft/rectangulotopes.tex
strong submodular

--- 2024-04-03 draft/rectangulotopes.tex
@@
 \usepackage{todonotes}
-\newcommand{\vincent}[1]{\todo[color=blue!30]{\rm #1 \\ \hfill --- V.}}
-\newcommand{\Vincent}[1]{\todo[inline, color=blue!30]{\rm #1 \\ \hfill --- V.}}
+\newcommand{\vincent}[1]{\todo[size=\scriptsize, color=blue!30]{\rm #1 \\ \hfill --- V.}}
+\newcommand{\Vincent}[1]{\todo[inline, size=\scriptsize, color=blue!30]{\rm #1 \\ \hfill --- V.}}
 \newcommand{\jean}[2][]{\todo[size=\scriptsize, color=orange!30,#1]{\rm #2 \\ \hfill --- J.}}
@@
 
+\jean{You may want to use another notation here, feel free.}
+
 Recall that every deformed permutahedra can be defined as
@@
 \begin{lemma}
-  \label{lem:strongsum}
+  \label{lem:strongMinkowski}
   The strong rectangulotope $\SRP(n)$ is realized by the Minkowski sum of all shard polytopes of yin and yang arcs.
@@
 
-\jean{Compute the submodular functions of each yin and yang shard polytopes and sum them.}
+\jean{That seems to work.}
+
+The submodular function corresponding to our realization of $\SRP(n)$ can be obtained as before by summing the functions defining the yin and yang shard polytopes in Lemmas~\ref{lem:yinsp} and~\ref{lem:yangsp}.
+
+\begin{lemma}
+  The strong rectangulotope $\SRP(n)$ is realized by the following submodular function:
+  \[
+  f(S) := |\{ I, J : I\cap S\not=\emptyset \wedge J\cap ([n]\setminus S) \not= \emptyset \}| +
+  |\{ I, J : J\cap S\not=\emptyset \wedge I\cap ([n]\setminus S) \not= \emptyset \}|,
+  \]
+  where $I, J$ are pairs of consecutive and nonempty intervals of $[n]$.
+\end{lemma}
+
+\begin{proof}
+  From Lemmas~\ref{lem:strongMinkowski} and~\ref{lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the yin and yang shard polytope in Lemmas~\ref{lem:yinsp} and~\ref{lem:yangsp}.
+For a yin arc $A$ defined by consecutive intervals $I, J$ of $[n]$, we have that its shard polytope $\SP(A)$ is realized by the submodular function
+  \[
+  \yin{f}_{I,J}(S) := \llbracket S\cap I\not=\emptyset \wedge S\not\supseteq J \rrbracket .
+  \]
+  Similarly, for a yang arc, the shard polytope $\SP(A)$ is defined by the following submodular function:
+  \[
+  \yang{f}_{I,J}(S) := \llbracket S\cap J\not=\emptyset \wedge S\not\supseteq I \rrbracket .
+  \]
+  Their sum
+  \[
+  f(S) := \sum_{I,J} \yin{f}_{I,J}(S) + \yang{f}_{I,J}(S)
+  \]
+  is therefore as claimed.
+\end{proof}
 

### 2024-04-03 programmation/submodular.py

@@ -2,6 +2,8 @@
 
 from itertools import *
 
+## SUBMODULAR FUNCTIONS
+
 # simplex for a Loday arc I
 # 1 if S and I intersects, 0 otherwise
 def f_loday(S, I):
@@ -10,17 +12,34 @@ def f_loday(S, I):
 # anti-simplex for an anti-Loday arc I
 # -1 if I is a subset of S, 0 otherwise
 def f_antiloday(S, I):
-    return - int(bool(set(I).issubset(set(S))))
+    return - int(set(I).issubset(set(S)))
 
-# simplex + antisimplex
+# simplex + antisimplex (cuboctahedron in 3d)
 # 1 if S and I intersects but I is not a subset of S, 0 otherwise
 def f_cuboc(S, I):
-    return int(bool(set(S) & set (I)) and not bool(set(I).issubset(set(S))))
+    return int(bool(set(S) & set (I)) and not set(I).issubset(set(S)))
 
 # weak rectangulation
 # number of I such that S and I intersects but I is not a subset of S
 def f_weakrect(S, n):
-    return sum([int(bool(set(S) & set (range(i, j+1))) and not bool(set(range(i, j+1)).issubset(set(S)))) for i in range(n-1) for j in range(i+1, n)])
+    return sum([int(bool(set(S) & set (range(i, j+1))) and not set(range(i, j+1)).issubset(set(S))) for i in range(n-1) for j in range(i+1, n)])
+
+# yin shard polytopes
+# 1 if and only if S intersects I but J is not a subset of S
+def f_yin(S, I, J):
+    return int(bool(set(S) & set(I)) and not (set(J).issubset(S)))
+
+# yin rectangulations
+# number of pairs I, J such that S intersects I but J is not a subset of S
+def f_yinrect(S, n):
+    return sum([int (bool(set(S) & set(range(i, j+1))) and not (set(range(j+1, k+1)).issubset(S))) for i in range(n-1) for j in range(i, n-1) for k in range(j+1, n)])
+
+# yang shard polytopes
+# 1 if and only if S intersects J but I is not a subset of S
+def f_yang(S, I, J):
+    return int(bool(set(S) & set(J)) and not (set(I).issubset(S)))
+
+## GENERIC FUNCTIONS
 
 # powerset recipe
 def powerset(iterable):
@@ -33,12 +52,9 @@ def submod_polytope(f, n):
     all_ineq = []
[…]

### 2024-04-04 draft/rectangulotopes.tex
minor improvements section quotientopes

@@
 % weak and strong equivalence on S_n
-\newcommand{\baxtereq}{\asymp}%{\stackrel{\frown}{\smile}}
-\newcommand{\strongeq}{\simeq}% % J: I would like to stack \backsim and \sim instead
+\newcommand{
+\weakeq}{\asymp}
+%\makeatletter
+%\DeclareRobustCommand{\strongeq}{%
+%  \mathbin{\mathpalette\scaletoplus@{\mkern1mu \smash{\begin{smallmatrix} \backsim \\[-.3cm] \sim \end{smallmatrix}} \mkern1mu}}%
+%}
+%\newcommand{\scaletoplus@}[2]{%
+%  \vcenter{%
+%    \hbox{\sbox\z@{$\m@th#1+$}\resizebox{\wd\z@}{!}{$\m@th#1#2$}}%
+%  }%
+%}
+%\makeatother
+\newcommand{\strongeq}{\mathbin{\smash{\begin{smallmatrix} \backsim \\[-.3cm] \sim \end{smallmatrix}}}}% % J: I would like to stack \backsim and \sim instead
 
@@
 The \defn{inversion set} of~$\sigma \in \f{S}_n$ is~$\inv(\sigma) \eqdef \set{(\sigma_i, \sigma_j)}{1 \le i < j \le n \text{ and } \sigma_i > \sigma_j}$.
-The \defn{weak order} is the lattice on~$\f{S}_n$ defined by the inclusion of their inversion sets.
-Note that the cover relations in the weak order are given by the transposition of two adjacent letters.
+The \defn{weak order} is the lattice on the permutations of~$\f{S}_n$ defined by the inclusion of their inversion sets.
+Note that the cover relations in the weak order are given by the transpositions of two adjacent letters.
 %The weak order is known to be a congruence uniform lattice~\cite{}.
@@
 We represent an arc by an abscissa monotone curve wiggling around the horizontal axis, starting at point~$a$ and ending at point~$b$, and passing above the points of~$A$ and below the points of~$B$.
-A \defn{noncrossing arc diagram} on~$[n]$ is a collection of arcs on~$[n]$ where any two arcs do not cross in their interior and have distinct left endpoints and distinct right endpoints (but the right endpoint of an arc can be the right endpoint of another arc).
+A \defn{noncrossing arc diagram} on~$[n]$ is a collection of arcs on~$[n]$ where any two arcs do not cross in their interior and have distinct left endpoints and distinct right endpoints (but the right endpoint of an arc can be the left endpoint of another arc).
 
@@
 In fact, all the congruence classes of~$\equiv$ are intervals of the weak order, and the quotient~$\f{S}_n/{\equiv}$ is isomorphic (as a poset) to the subposet of the weak order induced by permutations which are minimal in their class.
-Moreover, the latter are precisely the permutations whose noncrossing arc diagrams only use arcs corresponding to join irreducible permutations which are minimal in their class -- we denote by~$\c{A}_\equiv$ this set of arcs.
+Moreover, the latter are precisely the permutations whose noncrossing arc diagrams only use arcs corresponding to join irreducible permutations which are minimal in their class, and we denote by~$\c{A}_\equiv$ this set of arcs.
 In other words, the classes of~$\equiv$ are in bijection with noncrossing arc diagrams using only arcs in~$\c{A}_\equiv$.
@@
 
-Moreover, the sets of arcs of the form~$\c{A}_\equiv$ are easily described.
+%Moreover, the sets of arcs of the form~$\c{A}_\equiv$ are easily described.
 An arc~$(a, b, A, B)$ is a \defn{subarc} of an arc~$(a', b', A', B')$ if~$a' \le a < b \le b$ and~$A \subseteq A'$ while~$B \subseteq B'$.
 An \defn{arc ideal} is a subset of arcs closed by subarcs.
-The map~${\equiv} \mapsto \c{A}_\equiv$ is a bijection between the lattice congruences of the weak order on~$\f{S}_n$ and the arc ideals.
-(In other words, the lattice of congruences of the weak order is distributive, and its poset of join irreducibles is isomorphic to the subarc order).
+The map~${\equiv} \mapsto \c{A}_\equiv$ is a bijection between the lattice congruences of the weak order and the arc ideals.
+In other words, the lattice of congruences of the weak order is distributive, and its poset of join irreducibles is isomorphic to the subarc order.
 
@@
 
-The \defn{braid arrangement} is the arrangement of hyperplanes~$\set{\b{x} \in \R^n}{x_a = x_b}$ for all~$1 \le a < b \le n$.
+The \defn{braid arrangement} is the hyperplane arrangement consisting of the hyperplanes~$\set{\b{x} \in \R^n}{x_a = x_b}$ for all~$1 \le a < b \le n$.
 It has a chamber for each permutation~$\sigma$ of~$[n]$, given by the set of vectors whose coordinates are ordered as~$\sigma$.
@@
 %$\Sigma(a, b, A, B) \eqdef \set{\b{x} \in \R^n}{x_a = x_b \text{ and } x_{a'} < x_a \text{ for all } a' \in A \text{ and }  x_b < x_{b'} \text{ for all } b' \in B}$.
+
 The \defn{quotient fan} of a lattice congruence~$\equiv$ of the weak order is the polyhedral fan~$\c{F}_\equiv$ where
@@
 (These two descriptions are equivalent.)
-By construction, the dual graph of the braid fan~$\c{F}_\equiv$ is isomorphic to the cover graph of the quotient~$\f{S}_n/{\equiv}$.
+By construction, the braid fan refines the quotient fan~$\c{F}_\equiv$, and the dual graph of the quotient fan~$\c{F}_\equiv$ is isomorphic to the cover graph of the quotient~$\f{S}_n/{\equiv}$.
 
@@
 It was shown in~\cite{PadrolPilaudRitter} that any Minkowski sum of positive scalings of the shard polytopes~$\SP(\alpha)$ for all arcs~$\alpha$ in the arc ideal~$\c{A}_\equiv$ is a quotientope for~$\equiv$.
+
+By definition, any quotientope is a generalized (or deformed) permutahedron, and can thus be written as
+  \[
+  \Bigset{x \in \R^n}{\sum_{i \in S} x_i \le f(S)\ \forall \varnothing \ne S \subseteq [n] \text { and } \sum_{i \in [n]} x_i = f([n])},
+  \]
+where~$f$ is a \defn{submodular} set function $f:2^{[n]}\mapsto \R$ (\ie such that~$f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$).
 
@@
   The subsets $\mathcal{L}(P_w(R))$ of $\f{S}_n$, for all rectangulations $R$ on $n$ rectangles, 
-  are equivalence classes of a congruence relation $\baxtereq$ on the weak Bruhat order on $\f{S}_n$.
+  are equivalence classes of a congruence relation $
+\weakeq$ on the weak Bruhat order on $\f{S}_n$.
   Furthermore:
@@
 We call this congruence the \defn{weak rectangulation congruence} (it is sometimes referred to as the \defn{Baxter congruence}).
-The \defn{weak rectangulotope} $\WRP(n)$ is the quotientope of the weak rectangulation congruence $\baxtereq$ on the weak Bruhat order on $\f{S}_n$.
+The \defn{weak rectangulotope} $\WRP(n)$ is the quotientope of the weak rectangulation congruence $
+\weakeq$ on the weak Bruhat order on $\f{S}_n$.
 
@@
 \begin{lemma}
-The weak rectangulation congruence $\baxtereq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that do not cross the horizontal line.
+The weak rectangulation congruence $
+\weakeq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that do not cross the horizontal line.
 \end{lemma}
@@
 This, in passing, gives us an alternative characterization of the weak rectangulation congruence:
-$\pi\baxtereq\sigma$ if and only if for any interval $I$ of $[n]$,
+$\pi
+\weakeq\sigma$ if and only if for any interval $I$ of $[n]$,
 %the maximum and the minimum of $\pi^{-1}$ and $\sigma^{-1}$ on $I$ have the same index.

### 2024-04-04 draft/rectangulotopes.tex
proposition change thm 1

@@
 In particular, it skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.
+
+\Vincent{I would change this theorem...}
 In 2004, Loday published the following elegant realization of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex given the corresponding binary tree.
@@
   where $\ell^T_i$ and $r^T_i$ respectively denote the number of leaves in the left and right subtrees of $i$ in $T$.
+\end{theorem}
+
+\Vincent{... as follows}
+
+\begin{theorem}
+  \label{thm:loday}
+  The associahedron is realized by the polytope~$\Asso (n)$ in~$\R^n$ defined equivalently as
+  \begin{enumerate}[(i)]
+  \item the convex hull of the points~$\sum_{i\in [n]} \ell^T_i\cdot r^T_i \cdot \mathbf{e}_i$ for all binary trees~$T$ with $n$ internal nodes, where $\ell^T_i$ and $r^T_i$ denote the number of leaves in the left and right subtrees of $i$ in $T$,
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{S \cap I \ne \varnothing}$.
+  \end{enumerate}
 \end{theorem}
@@
 
-By definition, any quotientope is a generalized (or deformed) permutahedron, and can thus be written as
+By definition, any quotientope is a deformed permutahedron, and can thus be written as
   \[
@@
 where~$f$ is a \defn{submodular} set function $f:2^{[n]}\mapsto \R$ (\ie such that~$f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$).
+Recall that the submodular function of a Minkowski sum of deformed permutahedra is just the sum of the submodular functions of the Minkowski summands.
 

### 2024-04-04 draft/rectangulotopes.tex
improve 1.1

--- 2024-04-04 draft/rectangulotopes.tex
@@
 It is an $(n-1)$-dimensional polytope, whose facets are indexed by the proper nonempty subsets of $[n]$, and whose skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
-The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{generalized permutahedra}~\cite{Postnikov} (or \defn{deformed permutahedra}).
+The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{generalized permutahedra}~\cite{Postnikov} (that we prefer to call \defn{deformed permutahedra}).
 
-Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined by Tamari~\cite{T51}, then studied by Stasheff in a topological context~\cite{S63}, that has since then found applications in various areas of mathematics, including topology, operads, cluster algebras, combinatorial Hopf algebras, and physics~\cite{MR4675114}.
+Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined combinatorially by D.~Tamari~\cite{T51} and by J.~Stasheff in a topological context~\cite{S63}.
+It has since then found applications in various areas of mathematics, including topology, operads, cluster algebras, combinatorial Hopf algebras, and physics~\cite{PilaudSantosZiegler}.
 It was first defined as a combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894}.
 The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
-In particular, it skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.
-In 2004, Loday published the following elegant realization of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex given the corresponding binary tree.
+In particular, its skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.
+In 2004, J.-L.~Loday published the following elegant description of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex given the corresponding binary tree.
+We note that the inequality description of the same polytope was actually provided in 1993 by S.~Shnider and S.~Sternberg~\cite{ShniderSternberg}, but Loday's vertex description largely popularized this realization.
 
@@
   \label{thm:loday}
-  The $(n-1)$-dimensional associahedron is realized by the convex hull~$\Asso (n)$ of the points
-  $\sum_{i\in [n]} w^T_i \cdot \mathbf{e}_i$ for all binary trees $T$ on $n$ internal nodes, with
-  \[
-  w^T_i \eqdef \ell^T_i\cdot r^T_i,
-  \]
-  where $\ell^T_i$ and $r^T_i$ respectively denote the number of leaves in the left and right subtrees of $i$ in $T$.
+  The associahedron is realized by the polytope~$\Asso (n)$ in~$\R^n$ defined equivalently as
+  \begin{enumerate}[(i)]
+  \item the convex hull of the points~$\sum_{i\in [n]} \ell^T_i\cdot r^T_i \cdot \mathbf{e}_i$ for all binary trees~$T$ with $n$ internal nodes, where $\ell^T_i$ and $r^T_i$ denote the number of leaves in the left and right subtrees of $i$ in $T$, see~\cite{MR2108555},
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in I} x_i \le \#\set{J \text{ interval of } [n]}{I \cap J \ne \varnothing}$ for all intervals~$I \subseteq [n]$, see~\cite{ShniderSternberg}.
+  \end{enumerate}
 \end{theorem}
@@
 The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} relation called the \defn{sylvester congruence}.
-Pilaud and Santos~\cite{MR3964495}, answering a question of Reading~\cite{MR2142177}, proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more precisely whose normal fan is the \defn{quotient fan} defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
-Those polytopes are generalized permutahedra that they called \defn{quotientopes}.
+Answering a question of N.~Reading~\cite{MR2142177}, V.~Pilaud and F.~Santos~\cite{MR3964495} proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more precisely whose normal fan is the \defn{quotient fan} defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
+Those polytopes are deformed permutahedra that they called \defn{quotientopes}.
 Associahedra are therefore the quotientopes of a congruence of the weak Bruhat order whose classes are in bijection with triangulations.
@@
 %, in particular in the context of electronic engineering applications~\cite{}.
+There are two relevant lattice congruences related with rectangulations.
 
-The relevant lattice congruences are the \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) on one hand, studied by Law and Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}) and \defn{Baxter permutations}, and the \defn{strong rectangulation congruence}, whose classes are in bijection with \defn{strong rectangulations}~\cite{MR2864445}.
-The latter is mentioned by Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by Asinowski, Cardinal, Felsner, and Fusy~\cite{ACFF24}.
+On the one hand, the \defn{weak rectangulation congruence} (or \defn{Baxter congruence}), studied by S.~Law and N.~Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}), with \defn{twin binary trees}~\cite{Giraudo}, and with \defn{Baxter permutations}.
+\vincent{Give more Baxter references here.}
+The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
+The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
 
-One of our main results is an elementary description of the vertex coordinates of a realization of both families of quotientopes, in the spirit of Loday's realization in Theorem~\ref{thm:loday}.
-The quotientopes corresponding to the weak rectangulation congruence will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
-The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
-Similarly, the quotientopes corresponding to the strong rectangulation congruence will be called the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
+On the other hand, the \defn{strong rectangulation congruence}, whose classes are in bijection with \defn{strong rectangulations}~\cite{MR2864445}.
+The latter is mentioned by N.~Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
+The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
 The skeleton of $\SRP(n)$ is isomorphic to the flip graph of strong rectangulations of size $n$ defined by Meehan~\cite{MR3697823} (see also~\cite{ACFF24}).
 
-Rectangulotopes enrich the family of natural quotientopes, alongside associahedra and permutreehedra~\cite{MR3856522}, and are instructive examples of application of the theory of lattice congruences to concrete combinatorial objects.
-
-% jean:{Vincent, you may want to say more about permutreehedra, incl. the Loday coordinates.}
+Rectangulotopes enrich the family of natural quotientopes, alongside associahedra, permutreehedra~\cite{MR3856522}, brick polytopes~\cite{Pilaud-brickAlgebra}, and certain graphical zonotopes~\cite{}, providing instructive examples of application of the theory of lattice congruences to concrete combinatorial objects.
+Our main results are elementary vertex and facet descriptions of both families of quotientopes.
+For vertices, we describe the coordinates of the vertex corresponding to a weak or strong rectangulation with simple product formulas in the spirit of J.-L.~Loday's realization in~\cref{thm:loday}.
+As mentioned in~\cite{PilaudSantosZiegler}, there is no such formula for arbitrary quotientopes, which makes rectangulotopes join the restricted club of Loday type quotientopes, jointly with associahedra and permutreehedra~\cite{MR3856522}.
+For facets, we show that both weak and strong rectangulotopes have $2^n-2$ facets (the maximal possible number of facets of a deformed permutahedron), and provide simple formulas for the right hand side of the inequality corresponding to a given proper nonempty subset of~$[n]$.
+We now give the necessary definitions for precisely stating these vertex and facet descriptions of the weak and strong rectangulotopes.
 
 \subsection{Weak and strong rectangulations}
-
-We give the necessary definitions for stating our results on the vertex description of the weak and strong rectangulotopes.
 
@@
 
-Section~\ref{sec:quotientopes} is dedicated to the background on quotientopes and their realizations as Minkowski sums of \defn{shard polytopes}, due to Padrol, Pilaud, and Ritter~\cite{MR4584712}, which will be our main tool throughout.
-Section~\ref{sec:wr} details the case of weak rectangulations, while Section~\ref{sec:sr} deals with strong rectangulations.
+\cref{sec:quotientopes} is dedicated to the background on quotientopes and their realizations as Minkowski sums of \defn{shard polytopes}, due to Padrol, Pilaud, and Ritter~\cite{MR4584712}, which will be our main tool throughout.
+\cref{sec:wr} details the case of weak rectangulations, while \cref{sec:sr} deals with strong rectangulations.
 In both cases, in addition to the vertex coordinates, we also give the submodular functions defining the polytopes.
@@
 
-We now briefly recall some results on the weak order and its quotients, both from a lattice and geometric perspectives.
+We now briefly recall some results on the weak Bruhat order and its quotients, both from a lattice and geometric perspectives.
 We refer to \cite{Reading-arcDiagrams, Reading-Chapters, PadrolPilaudRitter} for details.
@@
 The \defn{inversion set} of~$\sigma \in \f{S}_n$ is~$\inv(\sigma) \eqdef \set{(\sigma_i, \sigma_j)}{1 \le i < j \le n \text{ and } \sigma_i > \sigma_j}$.
-The \defn{weak order} is the lattice on the permutations of~$\f{S}_n$ defined by the inclusion of their inversion sets.
-Note that the cover relations in the weak order are given by the transpositions of two adjacent letters.
-%The weak order is known to be a congruence uniform lattice~\cite{}.
+The \defn{weak Bruhat order}\footnote{The weak Bruhat order is usually just referred to as the weak order.
+In this paper, we prefer to use weak Bruhat order to avoid any confusion with the weak poset defined on weak rectangulations.} is the lattice on the permutations of~$\f{S}_n$ defined by the inclusion of their inversion sets.
+Note that the cover relations in the weak Bruhat order are given by the transpositions of two adjacent letters.
+%The weak Bruhat order is known to be a congruence uniform lattice~\cite{}.
 %In particular, it is join semidistributive, so that any permutation admits a canonical join representation.
@@
 This can also been visualized by representing the table~$(\sigma_j,j)$ of the permutation~$\sigma$, drawing the segments corresponding to the descents of~$\sigma$, and letting all points fall on the horizontal axis, allowing the segments to bend but not to cross each other nor to pass through a point.
-The single arcs correspond to permutations with a single descent, that is, to join irreducible permutations of the weak order.
-In general, the noncrossing arc diagram of a permutation~$\sigma$ actually encodes the canonical join representation of~$\sigma$ in the weak order (which was known to exist, as the weak order is join semidistributive).
+The single arcs correspond to permutations with a single descent, that is, to join irreducible permutations of the weak Bruhat order.
+In general, the noncrossing arc diagram of a permutation~$\sigma$ actually encodes the canonical join representation of~$\sigma$ in the weak Bruhat order (which was known to exist, as the weak Bruhat order is join semidistributive).
 See~\cite{Reading-arcDiagrams} for details.
@@
 
-A \defn{lattice congruence} of the weak order is an equivalence relation~$\equiv$ on~$\f{S}_n$ that respects the meet and join operations, \ie such that $x \equiv x'$ and~$y \equiv y'$ implies $x \meet y \, \equiv \, x' \meet y'$ and~$x \join y \, \equiv \, x' \join y'$.
+A \defn{lattice congruence} of the weak Bruhat order is an equivalence relation~$\equiv$ on~$\f{S}_n$ that respects the meet and join operations, \ie such that $x \equiv x'$ and~$y \equiv y'$ implies $x \meet y \, \equiv \, x' \meet y'$ and~$x \join y \, \equiv \, x' \join y'$.
 The \defn{lattice quotient}~$\f{S}_n/{\equiv}$ is the lattice on the congruence classes of~$\equiv$ where~$X \le Y$ if and only if there exist~$x \in X$ and~$y \in Y$ such that~$x \le y$, and~$X \meet Y$ (resp.~$X \join Y$) is the congruence class of~$x \meet y$ (resp.~$x \join y$) for any~$x \in X$~and~$y \in Y$.
@@
 Respecting the meet and join operation is a strong condition that imposes additional structure on~$\equiv$.
-In fact, all the congruence classes of~$\equiv$ are intervals of the weak order, and the quotient~$\f{S}_n/{\equiv}$ is isomorphic (as a poset) to the subposet of the weak order induced by permutations which are minimal in their class.
+In fact, all the congruence classes of~$\equiv$ are intervals of the weak Bruhat order, and the quotient~$\f{S}_n/{\equiv}$ is isomorphic (as a poset) to the subposet of the weak Bruhat order induced by permutations which are minimal in their class.
 Moreover, the latter are precisely the permutations whose noncrossing arc diagrams only use arcs corresponding to join irreducible permutations which are minimal in their class, and we denote by~$\c{A}_\equiv$ this set of arcs.
@@
 An \defn{arc ideal} is a subset of arcs closed by subarcs.
-The map~${\equiv} \mapsto \c{A}_\equiv$ is a bijection between the lattice congruences of the weak order and the arc ideals.
-In other words, the lattice of congruences of the weak order is distributive, and its poset of join irreducibles is isomorphic to the subarc order.
+The map~${\equiv} \mapsto \c{A}_\equiv$ is a bijection between the lattice congruences of the weak Bruhat order and the arc ideals.
+In other words, the lattice of congruences of the weak Bruhat order is distributive, and its poset of join irreducibles is isomorphic to the subarc order.
 
@@
 
-The \defn{quotient fan} of a lattice congruence~$\equiv$ of the weak order is the polyhedral fan~$\c{F}_\equiv$ where
+The \defn{quotient fan} of a lattice congruence~$\equiv$ of the weak Bruhat order is the polyhedral fan~$\c{F}_\equiv$ where
 \begin{itemize}
@@
 
-A \defn{quotientope} for a lattice congruence~$\equiv$ of the weak order is a polytope whose normal fan is the quotient fan~$\c{F}_\equiv$.
+A \defn{quotientope} for a lattice congruence~$\equiv$ of the weak Bruhat order is a polytope whose normal fan is the quotient fan~$\c{F}_\equiv$.
 In particular, the skeleton of a quotientope is isomorphic to the cover graph of the quotient~$\f{S}_n/{\equiv}$.
@@
 where $f$ is a \defn{submodular} set function $f:2^{[n]}\mapsto \R$.
-The description of the weak rectangulotope $\WRP(n)$ as a Minkowski sum in Lemma~\ref{lem:weakMinkowski} allows us to easily compute the submodular function defining it.
+The description of the weak rectangulotope $\WRP(n)$ as a Minkowski sum in \cref{lem:weakMinkowski} allows us to easily compute the submodular function defining it.
 
@@
 \begin{proof}
-  From Lemmas~\ref{lem:weakMinkowski} and~\ref{lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the Loday and anti-Loday shard polytope in Lemmas~\ref{lem:lodaysp} and~\ref{lem:antilodaysp}.
+  From \cref{lem:weakMinkowski,lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the Loday and anti-Loday shard polytope in \cref{lem:lodaysp,lem:antilodaysp}.
 For a Loday arc $A$ defined by an interval $I$ of $[n]$, we have that its shard polytope $\SP(A)$ is realized by the submodular function
@@
 Recall that in a Minkowski sum of polytopes, the vertex that is extremal in a generic direction is the sum of the vertices of the summands that are extremal in this direction.
-We therefore need to understand which vertices of the translates of the shard polytopes defined in Lemmas~\ref{lem:lodaysp} and~\ref{lem:antilodaysp} are extremal for a direction given by a permutation.
+We therefore need to understand which vertices of the translates of the shard polytopes defined in \cref{lem:lodaysp,lem:antilodaysp} are extremal for a direction given by a permutation.
 
@@
 A \defn{yang arc} is defined similarly, except that the arc goes below the elements of $I$ and above the elements of $J$.
+Note that all up arcs and down arcs are both yin arcs and yang arcs.
 
@@
 
-Since the combinatorial type of a Minkowski sum is invariant to translation of the summands, we can use the polytopes defined in Lemmas~\ref{lem:yinsp} and~\ref{lem:yangsp} as a definition of the yin and yang shard polytopes.
+Since the combinatorial type of a Minkowski sum is invariant to translation of the summands, we can use the polytopes defined in \cref{lem:yinsp,lem:yangsp} as a definition of the yin and yang shard polytopes.
 Also note that since Loday and anti-Loday shard polytopes are, up to a translation, both yin and yang, each Loday and anti-Loday shard polytope is summed twice in the proposed realization.
@@
 
-Lemma~\ref{lem:strongsum} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.
+\cref{lem:strongsum} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.
 
@@
 
-The submodular function corresponding to our realization of $\SRP(n)$ can be obtained as before by summing the functions defining the yin and yang shard polytopes in Lemmas~\ref{lem:yinsp} and~\ref{lem:yangsp}.
+The submodular function corresponding to our realization of $\SRP(n)$ can be obtained as before by summing the functions defining the yin and yang shard polytopes in \cref{lem:yinsp,lem:yangsp}.
 
@@
 \begin{proof}
-  From Lemmas~\ref{lem:strongMinkowski} and~\ref{lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the yin and yang shard polytope in Lemmas~\ref{lem:yinsp} and~\ref{lem:yangsp}.
+  From \cref{lem:strongMinkowski,lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the yin and yang shard polytope in \cref{lem:yinsp,lem:yangsp}.
 For a yin arc $A$ defined by consecutive intervals $I, J$ of $[n]$, we have that its shard polytope $\SP(A)$ is realized by the submodular function
@@
 Recall that in a Minkowski sum of polytopes, the vertex that is extremal in a generic direction is the sum of the vertices of the summands that are extremal in this direction.
-We therefore need to understand which vertices of the translates of the yin shard polytopes defined in Lemmas~\ref{lem:yinsp} are extremal for a direction given by a permutation.
+We therefore need to understand which vertices of the translates of the yin shard polytopes defined in \cref{lem:yinsp} are extremal for a direction given by a permutation.
 
@@
   Let $\pi$ be the 2-clumped permutation corresponding to $R$, hence the minimal element of $\mathcal{L}(P_s(R))$ in the weak Bruhat order.
-  From Lemma~\ref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc $A$ whose shard polytope $\SP(A)$ contributes to $\mathbf{e}_i-\mathbf{e}_j$ only if
+  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc $A$ whose shard polytope $\SP(A)$ contributes to $\mathbf{e}_i-\mathbf{e}_j$ only if
   $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.

### 2024-04-04 draft/rectangulotopes.tex
abstract

@@
   Rectangulations are decompositions of a square into finitely many axis-aligned rectangles.
-  We describe realizations of $(n-1)$-dimensional polytopes associated with rectangulations composed of $n$ rectangles.
+  We describe realizations of $(n-1)$-dimensional polytopes associated with two families of rectangulations composed of $n$ rectangles.
   They are defined as quotientopes of natural congruence relations on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
-  We give simple geometric realizations of these polytopes, including elementary formulas for computing the coordinates of each vertex of the polytope, in the spirit of Loday's realization of the associahedron.
+  We give simple vertex and facet descriptions of these polytopes, in particular elementary formulas for computing the coordinates of the vertex corresponding to each rectangulation, in the spirit of J.-L.~Loday's realization of the associahedron.
 \end{abstract}
@@
 
-\cref{sec:quotientopes} is dedicated to the background on quotientopes and their realizations as Minkowski sums of \defn{shard polytopes}, due to Padrol, Pilaud, and Ritter~\cite{MR4584712}, which will be our main tool throughout.
+\cref{sec:quotientopes} is dedicated to the background on quotientopes and their realizations as Minkowski sums of \defn{shard polytopes}, due to A.~Padrol, V.~Pilaud, and J.~Ritter~\cite{MR4584712}, which will be our main tool throughout.
 \cref{sec:wr} details the case of weak rectangulations, while \cref{sec:sr} deals with strong rectangulations.

### 2024-04-04 draft/rectangulotopes.tex
improve statements section 1.4

@@
 The encoding of combinatorial objects in the form of a polyhedral structure is a recurrent theme in geometric and algebraic combinatorics.
-An elementary yet striking example of this idea is the \defn{permutahedron} $\Perm (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \pi^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\pi\in\f{S}_n$.
+An elementary yet striking example of this idea is the \defn{permutahedron} $\Perm (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \sigma^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\sigma \in \f{S}_n$.
 It is an $(n-1)$-dimensional polytope, whose facets are indexed by the proper nonempty subsets of $[n]$, and whose skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
@@
   \label{thm:loday}
-  The associahedron is realized by the polytope~$\Asso (n)$ in~$\R^n$ defined equivalently as
+  The associahedron is realized by the polytope~$\Asso (n) \subset \R^n$ defined equivalently as
   \begin{enumerate}[(i)]
   \item the convex hull of the points~$\sum_{i\in [n]} \ell^T_i\cdot r^T_i \cdot \mathbf{e}_i$ for all binary trees~$T$ with $n$ internal nodes, where $\ell^T_i$ and $r^T_i$ denote the number of leaves in the left and right subtrees of $i$ in $T$, see~\cite{MR2108555},
-  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in I} x_i \le \#\set{J \text{ interval of } [n]}{I \cap J \ne \varnothing}$ for all intervals~$I \subseteq [n]$, see~\cite{ShniderSternberg}.
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in I} x_i \le \#\set{J \text{ interval of } [n]}{I \cap J \ne \varnothing}$ for all intervals~$\varnothing \ne I \subsetneq [n]$, see~\cite{ShniderSternberg}.
   \end{enumerate}
@@
 Those polytopes are deformed permutahedra that they called \defn{quotientopes}.
-Associahedra are therefore the quotientopes of a congruence of the weak Bruhat order whose classes are in bijection with triangulations.
+Associahedra are therefore the quotientopes of the sylvester congruence of the weak Bruhat order whose classes are in bijection with triangulations.
 
@@
 %, in particular in the context of electronic engineering applications~\cite{}.
+\vincent{More citations on electronic engineering...}
 There are two relevant lattice congruences related with rectangulations.
-
+%
 On the one hand, the \defn{weak rectangulation congruence} (or \defn{Baxter congruence}), studied by S.~Law and N.~Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}), with \defn{twin binary trees}~\cite{Giraudo}, and with \defn{Baxter permutations}.
@@
 The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
-
+%
 On the other hand, the \defn{strong rectangulation congruence}, whose classes are in bijection with \defn{strong rectangulations}~\cite{MR2864445}.
@@
 
-Rectangulotopes enrich the family of natural quotientopes, alongside associahedra, permutreehedra~\cite{MR3856522}, brick polytopes~\cite{Pilaud-brickAlgebra}, and certain graphical zonotopes~\cite{}, providing instructive examples of application of the theory of lattice congruences to concrete combinatorial objects.
+Rectangulotopes enrich the family of natural quotientopes, alongside associahedra, permutreehedra~\cite{MR3856522}, certain brick polytopes~\cite{PilaudSantos-brickPolytopes, PilaudStump-brickPolytopes, Pilaud-brickAlgebra}, and certain graphical zonotopes~\cite{Pilaud-brickAlgebra, Pilaud-acyclicReorientationLattices}, providing instructive examples of application of the theory of lattice congruences to concrete combinatorial objects.
 Our main results are elementary vertex and facet descriptions of both families of quotientopes.
@@
 
-Our first result is a concise formula for the coordinates of the vertices of the  weak rectangulotopes $\WRP(n)$, that consists of applying Loday's formula on each of the source and target trees of the rectangulation.
+Our first results are simple vertex and facet descriptions of the weak rectangulotopes~$\WRP(n)$.
+For the vertices, we provide a concise formula for the coordinates, that consists of applying J.-L.~Loday's formula on each of the source and target trees of the rectangulation.
+For the facets, we combine the right hand sides of two opposite associahedra to obtain those of the weak rectangulotopes.
 This materializes the result of Law and Reading that the weak rectangulotopes are Minkowski sums of two opposite associahedra~\cite{MR2871762}.
@@
 \begin{theorem}
-  The $(n-1)$-dimensional weak rectangulotope is realized by the convex hull $\WRP (n)$ of the points
+  The weak rectangulotope is realized by the polytope~$\WRP (n) \subset \R^n\!$ defined equivalently~as
+  \begin{enumerate}[(i)]
+  \item the convex hull of the points
   $\sum_{i\in [n]} (\loday{w}^R_i - \antiloday{w}^R_i)\cdot \mathbf{e}_i$ for all weak rectangulations $R$ on $n$ rectangles,
@@
   \]
-  where~$S$ and~$T$ are the source and target trees of the rectangulation~$R$, and $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$.
+  where~$S$ and~$T$ are the source and target trees of the rectangulation~$R$, and $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{I \not\subseteq S \text{ and } I \not\subseteq [n] \ssm S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$.
+  \end{enumerate}
 \end{theorem}
@@
   For instance, the $4$th coordinate is $h^T_4 \cdot v^T_4 - h^S_4 \cdot v^S_4 = 1 \cdot 5 - 1 \cdot 3 = 2$.
+  \vincent{give also a relevant example of facet...}
 \end{example}
@@
 \begin{theorem}
-  The $(n-1)$-dimensional strong rectangulotope is realized by the convex hull~$\SRP (n)$ of the points
+  The $(n-1)$-dimensional strong rectangulotope is realized by the polytope~$\SRP (n)$ defined equivalently as
+  \begin{enumerate}[(i)]
+  \item convex hull of the points
   \[
@@
   \end{itemize}
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ in the sets denote consecutive intervals of~$[n]$.
+  \end{enumerate}
 \end{theorem}
@@
   For instance, the $4$th coordinate is $???$.
-
+  \vincent{give also a relevant example of facet...}
 \end{example}
@@
 
-This work was initiated at the Workshop on Combinatorics, Algorithms, and Geometry organized by Torsten M\"utze on March 4-8, 2024 in Dresden, Germany.
+This work was initiated at the Workshop on Combinatorics, Algorithms, and Geometry held on March 4-8, 2024 in Dresden, Germany.
 The authors thank Namrata and Torsten M\"utze for the organization and the other participants of the workshop for the stimulating interactions, notably on other combinatorial aspects of rectangulations.

### 2024-04-04 programmation/submodular.py
improved submodular.py

@@ -4,40 +4,49 @@ from itertools import *
 
 ## SUBMODULAR FUNCTIONS
 
-# simplex for a Loday arc I
-# 1 if S and I intersects, 0 otherwise
+# simplex for an up arc I
+# 1 if S intersects I, 0 otherwise
 def f_loday(S, I):
     return int(bool(set(S) & set (I)))
 
-# anti-simplex for an anti-Loday arc I
+# anti-simplex for a down arc I
 # -1 if I is a subset of S, 0 otherwise
 def f_antiloday(S, I):
     return - int(set(I).issubset(set(S)))
 
 # simplex + antisimplex (cuboctahedron in 3d)
-# 1 if S and I intersects but I is not a subset of S, 0 otherwise
+# 1 if S intersects I but I is not a subset of S, 0 otherwise
 def f_cuboc(S, I):
     return int(bool(set(S) & set (I)) and not set(I).issubset(set(S)))
 
-# weak rectangulation
-# number of I such that S and I intersects but I is not a subset of S
+# weak rectangulotopes
+# number of I such that S intersects I but I is not a subset of S
 def f_weakrect(S, n):
     return sum([int(bool(set(S) & set (range(i, j+1))) and not set(range(i, j+1)).issubset(set(S))) for i in range(n-1) for j in range(i+1, n)])
 
 # yin shard polytopes
-# 1 if and only if S intersects I but J is not a subset of S
+# 1 if and only if S intersects I, and J is not a subset of S
 def f_yin(S, I, J):
     return int(bool(set(S) & set(I)) and not (set(J).issubset(S)))
 
-# yin rectangulations
-# number of pairs I, J such that S intersects I but J is not a subset of S
+# yin rectangulotopes
+# number of pairs I, J such that S intersects I, and J is not a subset of S
 def f_yinrect(S, n):
     return sum([int (bool(set(S) & set(range(i, j+1))) and not (set(range(j+1, k+1)).issubset(S))) for i in range(n-1) for j in range(i, n-1) for k in range(j+1, n)])
 
 # yang shard polytopes
-# 1 if and only if S intersects J but I is not a subset of S
+# 1 if and only if S intersects J, and I is not a subset of S
 def f_yang(S, I, J):
     return int(bool(set(S) & set(J)) and not (set(I).issubset(S)))
+
+# yang rectangulotopes
[…]

### 2024-04-04 draft/rectangulotopes.tex

@@
 The encoding of combinatorial objects in the form of a polyhedral structure is a recurrent theme in geometric and algebraic combinatorics.
-An elementary yet striking example of this idea is the \defn{permutahedron} $\Perm (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \sigma^{-1}(i)\cdot \mathbf{e}_i$, for all permutations $\sigma \in \f{S}_n$.
+An elementary yet striking example of this idea is the \defn{permutahedron} $\Perm (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \sigma^{-1}(i)\cdot \b{e}_i$, for all permutations $\sigma \in \f{S}_n$.
 It is an $(n-1)$-dimensional polytope, whose facets are indexed by the proper nonempty subsets of $[n]$, and whose skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
@@
   \begin{enumerate}[(i)]
-  \item the convex hull of the points~$\sum_{i\in [n]} \ell^T_i\cdot r^T_i \cdot \mathbf{e}_i$ for all binary trees~$T$ with $n$ internal nodes, where $\ell^T_i$ and $r^T_i$ denote the number of leaves in the left and right subtrees of $i$ in $T$, see~\cite{MR2108555},
+  \item the convex hull of the points~$\sum_{i\in [n]} \ell^T_i\cdot r^T_i \cdot \b{e}_i$ for all binary trees~$T$ with $n$ internal nodes, where $\ell^T_i$ and $r^T_i$ denote the number of leaves in the left and right subtrees of $i$ in $T$, see~\cite{MR2108555},
   \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in I} x_i \le \#\set{J \text{ interval of } [n]}{I \cap J \ne \varnothing}$ for all intervals~$\varnothing \ne I \subsetneq [n]$, see~\cite{ShniderSternberg}.
@@
   \item the convex hull of the points
-  $\sum_{i\in [n]} (\loday{w}^R_i - \antiloday{w}^R_i)\cdot \mathbf{e}_i$ for all weak rectangulations $R$ on $n$ rectangles,
+  $\sum_{i\in [n]} (\loday{w}^R_i - \antiloday{w}^R_i)\cdot \b{e}_i$ for all weak rectangulations $R$ on $n$ rectangles,
   with
@@
   \[
-  \sum_{i,j\in [n], i< j} (\yin{w}^R_{i,j} - \yang{w}^R_{i,j})\cdot (\mathbf{e}_i - \mathbf{e}_j),
+  \sum_{i,j\in [n], i< j} (\yin{w}^R_{i,j} - \yang{w}^R_{i,j})\cdot (\b{e}_i - \b{e}_j),
   \]
@@
   \[
-  \Bigset{x \in \R^n}{\sum_{i \in S} x_i \le f(S)\ \forall \varnothing \ne S \subseteq [n] \text { and } \sum_{i \in [n]} x_i = f([n])},
+  \Bigset{x \in \R^n}{ \sum_{i \in [n]} x_i = f([n]) \text{ and } \sum_{i \in S} x_i \le f(S) \text{ for all } \varnothing \ne S \subseteq [n]},
   \]
@@
 \[
-\mathcal{L}(P_w(R)) \eqdef \{\pi\in\f{S}_n : i\prec_w j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
+\mathcal{L}(P_w(R)) \eqdef \bigset{\pi\in\f{S}_n }{ i\prec_w j\implies \pi^{-1}(i) < \pi^{-1}(j)}.
 \]
@@
   \label{lem:lodaysp}
-  Given a Loday arc $A$ defined by an interval $I$, the Loday shard polytope $\SP(A)$ is a translate of the standard simplex $\conv \{ \mathbf{e}_i : i\in I \}$.
+  Given a Loday arc $A$ defined by an interval $I$, the Loday shard polytope $\SP(A)$ is a translate of the standard simplex $\conv \set{ \b{e}_i }{ i\in I }$.
 \end{lemma}
@@
   \label{lem:antilodaysp}
-  Given an anti-Loday arc $A$ defined by an interval $I$, the anti-Loday shard polytope $\SP(A)$ is a translate of the negative standard simplex $\conv \{ - \mathbf{e}_i : i\in I \}$.
+  Given an anti-Loday arc $A$ defined by an interval $I$, the anti-Loday shard polytope $\SP(A)$ is a translate of the negative standard simplex $\conv \set{ - \b{e}_i }{ i\in I }$.
 \end{lemma}
@@
 \vincent{Say also the minkowski sum of all faces of the standard simplex corresponding to intervals, and all their opposite.}
+\vincent{Say that this was already known by Law and Reading.}
 
@@
 
-\jean{You may want to use another notation here, feel free.}
-
-Recall that every deformed permutahedra can be defined as
-  \[
-  \{x\in\R^n : \sum_{i\in S} x_i \leq f(S)\ \forall S\subseteq [n], S\not=\emptyset, \text { and } \sum_{i\in [n]} x_i = f([n])\},
-  \]
-where $f$ is a \defn{submodular} set function $f:2^{[n]}\mapsto \R$.
-The description of the weak rectangulotope $\WRP(n)$ as a Minkowski sum in \cref{lem:weakMinkowski} allows us to easily compute the submodular function defining it.
+The description of the weak rectangulotope~$\WRP(n)$ as a Minkowski sum in \cref{lem:weakMinkowski} allows us to easily compute the corresponding submodular function.
 
@@
   \[
-  f(S) := |\{ I\text{ interval of } [n] : I\cap S\not=\emptyset \text{ and } I\cap ([n]\setminus S) \not= \emptyset \}|.
+  f(S) \eqdef \#\bigset{ I\text{ interval of } [n] }{ I\cap S\not=\emptyset \text{ and } I\cap ([n]\ssm S) \not= \emptyset }.
   \]
@@
   Consider a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by the submodular functions $f_1,f_2,\ldots ,f_k$.
-Then the Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f:=\sum_i f_i$.
+Then the Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f \eqdef \sum_i f_i$.
 \end{lemma}
@@
   \[
-  \loday{f}_I(S) := \llbracket S\cap I\not=\emptyset \rrbracket .
+  \loday{f}_I(S) \eqdef \llbracket S\cap I\not=\emptyset \rrbracket .
   \]
@@
   \[
-  \antiloday{f}_I(S) := - \llbracket I\subseteq S \rrbracket .
+  \antiloday{f}_I(S) \eqdef - \llbracket I\subseteq S \rrbracket .
   \]
@@
   \[
-  f(S) := \sum_{I\text{ interval of }[n]} \loday{f}_I(S) + \antiloday{f}_I(S)
+  f(S) \eqdef \sum_{I\text{ interval of }[n]} \loday{f}_I(S) + \antiloday{f}_I(S)
   \]
@@
   The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
-  is $\mathbf{e}_i$, where $i \eqdef \arg\max_{j\in I} \pi^{-1}(j)$.
+  is $\b{e}_i$, where $i \eqdef \arg\max_{j\in I} \pi^{-1}(j)$.
   Similarly, if $A$ is an anti-Loday arc defined by $I$, the unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
-  is $-\mathbf{e}_i$, where $i \eqdef \arg\min_{j\in I} \pi^{-1}(j)$.
+  is $-\b{e}_i$, where $i \eqdef \arg\min_{j\in I} \pi^{-1}(j)$.
 \end{lemma}
@@
   Let $R$ be a rectangulation on $n$ rectangles and $p(R)$ the vertex of $\WRP(n)$ associated with $R$.
-  Given $i\in [n]$, the number of Loday arcs $A$ such that $\mathbf{e}_i$ is the extremal vertex of $\SP(A)$ contributing to $p(R)$ is
+  Given $i\in [n]$, the number of Loday arcs $A$ such that $\b{e}_i$ is the extremal vertex of $\SP(A)$ contributing to $p(R)$ is
   \[
@@
 \[
-\mathcal{L}(P_s(R)) \eqdef \{\pi\in\f{S}_n : i\prec_s j\implies \pi^{-1}(i) < \pi^{-1}(j)\}.
+\mathcal{L}(P_s(R)) \eqdef \bigset{\pi\in\f{S}_n }{ i\prec_s j\implies \pi^{-1}(i) < \pi^{-1}(j)}.
 \]
@@
   Given a yin arc $A=(I,J)$, the yin shard polytope $\SP(A)$ is defined as the convex hull of the points
-  $\{\mathbf{0} \}\cup \{ \mathbf{e}_i - \mathbf{e}_j : (i,j)\in I\times J\}$.
+  $\{\b{0}\} \cup \set{\b{e}_i - \b{e}_j}{(i,j)\in I\times J}$.
 \end{lemma}
@@
   Given a yang arc $A=(I,J)$, the yang shard polytope $\SP(A)$ is a translate of the convex hull of the points
-  $\{\mathbf{0} \}\cup \{ \mathbf{e}_j - \mathbf{e}_i : (i,j)\in I\times J\}$.
+  $\{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{(i,j)\in I\times J}$.
 \end{lemma}
@@
 Note that yin shard polytopes are translates of the Loday shard polytopes when $|J|=1$, and of the anti-Loday shard polytopes when $|I|=1$.
-similarly, Loday and anti-Loday shard polytopes are translates of some yang shard polytopes.
+Similarly, Loday and anti-Loday shard polytopes are translates of some yang shard polytopes.
 
-The following is a direct application of the main result of Padrol, Pilaud, and Ritter~\cite{MR4584712}.
+The following is a direct application of the main result of A.~Padrol, V.~Pilaud, and J.~Ritter~\cite{MR4584712}.
 
@@
   \[
-  f(S) := |\{ I, J : I\cap S\not=\emptyset \wedge J\cap ([n]\setminus S) \not= \emptyset \}| +
-  |\{ I, J : J\cap S\not=\emptyset \wedge I\cap ([n]\setminus S) \not= \emptyset \}|,
+  f(S) \eqdef \# \bigset{ I, J }{ I\cap S\not=\emptyset \wedge J\cap ([n]\ssm S) \not= \emptyset } +
+  \#\bigset{ I, J }{ J\cap S\not=\emptyset \wedge I\cap ([n]\ssm S) \not= \emptyset },
   \]
@@
   \[
-  \yin{f}_{I,J}(S) := \llbracket S\cap I\not=\emptyset \wedge S\not\supseteq J \rrbracket .
+  \yin{f}_{I,J}(S) \eqdef \llbracket S\cap I\not=\emptyset \wedge S\not\supseteq J \rrbracket .
   \]
@@
   \[
-  \yang{f}_{I,J}(S) := \llbracket S\cap J\not=\emptyset \wedge S\not\supseteq I \rrbracket .
+  \yang{f}_{I,J}(S) \eqdef \llbracket S\cap J\not=\emptyset \wedge S\not\supseteq I \rrbracket .
   \]
@@
   \[
-  f(S) := \sum_{I,J} \yin{f}_{I,J}(S) + \yang{f}_{I,J}(S)
+  f(S) \eqdef \sum_{I,J} \yin{f}_{I,J}(S) + \yang{f}_{I,J}(S)
   \]
@@
   The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
-  is $\mathbf{e}_i-\mathbf{e}_j$ if $\pi^{-1}(i)>\pi^{-1}(j)$, and $\mathbf{0}$ otherwise.
+  is $\b{e}_i-\b{e}_j$ if $\pi^{-1}(i)>\pi^{-1}(j)$, and $\b{0}$ otherwise.
 \end{lemma}
@@
   Let $R$ be a rectangulation on $n$ rectangles, $([n],\prec_s)$ the associated strong order, and $p(R)$ the vertex of $\SRP(n)$ corresponding to $R$.
-  Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $A$ such that $\mathbf{e}_i-\mathbf{e}_j$
+  Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $A$ such that $\b{e}_i-\b{e}_j$
   is the extremal vertex of $\SP(A)$ contributing to $p(R)$ is
@@
   Let $\pi$ be the 2-clumped permutation corresponding to $R$, hence the minimal element of $\mathcal{L}(P_s(R))$ in the weak Bruhat order.
-  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc $A$ whose shard polytope $\SP(A)$ contributes to $\mathbf{e}_i-\mathbf{e}_j$ only if
+  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc $A$ whose shard polytope $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ only if
   $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.
-  Provided this holds, the set of yin arcs $A$ such that $\SP(A)$ contributes to $\mathbf{e}_i-\mathbf{e}_j$ are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that $i$ is the index of the maximum of $\pi^{-1}$ in $I$, and $j$ is the index of the maximum of $\pi^{-1}$ in $J$.
+  Provided this holds, the set of yin arcs $A$ such that $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that $i$ is the index of the maximum of $\pi^{-1}$ in $I$, and $j$ is the index of the maximum of $\pi^{-1}$ in $J$.
   %$i = \arg\max_{k\in I} \pi^{-1}(k)$ and $j = \arg\min_{k\in J} \pi^{-1}(k)$.

### 2024-04-04 draft/rectangulotopes.tex
improve Section 3

@@
 \newcommand{\h}{\widehat} % hat letters
-
-% hello
 
@@
 % Loday and anti-Loday accents (?)
+\makeatletter
+\newcommand{\oset}[3][0ex]{%
+  \mathrel{\mathop{#3}\limits^{
+    \vbox to#1{\kern-2\ex@
+    \hbox{$\scriptstyle#2$}\vss}}}}
+%
+\newcommand{\uset}[3][0ex]{%
+  \mathrel{\mathop{#3}\limits_{
+    \vbox to#1{\kern-7\ex@
+    \hbox{$\scriptstyle#2$}\vss}}}}
+\makeatother
+%
 \newcommand{\loday}[1]{\smash{\overset{\frown}{#1}}}
 \newcommand{\antiloday}[1]{\smash{\overset{\smile}{#1}}}
+\newcommand{\upArc}[1]{\smash{\raisebox{.05cm}{$\uset[0ex]{#1}{\frown}$}}}
+\newcommand{\downArc}[1]{\smash{\raisebox{-.05cm}{$\oset[.2ex]{#1}{\smile}$}}}
 % yin and yang accents (?)
@@
 Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$.
-We denote by %$i|j$ and $\frac{\ j\ }i$
-$i \, \backslash \, j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$. %, respectively.
+We denote by $i \, \backslash \, j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$.
 We also use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.
@@
   \end{itemize}
-  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ in the sets denote consecutive intervals of~$[n]$.
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ in the sets denote nonempty and consecutive intervals of~$[n]$.
   \end{enumerate}
@@
 
-Given a rectangulation $R$, the two trees $T(R)$ and $S(R)$, where the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top,
-define two partial orders on $[n]$.
-We define the \defn{weak poset} $P_w(R)=([n],\prec_w)$ of a rectangulation $R$ as the transitive closure of the union of the two partial orders defined by the intertwining binary trees $T(R)$ and $S(R)$~\cite{MR4014603}.
-The weak poset is a two-dimensional lattice, the minimum of which is the root of $S(R)$ and the maximum is the root of $T(R)$.
-We define the set $\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
+Given a rectangulation~$R$, the source and target trees~$S(R)$ and~$T(R)$ define two partial orders on $[n]$ (the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top).
+The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of the two partial orders defined by the intertwining binary trees~$S(R)$ and~$T(R)$~\cite{MR4014603}.
+The weak poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
+We define the set~$\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
 \[
@@
 
-\begin{theorem}
+\begin{theorem}[\cite{LawReading}]
   The subsets $\mathcal{L}(P_w(R))$ of $\f{S}_n$, for all rectangulations $R$ on $n$ rectangles, 
-  are equivalence classes of a congruence relation $
-\weakeq$ on the weak Bruhat order on $\f{S}_n$.
+  are equivalence classes of a congruence relation $\weakeq$ on the weak Bruhat order on $\f{S}_n$.
   Furthermore:
   \begin{itemize}
-  \item The congruence classes are one-to-one with weak rectangulations,  
+  \item the congruence classes are one-to-one with weak rectangulations,  
   \item the minimal elements of each congruence class are the twisted Baxter permutations,
@@
 
-\subsection{Loday and anti-Loday arcs}
+\subsection{Up and down arcs}
 
@@
 
-% jean: From this, do we directly have a description of the face lattice of $\WRP(n)$?}
+\subsection{Shard polytopes of up and down arcs}
 
-\subsection{Shard polytopes of Loday and anti-Loday arcs}
-
-Therefore, the quotientope is realized by a Minkowski sum of the shard polytopes defined by two families of arcs: the \defn{Loday arcs} and the \defn{anti-Loday arcs}, that go respectively above and below all elements of an interval $I$ of $[n]$.
+Therefore, the quotientope is realized by a Minkowski sum of the shard polytopes defined by two families of arcs: the \defn{up arcs}~$\upArc{I}$ and the \defn{down arcs}~$\downArc{I}$, that go respectively above and below all elements of an interval $I$ of $[n]$.
 
@@
   \label{lem:lodaysp}
-  Given a Loday arc $A$ defined by an interval $I$, the Loday shard polytope $\SP(A)$ is a translate of the standard simplex $\conv \set{ \b{e}_i }{ i\in I }$.
+%  Given an up arc~$\alpha$ defined by an interval $I$, the up shard polytope $\SP(\alpha)$ is a translate of the standard simplex~$\conv \set{ \b{e}_i }{ i\in I }$.
+  The up shard polytope $\SP(\upArc{I})$ is a translate of the simplex~$\conv \set{ \b{e}_i }{ i\in I }$.
 \end{lemma}
@@
   \label{lem:antilodaysp}
-  Given an anti-Loday arc $A$ defined by an interval $I$, the anti-Loday shard polytope $\SP(A)$ is a translate of the negative standard simplex $\conv \set{ - \b{e}_i }{ i\in I }$.
+%  Given a down arc~$\alpha$ defined by an interval $I$, the down shard polytope $\SP(\alpha)$ is a translate of the negative standard simplex~$\conv \set{ - \b{e}_i }{ i\in I }$.
+  The down shard polytope $\SP(\downArc{I})$ is a translate of the negative simplex~$\conv \set{ - \b{e}_i }{ i\in I }$.
 \end{lemma}
@@
   \label{lem:weakMinkowski}
-  The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all Loday and anti-Loday shard polytopes.
-  In particular, it is the Minkowski sum of two opposite Loday associahedra.
+  The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all up and down shard polytopes.
+  In particular, it is the Minkowski sum of two opposite associahedra of J.-L.~Loday~\cite{LawReading}.
 \end{lemma}
-
-\vincent{Say also the minkowski sum of all faces of the standard simplex corresponding to intervals, and all their opposite.}
-\vincent{Say that this was already known by Law and Reading.}
 
@@
   \[
-  f(S) \eqdef \#\bigset{ I\text{ interval of } [n] }{ I\cap S\not=\emptyset \text{ and } I\cap ([n]\ssm S) \not= \emptyset }.
+  f_{\weakeq}(S) \eqdef \#\bigset{ I\text{ interval of } [n] }{ I \not\subseteq S \text{ and } I \not\subseteq [n]\ssm S }.
   \]
@@
 
-The proof relies on the following simple lemma stating that the submodular function of a Minkowski sum of deformed permutahedra is the sum of the submodular function of the summands.
-
-\begin{lemma}
-  \label{lem:submodsum}
-  Consider a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by the submodular functions $f_1,f_2,\ldots ,f_k$.
+%The proof relies on the following simple lemma stating that the submodular function of a Minkowski sum of deformed permutahedra is the sum of the submodular function of the summands.
+%
+%\begin{lemma}
+%  \label{lem:submodsum}
+%  Consider a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by the submodular functions $f_1,f_2,\ldots ,f_k$.
 Then the Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f \eqdef \sum_i f_i$.
-\end{lemma}
+%\end{lemma}
 
 \begin{proof}
-  From \cref{lem:weakMinkowski,lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the Loday and anti-Loday shard polytope in \cref{lem:lodaysp,lem:antilodaysp}.
-For a Loday arc $A$ defined by an interval $I$ of $[n]$, we have that its shard polytope $\SP(A)$ is realized by the submodular function
+  %From \cref{lem:weakMinkowski,lem:submodsum},
+  The submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the up and down shard polytope in \cref{lem:lodaysp,lem:antilodaysp}.
+The shard polytope~$\SP(\upArc{I})$ of an up arc~$\upArc{I}$ is given by the submodular function
   \[
-  \loday{f}_I(S) \eqdef \llbracket S\cap I\not=\emptyset \rrbracket .
+  f_{\upArc{I}}(S) \eqdef \llbracket S\cap I\not=\emptyset \rrbracket .
   \]
   (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and 0 otherwise.)
-  Similarly, for an anti-Loday arc $A$ defined by an interval $I$, the shard polytope $\SP(A)$ is defined by the following submodular function:
+  Similarly, the shard polytope~$\SP(\downArc{I})$ of a down arc~$\downArc{I}$ is given by the submodular function
   \[
-  \antiloday{f}_I(S) \eqdef - \llbracket I\subseteq S \rrbracket .
+  f_{\downArc{I}}(S) \eqdef - \llbracket I\subseteq S \rrbracket .
   \]
@@
   \[
-  f(S) \eqdef \sum_{I\text{ interval of }[n]} \loday{f}_I(S) + \antiloday{f}_I(S)
+  f_{\weakeq}(S) = \sum_{I\text{ interval of }[n]} f_{\upArc{I}}(S) + f_{\downArc{I}}(S)
   \]
-  is therefore exactly the number of intervals of $I$ that intersect $S$ but are not contained in $S$, hence the number of intervals intersecting both $S$ and its complement.
+  is therefore exactly the number of intervals of $I$ that intersect $S$ but are not contained in $S$, hence the number of intervals contained neither in~$S$ nor in its complement.
 \end{proof}
@@
 
-\begin{theorem}
+\begin{theorem}[\cite{Reading}]
   The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ on $n$ rectangles, 
@@
   \begin{itemize}
-  \item The congruence classes are one-to-one with strong rectangulations,  
+  \item the congruence classes are one-to-one with strong rectangulations,  
   \item the minimal elements of each congruence class are the 2-clumped permutations,

### 2024-04-04 draft/rectangulotopes.tex
Section 3.5

@@
 
+The \defn{Minkowski sum} of two polytopes~$\polytope{P}$ and~$\polytope{Q}$ is~$\polytope{P} + \polytope{Q} \eqdef \set{\b{p} + \b{q}}{\b{p} \in \polytope{P} \text{ and } \b{q} \in \polytope{Q}}$.
+Recall that for any vector~$\b{v} \ne \b{0}$, the face of~$\polytope{P} + \polytope{Q}$ maximizing the scalar product with~$\b{v}$ is the Minkowski sum of the faces of~$\polytope{P}$ and~$\polytope{Q}$ maximizing the scalar product with~$\b{v}$.
+In particular, if~$\b{v}$ is a generic direction, the vertex of~$\polytope{P} + \polytope{Q}$ extremal in direction~$\b{v}$ is just the sum of the vertices of~$\polytope{P}$ and~$\polytope{Q}$ that are extremal in direction~$\b{v}$.
+
 Consider an arc~$\alpha \eqdef (a, b, A, B)$ on~$[n]$.
@@
 
-For every rectangulation $R$, we wish to give the coordinates of a point $p(R)\in\R^n$ such that $\WRP(n)$ is the convex hull of the points $p(R)$ for all rectangulations $R$.
+For every weak rectangulation $R$, we wish to give the coordinates of a point $\b{p}(R)\in\R^n$ such that $\WRP(n)$ is the convex hull of the points~$\b{p}(R)$ for all weak rectangulations~$R$.
 Recall that in a Minkowski sum of polytopes, the vertex that is extremal in a generic direction is the sum of the vertices of the summands that are extremal in this direction.
@@
   \label{lem:lodaymax}
-  Let $\pi\in\f{S}_n$ and consider a Loday arc $A$ defined by an interval $I$.
-  The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
-  is $\b{e}_i$, where $i \eqdef \arg\max_{j\in I} \pi^{-1}(j)$.
-  Similarly, if $A$ is an anti-Loday arc defined by $I$, the unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
-  is $-\b{e}_i$, where $i \eqdef \arg\min_{j\in I} \pi^{-1}(j)$.
+  Let $\pi \in \f{S}_n$ and $I$ be a subset of~$[n]$.
+  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\pi^{-1}$ is $\b{e}_i$, where $i \eqdef \arg\max_{j\in I} \pi^{-1}(j)$ (resp.~where~$i \eqdef \arg\min_{j\in I} \pi^{-1}(j)$).
 \end{lemma}
@@
 \begin{lemma}
-  Let $R$ be a rectangulation on $n$ rectangles and $p(R)$ the vertex of $\WRP(n)$ associated with $R$.
-  Given $i\in [n]$, the number of Loday arcs $A$ such that $\b{e}_i$ is the extremal vertex of $\SP(A)$ contributing to $p(R)$ is
+  Let $R$ be a rectangulation on~$n$ rectangles, with source and target trees~$S$ and~$T$, and let $\b{p}(R)$ the vertex of $\WRP(n)$ associated with $R$.
+  Given $i\in [n]$, the number of up arcs~$\upArc{I}$ such that $\b{e}_i$ is the extremal vertex of $\SP(\upArc{I})$ contributing to~$\b{p}(R)$ is
   \[
-  \loday{w}_i =  h(T(R), i)\cdot v(T(R),i),
+  \loday{w}_i =  h^{T}_i\cdot v^{T}_i,
   \]
-   where $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
+   where $h^T_i$ and $v^T_i$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
 \end{lemma}

### 2024-04-04 draft/rectangulotopes.tex
Sections 4.2 and 4.3

@@
 % yin and yang accents (?)
+\newcommand{\yin}[1]{\smash{\overset{\sim}{#1}}}
 \newcommand{\yang}[1]{\smash{\overset{\backsim}{#1}}}
-\newcommand{\yin}[1]{\smash{\overset{\sim}{#1}}}
+\newcommand{\yinArc}[2]{\smash{\raisebox{.05cm}{$\uset[0ex]{#1}{\frown}$}\hspace{-.3ex}\raisebox{-.03cm}{$\oset[.2ex]{#2}{\smile}$}}}
+\newcommand{\yangArc}[2]{\smash{\raisebox{-.03cm}{$\oset[.2ex]{#1}{\smile}$}\hspace{-.3ex}\raisebox{.05cm}{$\uset[0ex]{#2}{\frown}$}}}
 % weak and strong equivalence on S_n
-\newcommand{
-\weakeq}{\asymp}
-%\makeatletter
-%\DeclareRobustCommand{\strongeq}{%
-%  \mathbin{\mathpalette\scaletoplus@{\mkern1mu \smash{\begin{smallmatrix} \backsim \\[-.3cm] \sim \end{smallmatrix}} \mkern1mu}}%
-%}
-%\newcommand{\scaletoplus@}[2]{%
-%  \vcenter{%
-%    \hbox{\sbox\z@{$\m@th#1+$}\resizebox{\wd\z@}{!}{$\m@th#1#2$}}%
-%  }%
-%}
-%\makeatother
+\newcommand{\weakeq}{\asymp}
 \newcommand{\strongeq}{\mathbin{\smash{\begin{smallmatrix} \backsim \\[-.3cm] \sim \end{smallmatrix}}}}% % J: I would like to stack \backsim and \sim instead
@@
 
+These arcs can be classified into two families of arcs: the up arcs and the down arcs.
+For an interval $I$ of $[n]$ of size at least~$2$, the \defn{up arc}~$\upArc{I}$ (resp.~\defn{down arc}~$\downArc{I}$) is the arc that starts at~$\min(I)$ and ends at~$\max(I)$, and passes above (resp.~below) all remaining elements of~$I$, that is
+\begin{align*}
+\upArc{I} & \eqdef \big( \min(I), \max(I), I \ssm \{\min(I), \max(I)\}, \varnothing \big) \\
+\qquad\text{and}\qquad
+\downArc{I} & \eqdef \big( \min(I), \max(I), \varnothing, I \ssm \{\min(I), \max(I)\} \big).
+\end{align*}
+Note that the basic arc~$(i, i+1, \varnothing, \varnothing)$ is both an up and down arc.
+
 \subsection{Shard polytopes of up and down arcs}
 
-Therefore, the quotientope is realized by a Minkowski sum of the shard polytopes defined by two families of arcs: the \defn{up arcs}~$\upArc{I}$ and the \defn{down arcs}~$\downArc{I}$, that go respectively above and below all elements of an interval $I$ of $[n]$.
+Therefore, the quotientope is realized by a Minkowski sum of the shard polytopes of the up arcs and the down arcs.
+These quotientopes are very basic ones.
 
@@
 We call the congruence $\strongeq$ the \defn{strong rectangulation congruence}.
-The \defn{strong rectangulotope} $\SRP(n)$ is the quotientope of the strong rectangulation congruence $\strongeq$.
+The \defn{strong rectangulotope} $\SRP(n)$ is the quotientope of the strong rectangulation congruence~${\strongeq}$.
 
@@
 We classify the arcs into two families, depending on whether the arcs cross from above or below.
-A \defn{yin arc} is defined by two consecutive nonempty intervals $I$ and $J$ of $[n]$ such that the arc goes above all elements of $I$ and below all elements of $J$.
-A \defn{yang arc} is defined similarly, except that the arc goes below the elements of $I$ and above the elements of $J$.
+For two consecutive intervals~$I$ and~$J$ of~$[n]$, the \defn{yin arc}~$\yinArc{I}{J}$ (resp.~\defn{yang arc}~$\yangArc{I}{J}$) is the arc starts at~$\min(I)$ and ends at~$\max(J)$, and goes above (resp.~below) the remaining elements of $I$ and below (resp.~above) the remaining elements of $J$, that is,
+\begin{align*}
+\yinArc{I}{J} & \eqdef \big( \min(I), \max(J), I \ssm \{\min(I)\}, J \ssm \{\max(J)\} \big) \\
+\qquad\text{and}\qquad
+\yangArc{I}{J} & \eqdef \big( \min(I), \max(J), J \ssm \{\max(J)\}, I \ssm \{\min(I)\} \big).
+\end{align*}
 Note that all up arcs and down arcs are both yin arcs and yang arcs.
@@
   \label{lem:yinsp}
-  Given a yin arc $A=(I,J)$, the yin shard polytope $\SP(A)$ is defined as the convex hull of the points
-  $\{\b{0}\} \cup \set{\b{e}_i - \b{e}_j}{(i,j)\in I\times J}$.
+  The yin shard polytope $\SP(\yinArc{I}{J})$ is $\conv \{\b{0}\} \cup \set{\b{e}_i - \b{e}_j}{i \in I, \; j \in J}$.
 \end{lemma}
@@
   \label{lem:yangsp}
-  Given a yang arc $A=(I,J)$, the yang shard polytope $\SP(A)$ is a translate of the convex hull of the points
-  $\{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{(i,j)\in I\times J}$.
+  The yang shard polytope $\SP(\yangArc{I}{J})$ is a translate of~$\conv \{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{i \in I,\; j \in J}$.
 \end{lemma}
 
-Note that yin shard polytopes are translates of the Loday shard polytopes when $|J|=1$, and of the anti-Loday shard polytopes when $|I|=1$.
-Similarly, Loday and anti-Loday shard polytopes are translates of some yang shard polytopes.
+%Note that yin shard polytopes are translates of the up shard polytopes when $|J|=1$, and of the down shard polytopes when $|I|=1$.
+%Similarly, up and down shard polytopes are translates of some yang shard polytopes.
 
@@
   \label{lem:strongMinkowski}
-  The strong rectangulotope $\SRP(n)$ is realized by the Minkowski sum of all shard polytopes of yin and yang arcs.
+  The strong rectangulotope $\SRP(n)$ is realized by the Minkowski sum of all yin and yang shard polytopes.
 \end{lemma}
@@
 Since the combinatorial type of a Minkowski sum is invariant to translation of the summands, we can use the polytopes defined in \cref{lem:yinsp,lem:yangsp} as a definition of the yin and yang shard polytopes.
-Also note that since Loday and anti-Loday shard polytopes are, up to a translation, both yin and yang, each Loday and anti-Loday shard polytope is summed twice in the proposed realization.
+Also note that since the up and down shard polytopes are, up to a translation, both yin and yang, each up and down shard polytope is summed twice in the proposed realization.
 This, again, does not change the combinatorial type of the polytope.
 
-\cref{lem:strongsum} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.
+\cref{lem:strongMinkowski} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.
 
 \subsection{Submodular functions of strong rectangulotopes}
-
-\jean{That seems to work.}
 

### 2024-04-04 draft/rectangulotopes.tex
Section 4.4

@@
   \[
-  f(S) \eqdef \# \bigset{ I, J }{ I\cap S\not=\emptyset \wedge J\cap ([n]\ssm S) \not= \emptyset } +
-  \#\bigset{ I, J }{ J\cap S\not=\emptyset \wedge I\cap ([n]\ssm S) \not= \emptyset },
+  f(S) \eqdef \# \bigset{ I, J }{ I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S } +
+  \#\bigset{ I, J }{ I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S },
   \]
-  where $I, J$ are pairs of consecutive and nonempty intervals of $[n]$.
+  where $I, J$ are pairs of nonempty consecutive intervals of~$[n]$.
 \end{lemma}
@@
 \begin{proof}
-  From \cref{lem:strongMinkowski,lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the yin and yang shard polytope in \cref{lem:yinsp,lem:yangsp}.
-For a yin arc $A$ defined by consecutive intervals $I, J$ of $[n]$, we have that its shard polytope $\SP(A)$ is realized by the submodular function
+  The submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the yin and yang shard polytopes in \cref{lem:yinsp,lem:yangsp}.
+  The shard polytope~$\SP(\yinArc{I}{J})$ of a yin arc $\yinArc{I}{J}$ is given by the submodular function
   \[
-  \yin{f}_{I,J}(S) \eqdef \llbracket S\cap I\not=\emptyset \wedge S\not\supseteq J \rrbracket .
+  f_{\yinArc{I}{J}}(S) \eqdef \llbracket I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S \rrbracket .
   \]
-  Similarly, for a yang arc, the shard polytope $\SP(A)$ is defined by the following submodular function:
+  (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and 0 otherwise.)
+  Similarly, the shard polytope~$\SP(\yangArc{I}{J})$ of a yang arc~$\yangArc{I}{J}$ is given by the submodular function
   \[
-  \yang{f}_{I,J}(S) \eqdef \llbracket S\cap J\not=\emptyset \wedge S\not\supseteq I \rrbracket .
+  f_{\yangArc{I}{J}}(S) \eqdef \llbracket I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S \rrbracket .
   \]
@@
   \[
-  f(S) \eqdef \sum_{I,J} \yin{f}_{I,J}(S) + \yang{f}_{I,J}(S)
+  f_{\strongeq}(S) \eqdef \sum_{I,J} f_{\yinArc{I}{J}}(S) + f_{\yangArc{I}{J}}(S)
   \]

### 2024-04-04 draft/rectangulotopes.tex
labels

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \section{Introduction}
+\label{sec:intro}
+
+%%%%%%%%%%%%%
 
 \subsection{Polytopes of permutations, triangulations, and rectangulations}
+\label{subsec:permTrianRect}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Weak and strong rectangulations}
+\label{subsec:weakStrongRectangulations}
 
@@
 \end{itemize}
+
+%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Weak rectangulotopes}
+\label{subsec:weakRectangulotopes}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Strong rectangulotopes}
+\label{subsec:strongRectangulotopes}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Plan of the paper}
+\label{subsec:plan}
 
 \cref{sec:quotientopes} is dedicated to the background on quotientopes and their realizations as Minkowski sums of \defn{shard polytopes}, due to A.~Padrol, V.~Pilaud, and J.~Ritter~\cite{MR4584712}, which will be our main tool throughout.
-\cref{sec:wr} details the case of weak rectangulations, while \cref{sec:sr} deals with strong rectangulations.
+\cref{sec:weakRectangulotopes} details the case of weak rectangulations, while \cref{sec:strongRectangulotopes} deals with strong rectangulations.
 In both cases, in addition to the vertex coordinates, we also give the submodular functions defining the polytopes.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Weak order and noncrossing arc diagrams}
+\label{subsec:noncrossingArcDiagrams}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Quotients and arc ideals}
+\label{subsec:arcIdeals}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Quotient fans and shards}
+\label{subsec:quotientFans}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Quotientopes and shard polytopes}
+\label{subsec:quotientopes}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \section{Weak rectangulotopes}
-\label{sec:wr}
+\label{sec:weakRectangulotopes}
 
-%%%%%%%%%%%%
+%%%%%%%%%%%%%
 
 \subsection{The weak rectangulation congruence}
+\label{subsec:weakRectangulationCongruence}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Up and down arcs}
+\label{subsec:upDownArcs}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Shard polytopes of up and down arcs}
+\label{subsec:upDownShardPolytopes}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Submodular functions of weak rectangulotopes}
+\label{subsec:submodularWeakRectangulotopes}
 
@@
 
-\subsection{Loday-type coordinates for weak rectangulotopes}
+%%%%%%%%%%%%%
+
+\subsection{Loday coordinates of weak rectangulotopes}
+\label{subsec:LodayWeakRectangulotopes}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \section{Strong rectangulotopes}
-\label{sec:sr}
+\label{sec:strongRectangulotopes}
 
-%%%%%%%%%%%%
+%%%%%%%%%%%%%
+
 \subsection{The strong rectangulation congruence}
+\label{subsec:strongRectangulationCongruence}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Yin and yang arcs}
+\label{subsec:yinYangArcs}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Shard polytopes of yin and yang arcs}
+\label{subsec:yinYangShardPolytopes}
 
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Submodular functions of strong rectangulotopes}
+\label{subsec:submodularStrongRectangulotopes}
 
@@
 
-\subsection{Loday-type coordinates for strong rectangulotopes}
+%%%%%%%%%%%%%
 
-For every rectangulation $R$, we wish to give the coordinates of a point $p(R)\in\R^n$ such that $\SRP(n)$ is the convex hull of the points $p(R)$ for all rectangulations $R$.
+\subsection{Loday coordinates of strong rectangulotopes}
+\label{subsec:LodayStrongRectangulotopes}
+
+For every strong rectangulation $R$, we wish to give the coordinates of a point $\b{p}(R)\in\R^n$ such that $\SRP(n)$ is the convex hull of the points $\b{p}(R)$ for all strong rectangulations $R$.
 We first consider the contributions, to these coordinates, of the Minkowski summand $\SP(A)$ for all yin arcs $A$.
-Recall that in a Minkowski sum of polytopes, the vertex that is extremal in a generic direction is the sum of the vertices of the summands that are extremal in this direction.
-We therefore need to understand which vertices of the translates of the yin shard polytopes defined in \cref{lem:yinsp} are extremal for a direction given by a permutation.
+%Recall that in a Minkowski sum of polytopes, the vertex that is extremal in a generic direction is the sum of the vertices of the summands that are extremal in this direction.
+We need to understand which vertices of the translates of the yin shard polytopes defined in \cref{lem:yinsp} are extremal for a direction given by a permutation.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%
 \newpage
@@
 Properties/results we may need along the way.
+
+%%%%%%%%%%%%%
 
@@
 \end{lemma}
+
+%%%%%%%%%%%%%
 

### 2024-04-04 draft/rectangulotopes.tex
Section 4.5

@@
   \label{lem:lodaymax}
-  Let $\pi \in \f{S}_n$ and $I$ be a subset of~$[n]$.
-  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\pi^{-1}$ is $\b{e}_i$, where $i \eqdef \arg\max_{j\in I} \pi^{-1}(j)$ (resp.~where~$i \eqdef \arg\min_{j\in I} \pi^{-1}(j)$).
+  Let $\pi \in \f{S}_n$ and $I$ be an interval of~$[n]$ of size at least~$2$.
+  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~of the translated down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\pi^{-1}$ is $\b{e}_i$, where $i \eqdef \arg\max_{j\in I} \pi^{-1}(j)$ (resp.~where~$i \eqdef \arg\min_{j\in I} \pi^{-1}(j)$).
 \end{lemma}
@@
 
-For every strong rectangulation $R$, we wish to give the coordinates of a point $\b{p}(R)\in\R^n$ such that $\SRP(n)$ is the convex hull of the points $\b{p}(R)$ for all strong rectangulations $R$.
-We first consider the contributions, to these coordinates, of the Minkowski summand $\SP(A)$ for all yin arcs $A$.
-%Recall that in a Minkowski sum of polytopes, the vertex that is extremal in a generic direction is the sum of the vertices of the summands that are extremal in this direction.
-We need to understand which vertices of the translates of the yin shard polytopes defined in \cref{lem:yinsp} are extremal for a direction given by a permutation.
+As in \cref{subsec:LodayWeakRectangulotopes}, we need to understand which vertices of the translates of the yin and yang shard polytopes defined in \cref{lem:yinsp,lem:yangsp} are extremal for a direction given by a permutation.
 
@@
   \label{lem:yinminmax}
-  Let $\pi\in\f{S}_n$ and consider a yin arc $A=(I,J)$.
-  Let $i \eqdef \arg\max_{k\in I} \pi^{-1}(k)$ and $j \eqdef \arg\min_{k\in J} \pi^{-1}(k)$.
-  The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
-  is $\b{e}_i-\b{e}_j$ if $\pi^{-1}(i)>\pi^{-1}(j)$, and $\b{0}$ otherwise.
+  Let $\pi\in\f{S}_n$, let~$I$ and~$J$ be two consecutive intervals of~$[n]$, and let $i \eqdef \arg\max_{k\in I} \pi^{-1}(k)$ and $j \eqdef \arg\min_{k\in J} \pi^{-1}(k)$.
+  The unique vertex of the yin shard polytope~$\SP(\yinArc{I}{J})$ that is extreme with respect to the direction $\pi^{-1}$ is $\b{e}_i-\b{e}_j$ if $\pi^{-1}(i)>\pi^{-1}(j)$, and $\b{0}$ otherwise.
 \end{lemma}
@@
 \begin{lemma}
-  Let $R$ be a rectangulation on $n$ rectangles, $([n],\prec_s)$ the associated strong order, and $p(R)$ the vertex of $\SRP(n)$ corresponding to $R$.
-  Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $A$ such that $\b{e}_i-\b{e}_j$
-  is the extremal vertex of $\SP(A)$ contributing to $p(R)$ is
+  Let $R$ be a strong rectangulation on $n$ rectangles, $S$ and $T$ be its source and target trees, $([n],\prec_s)$ be its strong order, and $\b{p}(R)$ be the vertex of $\SRP(n)$ corresponding to $R$.
+  Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $A$ such that $\b{e}_i-\b{e}_j$ is the extremal vertex of $\SP(A)$ contributing to $p(R)$ is
   \[
-    h(T(R), i) \cdot cv (T(R), S(R), i, j)\cdot h(S(R), j) 
+    h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_j 
   \]
@@
 \begin{proof}
+  \vincent{I still need to work out this proof...}
   Let $\pi$ be the 2-clumped permutation corresponding to $R$, hence the minimal element of $\mathcal{L}(P_s(R))$ in the weak Bruhat order.
@@
 
-It remains to observe that the only case where this product is nonzero and $i,j$ is not an inversion in $\pi$ is when $i\backslash j$.
+It remains to observe that the only case where this product is nonzero and $i,j$ is not an inversion in $\pi$ is when $i \, \backslash \, j$.
 
 \begin{lemma}
-  Let $R$ be a rectangulation on $n$ rectangles and $([n],\prec_s)$ the associated strong order.
+  Let $R$ be a rectangulation on $n$ rectangles, $S$ and $T$ be its source and target trees, and $([n],\prec_s)$ be the associated strong order.
   Given a pair $i,j\in [n]$ such that $i<j$ and $i\prec_s j$, we have
   \[
-    cv (T(R), S(R), i, j) > 0 \implies i\backslash j.
+    cv^{T,S}_{i,j} > 0 \quad \Longrightarrow \quad i \, \backslash \, j.
   \]

### 2024-04-04 draft/rectangulotopes.tex
shuffled intro

--- 2024-04-04 draft/rectangulotopes.tex
@@
 % Loday and anti-Loday accents (?)
-\makeatletter
-\newcommand{\oset}[3][0ex]{%
-  \mathrel{\mathop{#3}\limits^{
-    \vbox to#1{\kern-2\ex@
-    \hbox{$\scriptstyle#2$}\vss}}}}
-%
-\newcommand{\uset}[3][0ex]{%
-  \mathrel{\mathop{#3}\limits_{
-    \vbox to#1{\kern-7\ex@
-    \hbox{$\scriptstyle#2$}\vss}}}}
-\makeatother
-%
 \newcommand{\loday}[1]{\smash{\overset{\frown}{#1}}}
 \newcommand{\antiloday}[1]{\smash{\overset{\smile}{#1}}}
-\newcommand{\upArc}[1]{\smash{\raisebox{.05cm}{$\uset[0ex]{#1}{\frown}$}}}
-\newcommand{\downArc}[1]{\smash{\raisebox{-.05cm}{$\oset[.2ex]{#1}{\smile}$}}}
 % yin and yang accents (?)
+\newcommand{\yang}[1]{\smash{\overset{\backsim}{#1}}}
 \newcommand{\yin}[1]{\smash{\overset{\sim}{#1}}}
-\newcommand{\yang}[1]{\smash{\overset{\backsim}{#1}}}
-\newcommand{\yinArc}[2]{\smash{\raisebox{.05cm}{$\uset[0ex]{#1}{\frown}$}\hspace{-.3ex}\raisebox{-.03cm}{$\oset[.2ex]{#2}{\smile}$}}}
-\newcommand{\yangArc}[2]{\smash{\raisebox{-.03cm}{$\oset[.2ex]{#1}{\smile}$}\hspace{-.3ex}\raisebox{.05cm}{$\uset[0ex]{#2}{\frown}$}}}
 % weak and strong equivalence on S_n
-\newcommand{\weakeq}{\asymp}
+\newcommand{
+\weakeq}{\asymp}
+%\makeatletter
+%\DeclareRobustCommand{\strongeq}{%
+%  \mathbin{\mathpalette\scaletoplus@{\mkern1mu \smash{\begin{smallmatrix} \backsim \\[-.3cm] \sim \end{smallmatrix}} \mkern1mu}}%
+%}
+%\newcommand{\scaletoplus@}[2]{%
+%  \vcenter{%
+%    \hbox{\sbox\z@{$\m@th#1+$}\resizebox{\wd\z@}{!}{$\m@th#1#2$}}%
+%  }%
+%}
+%\makeatother
 \newcommand{\strongeq}{\mathbin{\smash{\begin{smallmatrix} \backsim \\[-.3cm] \sim \end{smallmatrix}}}}% % J: I would like to stack \backsim and \sim instead
@@
 
-The goal of this paper is to describe the quotientopes of congruences of the weak Bruhat order whose classes are in bijection with (equivalence classes of) \defn{rectangulations}, defined as decompositions of a square into finitely many axis-aligned rectangles.
-The combinatorics of rectangulations has been studied for decades~\cite{MR2233287,MR2871762,MR2864445,MR2763051,MR3084577,MR3878132,MR4598046}, and the order-theoretic take on these structures has already proved useful.
-%, in particular in the context of electronic engineering applications~\cite{}.
-\vincent{More citations on electronic engineering...}
-There are two relevant lattice congruences related with rectangulations.
-%
-On the one hand, the \defn{weak rectangulation congruence} (or \defn{Baxter congruence}), studied by S.~Law and N.~Reading~\cite{MR2871762}, and whose classes are one-to-one with \defn{weak rectangulations} (or \defn{diagonal rectangulations}), with \defn{twin binary trees}~\cite{Giraudo}, and with \defn{Baxter permutations}.
-\vincent{Give more Baxter references here.}
+The goal of this paper is to describe the quotientopes of congruences of the weak Bruhat order whose classes are in bijection with equivalence classes of rectangulations, defined as decompositions of a square into rectangles.
+
+\subsection{Weak and strong rectangulations}
+
+The combinatorics of rectangulations has been studied for decades~\cite{MR2233287,MR2871762,MR2864445,MR2763051,MR3084577,MR3878132,MR4598046}.
+Interestingly, many combinatorial results were initially motivated by applications to the design of \defn{floorplans} for very large scale integrated circuits, and published in electrical engineering journals~\cite{}.
+The recent order-theoretic take on these structures has proved fruitful.
+
+We define a \defn{rectangulation} of size $n$ as a decomposition of the square into $n$ axis-parallel rectangles with disjoint interiors.
+The \defn{segments} of a rectangulation are the inclusionwise maximal line segments composed of edges of the rectangles, excluding the edges of the decomposed square.
+We suppose throughout that no four rectangles have a common vertex, hence that the rectangulations are \defn{generic} in that sense.
+In order to focus on the combinatorial structure of a rectangulation, we introduce two equivalence relations between rectangulations.
+First, we say that a rectangle $r$ is \defn{above} another rectangle $s$ (and $s$ is \defn{below} $r$) if either the bottom edge of $r$ lies in the same segment as the top edge of $s$, or if $r$ is above another rectangle $t$ that is above $s$.
+Similarly, a rectangle $r$ is \defn{on the left of} another rectangle $s$ (and $s$ is \defn{on the right of} $r$) if either the right edge of $r$ lies in the same segment as the left edge of $s$, or if $r$ is on the left of another rectangle $t$ that is on the left of $s$.
+
+Two rectangulations are said to be \defn{weakly equivalent} if there exists a bijection between their rectangles that preserves the above-below and left-right relations between the rectangles.
+Weak equivalence classes of rectangulations will be referred to as \defn{weak rectangulations}.
+(Note that this is a slight, yet convenient abuse of terminology, as weak rectangulations are really sets of rectangulations.)
+On the other hand, two rectangulations are said to be \defn{strongly equivalent} when there exists a bijection between their rectangles that not only preserves the above-below and left-right relations, but also the adjacency relation between rectangles.
+We will naturally refer to strong equivalence classes of rectangulations as \defn{strong rectangulations}.
+
+In order to make sense of these definitions, it is useful to consider \defn{wall slides} in a rectangulation: the local changes consisting of shifting a horizontal segment vertically, or a vertical segment horizontally, while extending or shortening the incident segments accordingly.
+Performing a wall slide in a rectangulation $R$ leads to a rectangulation that is always weakly equivalent to $R$.
+However, if the wall slide changes the adjacency relation between the rectangles, then the resulting rectangulation is not strongly equivalent to $R$ anymore.
+This is illustrated in Figure~\ref{fig:rectequiv}.
+A \defn{diagonal rectangulation} is a rectangulation in which every rectangle intersects the top-left to bottom-right diagonal of the square.
+It is simple to check that every weak rectangulation has a diagonal representative, hence weak rectangulations can also be thought of as diagonal rectangulations.
+
+\begin{figure}
+  \caption{\label{fig:rectequiv}Weak and strong equivalence between rectangulations.}
+\end{figure}
+
+\subsection{Weak and strong rectangulotopes}
+
+Weak and strong rectangulations of size $n$ have been shown to define congruences of the weak Bruhat order on $\f{S}_n$.
+The \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762}.
+Its classes are one-to-one with weak rectangulations, but also with \defn{twin binary trees}~\cite{Giraudo}, and with \defn{Baxter permutations}~\cite{}. % more baxter refs
 The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
-The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
-%
-On the other hand, the \defn{strong rectangulation congruence}, whose classes are in bijection with \defn{strong rectangulations}~\cite{MR2864445}.
+On the other hand, the \defn{strong rectangulation congruence}, whose classes are in bijection with strong rectangulations~\cite{MR2864445}.
 The latter is mentioned by N.~Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
 The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
-The skeleton of $\SRP(n)$ is isomorphic to the flip graph of strong rectangulations of size $n$ defined by Meehan~\cite{MR3697823} (see also~\cite{ACFF24}).
+
+An important motivation for studying these quotientopes is the informative structure of their skeleta, which are \defn{flip graphs} on the rectangulations.
+A flip is a local, reversible change in the structure of the rectangulation.
+Flips in rectangulations come in three flavors:
+\begin{description}
+\item[Simple flips] replace a vertical segment incident to exactly two rectangles by a horizontal segment, and vice-versa.
+\item[Pivoting] operations transform a pair of adjacent rectangles whose union is not a rectangle, changing a left-right pair into an above-below pair, and vice-versa.
+\item[Wall slide] flips are wall slides that modify the adjacency of the rectangles, removing exactly one adjacency and adding exactly one. These flips are only relevant in strong rectangulations.
+\end{description}
+
+The various types of flips are illustrated in Figure~\ref{fig:flips}.
+We refer to S.~Law and N.~Reading~\cite{MR2871762}, J.~Cardinal, V.~Sacrist\'an, and R.~Silveira~\cite{}, E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and \'E.~Fusy~\cite{ACFF24} for detailed descriptions of the flip graphs on both weak and strong rectangulations.
+Just like associahedra encode the rotation graphs on binary trees, the rectangulotopes directly yield flip graphs on rectangulations.
+
+\begin{lemma}
+  The skeleton of $\WRP(n)$ is isomorphic to the flip graph on weak rectangulations of size $n$.
+  Similarly, the skeleton of $\SRP(n)$ is isomorphic to the flip graph on strong rectangulations of size $n$.
+\end{lemma}
 
@@
 For facets, we show that both weak and strong rectangulotopes have $2^n-2$ facets (the maximal possible number of facets of a deformed permutahedron), and provide simple formulas for the right hand side of the inequality corresponding to a given proper nonempty subset of~$[n]$.
+
 We now give the necessary definitions for precisely stating these vertex and facet descriptions of the weak and strong rectangulotopes.
-
-\subsection{Weak and strong rectangulations}
-
-\begin{itemize}
-\item Terminology: rectangles, \defn{vertices}, \defn{segments}, \defn{generic} rectangulation
-\item \defn{Wall slides}
-\item Two equivalence classes: \defn{weak equivalence} preserving rectangle-segment adjacencies, and \defn{strong equivalence} preserving rectangle-rectangle adjacencies.
-\item Weak equivalence classes have a \defn{diagonal} representative.
-\item \defn{Flip graphs} on rectangulations
-\end{itemize}
 
@@
 Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$.
-We denote by $i \, \backslash \, j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$.
+We denote by %$i|j$ and $\frac{\ j\ }i$
+$i \, \backslash \, j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$. %, respectively.
 We also use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.
@@
   \end{itemize}
-  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ in the sets denote nonempty and consecutive intervals of~$[n]$.
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ in the sets denote consecutive intervals of~$[n]$.
   \end{enumerate}
@@
 
-The \defn{Minkowski sum} of two polytopes~$\polytope{P}$ and~$\polytope{Q}$ is~$\polytope{P} + \polytope{Q} \eqdef \set{\b{p} + \b{q}}{\b{p} \in \polytope{P} \text{ and } \b{q} \in \polytope{Q}}$.
-Recall that for any vector~$\b{v} \ne \b{0}$, the face of~$\polytope{P} + \polytope{Q}$ maximizing the scalar product with~$\b{v}$ is the Minkowski sum of the faces of~$\polytope{P}$ and~$\polytope{Q}$ maximizing the scalar product with~$\b{v}$.
-In particular, if~$\b{v}$ is a generic direction, the vertex of~$\polytope{P} + \polytope{Q}$ extremal in direction~$\b{v}$ is just the sum of the vertices of~$\polytope{P}$ and~$\polytope{Q}$ that are extremal in direction~$\b{v}$.
-
 Consider an arc~$\alpha \eqdef (a, b, A, B)$ on~$[n]$.
@@
 
+\begin{lemma}
+  \label{lem:submodsum}
+  Consider a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by the submodular functions $f_1,f_2,\ldots ,f_k$.
+Then the Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f \eqdef \sum_i f_i$.
+\end{lemma}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Given a rectangulation~$R$, the source and target trees~$S(R)$ and~$T(R)$ define two partial orders on $[n]$ (the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top).
-The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of the two partial orders defined by the intertwining binary trees~$S(R)$ and~$T(R)$~\cite{MR4014603}.
-The weak poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
-We define the set~$\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
+Given a rectangulation $R$, the two trees $T(R)$ and $S(R)$, where the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top,
+define two partial orders on $[n]$.
+We define the \defn{weak poset} $P_w(R)=([n],\prec_w)$ of a rectangulation $R$ as the transitive closure of the union of the two partial orders defined by the intertwining binary trees $T(R)$ and $S(R)$~\cite{MR4014603}.
+The weak poset is a two-dimensional lattice, the minimum of which is the root of $S(R)$ and the maximum is the root of $T(R)$.
+We define the set $\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
 \[
@@
 
-\begin{theorem}[\cite{LawReading}]
+\begin{theorem}
   The subsets $\mathcal{L}(P_w(R))$ of $\f{S}_n$, for all rectangulations $R$ on $n$ rectangles, 
-  are equivalence classes of a congruence relation $\weakeq$ on the weak Bruhat order on $\f{S}_n$.
+  are equivalence classes of a congruence relation $
+\weakeq$ on the weak Bruhat order on $\f{S}_n$.
   Furthermore:
   \begin{itemize}
-  \item the congruence classes are one-to-one with weak rectangulations,  
+  \item The congruence classes are one-to-one with weak rectangulations,  
   \item the minimal elements of each congruence class are the twisted Baxter permutations,
@@
 
-\subsection{Up and down arcs}
+\subsection{up and down arcs}
 
@@
 
-These arcs can be classified into two families of arcs: the up arcs and the down arcs.
-For an interval $I$ of $[n]$ of size at least~$2$, the \defn{up arc}~$\upArc{I}$ (resp.~\defn{down arc}~$\downArc{I}$) is the arc that starts at~$\min(I)$ and ends at~$\max(I)$, and passes above (resp.~below) all remaining elements of~$I$, that is
-\begin{align*}
-\upArc{I} & \eqdef \big( \min(I), \max(I), I \ssm \{\min(I), \max(I)\}, \varnothing \big) \\
-\qquad\text{and}\qquad
-\downArc{I} & \eqdef \big( \min(I), \max(I), \varnothing, I \ssm \{\min(I), \max(I)\} \big).
-\end{align*}
-Note that the basic arc~$(i, i+1, \varnothing, \varnothing)$ is both an up and down arc.
+% jean: From this, do we directly have a description of the face lattice of $\WRP(n)$?}
 
@@
 
-Therefore, the quotientope is realized by a Minkowski sum of the shard polytopes of the up arcs and the down arcs.
-These quotientopes are very basic ones.
+Therefore, the quotientope is realized by a Minkowski sum of the shard polytopes defined by two families of arcs: the \defn{up arcs} and the \defn{down arcs}, that go respectively above and below all elements of an interval $I$ of $[n]$.
 
@@
   \label{lem:lodaysp}
-%  Given an up arc~$\alpha$ defined by an interval $I$, the up shard polytope $\SP(\alpha)$ is a translate of the standard simplex~$\conv \set{ \b{e}_i }{ i\in I }$.
-  The up shard polytope $\SP(\upArc{I})$ is a translate of the simplex~$\conv \set{ \b{e}_i }{ i\in I }$.
+  Given a up arc $A$ defined by an interval $I$, the up shard polytope $\SP(A)$ is a translate of the standard simplex $\conv \set{ \b{e}_i }{ i\in I }$.
 \end{lemma}
@@
   \label{lem:antilodaysp}
-%  Given a down arc~$\alpha$ defined by an interval $I$, the down shard polytope $\SP(\alpha)$ is a translate of the negative standard simplex~$\conv \set{ - \b{e}_i }{ i\in I }$.
-  The down shard polytope $\SP(\downArc{I})$ is a translate of the negative simplex~$\conv \set{ - \b{e}_i }{ i\in I }$.
+  Given a down arc $A$ defined by an interval $I$, the down shard polytope $\SP(A)$ is a translate of the negative standard simplex $\conv \set{ - \b{e}_i }{ i\in I }$.
 \end{lemma}
@@
   The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all up and down shard polytopes.
-  In particular, it is the Minkowski sum of two opposite associahedra of J.-L.~Loday~\cite{LawReading}.
+  In particular, it is the Minkowski sum of two opposite Loday associahedra.
 \end{lemma}
+
+\vincent{Say also the minkowski sum of all faces of the standard simplex corresponding to intervals, and all their opposite.}
+\vincent{Say that this was already known by Law and Reading.}
 
@@
   \[
-  f_{\weakeq}(S) \eqdef \#\bigset{ I\text{ interval of } [n] }{ I \not\subseteq S \text{ and } I \not\subseteq [n]\ssm S }.
+  f(S) \eqdef \#\bigset{ I\text{ interval of } [n] }{ I\cap S\not=\emptyset \text{ and } I\cap ([n]\ssm S) \not= \emptyset }.
   \]
@@
 
-%The proof relies on the following simple lemma stating that the submodular function of a Minkowski sum of deformed permutahedra is the sum of the submodular function of the summands.
-%
-%\begin{lemma}
-%  \label{lem:submodsum}
-%  Consider a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by the submodular functions $f_1,f_2,\ldots ,f_k$.
-Then the Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f \eqdef \sum_i f_i$.
-%\end{lemma}
-
 \begin{proof}
-  %From \cref{lem:weakMinkowski,lem:submodsum},
-  The submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the up and down shard polytope in \cref{lem:lodaysp,lem:antilodaysp}.
-The shard polytope~$\SP(\upArc{I})$ of an up arc~$\upArc{I}$ is given by the submodular function
+  From \cref{lem:weakMinkowski,lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the up and down shard polytope in \cref{lem:lodaysp,lem:antilodaysp}.
+For a up arc $A$ defined by an interval $I$ of $[n]$, we have that its shard polytope $\SP(A)$ is realized by the submodular function
   \[
-  f_{\upArc{I}}(S) \eqdef \llbracket S\cap I\not=\emptyset \rrbracket .
+  \loday{f}_I(S) \eqdef \llbracket S\cap I\not=\emptyset \rrbracket .
   \]
   (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and 0 otherwise.)
-  Similarly, the shard polytope~$\SP(\downArc{I})$ of a down arc~$\downArc{I}$ is given by the submodular function
+  Similarly, for a down arc $A$ defined by an interval $I$, the shard polytope $\SP(A)$ is defined by the following submodular function:
   \[
-  f_{\downArc{I}}(S) \eqdef - \llbracket I\subseteq S \rrbracket .
+  \antiloday{f}_I(S) \eqdef - \llbracket I\subseteq S \rrbracket .
   \]
@@
   \[
-  f_{\weakeq}(S) = \sum_{I\text{ interval of }[n]} f_{\upArc{I}}(S) + f_{\downArc{I}}(S)
+  f(S) \eqdef \sum_{I\text{ interval of }[n]} \loday{f}_I(S) + \antiloday{f}_I(S)
   \]
-  is therefore exactly the number of intervals of $I$ that intersect $S$ but are not contained in $S$, hence the number of intervals contained neither in~$S$ nor in its complement.
+  is therefore exactly the number of intervals of $I$ that intersect $S$ but are not contained in $S$, hence the number of intervals intersecting both $S$ and its complement.
 \end{proof}
@@
 
-For every weak rectangulation $R$, we wish to give the coordinates of a point $\b{p}(R)\in\R^n$ such that $\WRP(n)$ is the convex hull of the points~$\b{p}(R)$ for all weak rectangulations~$R$.
+For every rectangulation $R$, we wish to give the coordinates of a point $p(R)\in\R^n$ such that $\WRP(n)$ is the convex hull of the points $p(R)$ for all rectangulations $R$.
 Recall that in a Minkowski sum of polytopes, the vertex that is extremal in a generic direction is the sum of the vertices of the summands that are extremal in this direction.
@@
   \label{lem:lodaymax}
-  Let $\pi \in \f{S}_n$ and $I$ be a subset of~$[n]$.
-  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\pi^{-1}$ is $\b{e}_i$, where $i \eqdef \arg\max_{j\in I} \pi^{-1}(j)$ (resp.~where~$i \eqdef \arg\min_{j\in I} \pi^{-1}(j)$).
+  Let $\pi\in\f{S}_n$ and consider a up arc $A$ defined by an interval $I$.
+  The unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
+  is $\b{e}_i$, where $i \eqdef \arg\max_{j\in I} \pi^{-1}(j)$.
+  Similarly, if $A$ is a down arc defined by $I$, the unique vertex of $\SP(A)$ that is extreme with respect to the direction $\pi^{-1}$
+  is $-\b{e}_i$, where $i \eqdef \arg\min_{j\in I} \pi^{-1}(j)$.
 \end{lemma}
@@
 \begin{lemma}
-  Let $R$ be a rectangulation on~$n$ rectangles, with source and target trees~$S$ and~$T$, and let $\b{p}(R)$ the vertex of $\WRP(n)$ associated with $R$.
-  Given $i\in [n]$, the number of up arcs~$\upArc{I}$ such that $\b{e}_i$ is the extremal vertex of $\SP(\upArc{I})$ contributing to~$\b{p}(R)$ is
+  Let $R$ be a rectangulation on $n$ rectangles and $p(R)$ the vertex of $\WRP(n)$ associated with $R$.
+  Given $i\in [n]$, the number of up arcs $A$ such that $\b{e}_i$ is the extremal vertex of $\SP(A)$ contributing to $p(R)$ is
   \[
-  \loday{w}_i =  h^{T}_i\cdot v^{T}_i,
+  \loday{w}_i =  h(T(R), i)\cdot v(T(R),i),
   \]
-   where $h^T_i$ and $v^T_i$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
+   where $h(T,i)$ and $v(T,i)$ denote respectively the number of leaves in the horizontal and vertical subtrees of $i$ in the tree $T$.
 \end{lemma}
@@
 
-\begin{theorem}[\cite{Reading}]
+\begin{theorem}
   The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ on $n$ rectangles, 
@@
   \begin{itemize}
-  \item the congruence classes are one-to-one with strong rectangulations,  
+  \item The congruence classes are one-to-one with strong rectangulations,  
   \item the minimal elements of each congruence class are the 2-clumped permutations,
@@
 We call the congruence $\strongeq$ the \defn{strong rectangulation congruence}.
-The \defn{strong rectangulotope} $\SRP(n)$ is the quotientope of the strong rectangulation congruence~${\strongeq}$.
+The \defn{strong rectangulotope} $\SRP(n)$ is the quotientope of the strong rectangulation congruence $\strongeq$.
 
@@
 We classify the arcs into two families, depending on whether the arcs cross from above or below.
-For two consecutive intervals~$I$ and~$J$ of~$[n]$, the \defn{yin arc}~$\yinArc{I}{J}$ (resp.~\defn{yang arc}~$\yangArc{I}{J}$) is the arc starts at~$\min(I)$ and ends at~$\max(J)$, and goes above (resp.~below) the remaining elements of $I$ and below (resp.~above) the remaining elements of $J$, that is,
-\begin{align*}
-\yinArc{I}{J} & \eqdef \big( \min(I), \max(J), I \ssm \{\min(I)\}, J \ssm \{\max(J)\} \big) \\
-\qquad\text{and}\qquad
-\yangArc{I}{J} & \eqdef \big( \min(I), \max(J), J \ssm \{\max(J)\}, I \ssm \{\min(I)\} \big).
-\end{align*}
+A \defn{yin arc} is defined by two consecutive nonempty intervals $I$ and $J$ of $[n]$ such that the arc goes above all elements of $I$ and below all elements of $J$.
+A \defn{yang arc} is defined similarly, except that the arc goes below the elements of $I$ and above the elements of $J$.
 Note that all up arcs and down arcs are both yin arcs and yang arcs.
@@
   \label{lem:yinsp}
-  The yin shard polytope $\SP(\yinArc{I}{J})$ is $\conv \{\b{0}\} \cup \set{\b{e}_i - \b{e}_j}{i \in I, \; j \in J}$.
+  Given a yin arc $A=(I,J)$, the yin shard polytope $\SP(A)$ is defined as the convex hull of the points
+  $\{\b{0}\} \cup \set{\b{e}_i - \b{e}_j}{(i,j)\in I\times J}$.
 \end{lemma}
@@
   \label{lem:yangsp}
-  The yang shard polytope $\SP(\yangArc{I}{J})$ is a translate of~$\conv \{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{i \in I,\; j \in J}$.
+  Given a yang arc $A=(I,J)$, the yang shard polytope $\SP(A)$ is a translate of the convex hull of the points
+  $\{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{(i,j)\in I\times J}$.
 \end{lemma}
 
-%Note that yin shard polytopes are translates of the up shard polytopes when $|J|=1$, and of the down shard polytopes when $|I|=1$.
-%Similarly, up and down shard polytopes are translates of some yang shard polytopes.
+Note that yin shard polytopes are translates of the up shard polytopes when $|J|=1$, and of the down shard polytopes when $|I|=1$.
+Similarly, up and down shard polytopes are translates of some yang shard polytopes.
 
@@
   \label{lem:strongMinkowski}
-  The strong rectangulotope $\SRP(n)$ is realized by the Minkowski sum of all yin and yang shard polytopes.
+  The strong rectangulotope $\SRP(n)$ is realized by the Minkowski sum of all shard polytopes of yin and yang arcs.
 \end{lemma}
@@
 Since the combinatorial type of a Minkowski sum is invariant to translation of the summands, we can use the polytopes defined in \cref{lem:yinsp,lem:yangsp} as a definition of the yin and yang shard polytopes.
-Also note that since the up and down shard polytopes are, up to a translation, both yin and yang, each up and down shard polytope is summed twice in the proposed realization.
+Also note that since up and down shard polytopes are, up to a translation, both yin and yang, each up and down shard polytope is summed twice in the proposed realization.
 This, again, does not change the combinatorial type of the polytope.
 
-\cref{lem:strongMinkowski} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.
+\cref{lem:strongsum} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.
 
 \subsection{Submodular functions of strong rectangulotopes}
+
+\jean{That seems to work.}
 
@@
   \[
-  f(S) \eqdef \# \bigset{ I, J }{ I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S } +
-  \#\bigset{ I, J }{ I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S },
+  f(S) \eqdef \# \bigset{ I, J }{ I\cap S\not=\emptyset \wedge J\cap ([n]\ssm S) \not= \emptyset } +
+  \#\bigset{ I, J }{ J\cap S\not=\emptyset \wedge I\cap ([n]\ssm S) \not= \emptyset },
   \]
-  where $I, J$ are pairs of nonempty consecutive intervals of~$[n]$.
+  where $I, J$ are pairs of consecutive and nonempty intervals of $[n]$.
 \end{lemma}
@@
 \begin{proof}
-  The submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the yin and yang shard polytopes in \cref{lem:yinsp,lem:yangsp}.
-  The shard polytope~$\SP(\yinArc{I}{J})$ of a yin arc $\yinArc{I}{J}$ is given by the submodular function
+  From \cref{lem:strongMinkowski,lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the yin and yang shard polytope in \cref{lem:yinsp,lem:yangsp}.
+For a yin arc $A$ defined by consecutive intervals $I, J$ of $[n]$, we have that its shard polytope $\SP(A)$ is realized by the submodular function
   \[
-  f_{\yinArc{I}{J}}(S) \eqdef \llbracket I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S \rrbracket .
+  \yin{f}_{I,J}(S) \eqdef \llbracket S\cap I\not=\emptyset \wedge S\not\supseteq J \rrbracket .
   \]
-  (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and 0 otherwise.)
-  Similarly, the shard polytope~$\SP(\yangArc{I}{J})$ of a yang arc~$\yangArc{I}{J}$ is given by the submodular function
+  Similarly, for a yang arc, the shard polytope $\SP(A)$ is defined by the following submodular function:
   \[
-  f_{\yangArc{I}{J}}(S) \eqdef \llbracket I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S \rrbracket .
+  \yang{f}_{I,J}(S) \eqdef \llbracket S\cap J\not=\emptyset \wedge S\not\supseteq I \rrbracket .
   \]
@@
   \[
-  f_{\strongeq}(S) \eqdef \sum_{I,J} f_{\yinArc{I}{J}}(S) + f_{\yangArc{I}{J}}(S)
+  f(S) \eqdef \sum_{I,J} \yin{f}_{I,J}(S) + \yang{f}_{I,J}(S)
   \]

### 2024-04-04 draft/rectangulotopes.tex

--- 2024-04-04 draft/rectangulotopes.tex
@@
 For facets, we show that both weak and strong rectangulotopes have $2^n-2$ facets (the maximal possible number of facets of a deformed permutahedron), and provide simple formulas for the right hand side of the inequality corresponding to a given proper nonempty subset of~$[n]$.
+
 We now give the necessary definitions for precisely stating these vertex and facet descriptions of the weak and strong rectangulotopes.
-
-%%%%%%%%%%%%%
-
-\subsection{Weak and strong rectangulations}
-\label{subsec:weakStrongRectangulations}
-
-\begin{itemize}
-\item Terminology: rectangles, \defn{vertices}, \defn{segments}, \defn{generic} rectangulation
-\item \defn{Wall slides}
-\item Two equivalence classes: \defn{weak equivalence} preserving rectangle-segment adjacencies, and \defn{strong equivalence} preserving rectangle-rectangle adjacencies.
-\item Weak equivalence classes have a \defn{diagonal} representative.
-\item \defn{Flip graphs} on rectangulations
-\end{itemize}
-
-%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%
-
 \subsection{Up and down arcs}
@@
 
-%%%%%%%%%%%%%
-
 \subsection{Shard polytopes of up and down arcs}
@@
   \label{lem:antilodaysp}
-%  Given a down arc~$\alpha$ defined by an interval $I$, the down shard polytope $\SP(\alpha)$ is a translate of the negative standard simplex~$\conv \set{ - \b{e}_i }{ i\in I }$.
   The down shard polytope $\SP(\downArc{I})$ is a translate of the negative simplex~$\conv \set{ - \b{e}_i }{ i\in I }$.
@@
 
-%The proof relies on the following simple lemma stating that the submodular function of a Minkowski sum of deformed permutahedra is the sum of the submodular function of the summands.
-%
-%\begin{lemma}
-%  \label{lem:submodsum}
-%  Consider a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by the submodular functions $f_1,f_2,\ldots ,f_k$.
-Then the Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f \eqdef \sum_i f_i$.
-%\end{lemma}
-
 \begin{proof}
-  %From \cref{lem:weakMinkowski,lem:submodsum},
   The submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the up and down shard polytope in \cref{lem:lodaysp,lem:antilodaysp}.

### 2024-04-04 draft/rectangulotopes.tex
first pictures

--- 2024-04-04 draft/rectangulotopes.tex
@@
 \usepackage{ulem}\normalem
+\usepackage{picins/picins}
 
@@
 On the other hand, the \defn{strong rectangulation congruence}, whose classes are in bijection with \defn{strong rectangulations}~\cite{MR2864445}.
-The latter is mentioned by N.~Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
+The latter is mentioned by N.~Reading~\cite{Reading-arcDiagrams} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
 The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
@@
 
+\begin{figure}
+	\capstart
+	\centerline{\includegraphics[scale=.6]{weakOrderLeft4} \; \includegraphics[scale=.6]{weakOrderCongruence4} \; \includegraphics[scale=.48]{TamariLattice4}}
+	\caption{The weak Bruhat order on~$\f{S}_4$ (left), the sylvester congruence~$\equiv_\textrm{sylv}$~(middle), and the Tamari lattice (right). \cite[Fig.~1 \& 2]{PilaudSantos-quotientopes}}
+	\label{fig:sylvesterCongruence}
+\end{figure}
+
 %%%%%%%%%%%%%
@@
 In this paper, we prefer to use weak Bruhat order to avoid any confusion with the weak poset defined on weak rectangulations.} is the lattice on the permutations of~$\f{S}_n$ defined by the inclusion of their inversion sets.
+See \cref{fig:sylvesterCongruence}\,(left).
 Note that the cover relations in the weak Bruhat order are given by the transpositions of two adjacent letters.
@@
 
+\parpic(4cm,2.5cm)(10pt, 150pt)[r][b]{\includegraphics[scale=.9]{arcDiagram}}{
 In~\cite{Reading-arcDiagrams}, N.~Reading defined an elegant bijection between permutations of~$[n]$ and noncrossing arc diagrams on~$[n]$.
-It sends a permutation~$\sigma$ to the noncrossing arc diagram with an arc~$(\sigma_{j+1}, \sigma_j, A_j, B_j)$ for each descent~$j \in [n-1]$ of~$\sigma$ (\ie with~$\sigma_j > \sigma_{j+1}$), where \linebreak $A_j \eqdef \set{\sigma_i}{1 \le i < j \text{ and } \sigma_{j+1} < \sigma_i < \sigma_j}$ and~$B_j \eqdef \set{\sigma_k}{j+1 < k \le n \text{ and } \sigma_{j+1} < \sigma_k < \sigma_j}$.
-\linebreak
-This can also been visualized by representing the table~$(\sigma_j,j)$ of the permutation~$\sigma$, drawing the segments corresponding to the descents of~$\sigma$, and letting all points fall on the horizontal axis, allowing the segments to bend but not to cross each other nor to pass through a point.
+It sends a permutation~$\sigma$ to the noncrossing arc diagram with an arc~$(\sigma_{j+1}, \sigma_j, A_j, B_j)$ for each descent~$j \in [n-1]$ of~$\sigma$ (\ie with~$\sigma_j > \sigma_{j+1}$), where
+
+\vspace{-.2cm}
+\begin{minipage}{10cm}
+\begin{align*}
+A_j & \eqdef \set{\sigma_i}{1 \le i < j \text{ and } \sigma_{j+1} < \sigma_i < \sigma_j} \\
+\text{and} \qquad
+B_j & \eqdef \set{\sigma_k}{j+1 < k \le n \text{ and } \sigma_{j+1} < \sigma_k < \sigma_j}.
+\end{align*}
+\end{minipage}
+
+\vspace{.2cm}
+\noindent
+As illustrated on the right, this can also been visualized by representing the table~$(\sigma_j,j)$ of the permutation~$\sigma$, drawing the segments corresponding to the descents of~$\sigma$, and letting all points fall on the horizontal axis, allowing the segments to bend but not to cross each other nor to pass through a point.
 The single arcs correspond to permutations with a single descent, that is, to join irreducible permutations of the weak Bruhat order.
@@
 See~\cite{Reading-arcDiagrams} for details.
+}
 
@@
 The \defn{lattice quotient}~$\f{S}_n/{\equiv}$ is the lattice on the congruence classes of~$\equiv$ where~$X \le Y$ if and only if there exist~$x \in X$ and~$y \in Y$ such that~$x \le y$, and~$X \meet Y$ (resp.~$X \join Y$) is the congruence class of~$x \meet y$ (resp.~$x \join y$) for any~$x \in X$~and~$y \in Y$.
+
+For instance, the \defn{sylvester congruence} is ... \vincent{todo}
+See \cref{fig:sylvesterCongruence}\,(middle).
+The quotient of the weak Bruhat order by the sylvester congruence is the classical \defn{Tamari lattice}.
+See \cref{fig:sylvesterCongruence}\,(right).
 
@@
 \label{subsec:quotientFans}
+
+\begin{figure}
+	\capstart
+	\centerline{\includegraphics[scale=.6]{braidFanLeft4} \qquad \raisebox{-.3cm}{\includegraphics[scale=.6]{sylvesterFan4}}}
+	\caption{The braid fan (left) and the sylvester fan (right). \cite[Fig.~1]{PilaudSantos-quotientopes}}
+	\label{fig:fans}
+\end{figure}
 
@@
 \label{subsec:quotientopes}
+
+\begin{figure}
+	\capstart
+	\centerline{\includegraphics[scale=.6]{permutahedronLeft4} \qquad \raisebox{-.4cm}{\includegraphics[scale=.6]{associahedron4}}}
+	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$. \cite[Fig.~1]{PilaudSantos-quotientopes}}
+	\label{fig:quotientopes}
+\end{figure}
 

### 2024-04-04 draft/rectangulotopes.tex

--- 2024-04-04 draft/rectangulotopes.tex
@@
 The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
+%The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
 On the other hand, the \defn{strong rectangulation congruence}, whose classes are in bijection with strong rectangulations~\cite{MR2864445}.
-The latter is mentioned by N.~Reading~\cite{MR3335492} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
+The latter is mentioned by N.~Reading~\cite{Reading-arcDiagrams} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
 The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.

### 2024-04-04 draft/rectangulotopes.tex
minor changes

@@
 Weak and strong rectangulations of size $n$ have been shown to define congruences of the weak Bruhat order on $\f{S}_n$.
-The \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762}.
-Its classes are one-to-one with weak rectangulations, but also with \defn{twin binary trees}~\cite{Giraudo}, and with \defn{Baxter permutations}~\cite{}. % more baxter refs
+
+The \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762} and revisited in~\cite[Thm.~1.1 \& Exm.~4.10]{Reading-arcDiagrams}.
+Its classes are in bijection with with weak rectangulations, but also with \defn{twin binary trees}~\cite{Giraudo}, and with \defn{Baxter permutations}~\cite{}. % more baxter refs
 The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
 %The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
-On the other hand, the \defn{strong rectangulation congruence}, whose classes are in bijection with strong rectangulations~\cite{MR2864445}.
-The latter is mentioned by N.~Reading~\cite{Reading-arcDiagrams} (see for instance in Theorem 1.2, and Example 4.11), and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
+
+The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams}, and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
+Its classes are in bijection with strong rectangulations.
 The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
@@
 \begin{lemma}
-  The skeleton of $\WRP(n)$ is isomorphic to the flip graph on weak rectangulations of size $n$.
-  Similarly, the skeleton of $\SRP(n)$ is isomorphic to the flip graph on strong rectangulations of size $n$.
+%  The skeleton of $\WRP(n)$ is isomorphic to the flip graph on weak rectangulations of size~$n$.
+%  Similarly, the skeleton of $\SRP(n)$ is isomorphic to the flip graph on strong rectangulations of size~$n$.
+  The skeleton of $\WRP(n)$ (resp.~of $\SRP(n)$) is isomorphic to the flip graph on weak (resp.~strong) rectangulations of size~$n$.
 \end{lemma}
@@
 
-Fix a rectangulation~$R$ with $n$ rectangles, and consider the directed graph~$D(R)$ whose vertex set is the set of all vertices of~$R$, and whose edges are obtained as follows.
+Fix a rectangulation~$R$ of size~$n$, and consider the directed graph~$D(R)$ whose vertex set is the set of all vertices of~$R$, and whose edges are obtained as follows.
 For each rectangle~$r$ of~$R$, we include in~$D(R)$ four edges joining the vertices of~$r$, where the horizontal edges are oriented from left to right, and the vertical edges are oriented from bottom to top.
@@
   \item the convex hull of the points
-  $\sum_{i\in [n]} (\loday{w}^R_i - \antiloday{w}^R_i)\cdot \b{e}_i$ for all weak rectangulations $R$ on $n$ rectangles,
+  $\sum_{i\in [n]} (\loday{w}^R_i - \antiloday{w}^R_i)\cdot \b{e}_i$ for all weak rectangulations $R$ of size~$n$,
   with
@@
   \]
-   for all strong rectangulations $R$ on $n$ rectangles, with
+   for all strong rectangulations $R$ of size~$n$, with
   \[
@@
 \begin{theorem}[\cite{LawReading}]
-  The subsets $\mathcal{L}(P_w(R))$ of $\f{S}_n$, for all rectangulations $R$ on $n$ rectangles, 
+  The subsets $\mathcal{L}(P_w(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, 
   are equivalence classes of a congruence relation $\weakeq$ on the weak Bruhat order on $\f{S}_n$.
@@
 \begin{lemma}
-  Let $R$ be a rectangulation on~$n$ rectangles, with source and target trees~$S$ and~$T$, and let $\b{p}(R)$ the vertex of $\WRP(n)$ associated with $R$.
+  Let $R$ be a rectangulation of size~$n$, with source and target trees~$S$ and~$T$, and let $\b{p}(R)$ the vertex of $\WRP(n)$ associated with $R$.
   Given $i\in [n]$, the number of up arcs~$\upArc{I}$ such that $\b{e}_i$ is the extremal vertex of $\SP(\upArc{I})$ contributing to~$\b{p}(R)$ is
@@
 \begin{theorem}[\cite{Reading}]
-  The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ on $n$ rectangles, 
+  The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, 
   are equivalence classes of a congruence relation $\strongeq$ on the weak Bruhat order on $\f{S}_n$.
@@
 \begin{lemma}
-  Let $R$ be a strong rectangulation on $n$ rectangles, $S$ and $T$ be its source and target trees, $([n],\prec_s)$ be its strong order, and $\b{p}(R)$ be the vertex of $\SRP(n)$ corresponding to $R$.
+  Let $R$ be a strong rectangulation of size~$n$, let~$S$ and $T$ be its source and target trees, let~$([n],\prec_s)$ be its strong order, and let~$\b{p}(R)$ be the vertex of $\SRP(n)$ corresponding to $R$.
   Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $A$ such that $\b{e}_i-\b{e}_j$ is the extremal vertex of $\SP(A)$ contributing to $p(R)$ is
@@
 \begin{lemma}
-  Let $R$ be a rectangulation on $n$ rectangles, $S$ and $T$ be its source and target trees, and $([n],\prec_s)$ be the associated strong order.
+  Let $R$ be a rectangulation of size~$n$, let $S$ and $T$ be its source and target trees, and let~$([n],\prec_s)$ be the associated strong order.
   Given a pair $i,j\in [n]$ such that $i<j$ and $i\prec_s j$, we have

### 2024-04-05 draft/rectangulotopes.tex
more pictures Section 2

@@
 It is an $(n-1)$-dimensional polytope, whose facets are indexed by the proper nonempty subsets of $[n]$, and whose skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
+See \cref{fig:quotientopes}\,(left) for an illustration when~$n = 4$.
 The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{generalized permutahedra}~\cite{Postnikov} (that we prefer to call \defn{deformed permutahedra}).
@@
 We note that the inequality description of the same polytope was actually provided in 1993 by S.~Shnider and S.~Sternberg~\cite{ShniderSternberg}, but Loday's vertex description largely popularized this realization.
+See \cref{fig:quotientopes}\,(right) for an illustration when~$n = 4$.
 
@@
 
-A natural object associated with the associahedron is the \defn{Tamari lattice}, the cover graph of which is isomorphic to the skeleton of the associahedron.
-The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} relation called the \defn{sylvester congruence}.
+A natural object associated with the associahedron is the \defn{Tamari lattice}, whose cover graph is isomorphic to the skeleton of the associahedron.
+The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} called the \defn{sylvester congruence}.
 Answering a question of N.~Reading~\cite{MR2142177}, V.~Pilaud and F.~Santos~\cite{MR3964495} proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more precisely whose normal fan is the \defn{quotient fan} defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
@@
 
-The goal of this paper is to describe the quotientopes of congruences of the weak Bruhat order whose classes are in bijection with equivalence classes of rectangulations, defined as decompositions of a square into rectangles.
+The goal of this paper is to describe the geometry of the quotientopes for two particular congruences of the weak Bruhat order whose classes are in bijection with equivalence classes of rectangulations, defined as decompositions of a square into rectangles.
 
@@
 \begin{figure}
-  \caption{\label{fig:rectequiv}Weak and strong equivalence between rectangulations.}
+	\caption{Weak and strong equivalence between rectangulations.}
+	\label{fig:rectequiv}
 \end{figure}
@@
 
-The \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762} and revisited in~\cite[Thm.~1.1 \& Exm.~4.10]{Reading-arcDiagrams}.
-Its classes are in bijection with with weak rectangulations, but also with \defn{twin binary trees}~\cite{Giraudo}, and with \defn{Baxter permutations}~\cite{}. % more baxter refs
-The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
+The \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762} and revisited in~\cite[Thm.~1.1 \& Exm.~4.10]{Reading-arcDiagrams} in terms of arc diagrams.
+Its classes are in bijection with weak rectangulations, but also with \defn{twin binary trees}~\cite{Giraudo}, and with \defn{Baxter permutations}~\cite{}. % more baxter refs
+The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by~$\WRP(n)$.
 %The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
 
-The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams}, and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
+The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams} in terms of arc diagrams, and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
 Its classes are in bijection with strong rectangulations.
-The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
+The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by~$\SRP(n)$.
 
@@
 
+\begin{figure}
+	\caption{Three types of flips between rectangulations.}
+	\label{fig:flip}
+\end{figure}
+
 \begin{lemma}
@@
   The skeleton of $\WRP(n)$ (resp.~of $\SRP(n)$) is isomorphic to the flip graph on weak (resp.~strong) rectangulations of size~$n$.
+  \vincent{I would leave that in the text, not in a proper thm environment...}
 \end{lemma}
@@
 
+\vspace{-.1cm}
 \parpic(4cm,2.5cm)(10pt, 150pt)[r][b]{\includegraphics[scale=.9]{arcDiagram}}{
@@
 
-For instance, the \defn{sylvester congruence} is ... \vincent{todo}
+For instance, the sets of linear extensions of binary trees (labeled in inorder and oriented toward their roots) are the classes of the \defn{sylvester congruence}~\cite{HivertNovelliThibon-binarySearchTrees}.
 See \cref{fig:sylvesterCongruence}\,(middle).
@@
 %Moreover, the sets of arcs of the form~$\c{A}_\equiv$ are easily described.
-An arc~$(a, b, A, B)$ is a \defn{subarc} of an arc~$(a', b', A', B')$ if~$a' \le a < b \le b$ and~$A \subseteq A'$ while~$B \subseteq B'$.
+\vspace{-.15cm}
+\parpic(5cm,2.8cm)(10pt, 90pt)[r][b]{\includegraphics[scale=.5]{subarcOrder}}{
+An arc~$(a, b, A, B)$ is a \defn{subarc} of an arc~$(a', b', A', B')$ if \linebreak $a' \le a < b \le b$ and~$A \subseteq A'$ while~$B \subseteq B'$.
 An \defn{arc ideal} is a subset of arcs closed by subarcs.
 The map~${\equiv} \mapsto \c{A}_\equiv$ is a bijection between the lattice congruences of the weak Bruhat order and the arc ideals.
-In other words, the lattice of congruences of the weak Bruhat order is distributive, and its poset of join irreducibles is isomorphic to the subarc order.
+In other words, the lattice of congruences of the weak Bruhat order is distributive, and its poset of join irreducibles is isomorphic to the subarc order, illustrated on the right for~$n = 4$.
+}
 
@@
 
-\begin{figure}
+\begin{figure}[b]
 	\capstart
 	\centerline{\includegraphics[scale=.6]{braidFanLeft4} \qquad \raisebox{-.3cm}{\includegraphics[scale=.6]{sylvesterFan4}}}
-	\caption{The braid fan (left) and the sylvester fan (right). \cite[Fig.~1]{PilaudSantos-quotientopes}}
+	\caption{The braid fan (left) and the sylvester fan (right). \cite[Fig.~1]{PilaudSantos-quotientopes} \&~\cite[Fig.~5]{PadrolPilaudRitter}}
 	\label{fig:fans}
@@
 It has a chamber for each permutation~$\sigma$ of~$[n]$, given by the set of vectors whose coordinates are ordered as~$\sigma$.
+See \cref{fig:fans}\,(left).
 
@@
 %$\Sigma(a, b, A, B) \eqdef \set{\b{x} \in \R^n}{x_a = x_b \text{ and } x_{a'} < x_a \text{ for all } a' \in A \text{ and }  x_b < x_{b'} \text{ for all } b' \in B}$.
-
-The \defn{quotient fan} of a lattice congruence~$\equiv$ of the weak Bruhat order is the polyhedral fan~$\c{F}_\equiv$ where
+In other words, the shards decompose each hyperplane~$x_a = x_b$ of the braid arrangement into~$2^{b-a-1}$ cones.
+
+The \defn{quotient fan} of a congruence~$\equiv$ of the weak Bruhat order is the polyhedral fan~$\c{F}_\equiv$~where
 \begin{itemize}
@@
 By construction, the braid fan refines the quotient fan~$\c{F}_\equiv$, and the dual graph of the quotient fan~$\c{F}_\equiv$ is isomorphic to the cover graph of the quotient~$\f{S}_n/{\equiv}$.
+For instance, \cref{fig:fans}\,(right) represents the \defn{sylvester fan} (the quotient fan of the sylvester congruence of \cref{fig:sylvesterCongruence}\,(middle)) for~$n = 4$, whose dual graph is the cover graph of the Tamari lattice of \cref{fig:sylvesterCongruence}\,(right).
 
@@
 	\centerline{\includegraphics[scale=.6]{permutahedronLeft4} \qquad \raisebox{-.4cm}{\includegraphics[scale=.6]{associahedron4}}}
-	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$. \cite[Fig.~1]{PilaudSantos-quotientopes}}
+	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$. \cite[Fig.~1]{PilaudSantos-quotientopes} \&~\cite[Fig.~5]{PadrolPilaudRitter}}
 	\label{fig:quotientopes}
@@
 In particular, the skeleton of a quotientope is isomorphic to the cover graph of the quotient~$\f{S}_n/{\equiv}$.
+For instance, the associahedron~$\Asso(n)$ (\cref{fig:quotientopes}\,(right) when~$n = 4$)  is a quotientope for the sylvester congruence (\cref{fig:sylvesterCongruence}\,(middle) when~$n = 4$), its normal fan is the sylvester fan (\cref{fig:fans}\,(right) when~$n = 4$), and its skeleton is the cover graph of the Tamari lattice (\cref{fig:sylvesterCongruence}\,(right) when~$n = 4$).
 The existence of quotientopes was first proved in \cite{PilaudSantos} and later better understood in~\cite{PadrolPilaudRitter} using Minkowski sums of shard polytopes.
@@
 In particular, if~$\b{v}$ is a generic direction, the vertex of~$\polytope{P} + \polytope{Q}$ extremal in direction~$\b{v}$ is just the sum of the vertices of~$\polytope{P}$ and~$\polytope{Q}$ that are extremal in direction~$\b{v}$.
+Moreover, the normal fan of the Minkowski sum~$\polytope{P} + \polytope{Q}$ is the common refinement of the normal fans of~$\polytope{P}$ and~$\polytope{Q}$.
 
@@
 
+\begin{figure}
+	\capstart
+	\centerline{\includegraphics[scale=1]{shardPolytopes}}
+	\caption{The shard polytopes of the up arc~$(1, 4, \{2, 3\}, \varnothing)$, down arc~$(1, 4, \varnothing, \{2, 3\})$, yin arc~$(1, 4, \{2\}, \{3\})$, and yang arc~$(1, 4, \{3\}, \{2\})$.
+The vertices of the shard polytope of an arc~$\alpha \eqdef (a, b, A, B)$ are labeled by the corresponding $\alpha$-alternating matchings, where we use solid dots~$\bullet$ for elements in~$\{a\} \cup A$ and hollow dots~$\circ$ for elements in~$B \cup \{b\}$.
+The corresponding vertex coordinates are directly read replacing~$\bullet$ by~$1$ and~$\circ$ by~$-1$.
+For instance, the vertex labeled~${\bullet \cdot \cdot \,\circ}$ has coordinates~$(1,0,0,-1)$.
+Adapted from~\cite[Fig.~10]{PadrolPilaudRitter}.}
+	\label{fig:shardPolytiopes}
+\end{figure}
+
 By definition, any quotientope is a deformed permutahedron, and can thus be written as
@@
   \label{lem:submodsum}
+  \vincent{I would not make this one so visible in a proper thm environment.}
   Consider a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by the submodular functions $f_1,f_2,\ldots ,f_k$.
@@
 The description of the weak rectangulotope~$\WRP(n)$ as a Minkowski sum in \cref{lem:weakMinkowski} allows us to easily compute the corresponding submodular function.
+\vincent{Say here that all rays are preserved}
 

### 2024-04-05 draft/rectangulotopes.tex
filled in references

--- 2024-04-04 draft/rectangulotopes.tex
@@
 It is an $(n-1)$-dimensional polytope, whose facets are indexed by the proper nonempty subsets of $[n]$, and whose skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
-The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{generalized permutahedra}~\cite{Postnikov} (that we prefer to call \defn{deformed permutahedra}).
-
-Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined combinatorially by D.~Tamari~\cite{T51} and by J.~Stasheff in a topological context~\cite{S63}.
-It has since then found applications in various areas of mathematics, including topology, operads, cluster algebras, combinatorial Hopf algebras, and physics~\cite{PilaudSantosZiegler}.
-It was first defined as a combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894}.
+The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra}~\cite{MR2520477,Postnikov,MR4064768,MR4651496} (also call \defn{generalized permutahedra}, or \defn{polymatroids}).
+
+Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined by D.~Tamari~\cite{T51} and by J.~Stasheff~\cite{S63} in a topological context.
+It has since then been identified as a fundamental object in many other areas of mathematics, including operads, cluster algebras, combinatorial Hopf algebras, and physics (see for instance the recent survey from V.
+Pilaud, F.
+Santos and G.
+Ziegler~\cite{PilaudSantosZiegler} and references therein).
+It was first thought of as a purely combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894,MR2321739}.
 The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
@@
 
-A natural object associated with the associahedron is the \defn{Tamari lattice}, the cover graph of which is isomorphic to the skeleton of the associahedron.
-The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} relation called the \defn{sylvester congruence}.
+A natural object associated with the associahedron is the \defn{Tamari lattice}, the cover graph of which is isomorphic to the skeleton of the associahedron~\cite{MR0146227,MR3235205}.
+The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} relation called the \defn{sylvester congruence}~\cite{MR1654173,MR2142078}.
 Answering a question of N.~Reading~\cite{MR2142177}, V.~Pilaud and F.~Santos~\cite{MR3964495} proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more precisely whose normal fan is the \defn{quotient fan} defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
@@
 
-The combinatorics of rectangulations has been studied for decades~\cite{MR2233287,MR2871762,MR2864445,MR2763051,MR3084577,MR3878132,MR4598046}.
-Interestingly, many combinatorial results were initially motivated by applications to the design of \defn{floorplans} for very large scale integrated circuits, and published in electrical engineering journals~\cite{}.
-The recent order-theoretic take on these structures has proved fruitful.
+The combinatorics of rectangulations has been studied for nearly two decades~\cite{MR2233287,MR2763051,MR2871762,MR2864445,MR3084577,MR3192492,MR3878132,MR4598046}.
+Interestingly, several combinatorial results were initially motivated by applications to the design of \defn{floorplans} for very large scale integrated circuits, and published in electrical engineering journals, see for instance R.
+Fujimaki, Y.
+Inoue, and T.
+Takahashi~\cite{FT07,TF08,ITF09,FIT09}.
 
@@
 We suppose throughout that no four rectangles have a common vertex, hence that the rectangulations are \defn{generic} in that sense.
+
 In order to focus on the combinatorial structure of a rectangulation, we introduce two equivalence relations between rectangulations.
-First, we say that a rectangle $r$ is \defn{above} another rectangle $s$ (and $s$ is \defn{below} $r$) if either the bottom edge of $r$ lies in the same segment as the top edge of $s$, or if $r$ is above another rectangle $t$ that is above $s$.
+We say that a rectangle $r$ is \defn{above} another rectangle $s$ (and $s$ is \defn{below} $r$) if either the bottom edge of $r$ lies in the same segment as the top edge of $s$, or if $r$ is above another rectangle $t$ that is above $s$.
 Similarly, a rectangle $r$ is \defn{on the left of} another rectangle $s$ (and $s$ is \defn{on the right of} $r$) if either the right edge of $r$ lies in the same segment as the left edge of $s$, or if $r$ is on the left of another rectangle $t$ that is on the left of $s$.
-
 Two rectangulations are said to be \defn{weakly equivalent} if there exists a bijection between their rectangles that preserves the above-below and left-right relations between the rectangles.
@@
 This is illustrated in Figure~\ref{fig:rectequiv}.
-A \defn{diagonal rectangulation} is a rectangulation in which every rectangle intersects the top-left to bottom-right diagonal of the square.
+A \defn{diagonal rectangulation} is a rectangulation in which every rectangle intersects the top-left to bottom-right diagonal of the square, see Figure~\ref{fig:weakRectangulation}.
 It is simple to check that every weak rectangulation has a diagonal representative, hence weak rectangulations can also be thought of as diagonal rectangulations.
@@
 The \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762} and revisited in~\cite[Thm.~1.1 \& Exm.~4.10]{Reading-arcDiagrams}.
-Its classes are in bijection with with weak rectangulations, but also with \defn{twin binary trees}~\cite{Giraudo}, and with \defn{Baxter permutations}~\cite{}. % more baxter refs
+Its classes are in bijection with with weak rectangulations, but also with \defn{Baxter permutations}~\cite{MR0491652,MR0555815}, \defn{twin binary trees}~\cite{MR1417289,MR2914637}, and many other combinatorial families~\cite{MR2763051}.
 The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
-%The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
-
-The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams}, and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
+
+The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams}, and studied more recently by E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
 Its classes are in bijection with strong rectangulations.
@@
 The various types of flips are illustrated in Figure~\ref{fig:flips}.
-We refer to S.~Law and N.~Reading~\cite{MR2871762}, J.~Cardinal, V.~Sacrist\'an, and R.~Silveira~\cite{}, E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and \'E.~Fusy~\cite{ACFF24} for detailed descriptions of the flip graphs on both weak and strong rectangulations.
+We refer to S.~Law and N.~Reading~\cite{MR2871762}, J.~Cardinal, V.~Sacrist\'an, and R.~Silveira~\cite{MR3878132}, E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and \'E.~Fusy~\cite{ACFF24} for detailed descriptions of the flip graphs on both weak and strong rectangulations.
 Just like associahedra encode the rotation graphs on binary trees, the rectangulotopes directly yield flip graphs on rectangulations.
@@
 \begin{lemma}
-%  The skeleton of $\WRP(n)$ is isomorphic to the flip graph on weak rectangulations of size~$n$.
-%  Similarly, the skeleton of $\SRP(n)$ is isomorphic to the flip graph on strong rectangulations of size~$n$.
   The skeleton of $\WRP(n)$ (resp.~of $\SRP(n)$) is isomorphic to the flip graph on weak (resp.~strong) rectangulations of size~$n$.
@@
 	\centerline{\includegraphics[width=.5\textwidth]{weakRectangulation} \qquad \includegraphics[width=.5\textwidth]{weakRectangulationTrees}}
-	\caption{A weak rectangulation (left) and its pair of separated twin trees (right). Example from \cite{ACFF24}.}
+	\caption{A diagonal rectangulation (left) and its pair of twin binary trees (right). Example from \cite{ACFF24}.}
 	% The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.
@@
 	\centerline{\includegraphics[width=.5\textwidth]{strongRectangulation} \qquad \includegraphics[width=.5\textwidth]{strongRectangulationTrees}}
-	\caption{A strong rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{ACFF24}.}
+	\caption{A (non-diagonal) rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{ACFF24}.}
         % The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
@@
 We complete each node of~$S(R)$ and~$T(R)$ with a vertical (resp.~horizontal) leaf if it has no vertical (resp.~horizontal) child, see \cref{fig:strongRectangulation}.
+If the rectangulation $R$ is diagonal, we retrieve the well-studied pair of twin binary trees~\cite{MR1417289,MR2914637}.
 Recall that the \defn{inorder labeling} of a binary tree~$T$ is the labeling of the nodes of~$T$ such that the label of each node~$t$ of~$T$ is larger than all labels in the left subtree of~$t$ and larger than all labels in the right subtree of~$t$.
@@
 
-This enables to unambiguously label the rectangles of~$R$ by the inorder 
+This enables to unambiguously label the rectangles of~$R$ by the inorder.
 The resulting labeling coincides with the NW--SE labeling of~\cite{ACFF24}.
@@
 This materializes the result of Law and Reading that the weak rectangulotopes are Minkowski sums of two opposite associahedra~\cite{MR2871762}.
+
+\Jean{We need to make it clear whether $\WRP (n)$ is the proposed realization or just the combinatorial type.
+I was thinking the latter.
+Then we can change the sentence ``The weak rectangulotope is realized by the polytope~$\WRP (n)$...'' by ``The weak rectangulotope $\WRP (n)$ is realized by:~''.
+Otherwise, refrain from using $\WRP (n)$ before here.}
 
@@
 We now briefly recall some results on the weak Bruhat order and its quotients, both from a lattice and geometric perspectives.
-We refer to \cite{Reading-arcDiagrams, Reading-Chapters, PadrolPilaudRitter} for details.
+We refer to \cite{Reading-arcDiagrams, MR3645055, MR3645056, MR4584712} for details.
+% J: Reading-Chapters: which one is this ?
 
@@
 	\centerline{\includegraphics[scale=.6]{weakOrderLeft4} \; \includegraphics[scale=.6]{weakOrderCongruence4} \; \includegraphics[scale=.48]{TamariLattice4}}
-	\caption{The weak Bruhat order on~$\f{S}_4$ (left), the sylvester congruence~$\equiv_\textrm{sylv}$~(middle), and the Tamari lattice (right). \cite[Fig.~1 \& 2]{PilaudSantos-quotientopes}}
+	\caption{The weak Bruhat order on~$\f{S}_4$ (left), the sylvester congruence~$\equiv_\textrm{sylv}$~(middle), and the Tamari lattice (right). \cite[Fig.~1 \& 2]{MR3964495}}
 	\label{fig:sylvesterCongruence}
@@
 	\centerline{\includegraphics[scale=.6]{braidFanLeft4} \qquad \raisebox{-.3cm}{\includegraphics[scale=.6]{sylvesterFan4}}}
-	\caption{The braid fan (left) and the sylvester fan (right). \cite[Fig.~1]{PilaudSantos-quotientopes}}
+	\caption{The braid fan (left) and the sylvester fan (right). \cite[Fig.~1]{MR3964495}}
 	\label{fig:fans}
@@
 	\centerline{\includegraphics[scale=.6]{permutahedronLeft4} \qquad \raisebox{-.4cm}{\includegraphics[scale=.6]{associahedron4}}}
-	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$. \cite[Fig.~1]{PilaudSantos-quotientopes}}
+	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$. \cite[Fig.~1]{MR3964495}}
 	\label{fig:quotientopes}
@@
 In particular, the skeleton of a quotientope is isomorphic to the cover graph of the quotient~$\f{S}_n/{\equiv}$.
-The existence of quotientopes was first proved in \cite{PilaudSantos} and later better understood in~\cite{PadrolPilaudRitter} using Minkowski sums of shard polytopes.
+The existence of quotientopes was first proved in \cite{MR3964495} and later better understood in~\cite{MR4584712} using Minkowski sums of shard polytopes.
 
@@
 The \defn{shard polytope} of~$\alpha$ is the convex hull~$\SP(\alpha)$ of the characteristic vectors of all $\alpha$-alternating matchings.
-It was shown in~\cite{PadrolPilaudRitter} that any Minkowski sum of positive scalings of the shard polytopes~$\SP(\alpha)$ for all arcs~$\alpha$ in the arc ideal~$\c{A}_\equiv$ is a quotientope for~$\equiv$.
+It was shown in~\cite{MR4584712} that any Minkowski sum of positive scalings of the shard polytopes~$\SP(\alpha)$ for all arcs~$\alpha$ in the arc ideal~$\c{A}_\equiv$ is a quotientope for~$\equiv$.
 
@@
 
-Given a rectangulation~$R$, the source and target trees~$S(R)$ and~$T(R)$ define two partial orders on $[n]$ (the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top).
-The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of the two partial orders defined by the intertwining binary trees~$S(R)$ and~$T(R)$~\cite{MR4014603}.
+xGiven a rectangulation~$R$, the source and target trees~$S(R)$ and~$T(R)$ define two partial orders on $[n]$ (the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top).
+The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of the two partial orders defined by the intertwining binary trees~$S(R)$ and~$T(R)$.
+These posets were characterized (and called Baxter posets) by E.~Meehan~\cite{MR4014603}.
 The weak poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
@@
 
-\begin{theorem}[\cite{LawReading}]
+\begin{theorem}[\cite{MR2871762}]
   The subsets $\mathcal{L}(P_w(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, 
@@
   The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all up and down shard polytopes.
-  In particular, it is the Minkowski sum of two opposite associahedra of J.-L.~Loday~\cite{LawReading}.
+  In particular, it is the Minkowski sum of two opposite associahedra of J.-L.~Loday~\cite{MR2871762}.
 \end{lemma}
@@
 
-\begin{theorem}[\cite{Reading}]
+\begin{theorem}[\cite{MR2864445,ACFF24}]
   The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, 

### 2024-04-05 draft/rectangulotopes.tex

@@
 It is an $(n-1)$-dimensional polytope, whose facets are indexed by the proper nonempty subsets of $[n]$, and whose skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
+See \cref{fig:quotientopes}\,(left) for an illustration when~$n = 4$.
 The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra}~\cite{MR2520477,Postnikov,MR4064768,MR4651496} (also call \defn{generalized permutahedra}, or \defn{polymatroids}).
@@
 Its classes are in bijection with strong rectangulations.
-The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by $\SRP(n)$.
+The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by~$\SRP(n)$.
 
@@
 
+\begin{figure}
+	\caption{Three types of flips between rectangulations.}
+	\label{fig:flip}
+\end{figure}
+
 \begin{lemma}
   The skeleton of $\WRP(n)$ (resp.~of $\SRP(n)$) is isomorphic to the flip graph on weak (resp.~strong) rectangulations of size~$n$.
+  \vincent{I would leave that in the text, not in a proper thm environment...}
 \end{lemma}
@@
 
-\begin{figure}
+\begin{figure}[b]
 	\capstart
 	\centerline{\includegraphics[scale=.6]{braidFanLeft4} \qquad \raisebox{-.3cm}{\includegraphics[scale=.6]{sylvesterFan4}}}
-	\caption{The braid fan (left) and the sylvester fan (right). \cite[Fig.~1]{MR3964495}}
+	\caption{The braid fan (left) and the sylvester fan (right). \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
 	\label{fig:fans}
@@
 	\centerline{\includegraphics[scale=.6]{permutahedronLeft4} \qquad \raisebox{-.4cm}{\includegraphics[scale=.6]{associahedron4}}}
-	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$. \cite[Fig.~1]{MR3964495}}
+	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$. \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
 	\label{fig:quotientopes}
@@
 In particular, the skeleton of a quotientope is isomorphic to the cover graph of the quotient~$\f{S}_n/{\equiv}$.
+For instance, the associahedron~$\Asso(n)$ (\cref{fig:quotientopes}\,(right) when~$n = 4$)  is a quotientope for the sylvester congruence (\cref{fig:sylvesterCongruence}\,(middle) when~$n = 4$), its normal fan is the sylvester fan (\cref{fig:fans}\,(right) when~$n = 4$), and its skeleton is the cover graph of the Tamari lattice (\cref{fig:sylvesterCongruence}\,(right) when~$n = 4$).
 The existence of quotientopes was first proved in \cite{MR3964495} and later better understood in~\cite{MR4584712} using Minkowski sums of shard polytopes.

### 2024-04-05 draft/rectangulotopes.tex
corrected one ref

@@
 For instance, the vertex labeled~${\bullet \cdot \cdot \,\circ}$ has coordinates~$(1,0,0,-1)$.
-Adapted from~\cite[Fig.~10]{PadrolPilaudRitter}.}
+Adapted from~\cite[Fig.~10]{MR4584712}.}
 	\label{fig:shardPolytiopes}
@@
 
-xGiven a rectangulation~$R$, the source and target trees~$S(R)$ and~$T(R)$ define two partial orders on $[n]$ (the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top).
+Given a rectangulation~$R$, the source and target trees~$S(R)$ and~$T(R)$ define two partial orders on $[n]$ (the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top).
 The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of the two partial orders defined by the intertwining binary trees~$S(R)$ and~$T(R)$.

### 2024-04-05 draft/rectangulotopes.tex
corrected one ref

@@
 The various types of flips are illustrated in Figure~\ref{fig:flips}.
-We refer to S.~Law and N.~Reading~\cite{MR2871762}, J.~Cardinal, V.~Sacrist\'an, and R.~Silveira~\cite{MR3878132}, E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and \'E.~Fusy~\cite{ACFF24} for detailed descriptions of the flip graphs on both weak and strong rectangulations.
+We refer to \cite{MR2871762,MR3878132,MR3697823,ACFF24} for detailed descriptions of the flip graphs on both weak and strong rectangulations.
 Just like associahedra encode the rotation graphs on binary trees, the rectangulotopes directly yield flip graphs on rectangulations.

### 2024-04-05 programmation/submodular.py
improved submodular.py

@@ -57,41 +57,41 @@ def powerset(iterable):
     return chain.from_iterable(combinations(s, r) for r in range(len(s)+1))
 
 # compute the polyhedron for a given submod function
+# f takes two arguments : subset S and size of ground set n
 def submod_polytope(f, n):
-    all_ineq = [ (f(S),) + tuple(-1 if x in S else 0 for x in range(n)) for S in powerset(range(n)) ]
-    the_eq = [ (f(range(n)),) + tuple(-1 for i in range(n)) ]
-    return Polyhedron(ieqs = all_ineq, eqns = the_eq)
+    all_ineq = [ (f(S, n),) + tuple(-1 if x in S else 0 for x in range(n)) for S in powerset(range(n)) ]
+    return Polyhedron(ieqs = all_ineq, eqns = [all_ineq[-1]]) # the last ineq is an eq
 
 ## RECTANGULOTOPES
 
 # associahedron as a Minkowski sum
 def assoc(n):
-    return sum([submod_polytope(lambda S : f_loday(S, range(i, j+1)), n) for i in range(n-1) for j in range(i+1, n)])
+    return sum([submod_polytope(lambda S, m : f_loday(S, range(i, j+1)), n) for i in range(n-1) for j in range(i+1, n)])
 
 # weak rectangulation polytope (diagonal / Baxter)
 # either sum all lodays and antilodays, or the cubocs
 def wrp(n):
-    return sum([submod_polytope(lambda S : f_cuboc(S, I), n) for i in range(n-1) for j in range(i+1, n)])
+    return sum([submod_polytope(lambda S, m : f_cuboc(S, range(i, j+1)), n) for i in range(n-1) for j in range(i+1, n)])
 
 # weak rectangulotope via the summed submodular function f_weakrect 
 def wrp2(n):
-    return submod_polytope(lambda S : f_weakrect(S, n), n)
+    return submod_polytope(f_weakrect, n)
 
 # yin rectangulotopes via yin shards
 def yrp(n):
-    return sum([submod_polytope(lambda S : f_yin(S, range(i, j+1), range(j+1, k+1)), n) for i in range(n-1) for j in range(i, n-1) for k in range(j+1, n)])
+    return sum([submod_polytope(lambda S, m : f_yin(S, range(i, j+1), range(j+1, k+1)), n) for i in range(n-1) for j in range(i, n-1) for k in range(j+1, n)])
 
[…]

### 2024-04-05 draft/rectangulotopes.tex
pictures rectangulotopes

@@
 See \cref{fig:quotientopes}\,(left) for an illustration when~$n = 4$.
-The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra}~\cite{MR2520477,Postnikov,MR4064768,MR4651496} (also call \defn{generalized permutahedra}, or \defn{polymatroids}).
-
-Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined by D.~Tamari~\cite{T51} and by J.~Stasheff~\cite{S63} in a topological context.
-It has since then been identified as a fundamental object in many other areas of mathematics, including operads, cluster algebras, combinatorial Hopf algebras, and physics (see for instance the recent survey from V.
-Pilaud, F.
-Santos and G.
-Ziegler~\cite{PilaudSantosZiegler} and references therein).
-It was first thought of as a purely combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894,MR2321739}.
+The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{generalized permutahedra}~\cite{Postnikov} (that we prefer to call \defn{deformed permutahedra}).
+
+Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined combinatorially by D.~Tamari~\cite{T51} and by J.~Stasheff in a topological context~\cite{S63}.
+It has since then found applications in various areas of mathematics, including topology, operads, cluster algebras, combinatorial Hopf algebras, and physics~\cite{PilaudSantosZiegler}.
+It was first defined as a combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894}.
 The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
@@
 
-A natural object associated with the associahedron is the \defn{Tamari lattice}, the cover graph of which is isomorphic to the skeleton of the associahedron~\cite{MR0146227,MR3235205}.
-The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} relation called the \defn{sylvester congruence}~\cite{MR1654173,MR2142078}.
+A natural object associated with the associahedron is the \defn{Tamari lattice}, whose cover graph is isomorphic to the skeleton of the associahedron.
+The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} called the \defn{sylvester congruence}.
 Answering a question of N.~Reading~\cite{MR2142177}, V.~Pilaud and F.~Santos~\cite{MR3964495} proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more precisely whose normal fan is the \defn{quotient fan} defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
@@
 
-The combinatorics of rectangulations has been studied for nearly two decades~\cite{MR2233287,MR2763051,MR2871762,MR2864445,MR3084577,MR3192492,MR3878132,MR4598046}.
-Interestingly, several combinatorial results were initially motivated by applications to the design of \defn{floorplans} for very large scale integrated circuits, and published in electrical engineering journals, see for instance R.
-Fujimaki, Y.
-Inoue, and T.
-Takahashi~\cite{FT07,TF08,ITF09,FIT09}.
+The combinatorics of rectangulations has been studied for decades~\cite{MR2233287,MR2871762,MR2864445,MR2763051,MR3084577,MR3878132,MR4598046}.
+Interestingly, many combinatorial results were initially motivated by applications to the design of \defn{floorplans} for very large scale integrated circuits, and published in electrical engineering journals~\cite{}.
+The recent order-theoretic take on these structures has proved fruitful.
 
@@
 We suppose throughout that no four rectangles have a common vertex, hence that the rectangulations are \defn{generic} in that sense.
-
 In order to focus on the combinatorial structure of a rectangulation, we introduce two equivalence relations between rectangulations.
-We say that a rectangle $r$ is \defn{above} another rectangle $s$ (and $s$ is \defn{below} $r$) if either the bottom edge of $r$ lies in the same segment as the top edge of $s$, or if $r$ is above another rectangle $t$ that is above $s$.
+First, we say that a rectangle $r$ is \defn{above} another rectangle $s$ (and $s$ is \defn{below} $r$) if either the bottom edge of $r$ lies in the same segment as the top edge of $s$, or if $r$ is above another rectangle $t$ that is above $s$.
 Similarly, a rectangle $r$ is \defn{on the left of} another rectangle $s$ (and $s$ is \defn{on the right of} $r$) if either the right edge of $r$ lies in the same segment as the left edge of $s$, or if $r$ is on the left of another rectangle $t$ that is on the left of $s$.
+
 Two rectangulations are said to be \defn{weakly equivalent} if there exists a bijection between their rectangles that preserves the above-below and left-right relations between the rectangles.
@@
 This is illustrated in Figure~\ref{fig:rectequiv}.
-A \defn{diagonal rectangulation} is a rectangulation in which every rectangle intersects the top-left to bottom-right diagonal of the square, see Figure~\ref{fig:weakRectangulation}.
+A \defn{diagonal rectangulation} is a rectangulation in which every rectangle intersects the top-left to bottom-right diagonal of the square.
 It is simple to check that every weak rectangulation has a diagonal representative, hence weak rectangulations can also be thought of as diagonal rectangulations.
@@
 
-The \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762} and revisited in~\cite[Thm.~1.1 \& Exm.~4.10]{Reading-arcDiagrams}.
-Its classes are in bijection with with weak rectangulations, but also with \defn{Baxter permutations}~\cite{MR0491652,MR0555815}, \defn{twin binary trees}~\cite{MR1417289,MR2914637}, and many other combinatorial families~\cite{MR2763051}.
-The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
-
-The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams}, and studied more recently by E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
+The \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762} and revisited in~\cite[Thm.~1.1 \& Exm.~4.10]{Reading-arcDiagrams} in terms of arc diagrams.
+Its classes are in bijection with weak rectangulations, but also with \defn{twin binary trees}~\cite{Giraudo}, and with \defn{Baxter permutations}~\cite{}. % more baxter refs
+The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by~$\WRP(n)$.
+%The skeleton of $\WRP(n)$ is isomorphic to the flip graph of weak rectangulations of size $n$ defined by Law and Reading~\cite{MR2871762}.
+
+The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams} in terms of arc diagrams, and studied more recently by A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
 Its classes are in bijection with strong rectangulations.
@@
 The various types of flips are illustrated in Figure~\ref{fig:flips}.
-We refer to \cite{MR2871762,MR3878132,MR3697823,ACFF24} for detailed descriptions of the flip graphs on both weak and strong rectangulations.
+We refer to S.~Law and N.~Reading~\cite{MR2871762}, J.~Cardinal, V.~Sacrist\'an, and R.~Silveira~\cite{}, E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and \'E.~Fusy~\cite{ACFF24} for detailed descriptions of the flip graphs on both weak and strong rectangulations.
 Just like associahedra encode the rotation graphs on binary trees, the rectangulotopes directly yield flip graphs on rectangulations.
@@
 \begin{lemma}
+%  The skeleton of $\WRP(n)$ is isomorphic to the flip graph on weak rectangulations of size~$n$.
+%  Similarly, the skeleton of $\SRP(n)$ is isomorphic to the flip graph on strong rectangulations of size~$n$.
   The skeleton of $\WRP(n)$ (resp.~of $\SRP(n)$) is isomorphic to the flip graph on weak (resp.~strong) rectangulations of size~$n$.
@@
 	\centerline{\includegraphics[width=.5\textwidth]{weakRectangulation} \qquad \includegraphics[width=.5\textwidth]{weakRectangulationTrees}}
-	\caption{A diagonal rectangulation (left) and its pair of twin binary trees (right). Example from \cite{ACFF24}.}
+	\caption{A weak rectangulation (left) and its pair of separated twin trees (right). Example from \cite{ACFF24}.}
 	% The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.
@@
 	\centerline{\includegraphics[width=.5\textwidth]{strongRectangulation} \qquad \includegraphics[width=.5\textwidth]{strongRectangulationTrees}}
-	\caption{A (non-diagonal) rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{ACFF24}.}
+	\caption{A strong rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{ACFF24}.}
         % The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
@@
 We complete each node of~$S(R)$ and~$T(R)$ with a vertical (resp.~horizontal) leaf if it has no vertical (resp.~horizontal) child, see \cref{fig:strongRectangulation}.
-If the rectangulation $R$ is diagonal, we retrieve the well-studied pair of twin binary trees~\cite{MR1417289,MR2914637}.
 Recall that the \defn{inorder labeling} of a binary tree~$T$ is the labeling of the nodes of~$T$ such that the label of each node~$t$ of~$T$ is larger than all labels in the left subtree of~$t$ and larger than all labels in the right subtree of~$t$.
@@
 
-This enables to unambiguously label the rectangles of~$R$ by the inorder.
+This enables to unambiguously label the rectangles of~$R$ by the inorder 
 The resulting labeling coincides with the NW--SE labeling of~\cite{ACFF24}.
@@
 \label{subsec:weakRectangulotopes}
+
+\begin{figure}
+	\centerline{\includegraphics[scale=1.5]{weakRectangulotopeLabeled}}
+	\caption{The strong rectangulotope~$\WRP(4)$.}
+        \label{fig:weakRectangulotope}
+\end{figure}
 
@@
 This materializes the result of Law and Reading that the weak rectangulotopes are Minkowski sums of two opposite associahedra~\cite{MR2871762}.
-
-\Jean{We need to make it clear whether $\WRP (n)$ is the proposed realization or just the combinatorial type.
-I was thinking the latter.
-Then we can change the sentence ``The weak rectangulotope is realized by the polytope~$\WRP (n)$...'' by ``The weak rectangulotope $\WRP (n)$ is realized by:~''.
-Otherwise, refrain from using $\WRP (n)$ before here.}
 
@@
 
+For instance, the weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotopeLabeled}.
+
 \begin{example}
@@
 \label{subsec:strongRectangulotopes}
+
+\begin{figure}
+	\centerline{\includegraphics[scale=1.5]{strongRectangulotopeLabeled}}
+	\caption{The strong rectangulotope~$\SRP(4)$.}
+        \label{fig:strongRectangulotope}
+\end{figure}
 
@@
 
+For instance, the strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotopeLabeled}.
+
 \begin{example}
@@
 We now briefly recall some results on the weak Bruhat order and its quotients, both from a lattice and geometric perspectives.
-We refer to \cite{Reading-arcDiagrams, MR3645055, MR3645056, MR4584712} for details.
-% J: Reading-Chapters: which one is this ?
+We refer to \cite{Reading-arcDiagrams, Reading-Chapters, PadrolPilaudRitter} for details.
 
@@
 	\centerline{\includegraphics[scale=.6]{weakOrderLeft4} \; \includegraphics[scale=.6]{weakOrderCongruence4} \; \includegraphics[scale=.48]{TamariLattice4}}
-	\caption{The weak Bruhat order on~$\f{S}_4$ (left), the sylvester congruence~$\equiv_\textrm{sylv}$~(middle), and the Tamari lattice (right). \cite[Fig.~1 \& 2]{MR3964495}}
+	\caption{The weak Bruhat order on~$\f{S}_4$ (left), the sylvester congruence~$\equiv_\textrm{sylv}$~(middle), and the Tamari lattice (right). \cite[Fig.~1 \& 2]{PilaudSantos-quotientopes}}
 	\label{fig:sylvesterCongruence}
@@
 	\centerline{\includegraphics[scale=.6]{braidFanLeft4} \qquad \raisebox{-.3cm}{\includegraphics[scale=.6]{sylvesterFan4}}}
-	\caption{The braid fan (left) and the sylvester fan (right). \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
+	\caption{The braid fan (left) and the sylvester fan (right). \cite[Fig.~1]{PilaudSantos-quotientopes} \&~\cite[Fig.~5]{PadrolPilaudRitter}}
 	\label{fig:fans}
@@
 	\centerline{\includegraphics[scale=.6]{permutahedronLeft4} \qquad \raisebox{-.4cm}{\includegraphics[scale=.6]{associahedron4}}}
-	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$. \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
+	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$. \cite[Fig.~1]{PilaudSantos-quotientopes} \&~\cite[Fig.~5]{PadrolPilaudRitter}}
 	\label{fig:quotientopes}
@@
 For instance, the associahedron~$\Asso(n)$ (\cref{fig:quotientopes}\,(right) when~$n = 4$)  is a quotientope for the sylvester congruence (\cref{fig:sylvesterCongruence}\,(middle) when~$n = 4$), its normal fan is the sylvester fan (\cref{fig:fans}\,(right) when~$n = 4$), and its skeleton is the cover graph of the Tamari lattice (\cref{fig:sylvesterCongruence}\,(right) when~$n = 4$).
-The existence of quotientopes was first proved in \cite{MR3964495} and later better understood in~\cite{MR4584712} using Minkowski sums of shard polytopes.
+The existence of quotientopes was first proved in \cite{PilaudSantos} and later better understood in~\cite{PadrolPilaudRitter} using Minkowski sums of shard polytopes.
 
@@
 The \defn{shard polytope} of~$\alpha$ is the convex hull~$\SP(\alpha)$ of the characteristic vectors of all $\alpha$-alternating matchings.
-It was shown in~\cite{MR4584712} that any Minkowski sum of positive scalings of the shard polytopes~$\SP(\alpha)$ for all arcs~$\alpha$ in the arc ideal~$\c{A}_\equiv$ is a quotientope for~$\equiv$.
+It was shown in~\cite{PadrolPilaudRitter} that any Minkowski sum of positive scalings of the shard polytopes~$\SP(\alpha)$ for all arcs~$\alpha$ in the arc ideal~$\c{A}_\equiv$ is a quotientope for~$\equiv$.
 
@@
 For instance, the vertex labeled~${\bullet \cdot \cdot \,\circ}$ has coordinates~$(1,0,0,-1)$.
-Adapted from~\cite[Fig.~10]{MR4584712}.}
+Adapted from~\cite[Fig.~10]{PadrolPilaudRitter}.}
 	\label{fig:shardPolytiopes}
@@
 Given a rectangulation~$R$, the source and target trees~$S(R)$ and~$T(R)$ define two partial orders on $[n]$ (the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top).
-The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of the two partial orders defined by the intertwining binary trees~$S(R)$ and~$T(R)$.
-These posets were characterized (and called Baxter posets) by E.~Meehan~\cite{MR4014603}.
+The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of the two partial orders defined by the intertwining binary trees~$S(R)$ and~$T(R)$~\cite{MR4014603}.
 The weak poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
@@
 
-\begin{theorem}[\cite{MR2871762}]
+\begin{theorem}[\cite{LawReading}]
   The subsets $\mathcal{L}(P_w(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, 
@@
   The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all up and down shard polytopes.
-  In particular, it is the Minkowski sum of two opposite associahedra of J.-L.~Loday~\cite{MR2871762}.
+  In particular, it is the Minkowski sum of two opposite associahedra of J.-L.~Loday~\cite{LawReading}.
 \end{lemma}
@@
 
-\begin{theorem}[\cite{MR2864445,ACFF24}]
+\begin{theorem}[\cite{Reading}]
   The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, 

### 2024-04-08 draft/rectangulotopes.tex
minor improvements

--- 2024-04-05 draft/rectangulotopes.tex
@@
   We describe realizations of $(n-1)$-dimensional polytopes associated with two families of rectangulations composed of $n$ rectangles.
-  They are defined as quotientopes of natural congruence relations on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
+  They are defined as quotientopes of natural lattice congruence on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
   We give simple vertex and facet descriptions of these polytopes, in particular elementary formulas for computing the coordinates of the vertex corresponding to each rectangulation, in the spirit of J.-L.~Loday's realization of the associahedron.
@@
 The encoding of combinatorial objects in the form of a polyhedral structure is a recurrent theme in geometric and algebraic combinatorics.
-An elementary yet striking example of this idea is the \defn{permutahedron} $\Perm (n)$, defined as the convex hull of the points $\sum_{i\in [n]} \sigma^{-1}(i)\cdot \b{e}_i$, for all permutations $\sigma \in \f{S}_n$.
-It is an $(n-1)$-dimensional polytope, whose facets are indexed by the proper nonempty subsets of $[n]$, and whose skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
+An elementary yet striking example of this idea is the \defn{permutahedron} $\Perm (n)$.
+It is the $(n-1)$-dimensional polytope defined either as the convex hull of the points $\sum_{i\in [n]} \sigma^{-1}(i)\cdot \b{e}_i$ for all permutations $\sigma \in \f{S}_n$, or as the intersection of the hyperplane of~$\R^n$ given by~$\sum_{i \in [n]} x_i = \smash{\binom{n+1}{2}}$ with the halfspaces given by~$\sum_{i \in I} x_i \ge \smash{\binom{|I|+1}{2}}$ for all nonempty proper subsets~$\varnothing \ne I \subsetneq [n]$.
+Its skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
 See \cref{fig:quotientopes}\,(left) for an illustration when~$n = 4$.
@@
 
-Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined by D.~Tamari~\cite{T51} and by J.~Stasheff~\cite{S63} in a topological context.
+Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined combinatorially by D.~Tamari~\cite{T51} and by J.~Stasheff~\cite{S63} in a topological context.
 It has since then been identified as a fundamental object in many other areas of mathematics, including operads, cluster algebras, combinatorial Hopf algebras, and physics (see for instance the recent survey from V.
@@
 In particular, its skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.
-In 2004, J.-L.~Loday published the following elegant description of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex given the corresponding binary tree.
-We note that the inequality description of the same polytope was actually provided in 1993 by S.~Shnider and S.~Sternberg~\cite{ShniderSternberg}, but Loday's vertex description largely popularized this realization.
+In~2004, J.-L.~Loday published the following elegant description of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex given the corresponding binary tree.
+We note that the inequality description of the same polytope was actually provided in~1993 by S.~Shnider and S.~Sternberg~\cite{ShniderSternberg}, but Loday's vertex description largely popularized this realization.
 See \cref{fig:quotientopes}\,(right) for an illustration when~$n = 4$.
@@
 
-A natural object associated with the associahedron is the \defn{Tamari lattice}, the cover graph of which is isomorphic to the skeleton of the associahedron~\cite{MR0146227,MR3235205}.
-The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} relation called the \defn{sylvester congruence}~\cite{MR1654173,MR2142078}.
+A natural object associated with the associahedron is the \defn{Tamari lattice}, whose cover graph is isomorphic to the skeleton of the associahedron~\cite{MR0146227,MR3235205}.
+The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} called the \defn{sylvester congruence}~\cite{MR1654173,MR2142078}.
 Answering a question of N.~Reading~\cite{MR2142177}, V.~Pilaud and F.~Santos~\cite{MR3964495} proved that with \emph{every} lattice congruence of the weak Bruhat order, one can associate a polytope whose skeleton is the cover graph of the lattice quotient, and more precisely whose normal fan is the \defn{quotient fan} defined by gluing the cones of the \defn{braid fan} (the type $A$ Coxeter arrangement) that belong to the same congruence class.
@@
 In order to focus on the combinatorial structure of a rectangulation, we introduce two equivalence relations between rectangulations.
-We say that a rectangle $r$ is \defn{above} another rectangle $s$ (and $s$ is \defn{below} $r$) if either the bottom edge of $r$ lies in the same segment as the top edge of $s$, or if $r$ is above another rectangle $t$ that is above $s$.
-Similarly, a rectangle $r$ is \defn{on the left of} another rectangle $s$ (and $s$ is \defn{on the right of} $r$) if either the right edge of $r$ lies in the same segment as the left edge of $s$, or if $r$ is on the left of another rectangle $t$ that is on the left of $s$.
+We say that a rectangle~$r$ is \defn{above} another rectangle~$s$ (and~$s$ is \defn{below}~$r$) if either the bottom edge of~$r$ lies in the same segment as the top edge of~$s$, or if~$r$ is above another rectangle~$t$ that is above~$s$.
+Similarly, a rectangle~$r$ is \defn{on the left of} another rectangle $s$ (and $s$ is \defn{on the right of} $r$) if either the right edge of $r$ lies in the same segment as the left edge of $s$, or if $r$ is on the left of another rectangle $t$ that is on the left of $s$.
 Two rectangulations are said to be \defn{weakly equivalent} if there exists a bijection between their rectangles that preserves the above-below and left-right relations between the rectangles.
@@
 
-The \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762} and revisited in~\cite[Thm.~1.1 \& Exm.~4.10]{Reading-arcDiagrams}.
-Its classes are in bijection with with weak rectangulations, but also with \defn{Baxter permutations}~\cite{MR0491652,MR0555815}, \defn{twin binary trees}~\cite{MR1417289,MR2914637}, and many other combinatorial families~\cite{MR2763051}.
+The \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762} and revisited in~\cite[Thm.~1.1 \& Exm.~4.10]{Reading-arcDiagrams} in terms of arc diagrams.
+Its classes are in bijection with weak rectangulations, but also with \defn{Baxter permutations}~\cite{MR0491652,MR0555815}, with \defn{twin binary trees}~\cite{MR1417289,MR2914637}, and with many other combinatorial families~\cite{MR2763051}.
 The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
 
-The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams}, and studied more recently by E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
+The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams} in terms of arc diagrams, and studied more recently by E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
 Its classes are in bijection with strong rectangulations.
@@
 The resulting labeling coincides with the NW--SE labeling of~\cite{ACFF24}.
-From now on, the labels of the rectangles in a rectangulations are the inorder labels, and the two trees $T(R)$ and $S(R)$ are defined on the vertex set $[n]$ accordingly.
+From now on, the labels of the rectangles in a rectangulation are the inorder labels, and the two trees $T(R)$ and $S(R)$ are defined on the vertex set $[n]$ accordingly.
 For each $i\in [n]$, we refer to its two subtrees in $T(R)$ as the \defn{horizontal} and \defn{vertical subtrees}, and similarly for $S(R)$.
@@
 For the facets, we combine the right hand sides of two opposite associahedra to obtain those of the weak rectangulotopes.
-This materializes the result of Law and Reading that the weak rectangulotopes are Minkowski sums of two opposite associahedra~\cite{MR2871762}.
+These two descriptions materialize the result of Law and Reading that the weak rectangulotopes are Minkowski sums of two opposite associahedra~\cite{MR2871762}.
 
@@
 
+For instance, the weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
+
 \begin{example}
@@
   \item $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
-  \item $ch^{T,S}_{i,j}$ (resp.~$cv^{T,S}_{i,j}$) denote the number of common leaves of the horizontal (resp.~vertical) subtree of $i$ in $T$ and the horizontal (resp.~vertical) subtree of $j$ in $S$.
+  \item $ch^{T,S}_{i,j}$ (resp.~$cv^{T,S}_{i,j}$) denote the number of common leaves of the horizontal (resp.~vertical) subtree of $i$ in $T$ and the horizontal (resp.~vertical) subtree of $j$ in $S$,
   \end{itemize}
-  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ in the sets denote nonempty and consecutive intervals of~$[n]$.
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
   \end{enumerate}
@@
 % jean:{Is there better way to write this down/compute? Here we do not take advantage of the fact that at most one of the two factors $\yin{w}_{i,j}$ and $\yang{w}_{i,j}$ is nonzero.}
+
+For instance, the strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotope}.
 
@@
 \cref{sec:weakRectangulotopes} details the case of weak rectangulations, while \cref{sec:strongRectangulotopes} deals with strong rectangulations.
-In both cases, in addition to the vertex coordinates, we also give the submodular functions defining the polytopes.
+In both cases, we provide both the vertex and facet descriptions of these polytopes.
 
@@
 The \defn{inversion set} of~$\sigma \in \f{S}_n$ is~$\inv(\sigma) \eqdef \set{(\sigma_i, \sigma_j)}{1 \le i < j \le n \text{ and } \sigma_i > \sigma_j}$.
-The \defn{weak Bruhat order}\footnote{The weak Bruhat order is usually just referred to as the weak order.
+The \defn{weak Bruhat order}\footnote{The weak Bruhat order is usually just referred to as the weak order~\cite{Humphrey, BjornerBrenti}.
 In this paper, we prefer to use weak Bruhat order to avoid any confusion with the weak poset defined on weak rectangulations.} is the lattice on the permutations of~$\f{S}_n$ defined by the inclusion of their inversion sets.
@@
 
+For instance, the join irreducible permutations minimal in their sylvester congruence class are precisely the \defn{up arcs}, \ie the arcs of the form~$(a, b, {]a,b[}, \varnothing)$ (similarly, we call \defn{down arcs} the arcs of the form~$(a, b, \varnothing, {]a,b[})$
+).
+The corresponding noncrossing arc diagrams are famously known as noncrossing partitions.
+
 %Moreover, the sets of arcs of the form~$\c{A}_\equiv$ are easily described.
-\vspace{-.15cm}
+\vspace{-.1cm}
 \parpic(5cm,2.8cm)(10pt, 90pt)[r][b]{\includegraphics[scale=.5]{subarcOrder}}{
@@
 It has a chamber for each permutation~$\sigma$ of~$[n]$, given by the set of vectors whose coordinates are ordered as~$\sigma$.
+It has a ray for each nonempty proper subset~$\varnothing \ne J \subsetneq [n]$.
+(To be more precise when speaking about rays, we should actually consider the essentialization of the braid arrangement obtained by slicing it with a hyperplane normal to its lineality space, which is generated by the vector~$(1, \dots, 1)$.)
 See \cref{fig:fans}\,(left).
@@
 
+It was shown in~\cite[Lem.~23]{AlbertinPilaudRitter} that the ray of the braid arrangement corresponding to a nonempty proper subset~$\varnothing \ne J \subsetneq [n]$ is preserved in the quotient fan~$\c{F}_\equiv$ if and only if the arc ideal~$\c{A}_\equiv$ contains the $|J|-1$ down arcs joining two consecutive elements of~$J$ and the~$n-|J|-1$ up arcs joining two consecutive elements of~$[n] \ssm J$.
+An immediate consequence is that all $2^n-2$ rays of the braid arrangement are preserved in the quotient fan~$\c{F}_\equiv$ if and only if the arc ideal~$\c{A}_\equiv$ contains all up arcs and all down arcs (in other words, if it is above the weak rectangulation congruence in the lattice of congruences of the weak Bruhat order).
+
 %%%%%%%%%%%%%
@@
 	\centerline{\includegraphics[scale=.6]{permutahedronLeft4} \qquad \raisebox{-.4cm}{\includegraphics[scale=.6]{associahedron4}}}
-	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$. \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
+	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$ (right). \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
 	\label{fig:quotientopes}
@@
 
-The \defn{Minkowski sum} of two polytopes~$\polytope{P}$ and~$\polytope{Q}$ is~$\polytope{P} + \polytope{Q} \eqdef \set{\b{p} + \b{q}}{\b{p} \in \polytope{P} \text{ and } \b{q} \in \polytope{Q}}$.
-Recall that for any vector~$\b{v} \ne \b{0}$, the face of~$\polytope{P} + \polytope{Q}$ maximizing the scalar product with~$\b{v}$ is the Minkowski sum of the faces of~$\polytope{P}$ and~$\polytope{Q}$ maximizing the scalar product with~$\b{v}$.
+The \defn{Minkowski sum} of two polytopes~$\polytope{P}, \polytope{Q} \subset \R^n$ is~$\polytope{P} + \polytope{Q} \eqdef \set{\b{p} + \b{q}}{\b{p} \in \polytope{P} \text{ and } \b{q} \in \polytope{Q}} \subset \R^n$.
+Recall that for any vector~$\b{v} \ne \b{0}$ of~$\R^n$, the face of~$\polytope{P} + \polytope{Q}$ maximizing the scalar product with~$\b{v}$ is the Minkowski sum of the faces of~$\polytope{P}$ and~$\polytope{Q}$ maximizing the scalar product with~$\b{v}$.
 In particular, if~$\b{v}$ is a generic direction, the vertex of~$\polytope{P} + \polytope{Q}$ extremal in direction~$\b{v}$ is just the sum of the vertices of~$\polytope{P}$ and~$\polytope{Q}$ that are extremal in direction~$\b{v}$.

### 2024-04-08 draft/rectangulotopes.tex
tables and other minor mprovements

@@
 %\tikzexternalize
-\graphicspath{{figures/}{figures/nodes/}}
+\graphicspath{{figures/}{figures/rectangulations/}}
 \makeatletter\def\input@path{{figures/}}\makeatother
@@
 
-A natural object associated with the associahedron is the \defn{Tamari lattice}, whose cover graph is isomorphic to the skeleton of the associahedron~\cite{MR0146227,MR3235205}.
+A natural object associated with the associahedron is the \defn{Tamari lattice}, the cover graph of which is isomorphic to the skeleton of the associahedron~\cite{MR0146227,MR3235205}.
 The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} called the \defn{sylvester congruence}~\cite{MR1654173,MR2142078}.
@@
 
-In order to focus on the combinatorial structure of a rectangulation, we introduce two equivalence relations between rectangulations.
+In order to focus on the combinatorial structure of a rectangulation, we recall two equivalence relations between rectangulations~\cite{??}.
 We say that a rectangle~$r$ is \defn{above} another rectangle~$s$ (and~$s$ is \defn{below}~$r$) if either the bottom edge of~$r$ lies in the same segment as the top edge of~$s$, or if~$r$ is above another rectangle~$t$ that is above~$s$.
-Similarly, a rectangle~$r$ is \defn{on the left of} another rectangle $s$ (and $s$ is \defn{on the right of} $r$) if either the right edge of $r$ lies in the same segment as the left edge of $s$, or if $r$ is on the left of another rectangle $t$ that is on the left of $s$.
+We define \defn{on the left of} and \defn{on the right of} similarly.
 Two rectangulations are said to be \defn{weakly equivalent} if there exists a bijection between their rectangles that preserves the above-below and left-right relations between the rectangles.
@@
 However, if the wall slide changes the adjacency relation between the rectangles, then the resulting rectangulation is not strongly equivalent to $R$ anymore.
-This is illustrated in Figure~\ref{fig:rectequiv}.
-A \defn{diagonal rectangulation} is a rectangulation in which every rectangle intersects the top-left to bottom-right diagonal of the square, see Figure~\ref{fig:weakRectangulation}.
+This is illustrated in \cref{fig:rectequiv}.
+A \defn{diagonal rectangulation} is a rectangulation in which every rectangle intersects the top-left to bottom-right diagonal of the square, see \cref{fig:weakRectangulation}.
 It is simple to check that every weak rectangulation has a diagonal representative, hence weak rectangulations can also be thought of as diagonal rectangulations.
@@
 The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
+For~$n = 4$, the weak rectangulation lattice is represented in \cref{fig:weakRectangulationLattice} and the weak rectangulotope is represented in \cref{fig:weakRectangulotope}.
+
+\begin{figure}
+	\centerline{\includegraphics[scale=1.05]{weakRectangulationLattice}}
+	\caption{The weak rectangulation lattice.}
+	\label{fig:weakRectangulationLattice}
+\end{figure}
+
+\begin{figure}
+	\centerline{\includegraphics[scale=1.05]{weakRectangulotopeLabeled}}
+	\caption{The strong rectangulotope~$\WRP(4)$.}
+        \label{fig:weakRectangulotope}
+\end{figure}
 
@@
 The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by~$\SRP(n)$.
+For~$n = 4$, the strong rectangulation lattice is represented in \cref{fig:strongRectangulationLattice} and the strong rectangulotope is represented in \cref{fig:strongRectangulotope}.
+
+\begin{figure}
+	\centerline{\includegraphics[scale=1.05]{strongRectangulationLattice}}
+	\caption{The strong rectangulation lattice.}
+	\label{fig:strongRectangulationLattice}
+\end{figure}
+
+\begin{figure}
+	\centerline{\includegraphics[scale=1.05]{strongRectangulotopeLabeled}}
+	\caption{The strong rectangulotope~$\SRP(4)$.}
+        \label{fig:strongRectangulotope}
+\end{figure}
 
@@
 Flips in rectangulations come in three flavors:
-\begin{description}
-\item[Simple flips] replace a vertical segment incident to exactly two rectangles by a horizontal segment, and vice-versa.
-\item[Pivoting] operations transform a pair of adjacent rectangles whose union is not a rectangle, changing a left-right pair into an above-below pair, and vice-versa.
-\item[Wall slide] flips are wall slides that modify the adjacency of the rectangles, removing exactly one adjacency and adding exactly one. These flips are only relevant in strong rectangulations.
-\end{description}
-
-The various types of flips are illustrated in Figure~\ref{fig:flips}.
+\begin{itemize}
+\item \defn{Simple flips} replace a vertical segment incident to exactly two rectangles by a horizontal segment, and vice-versa,
+\item \defn{Pivoting operations} transform a pair of adjacent rectangles whose union is not a rectangle, changing a left-right pair into an above-below pair, and vice-versa,
+\item \defn{Wall slide flips} are wall slides that modify the adjacency of the rectangles, removing exactly one adjacency and adding exactly one.
+These flips are only relevant in strong rectangulations.
+\end{itemize}
+
+The various types of flips are illustrated in \cref{fig:flips}.
 We refer to \cite{MR2871762,MR3878132,MR3697823,ACFF24} for detailed descriptions of the flip graphs on both weak and strong rectangulations.
 Just like associahedra encode the rotation graphs on binary trees, the rectangulotopes directly yield flip graphs on rectangulations.
+Namely, the skeleton of $\WRP(n)$ (resp.~of $\SRP(n)$) is isomorphic to the flip graph on weak (resp.~strong) rectangulations of size~$n$.
 
@@
 
-\begin{lemma}
-  The skeleton of $\WRP(n)$ (resp.~of $\SRP(n)$) is isomorphic to the flip graph on weak (resp.~strong) rectangulations of size~$n$.
-  \vincent{I would leave that in the text, not in a proper thm environment...}
-\end{lemma}
-
 Rectangulotopes enrich the family of natural quotientopes, alongside associahedra, permutreehedra~\cite{MR3856522}, certain brick polytopes~\cite{PilaudSantos-brickPolytopes, PilaudStump-brickPolytopes, Pilaud-brickAlgebra}, and certain graphical zonotopes~\cite{Pilaud-brickAlgebra, Pilaud-acyclicReorientationLattices}, providing instructive examples of application of the theory of lattice congruences to concrete combinatorial objects.
 Our main results are elementary vertex and facet descriptions of both families of quotientopes.
-For vertices, we describe the coordinates of the vertex corresponding to a weak or strong rectangulation with simple product formulas in the spirit of J.-L.~Loday's realization in~\cref{thm:loday}.
+For vertices, we describe the coordinates of the vertex corresponding to a weak or strong rectangulation with simple product formulas in the spirit of J.-L.~Loday's realization in~\cref{thm:loday}\,(i).
 As mentioned in~\cite{PilaudSantosZiegler}, there is no such formula for arbitrary quotientopes, which makes rectangulotopes join the restricted club of Loday type quotientopes, jointly with associahedra and permutreehedra~\cite{MR3856522}.
@@
 
-\begin{figure}
-	\centerline{\includegraphics[scale=1.5]{weakRectangulotopeLabeled}}
-	\caption{The strong rectangulotope~$\WRP(4)$.}
-        \label{fig:weakRectangulotope}
-\end{figure}
-
 Our first results are simple vertex and facet descriptions of the weak rectangulotopes~$\WRP(n)$.
@@
 \begin{theorem}
+\label{thm:weakRectangulotope}
   The weak rectangulotope is realized by the polytope~$\WRP (n) \subset \R^n\!$ defined equivalently~as
@@
 For instance, the weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
+The coordinates of the vertices corresponding to each weak rectangulation are gathered in \cref{tab:verticesWeakRectangulotope}.
+\vincent{add table for the facets}
+
+\begin{table}[p]
+	\centerline{
+	\begin{tabular}{cccccccc}
+		\includegraphics[scale=1.5]{rectangulation1} &
+		\includegraphics[scale=1.5]{rectangulation2} &
+		\includegraphics[scale=1.5]{rectangulation3} &
+		\includegraphics[scale=1.5]{rectangulation4} &
+		\includegraphics[scale=1.5]{rectangulation5} &
+		\includegraphics[scale=1.5]{rectangulation6} &
+		\\
+		$(-3, -1, 1, 3)$ &
+		$(-3, -1, 5, -1)$ &
+		$(-3, 3, -3, 3)$ &
+		$(1, -5, 1, 3)$ &
+		$(-3, 0, 5, -2)$ &
+		$(-3, 5, -3, 1)$ &
+		\\[.2cm]
+		\includegraphics[scale=1.5]{rectangulation8} &
+		\includegraphics[scale=1.5]{rectangulation9} &
+		\includegraphics[scale=1.5]{rectangulation10} &
+		\includegraphics[scale=1.5]{rectangulation11} &
+		\includegraphics[scale=1.5]{rectangulation12} &
+		\includegraphics[scale=1.5]{rectangulation13} &
+		\\
+		$(-1, 3, -5, 3)$ &
+		$(2, -5, 0, 3)$ &
+		$(-3, 5, 0, -2)$ &
+		$(-2, 0, 5, -3)$ &
+		$(1, -5, 5, -1)$ &
+		$(-1, 5, -5, 1)$ &
+		\\[.2cm]
+		\includegraphics[scale=1.5]{rectangulation14} &
+		\includegraphics[scale=1.5]{rectangulation15} &
+		\includegraphics[scale=1.5]{rectangulation16} &
+		\includegraphics[scale=1.5]{rectangulation17} &
+		\includegraphics[scale=1.5]{rectangulation19} &
+		\includegraphics[scale=1.5]{rectangulation20} &
+		\\
+		$(3, -5, 0, 2)$ &
+		$(2, 0, -5, 3)$ &
+		$(-2, 5, 0, -3)$ &
+		$(1, -3, 5, -3)$ &
+		$(3, -5, 3, -1)$ &
+		$(3, 0, -5, 2)$ &
+		\\[.2cm]
+		&		
+		\includegraphics[scale=1.5]{rectangulation21} &
+		\includegraphics[scale=1.5]{rectangulation22} &
+		\includegraphics[scale=1.5]{rectangulation23} &
+		\includegraphics[scale=1.5]{rectangulation24} &
+		\\
+		&
+		$(-1, 5, -1, -3)$ &
+		$(3, -3, 3, -3)$ &
+		$(3, 1, -5, 1)$ &
+		$(3, 1, -1, -3)$ &
+	\end{tabular}
+	}
+	\caption{The vertices of the weak rectangulotope~$\WRP(4)$ of \cref{fig:weakRectangulotope}.}
+	\label{tab:verticesWeakRectangulotope}
+\end{table}
 
@@
 \label{subsec:strongRectangulotopes}
-
-\begin{figure}
-	\centerline{\includegraphics[scale=1.5]{strongRectangulotopeLabeled}}
-	\caption{The strong rectangulotope~$\SRP(4)$.}
-        \label{fig:strongRectangulotope}
-\end{figure}
 
@@
 \begin{theorem}
+\label{thm:strongRectangulotope}
   The $(n-1)$-dimensional strong rectangulotope is realized by the polytope~$\SRP (n)$ defined equivalently as
@@
 For instance, the strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotope}.
+The coordinates of the vertices corresponding to each strong rectangulation are gathered in \cref{tab:verticesStrongRectangulotope}.
+\vincent{add table for the facets}
+
+\begin{table}
+	\centerline{
+	\begin{tabular}{cccccccc}
+		\includegraphics[scale=1.5]{rectangulation1} &
+		\includegraphics[scale=1.5]{rectangulation2} &
+		\includegraphics[scale=1.5]{rectangulation3} &
+		\includegraphics[scale=1.5]{rectangulation4} &
+		\includegraphics[scale=1.5]{rectangulation5} &
+		\includegraphics[scale=1.5]{rectangulation6} &
+		\\
+		$(-6, -2, 2, 6)$ &
+		$(-6, -2, 9, -1)$ &
+		$(-6, 5, -5, 6)$ &
+		$(1, -9, 2, 6)$ &
+		$(-6, 1, 9, -4)$ &
+		$(-6, 9, -5, 2)$ &
+		\\[.2cm]
+		\includegraphics[scale=1.5]{rectangulation8} &
+		\includegraphics[scale=1.5]{rectangulation9} &
+		\includegraphics[scale=1.5]{rectangulation10} &
+		\includegraphics[scale=1.5]{rectangulation11} &
+		\includegraphics[scale=1.5]{rectangulation12} &
+		\includegraphics[scale=1.5]{rectangulation13} &
+		\\
+		$(-2, 5, -9, 6)$ &
+		$(4, -9, -1, 6)$ &
+		$(-6, 9, 1, -4)$ &
+		$(-4, 1, 9, -6)$ &
+		$(2, -9, 9, -2)$ &
+		$(-2, 9, -9, 2)$ &
+		\\[.2cm]
+		\includegraphics[scale=1.5]{rectangulation14} &
+		\includegraphics[scale=1.5]{rectangulation15} &
+		\includegraphics[scale=1.5]{rectangulation16} &
+		\includegraphics[scale=1.5]{rectangulation17} &
+		\includegraphics[scale=1.5]{rectangulation19} &
+		\includegraphics[scale=1.5]{rectangulation20} &
+		\\
+		$(6, -9, -1, 4)$ &
+		$(4, -1, -9, 6)$ &
+		$(-4, 9, 1, -6)$ &
+		$(2, -5, 9, -6)$ &
+		$(6, -9, 5, -2)$ &
+		$(6, -1, -9, 4)$ &
+		\\[.2cm]
+		\includegraphics[scale=1.5]{rectangulation21} &
+		\includegraphics[scale=1.5]{rectangulation22} &
+		\includegraphics[scale=1.5]{rectangulation23} &
+		\includegraphics[scale=1.5]{rectangulation24} &
+		\includegraphics[scale=1.5]{rectangulation7} &
+		\includegraphics[scale=1.5]{rectangulation18}
+		\\
+		$(-1, 9, -2, -6)$ &
+		$(6, -5, 5, -6)$ &
+		$(6, 2, -9, 1)$ &
+		$(6, 2, -2, -6)$ &
+		$(1, -9, 9, -1)$ &
+		$(-1, 9, -9, 1)$
+	\end{tabular}
+	}
+	\caption{The vertices of the weak rectangulotope~$\WRP(4)$ of \cref{fig:weakRectangulotope}.}
+	\label{tab:verticesWeakRectangulotope}
+\end{table}
 
@@
   \[
-  ???
-  \]
-  For instance, the $4$th coordinate is $???$.
+  (-41, 5, 262, 24, -110, 50, -525, -50, 184, 3, -88, 450, 35, -221, -5, 27).
+  \]
+  For instance, the contribution of~$\b{e}_5 - \b{e}_7$ is ...
   \vincent{give also a relevant example of facet...}
 \end{example}
+
+We think of the formula of \cref{thm:strongRectangulotope}\,(i) as the analogue of J.-L.-Loday's product formula of \cref{thm:loday}\,(i) for the coordinates of the vertices of the associahedron.
+We note however that our formula is quadratic in~$n$.
+It remains unclear to us if there is a simple linear formula to compute the coordinates of the vertex of~$\SRP(n)$ corresponding to a given strong rectangulation.
 
@@
 
-\begin{figure}
-	\centerline{\includegraphics[scale=.9]{weakRectangulationLattice}}
-	\caption{The weak rectangulation lattice.}
-	\label{fig:weakRectangulationLattice}
-\end{figure}
-
-\begin{figure}
-	\centerline{\includegraphics[scale=.9]{strongRectangulationLattice}}
-	\caption{The strong rectangulation lattice.}
-	\label{fig:strongRectangulationLattice}
-\end{figure}
-
 \end{document}

### 2024-04-08 draft/rectangulotopes.tex
first facets

@@
   where~$S$ and~$T$ are the source and target trees of the rectangulation~$R$, and $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
-  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{I \not\subseteq S \text{ and } I \not\subseteq [n] \ssm S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$.
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{I \not\subseteq S \text{ and } I \not\subseteq [n] \ssm S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$.
   \end{enumerate}
@@
 For instance, the weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
-The coordinates of the vertices corresponding to each weak rectangulation are gathered in \cref{tab:verticesWeakRectangulotope}.
-\vincent{add table for the facets}
+The coordinates of the vertex corresponding to each weak rectangulation are gathered in \cref{tab:verticesWeakRectangulotope}.
+The facet inequalities corresponding to each nonempty proper subset of~$[n]$ are gathered in \cref{tab:facetsWeakRectangulotope}.
 
@@
 	\centerline{
-	\begin{tabular}{cccccccc}
+	\begin{tabular}{cccccc}
 		\includegraphics[scale=1.5]{rectangulation1} &
@@
 		\includegraphics[scale=1.5]{rectangulation5} &
-		\includegraphics[scale=1.5]{rectangulation6} &
+		\includegraphics[scale=1.5]{rectangulation6}
 		\\
@@
 		$(-3, 0, 5, -2)$ &
-		$(-3, 5, -3, 1)$ &
+		$(-3, 5, -3, 1)$
 		\\[.2cm]
@@
 		\includegraphics[scale=1.5]{rectangulation12} &
-		\includegraphics[scale=1.5]{rectangulation13} &
+		\includegraphics[scale=1.5]{rectangulation13}
 		\\
@@
 		$(1, -5, 5, -1)$ &
-		$(-1, 5, -5, 1)$ &
+		$(-1, 5, -5, 1)$
 		\\[.2cm]
@@
 		\includegraphics[scale=1.5]{rectangulation19} &
-		\includegraphics[scale=1.5]{rectangulation20} &
+		\includegraphics[scale=1.5]{rectangulation20}
 		\\
@@
 		$(3, -5, 3, -1)$ &
-		$(3, 0, -5, 2)$ &
+		$(3, 0, -5, 2)$
 		\\[.2cm]
@@
 
+\begin{table}
+	\centerline{
+	\begin{tabular}{c|cccccccccccccc}
+		$J$ &
+		$1$ &
+		$2$ &
+		$3$ &
+		$4$ &
+		$12$ &
+		$13$ &
+		$14$ &
+		$23$ &
+		$24$ &
+		$34$ &
+		$123$ &
+		$124$ &
+		$134$ &
+		$234$
+		\\
+		\hline
+		$f_{\weakeq}(J)$ &
+		$3$ &
+		$5$ &
+		$5$ &
+		$3$ &
+		$4$ &
+		$6$ &
+		$5$ &
+		$5$ &
+		$6$ &
+		$4$ &
+		$3$ &
+		$5$ &
+		$5$ &
+		$3$
+	\end{tabular}
+	}
+	\caption{The facets of the weak rectangulotope~$\WRP(4)$ of \cref{fig:weakRectangulotope}.}
+	\label{tab:facetsWeakRectangulotope}
+\end{table}
+
 \begin{example}
@@
   \end{itemize}
-  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
   \end{enumerate}
@@
 For instance, the strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotope}.
-The coordinates of the vertices corresponding to each strong rectangulation are gathered in \cref{tab:verticesStrongRectangulotope}.
-\vincent{add table for the facets}
+The coordinates of the vertex corresponding to each strong rectangulation are gathered in \cref{tab:verticesStrongRectangulotope}.
+The facet inequalities corresponding to each nonempty proper subset of~$[n]$ are gathered in \cref{tab:facetsStrongRectangulotope}.
 
@@
 		$(-6, 1, 9, -4)$ &
-		$(-6, 9, -5, 2)$ &
+		$(-6, 9, -5, 2)$
 		\\[.2cm]
@@
 		$(2, -9, 9, -2)$ &
-		$(-2, 9, -9, 2)$ &
+		$(-2, 9, -9, 2)$
 		\\[.2cm]
@@
 		$(6, -9, 5, -2)$ &
-		$(6, -1, -9, 4)$ &
+		$(6, -1, -9, 4)$
 		\\[.2cm]
@@
 	}
-	\caption{The vertices of the weak rectangulotope~$\WRP(4)$ of \cref{fig:weakRectangulotope}.}
-	\label{tab:verticesWeakRectangulotope}
+	\caption{The vertices of the strong rectangulotope~$\SRP(4)$ of \cref{fig:strongRectangulotope}.}
+	\label{tab:verticesStrongRectangulotope}
+\end{table}
+
+\begin{table}
+	\centerline{
+	\begin{tabular}{cccccccc}
+	\end{tabular}
+	}
+	\caption{The facets of the strong rectangulotope~$\SRP(4)$ of \cref{fig:strongRectangulotope}.}
+	\label{tab:facetsStrongRectangulotope}
 \end{table}

### 2024-04-09 draft/rectangulotopes.tex
facets strong rectangulotope
minor changes

@@
 	\caption{Three types of flips between rectangulations.}
-	\label{fig:flip}
+	\label{fig:flips}
 \end{figure}
@@
   where~$S$ and~$T$ are the source and target trees of the rectangulation~$R$, and $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
-  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{I \not\subseteq S \text{ and } I \not\subseteq [n] \ssm S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$.
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by \linebreak $\sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{I \not\subseteq S \text{ and } I \not\subseteq [n] \ssm S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$.
   \end{enumerate}
@@
 
-For instance, the weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
+\begin{example}
+The weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
 The coordinates of the vertex corresponding to each weak rectangulation are gathered in \cref{tab:verticesWeakRectangulotope}.
@@
 \end{table}
+\end{example}
 
@@
   For instance, the $4$th coordinate is $h^T_4 \cdot v^T_4 - h^S_4 \cdot v^S_4 = 1 \cdot 5 - 1 \cdot 3 = 2$.
-  \vincent{give also a relevant example of facet...}
+  The facet of~$\WRP (16)$ corresponding to the subset~$\{3,4,7,8,9,10,14\}$ is given by the inequality
+  \[
+  x_3 + x_4 + x_7 + x_8 + x_9 + x_{10} + x_{14} \le ???
+  \]
+  \vincent{Compute and check...}
 \end{example}
@@
   \end{itemize}
-  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
+  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by \linebreak $\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
   \end{enumerate}
@@
 
-For instance, the strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotope}.
+\begin{example}
+The strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotope}.
 The coordinates of the vertex corresponding to each strong rectangulation are gathered in \cref{tab:verticesStrongRectangulotope}.
@@
 	\centerline{
-	\begin{tabular}{cccccccc}
+	\begin{tabular}{c|cccccccccccccc}
+		$J$ &
+		$1$ &
+		$2$ &
+		$3$ &
+		$4$ &
+		$12$ &
+		$13$ &
+		$14$ &
+		$23$ &
+		$24$ &
+		$34$ &
+		$123$ &
+		$124$ &
+		$134$ &
+		$234$
+		\\
+		\hline
+		$f_{\strongeq}(J)$ &
+		$6$ &
+		$9$ &
+		$9$ &
+		$6$ &
+		$8$ &
+		$11$ &
+		$10$ &
+		$10$ &
+		$11$ &
+		$8$ &
+		$6$ &
+		$9$ &
+		$9$ &
+		$6$
 	\end{tabular}
@@
   For instance, the contribution of~$\b{e}_5 - \b{e}_7$ is ...
-  \vincent{give also a relevant example of facet...}
+  \vincent{Compute and check...}
+  The facet of~$\SRP (16)$ corresponding to the subset~$\{3,4,7,8,9,10,14\}$ is given by the inequality
+  \[
+  x_3 + x_4 + x_7 + x_8 + x_9 + x_{10} + x_{14} \le ???
+  \]
+  \vincent{Compute and check...}
 \end{example}

### 2024-04-19 draft/rectangulotopes.tex
proof of coords for strong rect.topes

@@
 \newcommand{\verticalPattern}{\smash{\raisebox{-.25cm}{\includegraphics[scale=.5]{verticalPattern}}}} 
+% orders
+\newcommand{\tri}{\lhd}
+\newcommand{\btri}{\LHD}
 
@@
 See \cref{fig:quotientopes}\,(left) for an illustration when~$n = 4$.
-The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra}~\cite{MR2520477,Postnikov,MR4064768,MR4651496} (also call \defn{generalized permutahedra}, or \defn{polymatroids}).
+The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra}~\cite{MR2520477,Postnikov,MR4064768,MR4651496} (also called \defn{generalized permutahedra}, or \defn{polymatroids}).
 
@@
 
+A set function $f:2^{[n]}\mapsto \R$ is said to be \defn{submodular} when $f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$.
 By definition, any quotientope is a deformed permutahedron, and can thus be written as
@@
   \]
-where~$f$ is a \defn{submodular} set function $f:2^{[n]}\mapsto \R$ (\ie such that~$f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$).
-Recall that the submodular function of a Minkowski sum of deformed permutahedra is just the sum of the submodular functions of the Minkowski summands.
-
-\begin{lemma}
-  \label{lem:submodsum}
-  \vincent{I would not make this one so visible in a proper thm environment.}
-  Consider a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by the submodular functions $f_1,f_2,\ldots ,f_k$.
-Then the Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f \eqdef \sum_i f_i$.
-\end{lemma}
+where~$f$ is a submodular set function.
+We will need the following simple result: Given a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by submodular set functions $f_1,f_2,\ldots ,f_k$, their Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f \eqdef \sum_i f_i$.
 
@@
 %%%%%%%%%%%%%
-
+\subsection{The weak poset}
+\label{subsec:weakPoset}
+
+Given a rectangulation~$R$, let us consider the source and target trees~$S(R)$ and~$T(R)$, and orient the horizontal edges from left to right, and the vertical edges from bottom to top.
+This defines two partial orders on $[n]$.
+The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of these two partial orders defined by~$S(R)$ and~$T(R)$.
+Equivalently, the weak poset can be defined by considering the adjacency graph of the rectangles in a diagonal representative of the weak equivalence class of $R$, orienting its edges from left to right and from bottom to top, and taking the transitive closure of this directed acyclic graph.
+These posets were characterized (and called Baxter posets) by E.~Meehan~\cite{MR4014603}.
+The weak poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
+
+%%%%%%%%%%%%%
 \subsection{The weak rectangulation congruence}
@@
 
-Given a rectangulation~$R$, the source and target trees~$S(R)$ and~$T(R)$ define two partial orders on $[n]$ (the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top).
-The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of the two partial orders defined by the intertwining binary trees~$S(R)$ and~$T(R)$.
-These posets were characterized (and called Baxter posets) by E.~Meehan~\cite{MR4014603}.
-The weak poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
 We define the set~$\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
@@
 \begin{lemma}
-The weak rectangulation congruence $
-\weakeq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that do not cross the horizontal line.
+The weak rectangulation congruence $\weakeq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that do not cross the horizontal line.
 \end{lemma}
@@
 %%%%%%%%%%%%%
-
+\subsection{The strong poset}
+\label{subsec:strongPoset}
+
+Given a rectangulation $R$, we define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of $R$ by considering two relations on $[n]$~\cite{ACFF24}:
+\begin{enumerate}
+\item The \defn{adjacency poset} $([n],\tri)$ of $R$, defined by considering the adjacency graph of the rectangles of $R$ (identified by their inorder labels in $[n]$), orienting its edges from left to right and from bottom to top, and taking the transitive closure of this directed acyclic graph.
+\item The \defn{special relation} $([n],\btri)$ of $R$, in which $a\btri b$ if the two rectangles with inorder labels $a$ and $b$ are such that either (i) the right side of $b$ and the left side of $a$ lie on the same vertical segment, and the bottom-right corner of $b$ is above the top-left corner of $a$, or (ii) the top side of $b$ and the bottom side of rectangle $a$ lie on the same horizontal segment, and the top-left corner of $b$ lies on the right of the bottom-right corner of  $a$.
+\end{enumerate}
+The strong poset is the transitive closure of the union of the two relations $\tri$ and $\btri$. 
+The strong poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
+
+We will use the following statement.
+\begin{lemma}
+  \label{lem:sextw}
+  The strong poset $P_s(R)$ of a rectangulation $R$ is an extension of its weak poset $P_w(R)$.
+\end{lemma}
+\begin{proof}
+  Clearly, the strong poset is an extension of the adjacency poset $([n],\tri)$ of $R$, since it is obtained by adding pairs from the special relation.
+  It remains to observe that the adjacency poset of $R$ is an extension of the weak poset of $R$.
+  For this, we consider the diagonal representative $D$ of the weak equivalence class of $R$, and recall that the weak poset of $R$ is the adjacency poset of $D$.
+  By definition, $R$ can be obtained from $D$ by a sequence of wall slides.
+  Each wall slide in this sequence consists of either a horizontal segment being pulled down past another horizontal segment on its right, or a vertical segment being pushed right past another vertical segment below it.
+  Note that each of these we can only add new adjacencies, hence the adjacency poset of $R$ is always an extension of that of $D$. 
+\end{proof}
+
+%%%%%%%%%%%%%
 \subsection{The strong rectangulation congruence}
@@
 
-We define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of a rectangulation $R$, and $\mathcal{L}(P_s)$ as the set of permutations of $\f{S}_n$ corresponding to linear extensions of $P_s$:
+Let $\mathcal{L}(P_s)$ be the set of permutations of $\f{S}_n$ corresponding to linear extensions of the strong poset $P_s$:
 \[
@@
 \begin{theorem}[\cite{MR2864445,ACFF24}]
-  The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, 
-  are equivalence classes of a congruence relation $\strongeq$ on the weak Bruhat order on $\f{S}_n$.
+  The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, are equivalence classes of a congruence relation $\strongeq$ on the weak Bruhat order on $\f{S}_n$.
     Furthermore:
@@
 \begin{proof}
-  \vincent{I still need to work out this proof...}
-  Let $\pi$ be the 2-clumped permutation corresponding to $R$, hence the minimal element of $\mathcal{L}(P_s(R))$ in the weak Bruhat order.
-  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc $A$ whose shard polytope $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ only if
-  $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.
+  Let $\pi$ be any permutation corresponding to $R$, hence an element of $\mathcal{L}(P_s(R))$.
+  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc $A$ whose shard polytope $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ only if $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.
   Provided this holds, the set of yin arcs $A$ such that $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that $i$ is the index of the maximum of $\pi^{-1}$ in $I$, and $j$ is the index of the maximum of $\pi^{-1}$ in $J$.
-  %$i = \arg\max_{k\in I} \pi^{-1}(k)$ and $j = \arg\min_{k\in J} \pi^{-1}(k)$.
   We claim that the number of choices of the three endpoints defining such a pair of intervals $I$ and $J$ is the product of the
-  number of leaves of the subtree of $T(R)$ rooted at $i$, the number of common leaves to the vertical subtrees of $T(R)$ and $S(R)$ rooted
-  at $i$ and $j$, respectively, and the number of leaves of the subtree of $S(R)$ rooted at $j$.
+  number $h^T_i$ of leaves of the horizontal subtree of $T(R)$ rooted at $i$, the number $cv^{T,S}_{i,j}$ of common leaves to the vertical subtrees of $T(R)$ and $S(R)$ rooted
+  at $i$ and $j$, respectively, and the number $h^S_j$ of leaves of the horizontal subtree of $S(R)$ rooted at $j$.
+
+  We first show that the number of choices for the left endpoint of $I$ is equal to the number $h^T_i$ of leaves in the horizontal subtree of $i$ in $T(R)$.
+  Each such leaf determines a left endpoint for $I$, defined as the node of the tree that follows it in an inorder traversal.
+  Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
+  Since from Lemma~\ref{lem:sextw} the strong poset extends the order induced by $T(R)$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
+  Therefore, each such choice of left endpoint for $I$ preserves the fact that $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
+  Now consider the leaf of $T(R)$ that is immediately left of the leftmost leaf of the horizontal subtree of $i$.
+  This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T(R)$.
+  Therefore, we must have $\ell\succ_s i$, and $\pi^{-1}(\ell) > \pi^{-1}(i)$, and we cannot extend $I$ past this index on the left.
+  This shows that there are exactly $h^T_i$ choices, and not more, for the choice of the left endpoint of $I$.
+
+  The rest of the formula comes by symmetric arguments.
+  First, the right endpoint of $I$ can only be chosen among the leaves of the vertical subtree of $T(R)$ rooted at $i$.
+  Then the same reasoning holds for the two endpoints of $J$, replacing $i$ by $j$ and $T(R)$ by $S(R)$.
+  Finally, since the right endpoint of $I$ and the left endpoint of $J$ must be consecutive, the selected leaves must correspond to the same pair of successive indices of $[n]$.
+  These leaves are exactly the common leaves defined previously.
 \end{proof}

### 2024-04-19 draft/rectangulotopes.tex
some proofs

@@
 \usepackage[T1]{fontenc}
-\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont, verbatim, stackrel, stmaryrd}
+\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont, verbatim, stackrel, stmaryrd, subcaption}
 \usepackage{marginnote}
@@
   \label{thm:loday}
-  The associahedron is realized by the polytope~$\Asso (n) \subset \R^n$ defined equivalently as
+  % The associahedron is realized by the polytope~$\Asso (n) \subset \R^n$ defined equivalently as
+    The associahedron $\Asso (n)$ is realized by:
   \begin{enumerate}[(i)]
-  \item the convex hull of the points~$\sum_{i\in [n]} \ell^T_i\cdot r^T_i \cdot \b{e}_i$ for all binary trees~$T$ with $n$ internal nodes, where $\ell^T_i$ and $r^T_i$ denote the number of leaves in the left and right subtrees of $i$ in $T$, see~\cite{MR2108555},
-  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in I} x_i \le \#\set{J \text{ interval of } [n]}{I \cap J \ne \varnothing}$ for all intervals~$\varnothing \ne I \subsetneq [n]$, see~\cite{ShniderSternberg}.
+  \item the convex hull of the points
+    \[
+    \sum_{i\in [n]} \ell^T_i\cdot r^T_i \cdot \b{e}_i
+    \]
+    for all binary trees~$T$ with $n$ internal nodes, where $\ell^T_i$ and $r^T_i$ denote the number of leaves in the left and right subtrees of $i$ in $T$, see~\cite{MR2108555},
+  \item the intersection of the hyperplane defined by $\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by
+    \[
+    \sum_{i \in I} x_i \le \#\set{J \text{ interval of } [n]}{I \cap J \ne \varnothing}
+    \]
+    for all intervals~$\varnothing \ne I \subsetneq [n]$, see~\cite{ShniderSternberg}.
   \end{enumerate}
@@
 
-A natural object associated with the associahedron is the \defn{Tamari lattice}, the cover graph of which is isomorphic to the skeleton of the associahedron~\cite{MR0146227,MR3235205}.
+A natural object associated with the associahedron is the \defn{Tamari lattice}, whose cover graph is isomorphic to the skeleton of the associahedron~\cite{MR0146227,MR3235205}.
 The Tamari lattice is known to be the quotient of the weak Bruhat order by a \defn{lattice congruence} called the \defn{sylvester congruence}~\cite{MR1654173,MR2142078}.
@@
 The goal of this paper is to describe the geometry of the quotientopes for two particular congruences of the weak Bruhat order whose classes are in bijection with equivalence classes of rectangulations, defined as decompositions of a square into rectangles.
+
+\begin{figure}
+	\capstart
+	\centerline{\includegraphics[scale=.6]{permutahedronLeft4} \qquad \raisebox{-.4cm}{\includegraphics[scale=.6]{associahedron4}}}
+	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$ (right). \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
+	\label{fig:quotientopes}
+\end{figure}
 
@@
 
-In order to focus on the combinatorial structure of a rectangulation, we recall two equivalence relations between rectangulations~\cite{??}.
+In order to focus on the combinatorial structure of a rectangulation, we recall two equivalence relations between rectangulations (see for instance~\cite{ACFF24} and references therein).
 We say that a rectangle~$r$ is \defn{above} another rectangle~$s$ (and~$s$ is \defn{below}~$r$) if either the bottom edge of~$r$ lies in the same segment as the top edge of~$s$, or if~$r$ is above another rectangle~$t$ that is above~$s$.
@@
 However, if the wall slide changes the adjacency relation between the rectangles, then the resulting rectangulation is not strongly equivalent to $R$ anymore.
-This is illustrated in \cref{fig:rectequiv}.
-A \defn{diagonal rectangulation} is a rectangulation in which every rectangle intersects the top-left to bottom-right diagonal of the square, see \cref{fig:weakRectangulation}.
+A \defn{diagonal rectangulation} is a rectangulation in which every rectangle intersects the top-left to bottom-right diagonal of the square.
+Examples of rectangulations are given in \cref{fig:rectequiv}.
 It is simple to check that every weak rectangulation has a diagonal representative, hence weak rectangulations can also be thought of as diagonal rectangulations.
@@
 \begin{figure}
-	\caption{Weak and strong equivalence between rectangulations.}
+  \includegraphics[width=.7\textwidth]{rectangulations.pdf}
+	\caption{Three rectangulations. All three rectangulations are weakly equivalent, but only the first two are strongly equivalent. Only the third one is diagonal.}
 	\label{fig:rectequiv}
@@
 For~$n = 4$, the weak rectangulation lattice is represented in \cref{fig:weakRectangulationLattice} and the weak rectangulotope is represented in \cref{fig:weakRectangulotope}.
+
+The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams} in terms of arc diagrams, and studied more recently by E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
+Its classes are in bijection with strong rectangulations.
+The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by~$\SRP(n)$.
+For~$n = 4$, the strong rectangulation lattice is represented in \cref{fig:strongRectangulationLattice} and the strong rectangulotope is represented in \cref{fig:strongRectangulotope}.
 
@@
 \end{figure}
-
 \begin{figure}
 	\centerline{\includegraphics[scale=1.05]{weakRectangulotopeLabeled}}
-	\caption{The strong rectangulotope~$\WRP(4)$.}
+	\caption{The weak rectangulotope~$\WRP(4)$.}
         \label{fig:weakRectangulotope}
 \end{figure}
-
-The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams} in terms of arc diagrams, and studied more recently by E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
-Its classes are in bijection with strong rectangulations.
-The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by~$\SRP(n)$.
-For~$n = 4$, the strong rectangulation lattice is represented in \cref{fig:strongRectangulationLattice} and the strong rectangulotope is represented in \cref{fig:strongRectangulotope}.
-
 \begin{figure}
@@
 \end{figure}
-
 \begin{figure}
@@
 \item \defn{Simple flips} replace a vertical segment incident to exactly two rectangles by a horizontal segment, and vice-versa,
-\item \defn{Pivoting operations} transform a pair of adjacent rectangles whose union is not a rectangle, changing a left-right pair into an above-below pair, and vice-versa,
+\item \defn{Pivoting flips} transform a pair of adjacent rectangles whose union is not a rectangle, changing a left-right pair into an above-below pair, and vice-versa,
 \item \defn{Wall slide flips} are wall slides that modify the adjacency of the rectangles, removing exactly one adjacency and adding exactly one.
@@
 \begin{figure}
-	\caption{Three types of flips between rectangulations.}
+  \begin{center}
+  \begin{subfigure}{0.3\textwidth}
+    \begin{center}
+    \includegraphics[page=2, scale=.5]{flipGraph.pdf}
+    \caption{Simple flips.}
+    \end{center}
+    \end{subfigure}
+  \begin{subfigure}{0.3\textwidth}
+    \begin{center}
+    \includegraphics[page=1, scale=.5]{flipGraph.pdf}
+    \caption{Pivoting flips.}
+    \end{center}
+  \end{subfigure}
+    \begin{subfigure}{0.3\textwidth}
+    \begin{center}
+    \includegraphics[page=3, scale=.5]{flipGraph.pdf}
+    \caption{\label{fig:wallSlides}Wall slide flips.}
+    \end{center}
+    \end{subfigure}
+    \end{center}
+	\caption{The three types of flips between rectangulations. The flips are possible only if the grey areas do not intersect any segment of the rectangulation.}
 	\label{fig:flips}
@@
 
-% jean:{Here only give the definitions allowing to state the results.}
+\begin{figure}
+	\centerline{\includegraphics[width=.5\textwidth]{strongRectangulation} \qquad \includegraphics[width=.5\textwidth]{strongRectangulationTrees}}
+	\caption{A rectangulation $R$ (left) and its pair of binary trees $S(R)$ and $T(R)$ (right). Example from \cite{ACFF24}.}
+        % The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
+        \label{fig:strongRectangulation}
+\end{figure}
 
@@
 
-\begin{figure}
-	\centerline{\includegraphics[width=.5\textwidth]{strongRectangulation} \qquad \includegraphics[width=.5\textwidth]{strongRectangulationTrees}}
-	\caption{A (non-diagonal) rectangulation (left) and its pair of intertwining binary trees (right). Example from \cite{ACFF24}.}
-        % The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
-        \label{fig:strongRectangulation}
-\end{figure}
-
 Fix a rectangulation~$R$ of size~$n$, and consider the directed graph~$D(R)$ whose vertex set is the set of all vertices of~$R$, and whose edges are obtained as follows.
@@
 The \defn{source tree}~$S(R)$ (resp.~\defn{target tree}~$T(R)$) is the subgraph of~$D(R)$ induced by the sources~$\set{s(r)}{r \in [n]}$ (resp.~by the targets~$\set{t(r)}{r \in [n]}$).
-See \cref{fig:strongRectangulation}.
-
-\begin{lemma}
-The source tree~$S(R)$ (resp.~target tree~$T(R)$) is a binary tree, rooted at the bottom left (resp.~top right) corner of~$R$, and oriented from (resp.~towards) its root.
-\end{lemma}
+It is not difficult to check that the source tree~$S(R)$ (resp.~target tree~$T(R)$) is a binary tree, rooted at the bottom left (resp.~top right) corner of~$R$, and oriented from (resp.~towards) its root.
 
@@
 
-We complete each node of~$S(R)$ and~$T(R)$ with a vertical (resp.~horizontal) leaf if it has no vertical (resp.~horizontal) child, see \cref{fig:strongRectangulation}.
-If the rectangulation $R$ is diagonal, we retrieve the well-studied pair of twin binary trees~\cite{MR1417289,MR2914637}.
+We complete each node of~$S(R)$ and~$T(R)$ with a vertical (resp.~horizontal) leaf if it has no vertical (resp.~horizontal) child.
+An example is given in~\cref{fig:strongRectangulation}.
+If the rectangulation $R$ is diagonal, we retrieve the well-studied pair of twin binary trees~\cite{MR1417289,MR2914637}, see~\cref{fig:weakRectangulation}.
 Recall that the \defn{inorder labeling} of a binary tree~$T$ is the labeling of the nodes of~$T$ such that the label of each node~$t$ of~$T$ is larger than all labels in the left subtree of~$t$ and larger than all labels in the right subtree of~$t$.
-
-\begin{lemma}
+%\begin{lemma}
 For any rectangle~$r$ of~$R$, the inorder label of the source~$s(r)$ in~$S(R)$ coincides with the inorder label of the target~$t(r)$ in~$T(R)$.
-\end{lemma}
-
+%\end{lemma}
 This enables to unambiguously label the rectangles of~$R$ by the inorder.
@@
 
-\Jean{We need to make it clear whether $\WRP (n)$ is the proposed realization or just the combinatorial type.
-I was thinking the latter.
-Then we can change the sentence ``The weak rectangulotope is realized by the polytope~$\WRP (n)$...'' by ``The weak rectangulotope $\WRP (n)$ is realized by:~''.
-Otherwise, refrain from using $\WRP (n)$ before here.}
-
 \begin{theorem}
   \label{thm:weakRectangulotope}
-  The weak rectangulotope is realized by the polytope~$\WRP (n) \subset \R^n\!$ defined equivalently~as
+  % Jean: I chose to denote the *abstract* polytope by WRP(n), hence changed the next sentence accordingly.
+%The weak rectangulotope is realized by the polytope~$\WRP (n) \subset \R^n\!$ defined equivalently~as
+The weak rectangulotope $\WRP (n)$ is realized by:
   \begin{enumerate}[(i)]
   \item the convex hull of the points
-  $\sum_{i\in [n]} (\loday{w}^R_i - \antiloday{w}^R_i)\cdot \b{e}_i$ for all weak rectangulations $R$ of size~$n$,
-  with
-  \[
-%  \begin{split}
-%    \loday{w}^R_i \eqdef & h(T, i)\cdot v(T,i) \\
-%    \antiloday{w}^R_i \eqdef & h(S, i)\cdot v(S,i),
-%  \end{split}
+    \[
+    \sum_{i\in [n]} (\loday{w}^R_i - \antiloday{w}^R_i)\cdot \b{e}_i
+    \]
+    for all weak rectangulations $R$ of size~$n$, with
+  \[
     \loday{w}^R_i \eqdef h^T_i\cdot v^T_i
@@
   where~$S$ and~$T$ are the source and target trees of the rectangulation~$R$, and $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
-  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by \linebreak $\sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{I \not\subseteq S \text{ and } I \not\subseteq [n] \ssm S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$.
+\item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by
+  \[
+  \sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{I \not\subseteq S \text{ and } I \not\subseteq [n] \ssm S}
+  \]
+  for all subsets~$\varnothing \ne S \subsetneq [n]$.
   \end{enumerate}
@@
 
+\begin{comment}
 \begin{example}
@@
 \end{example}
+\end{comment}
 
@@
 \label{thm:strongRectangulotope}
-  The $(n-1)$-dimensional strong rectangulotope is realized by the polytope~$\SRP (n)$ defined equivalently as
+% The strong rectangulotope is realized by the polytope~$\SRP (n)$ defined equivalently as
+The strong rectangulotope~$\SRP (n)$ is realized by:
   \begin{enumerate}[(i)]
@@
   \[
-%  \begin{split}
-%    \yin{w}_{i,j} \eqdef & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot \llbracket \neg i\backslash j \rrbracket \\% \text { if not } i\backslash j, \text{and } 0\text{ otherwise.} \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
-%    \yang{w}_{i,j} \eqdef & v(S, i) \cdot ch (S, T, i, j)\cdot v(T, j) \cdot \llbracket \neg j\backslash i \rrbracket,
-%  \end{split}
     \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \neg \, i \, \backslash \, j \rrbracket
@@
   \end{itemize}
-  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by \linebreak $\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
+\item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by
+  \[
+  \sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}
+  \]
+  for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
   \end{enumerate}
@@
 The \defn{inversion set} of~$\sigma \in \f{S}_n$ is~$\inv(\sigma) \eqdef \set{(\sigma_i, \sigma_j)}{1 \le i < j \le n \text{ and } \sigma_i > \sigma_j}$.
-The \defn{weak Bruhat order}\footnote{The weak Bruhat order is usually just referred to as the weak order~\cite{Humphrey, BjornerBrenti}.
+The \defn{weak Bruhat order}\footnote{The weak Bruhat order is usually just referred to as the weak order~\cite{MR1066460, MR2133266}.
 In this paper, we prefer to use weak Bruhat order to avoid any confusion with the weak poset defined on weak rectangulations.} is the lattice on the permutations of~$\f{S}_n$ defined by the inclusion of their inversion sets.
@@
 
-For instance, the sets of linear extensions of binary trees (labeled in inorder and oriented toward their roots) are the classes of the \defn{sylvester congruence}~\cite{HivertNovelliThibon-binarySearchTrees}.
+For instance, the sets of linear extensions of binary trees (labeled in inorder and oriented toward their roots) are the classes of the \defn{sylvester congruence}~\cite{MR2142078}.
 See \cref{fig:sylvesterCongruence}\,(middle).
@@
 
-It was shown in~\cite[Lem.~23]{AlbertinPilaudRitter} that the ray of the braid arrangement corresponding to a nonempty proper subset~$\varnothing \ne J \subsetneq [n]$ is preserved in the quotient fan~$\c{F}_\equiv$ if and only if the arc ideal~$\c{A}_\equiv$ contains the $|J|-1$ down arcs joining two consecutive elements of~$J$ and the~$n-|J|-1$ up arcs joining two consecutive elements of~$[n] \ssm J$.
+It was shown in~\cite[Lem.~23]{MR4328906} that the ray of the braid arrangement corresponding to a nonempty proper subset~$\varnothing \ne J \subsetneq [n]$ is preserved in the quotient fan~$\c{F}_\equiv$ if and only if the arc ideal~$\c{A}_\equiv$ contains the $|J|-1$ down arcs joining two consecutive elements of~$J$ and the~$n-|J|-1$ up arcs joining two consecutive elements of~$[n] \ssm J$.
 An immediate consequence is that all $2^n-2$ rays of the braid arrangement are preserved in the quotient fan~$\c{F}_\equiv$ if and only if the arc ideal~$\c{A}_\equiv$ contains all up arcs and all down arcs (in other words, if it is above the weak rectangulation congruence in the lattice of congruences of the weak Bruhat order).
@@
 \label{subsec:quotientopes}
-
-\begin{figure}
-	\capstart
-	\centerline{\includegraphics[scale=.6]{permutahedronLeft4} \qquad \raisebox{-.4cm}{\includegraphics[scale=.6]{associahedron4}}}
-	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$ (right). \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
-	\label{fig:quotientopes}
-\end{figure}
 
@@
 The \defn{shard polytope} of~$\alpha$ is the convex hull~$\SP(\alpha)$ of the characteristic vectors of all $\alpha$-alternating matchings.
+We refer to Figure~\ref{fig:shardPolytopes} for an illustration of this definition.
 It was shown in~\cite{MR4584712} that any Minkowski sum of positive scalings of the shard polytopes~$\SP(\alpha)$ for all arcs~$\alpha$ in the arc ideal~$\c{A}_\equiv$ is a quotientope for~$\equiv$.
@@
 Adapted from~\cite[Fig.~10]{MR4584712}.}
-	\label{fig:shardPolytiopes}
+	\label{fig:shardPolytopes}
 \end{figure}
@@
 \]
+
+We recall that \emph{Baxter permutations} are permutations avoiding the vincular pattern $2\underline{41}3$ and $3\underline{14}2$, where the underlining indicates that the symbols must appear contiguously in the permutation.
+Similarly, \defn{twisted Baxter permutations} are permutations avoiding the patterns $2\underline{41}3$ and $3\underline{41}2$, and \defn{co-twisted Baxter permutations} are permutations avoiding the patterns $2\underline{14}3$ and $3\underline{14}2$.
+These permutation families are central in the theory of rectangulations and appear in many other contexts~\cite{MR2763051}.
 
@@
 
-Therefore, the quotientope is realized by a Minkowski sum of the shard polytopes of the up arcs and the down arcs.
-These quotientopes are very basic ones.
+From the results of section~\ref{subsec:quotientopes}, the quotientope is realized by a Minkowski sum of the shard polytopes of the up arcs and the down arcs.
+These quotientopes are simplices and anti-simplices (see Figure~\ref{fig:shardPolytopes}, left).
 
@@
 The description of the weak rectangulotope~$\WRP(n)$ as a Minkowski sum in \cref{lem:weakMinkowski} allows us to easily compute the corresponding submodular function.
-\vincent{Say here that all rays are preserved}
+The arc ideal of the weak rectangulation congruence contains exactly all up arcs and all down arcs, hence from the result of~\cite{MR4328906} mentioned in section~\ref{subsec:quotientFans}, all $2^n-2$ rays of the braid fan are preserved in the fan of the weak rectangulotope.
+This implies that every nonempty proper subset will define a facet of the weak rectangulotope, and the description of the weak rectangulotope by its submodular function is actually an irredundant facet description.
 
@@
   Let $\pi \in \f{S}_n$ and $I$ be an interval of~$[n]$ of size at least~$2$.
-  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~of the translated down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\pi^{-1}$ is $\b{e}_i$, where $i \eqdef \arg\max_{j\in I} \pi^{-1}(j)$ (resp.~where~$i \eqdef \arg\min_{j\in I} \pi^{-1}(j)$).
+  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~of the translated down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\pi^{-1}$ is $\b{e}_i$ (resp.~$-\b{e}_i$), where $i$ is the maximum (resp.
+the minimum) of $\pi^{-1}$ in $I$.
 \end{lemma}
@@
 
-\begin{lemma}
+The proof of the following lemma actually follows from Loday's realization in Theorem~\ref{thm:Loday}.
+For the sake of completeness, and because we will reuse the same reasoning in the next section, we give an elementary proof.
+
+\begin{lemma}
+  \label{lem:weakCoord}
   Let $R$ be a rectangulation of size~$n$, with source and target trees~$S$ and~$T$, and let $\b{p}(R)$ the vertex of $\WRP(n)$ associated with $R$.
@@
 \end{lemma}
+\begin{proof}
+  Let $\pi$ be any permutation in $\mathcal{L}(P_w(R))$.
+  From \cref{lem:lodaymax}, for an $i\in [n]$, a shard polytope $\SP(\upArc{I})$ contributes to $\b{e}_i$ only if $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
+  We claim that the number of intervals $I$ satisfying this condition is the given product.
+
+  We first show that the number of choices for the left endpoint of $I$ is equal to the number $h^T_i$ of leaves in the horizontal subtree of $i$ in $T(R)$.
+  Each such leaf determines a left endpoint for $I$, defined as the node of the tree that follows it in an inorder traversal.
+  Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
+  Since the weak poset extends the order induced by $T(R)$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
+  Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
+  Now consider the leaf of $T(R)$ that is immediately left of the leftmost leaf of the horizontal subtree of $i$.
+  This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T(R)$.
+  Therefore, we must have $\ell\succ_s i$, and $\pi^{-1}(\ell) > \pi^{-1}(i)$, and we cannot extend $I$ past this index on the left.
+  This shows that there are exactly $h^T_i$ choices, and not more, for the choice of the left endpoint of $I$.
+  A symmetric argument shows that there are exactly $v^{T}_i$ choices for the right endpoint of $I$.
+\end{proof}
+
+A symmetric analogue of Lemma~\ref{lem:weakCoord} holds for down shard polytopes, where the product contributes to $-\b{e}_i$ instead of $\b{e}_i$.
+This proves Theorem~\ref{thm:weakRectangulotope}.
 
@@
 The strong poset is the transitive closure of the union of the two relations $\tri$ and $\btri$. 
-The strong poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
-
-We will use the following statement.
+The strong poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$. The following statement will be used later.
+
 \begin{lemma}
@@
   It remains to observe that the adjacency poset of $R$ is an extension of the weak poset of $R$.
-  For this, we consider the diagonal representative $D$ of the weak equivalence class of $R$, and recall that the weak poset of $R$ is the adjacency poset of $D$.
+  For this, we consider a diagonal representative $D$ of the weak equivalence class of $R$, and recall that the weak poset of $R$ is the adjacency poset of $D$.
   By definition, $R$ can be obtained from $D$ by a sequence of wall slides.
   Each wall slide in this sequence consists of either a horizontal segment being pulled down past another horizontal segment on its right, or a vertical segment being pushed right past another vertical segment below it.
-  Note that each of these we can only add new adjacencies, hence the adjacency poset of $R$ is always an extension of that of $D$. 
+  (These are the flips described in~\cref{fig:wallSlides}, performed in bottom to top order on the figure.) 
+  Note that each of these we can only add new pairs in the adjacency poset, hence the adjacency poset of $R$ is always an extension of that of $D$. 
 \end{proof}
@@
 \]
+
+N.
+Reading~\cite{MR2864445} introduced \defn{2-clumped permutations}, defined as the permutations avoiding the vincular patterns $24\underline{51}3$, $42\underline{51}3$, $3\underline{51}24$, and $3\underline{51}42$.
+The \defn{co-2-clumped permutations} are defined similarly, as permutations avoiding the mirror images of these four patterns.
+He showed that 2-clumped permutations were in bijection with strong rectangulations.
 
@@
 
+The following lemmas are direct applications of the results from~\cite{MR4584712} given in section~\ref{subsec:quotientopes}.
+
 \begin{lemma}
@@
 
+We refer to Figure~\ref{fig:shardPolytopes} for examples of yin and yang shard polytopes.
 %Note that yin shard polytopes are translates of the up shard polytopes when $|J|=1$, and of the down shard polytopes when $|I|=1$.
 %Similarly, up and down shard polytopes are translates of some yang shard polytopes.
-
-The following is a direct application of the main result of A.~Padrol, V.~Pilaud, and J.~Ritter~\cite{MR4584712}.
 
@@
 \cref{lem:strongMinkowski} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.
+It can be checked that the equivalence classes of these congruences are counted by \defn{semi-Baxter permutations}, defined as permutations avoiding $2\underline{41}3$, hence only one of the two forbidden Baxter patterns.
+These equivalence classes are in turn in bijection with rectangulations that avoid one of the two patterns defining diagonal rectangulations, see~\cite{ACFF24}.
 
@@
 
-The submodular function corresponding to our realization of $\SRP(n)$ can be obtained as before by summing the functions defining the yin and yang shard polytopes in \cref{lem:yinsp,lem:yangsp}.
+From the results of section~\ref{subsec:quotientopes}, the submodular function corresponding to our realization of $\SRP(n)$ can be obtained as before by summing the functions defining the yin and yang shard polytopes in \cref{lem:yinsp,lem:yangsp}.
+Note again that all rays of the braid fan are preserved in the fan of the strong rectangulotope, hence our characterization of the strong rectangulotope is an irredundant facet description.
 
@@
   \]
-  if $i\succ_s j$, and $0$ otherwise.
+  if $i\succ_s j$, and $0$ otherwise, where
+  \begin{itemize}
+  \item $h^T_i$ denotes the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
+  \item $cv^{T,S}_{i,j}$ denotes the number of common leaves of the vertical subtree of $i$ in $T$ and the vertical subtree of $j$ in $S$.
+  \end{itemize}  
 \end{lemma}
 \begin{proof}
-  Let $\pi$ be any permutation corresponding to $R$, hence an element of $\mathcal{L}(P_s(R))$.
+  The proof is not much more involved that of Lemma~\ref{lem:weakCoord}.
+  Let $\pi$ be any permutation in $\mathcal{L}(P_s(R))$.
   From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc $A$ whose shard polytope $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ only if $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.
   Provided this holds, the set of yin arcs $A$ such that $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that $i$ is the index of the maximum of $\pi^{-1}$ in $I$, and $j$ is the index of the maximum of $\pi^{-1}$ in $J$.
-  We claim that the number of choices of the three endpoints defining such a pair of intervals $I$ and $J$ is the product of the
-  number $h^T_i$ of leaves of the horizontal subtree of $T(R)$ rooted at $i$, the number $cv^{T,S}_{i,j}$ of common leaves to the vertical subtrees of $T(R)$ and $S(R)$ rooted
-  at $i$ and $j$, respectively, and the number $h^S_j$ of leaves of the horizontal subtree of $S(R)$ rooted at $j$.
+  We claim that the number of choices of the three endpoints defining such a pair of intervals $I$ and $J$ is the given product.
 
@@
   Since from Lemma~\ref{lem:sextw} the strong poset extends the order induced by $T(R)$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
-  Therefore, each such choice of left endpoint for $I$ preserves the fact that $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
+  Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
   Now consider the leaf of $T(R)$ that is immediately left of the leftmost leaf of the horizontal subtree of $i$.
@@
 
+This concludes the proof of Theorem~\ref{thm:strongRectangulotope}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{comment}
 
@@
 \appendix
-\section{Material}
+\section{Unused Material}
 
@@
 
-\begin{comment}
 \newpage

### 2024-04-19 draft/rectangulotopes.tex
typos

@@
   \label{lem:weakCoord}
-  Let $R$ be a rectangulation of size~$n$, with source and target trees~$S$ and~$T$, and let $\b{p}(R)$ the vertex of $\WRP(n)$ associated with $R$.
+  Let $R$ be a rectangulation of size~$n$, with target tree~$T$, and let $\b{p}(R)$ the vertex of $\WRP(n)$ associated with $R$.
   Given $i\in [n]$, the number of up arcs~$\upArc{I}$ such that $\b{e}_i$ is the extremal vertex of $\SP(\upArc{I})$ contributing to~$\b{p}(R)$ is
@@
 
-  We first show that the number of choices for the left endpoint of $I$ is equal to the number $h^T_i$ of leaves in the horizontal subtree of $i$ in $T(R)$.
+  We first show that the number of choices for the left endpoint of $I$ is equal to the number $h^T_i$ of leaves in the horizontal subtree of $i$ in $T$.
   Each such leaf determines a left endpoint for $I$, defined as the node of the tree that follows it in an inorder traversal.
   Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
-  Since the weak poset extends the order induced by $T(R)$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
+  Since the weak poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
   Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
-  Now consider the leaf of $T(R)$ that is immediately left of the leftmost leaf of the horizontal subtree of $i$.
-  This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T(R)$.
+  Now consider the leaf of $T$ that is immediately left of the leftmost leaf of the horizontal subtree of $i$.
+  This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T$.
   Therefore, we must have $\ell\succ_s i$, and $\pi^{-1}(\ell) > \pi^{-1}(i)$, and we cannot extend $I$ past this index on the left.
@@
 
-A symmetric analogue of Lemma~\ref{lem:weakCoord} holds for down shard polytopes, where the product contributes to $-\b{e}_i$ instead of $\b{e}_i$.
+A symmetric analogue of Lemma~\ref{lem:weakCoord} holds for down shard polytopes, where the product contributes to $-\b{e}_i$ instead of $\b{e}_i$, and we consider the tree $S(R)$ instead of the tree $T(R)$.
 This proves Theorem~\ref{thm:weakRectangulotope}.
@@
 
-  We first show that the number of choices for the left endpoint of $I$ is equal to the number $h^T_i$ of leaves in the horizontal subtree of $i$ in $T(R)$.
+  We first show that the number of choices for the left endpoint of $I$ is equal to the number $h^T_i$ of leaves in the horizontal subtree of $i$ in $T$.
   Each such leaf determines a left endpoint for $I$, defined as the node of the tree that follows it in an inorder traversal.
   Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
-  Since from Lemma~\ref{lem:sextw} the strong poset extends the order induced by $T(R)$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
+  Since from Lemma~\ref{lem:sextw} the strong poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
   Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
-  Now consider the leaf of $T(R)$ that is immediately left of the leftmost leaf of the horizontal subtree of $i$.
-  This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T(R)$.
+  Now consider the leaf of $T$ that is immediately left of the leftmost leaf of the horizontal subtree of $i$.
+  This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T$.
   Therefore, we must have $\ell\succ_s i$, and $\pi^{-1}(\ell) > \pi^{-1}(i)$, and we cannot extend $I$ past this index on the left.
@@
   The rest of the formula comes by symmetric arguments.
-  First, the right endpoint of $I$ can only be chosen among the leaves of the vertical subtree of $T(R)$ rooted at $i$.
-  Then the same reasoning holds for the two endpoints of $J$, replacing $i$ by $j$ and $T(R)$ by $S(R)$.
+  First, the right endpoint of $I$ can only be chosen among the leaves of the vertical subtree of $T$ rooted at $i$.
+  Then the same reasoning holds for the two endpoints of $J$, replacing $i$ by $j$ and $T$ by $S$.
   Finally, since the right endpoint of $I$ and the left endpoint of $J$ must be consecutive, the selected leaves must correspond to the same pair of successive indices of $[n]$.

### 2024-04-19 draft/rectangulotopes.tex
minor change

--- 2024-04-19 draft/rectangulotopes.tex
@@
 \newcommand{\verticalPattern}{\smash{\raisebox{-.25cm}{\includegraphics[scale=.5]{verticalPattern}}}}
-% orders
-\newcommand{\tri}{\lhd}
-\newcommand{\btri}{\LHD}
 
@@
 See \cref{fig:quotientopes}\,(left) for an illustration when~$n = 4$.
-The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra}~\cite{MR2520477,Postnikov,MR4064768,MR4651496} (also called \defn{generalized permutahedra}, or \defn{polymatroids}).
+The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra}~\cite{MR2520477,Postnikov,MR4064768,MR4651496} (also call \defn{generalized permutahedra}, or \defn{polymatroids}).
 
@@
 \noindent
-As illustrated on the right, this can also been visualized by representing the table~$(\sigma_j,j)$ of the permutation~$\sigma$, drawing the segments corresponding to the descents of~$\sigma$, and letting all points fall on the horizontal axis, allowing the segments to bend but not to cross each other nor to pass through a point.
+As illustrated on the right with~$\sigma = 2531746$, this can also been visualized by representing the table~$(\sigma_j,j)$ of the permutation~$\sigma$, drawing the segments corresponding to the descents of~$\sigma$, and letting all points fall on the horizontal axis, allowing the segments to bend but not to cross each other nor to pass through a point.
 The single arcs correspond to permutations with a single descent, that is, to join irreducible permutations of the weak Bruhat order.
@@
 
-A set function $f:2^{[n]}\mapsto \R$ is said to be \defn{submodular} when $f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$.
 By definition, any quotientope is a deformed permutahedron, and can thus be written as
@@
   \]
-where~$f$ is a submodular set function.
-We will need the following simple result: Given a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by submodular set functions $f_1,f_2,\ldots ,f_k$, their Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f \eqdef \sum_i f_i$.
+where~$f$ is a \defn{submodular} set function $f:2^{[n]}\mapsto \R$ (\ie such that~$f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$).
+Recall that the submodular function of a Minkowski sum of deformed permutahedra is just the sum of the submodular functions of the Minkowski summands.
+
+\begin{lemma}
+  \label{lem:submodsum}
+  \vincent{I would not make this one so visible in a proper thm environment.}
+  Consider a collection $\polytope{P}_1, \polytope{P}_2,\ldots ,\polytope{P}_k$ of deformed permutahedra defined by the submodular functions $f_1,f_2,\ldots ,f_k$.
+Then the Minkowski sum $\sum_i \polytope{P}_i$ is defined by the submodular function $f \eqdef \sum_i f_i$.
+\end{lemma}
 
@@
 %%%%%%%%%%%%%
-\subsection{The weak poset}
-\label{subsec:weakPoset}
-
-Given a rectangulation~$R$, let us consider the source and target trees~$S(R)$ and~$T(R)$, and orient the horizontal edges from left to right, and the vertical edges from bottom to top.
-This defines two partial orders on $[n]$.
-The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of these two partial orders defined by~$S(R)$ and~$T(R)$.
-Equivalently, the weak poset can be defined by considering the adjacency graph of the rectangles in a diagonal representative of the weak equivalence class of $R$, orienting its edges from left to right and from bottom to top, and taking the transitive closure of this directed acyclic graph.
+
+\subsection{The weak rectangulation congruence}
+\label{subsec:weakRectangulationCongruence}
+
+Given a rectangulation~$R$, the source and target trees~$S(R)$ and~$T(R)$ define two partial orders on $[n]$ (the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top).
+The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of the two partial orders defined by the intertwining binary trees~$S(R)$ and~$T(R)$.
 These posets were characterized (and called Baxter posets) by E.~Meehan~\cite{MR4014603}.
 The weak poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
-
-%%%%%%%%%%%%%
-\subsection{The weak rectangulation congruence}
-\label{subsec:weakRectangulationCongruence}
-
 We define the set~$\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
@@
 \begin{lemma}
-The weak rectangulation congruence $\weakeq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that do not cross the horizontal line.
+The weak rectangulation congruence $
+\weakeq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that do not cross the horizontal line.
 \end{lemma}
@@
 %%%%%%%%%%%%%
-\subsection{The strong poset}
-\label{subsec:strongPoset}
-
-Given a rectangulation $R$, we define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of $R$ by considering two relations on $[n]$~\cite{ACFF24}:
-\begin{enumerate}
-\item The \defn{adjacency poset} $([n],\tri)$ of $R$, defined by considering the adjacency graph of the rectangles of $R$ (identified by their inorder labels in $[n]$), orienting its edges from left to right and from bottom to top, and taking the transitive closure of this directed acyclic graph.
-\item The \defn{special relation} $([n],\btri)$ of $R$, in which $a\btri b$ if the two rectangles with inorder labels $a$ and $b$ are such that either (i) the right side of $b$ and the left side of $a$ lie on the same vertical segment, and the bottom-right corner of $b$ is above the top-left corner of $a$, or (ii) the top side of $b$ and the bottom side of rectangle $a$ lie on the same horizontal segment, and the top-left corner of $b$ lies on the right of the bottom-right corner of  $a$.
-\end{enumerate}
-The strong poset is the transitive closure of the union of the two relations $\tri$ and $\btri$. 
-The strong poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
-
-We will use the following statement.
-\begin{lemma}
-  \label{lem:sextw}
-  The strong poset $P_s(R)$ of a rectangulation $R$ is an extension of its weak poset $P_w(R)$.
-\end{lemma}
-\begin{proof}
-  Clearly, the strong poset is an extension of the adjacency poset $([n],\tri)$ of $R$, since it is obtained by adding pairs from the special relation.
-  It remains to observe that the adjacency poset of $R$ is an extension of the weak poset of $R$.
-  For this, we consider the diagonal representative $D$ of the weak equivalence class of $R$, and recall that the weak poset of $R$ is the adjacency poset of $D$.
-  By definition, $R$ can be obtained from $D$ by a sequence of wall slides.
-  Each wall slide in this sequence consists of either a horizontal segment being pulled down past another horizontal segment on its right, or a vertical segment being pushed right past another vertical segment below it.
-  Note that each of these we can only add new adjacencies, hence the adjacency poset of $R$ is always an extension of that of $D$. 
-\end{proof}
-
-%%%%%%%%%%%%%
+
 \subsection{The strong rectangulation congruence}
@@
 
-Let $\mathcal{L}(P_s)$ be the set of permutations of $\f{S}_n$ corresponding to linear extensions of the strong poset $P_s$:
+We define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of a rectangulation $R$, and $\mathcal{L}(P_s)$ as the set of permutations of $\f{S}_n$ corresponding to linear extensions of $P_s$:
 \[
@@
 \begin{theorem}[\cite{MR2864445,ACFF24}]
-  The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, are equivalence classes of a congruence relation $\strongeq$ on the weak Bruhat order on $\f{S}_n$.
+  The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, 
+  are equivalence classes of a congruence relation $\strongeq$ on the weak Bruhat order on $\f{S}_n$.
     Furthermore:
@@
 \begin{proof}
-  Let $\pi$ be any permutation corresponding to $R$, hence an element of $\mathcal{L}(P_s(R))$.
-  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc $A$ whose shard polytope $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ only if $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.
+  \vincent{I still need to work out this proof...}
+  Let $\pi$ be the 2-clumped permutation corresponding to $R$, hence the minimal element of $\mathcal{L}(P_s(R))$ in the weak Bruhat order.
+  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc $A$ whose shard polytope $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ only if
+  $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.
   Provided this holds, the set of yin arcs $A$ such that $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that $i$ is the index of the maximum of $\pi^{-1}$ in $I$, and $j$ is the index of the maximum of $\pi^{-1}$ in $J$.
+  %$i = \arg\max_{k\in I} \pi^{-1}(k)$ and $j = \arg\min_{k\in J} \pi^{-1}(k)$.
   We claim that the number of choices of the three endpoints defining such a pair of intervals $I$ and $J$ is the product of the
-  number $h^T_i$ of leaves of the horizontal subtree of $T(R)$ rooted at $i$, the number $cv^{T,S}_{i,j}$ of common leaves to the vertical subtrees of $T(R)$ and $S(R)$ rooted
-  at $i$ and $j$, respectively, and the number $h^S_j$ of leaves of the horizontal subtree of $S(R)$ rooted at $j$.
-
-  We first show that the number of choices for the left endpoint of $I$ is equal to the number $h^T_i$ of leaves in the horizontal subtree of $i$ in $T(R)$.
-  Each such leaf determines a left endpoint for $I$, defined as the node of the tree that follows it in an inorder traversal.
-  Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
-  Since from Lemma~\ref{lem:sextw} the strong poset extends the order induced by $T(R)$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
-  Therefore, each such choice of left endpoint for $I$ preserves the fact that $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
-  Now consider the leaf of $T(R)$ that is immediately left of the leftmost leaf of the horizontal subtree of $i$.
-  This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T(R)$.
-  Therefore, we must have $\ell\succ_s i$, and $\pi^{-1}(\ell) > \pi^{-1}(i)$, and we cannot extend $I$ past this index on the left.
-  This shows that there are exactly $h^T_i$ choices, and not more, for the choice of the left endpoint of $I$.
-
-  The rest of the formula comes by symmetric arguments.
-  First, the right endpoint of $I$ can only be chosen among the leaves of the vertical subtree of $T(R)$ rooted at $i$.
-  Then the same reasoning holds for the two endpoints of $J$, replacing $i$ by $j$ and $T(R)$ by $S(R)$.
-  Finally, since the right endpoint of $I$ and the left endpoint of $J$ must be consecutive, the selected leaves must correspond to the same pair of successive indices of $[n]$.
-  These leaves are exactly the common leaves defined previously.
+  number of leaves of the subtree of $T(R)$ rooted at $i$, the number of common leaves to the vertical subtrees of $T(R)$ and $S(R)$ rooted
+  at $i$ and $j$, respectively, and the number of leaves of the subtree of $S(R)$ rooted at $j$.
 \end{proof}

### 2024-04-20 draft/rectangulotopes.tex
short proofs of arc ideal Lemmas

@@
   \begin{enumerate}[(i)]
-  \item convex hull of the points
+  \item the convex hull of the points
   \[
@@
 
-We recall that \emph{Baxter permutations} are permutations avoiding the vincular pattern $2\underline{41}3$ and $3\underline{14}2$, where the underlining indicates that the symbols must appear contiguously in the permutation.
+We recall that \defn{Baxter permutations} are permutations avoiding the vincular pattern $2\underline{41}3$ and $3\underline{14}2$, where the underlining indicates that the symbols must appear contiguously in the permutation.
 Similarly, \defn{twisted Baxter permutations} are permutations avoiding the patterns $2\underline{41}3$ and $3\underline{41}2$, and \defn{co-twisted Baxter permutations} are permutations avoiding the patterns $2\underline{14}3$ and $3\underline{14}2$.
@@
 \end{lemma}
+\begin{proof}
+  We consider the forbidden vincular patterns $2\underline{41}3$ and $3\underline{41}2$ that define the minimal elements of each congruence class of $\weakeq$.
+  We observe that every occurrence of a pattern in a permutation corresponds to an occurrence of an arc that cross the horizontal line at least once in the arc diagram of the permutation.
+  The twisted Baxter permutations are therefore the permutations whose arc diagrams are composed only of arcs that do not cross the horizontal line.
+\end{proof}
 
@@
 \end{proof}
+
+This proves the second part of Theorem~\ref{thm:weakRectangulotope}.
 
@@
 A symmetric analogue of Lemma~\ref{lem:weakCoord} holds for down shard polytopes, where the product contributes to $-\b{e}_i$ instead of $\b{e}_i$, and we consider the tree $S(R)$ instead of the tree $T(R)$.
-This proves Theorem~\ref{thm:weakRectangulotope}.
+This concludes the proof of Theorem~\ref{thm:weakRectangulotope}.
 
@@
 \end{lemma}
+\begin{proof}
+  We consider the forbidden vincular patterns $24\underline{51}3$, $42\underline{51}3$, $3\underline{51}24$, and $3\underline{51}42$ that define the 2-clumped permutations, the minimal elements of each congruence class of $\strongeq$.
+  We observe that every occurrence of a pattern in a permutation corresponds to an occurrence of an arc that cross the horizontal line at least twice in the arc diagram of the permutation.
+  The 2-clumped permutations are therefore the permutations whose arc diagrams are composed only of arcs that cross the horizontal line at most once.
+\end{proof}
 
@@
   \[
-  f(S) \eqdef \# \bigset{ I, J }{ I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S } +
+  f_{\strongeq}(S) \eqdef \# \bigset{ I, J }{ I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S } +
   \#\bigset{ I, J }{ I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S },
@@
 \end{proof}
+
+This proves the second part of Theorem~\ref{thm:strongRectangulotope}.
 

### 2024-04-21 draft/rectangulotopes.tex
typo
minor changes

--- 2024-04-19 draft/rectangulotopes.tex
@@
 	\capstart
-	\centerline{\includegraphics[scale=.6]{permutahedronLeft4} \qquad \raisebox{-.4cm}{\includegraphics[scale=.6]{associahedron4}}}
+	\centerline{\includegraphics[scale=.75]{permutahedronLeft4} \quad \raisebox{-.4cm}{\includegraphics[scale=.75]{associahedron4}}}
 	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$ (right). \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
@@
 \end{figure}
+
+%%%%%%%%%%%%%
 
@@
 \begin{figure}
-  \includegraphics[width=.7\textwidth]{rectangulations.pdf}
+  \includegraphics[width=.9\textwidth]{rectangulations.pdf}
 	\caption{Three rectangulations. All three rectangulations are weakly equivalent, but only the first two are strongly equivalent. Only the third one is diagonal.}
@@
 \end{figure}
+\vincent{Add one which is not weakly equivalent?}
+
+%%%%%%%%%%%%%
 
@@
 
+\begin{figure}
+	\centerline{\includegraphics[width=.4\textwidth]{strongRectangulation} \qquad \includegraphics[width=.4\textwidth]{strongRectangulationTrees}}
+	\caption{A rectangulation $R$ (left) and its pair of binary trees $S(R)$ and $T(R)$ (right). Example from \cite{ACFF24}.}
+        % The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
+        \label{fig:strongRectangulation}
+\end{figure}
+
+\begin{figure}
+	\centerline{\includegraphics[width=.4\textwidth]{weakRectangulation} \qquad \includegraphics[width=.4\textwidth]{weakRectangulationTrees}}
+	\caption{A diagonal rectangulation (left) and its pair of twin binary trees (right). Example from \cite{ACFF24}.}
+	% The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.
+	\label{fig:weakRectangulation}
+\end{figure}
+
 Rectangulotopes enrich the family of natural quotientopes, alongside associahedra, permutreehedra~\cite{MR3856522}, certain brick polytopes~\cite{PilaudSantos-brickPolytopes, PilaudStump-brickPolytopes, Pilaud-brickAlgebra}, and certain graphical zonotopes~\cite{Pilaud-brickAlgebra, Pilaud-acyclicReorientationLattices}, providing instructive examples of application of the theory of lattice congruences to concrete combinatorial objects.
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Source and target trees of a rectangulation}
 \label{subsec:sourceTargetTrees}
-
-\begin{figure}
-	\centerline{\includegraphics[width=.5\textwidth]{strongRectangulation} \qquad \includegraphics[width=.5\textwidth]{strongRectangulationTrees}}
-	\caption{A rectangulation $R$ (left) and its pair of binary trees $S(R)$ and $T(R)$ (right). Example from \cite{ACFF24}.}
-        % The corresponding $2$-clumped permutation is $[7,5,14,8,1,6,15,11,4,10,16,2,9,13,3,12]$ and its inverse is $[5, 12, 15, 9, 2, 6, 1, 4, 13, 10, 8, 16, 14, 3, 7, 11]$.
-        \label{fig:strongRectangulation}
-\end{figure}
-
-\begin{figure}
-	\centerline{\includegraphics[width=.5\textwidth]{weakRectangulation} \qquad \includegraphics[width=.5\textwidth]{weakRectangulationTrees}}
-	\caption{A diagonal rectangulation (left) and its pair of twin binary trees (right). Example from \cite{ACFF24}.}
-	% The corresponding twisted Baxter permutation is $[7, 5, 1, 14, 8, 6, 4, 2, 11, 10, 9, 3, 15, 16, 13, 12]$ and its inverse is $[3, 8, 12, 7, 2, 6, 1, 5, 11, 10, 9, 16, 15, 4, 13, 14]$.
-	\label{fig:weakRectangulation}
-\end{figure}
 
@@
 Hence, the bottom left corner of~$r$ is its source~$s(r)$, while its top right corner is its target~$t(r)$.
-Note that since~$R$ is generic, the sources~$\set{s(r)}{r \in [n]}$ and the targets~$\set{t(r)}{r \in [n]}$ are disjoint.
-The \defn{source tree}~$S(R)$ (resp.~\defn{target tree}~$T(R)$) is the subgraph of~$D(R)$ induced by the sources~$\set{s(r)}{r \in [n]}$ (resp.~by the targets~$\set{t(r)}{r \in [n]}$).
+Note that since~$R$ is generic, the sources~$\set{s(r)}{r \in R}$ and the targets~$\set{t(r)}{r \in R}$ are disjoint.
+The \defn{source tree}~$S(R)$ (resp.~\defn{target tree}~$T(R)$) is the subgraph of~$D(R)$ induced by the sources~$\set{s(r)}{r \in R}$ (resp.~by the targets~$\set{t(r)}{r \in R}$).
 It is not difficult to check that the source tree~$S(R)$ (resp.~target tree~$T(R)$) is a binary tree, rooted at the bottom left (resp.~top right) corner of~$R$, and oriented from (resp.~towards) its root.
@@
 %The weak rectangulotope is realized by the polytope~$\WRP (n) \subset \R^n\!$ defined equivalently~as
-The weak rectangulotope $\WRP (n)$ is realized by:
+The weak rectangulotope $\WRP (n)$ is realized by: \vincent{We loose the information that the two descriptions define the same polytope...}
   \begin{enumerate}[(i)]
@@
 
-\begin{example}
-The weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
-The coordinates of the vertex corresponding to each weak rectangulation are gathered in \cref{tab:verticesWeakRectangulotope}.
-The facet inequalities corresponding to each nonempty proper subset of~$[n]$ are gathered in \cref{tab:facetsWeakRectangulotope}.
-
-\begin{table}[p]
+\begin{table}
 	\centerline{
 	\begin{tabular}{cccccc}
-		\includegraphics[scale=1.5]{rectangulation1} &
-		\includegraphics[scale=1.5]{rectangulation2} &
-		\includegraphics[scale=1.5]{rectangulation3} &
-		\includegraphics[scale=1.5]{rectangulation4} &
-		\includegraphics[scale=1.5]{rectangulation5} &
-		\includegraphics[scale=1.5]{rectangulation6}
-		\\
+		\includegraphics[scale=2]{rectangulation1} &
+		\includegraphics[scale=2]{rectangulation2} &
+		\includegraphics[scale=2]{rectangulation3} &
+		\includegraphics[scale=2]{rectangulation4} &
+		\includegraphics[scale=2]{rectangulation5} &
+		\includegraphics[scale=2]{rectangulation6}
+		\\[-.1cm]
 		$(-3, -1, 1, 3)$ &
@@
 		\\[.2cm]
-		\includegraphics[scale=1.5]{rectangulation8} &
-		\includegraphics[scale=1.5]{rectangulation9} &
-		\includegraphics[scale=1.5]{rectangulation10} &
-		\includegraphics[scale=1.5]{rectangulation11} &
-		\includegraphics[scale=1.5]{rectangulation12} &
-		\includegraphics[scale=1.5]{rectangulation13}
-		\\
+		\includegraphics[scale=2]{rectangulation8} &
+		\includegraphics[scale=2]{rectangulation9} &
+		\includegraphics[scale=2]{rectangulation10} &
+		\includegraphics[scale=2]{rectangulation11} &
+		\includegraphics[scale=2]{rectangulation12} &
+		\includegraphics[scale=2]{rectangulation13}
+		\\[-.1cm]
 		$(-1, 3, -5, 3)$ &
@@
 		\\[.2cm]
-		\includegraphics[scale=1.5]{rectangulation14} &
-		\includegraphics[scale=1.5]{rectangulation15} &
-		\includegraphics[scale=1.5]{rectangulation16} &
-		\includegraphics[scale=1.5]{rectangulation17} &
-		\includegraphics[scale=1.5]{rectangulation19} &
-		\includegraphics[scale=1.5]{rectangulation20}
-		\\
+		\includegraphics[scale=2]{rectangulation14} &
+		\includegraphics[scale=2]{rectangulation15} &
+		\includegraphics[scale=2]{rectangulation16} &
+		\includegraphics[scale=2]{rectangulation17} &
+		\includegraphics[scale=2]{rectangulation19} &
+		\includegraphics[scale=2]{rectangulation20}
+		\\[-.1cm]
 		$(3, -5, 0, 2)$ &
@@
 		&		
-		\includegraphics[scale=1.5]{rectangulation21} &
-		\includegraphics[scale=1.5]{rectangulation22} &
-		\includegraphics[scale=1.5]{rectangulation23} &
-		\includegraphics[scale=1.5]{rectangulation24} &
-		\\
+		\includegraphics[scale=2]{rectangulation21} &
+		\includegraphics[scale=2]{rectangulation22} &
+		\includegraphics[scale=2]{rectangulation23} &
+		\includegraphics[scale=2]{rectangulation24} &
+		\\[-.1cm]
 		&
@@
 \end{table}
+
+\begin{table}
+	\centerline{
+	\begin{tabular}{cccccc}
+		\includegraphics[scale=2]{rectangulation1} &
+		\includegraphics[scale=2]{rectangulation2} &
+		\includegraphics[scale=2]{rectangulation3} &
+		\includegraphics[scale=2]{rectangulation4} &
+		\includegraphics[scale=2]{rectangulation5} &
+		\includegraphics[scale=2]{rectangulation6}
+		\\[-.1cm]
+		$(-6, -2, 2, 6)$ &
+		$(-6, -2, 9, -1)$ &
+		$(-6, 5, -5, 6)$ &
+		$(1, -9, 2, 6)$ &
+		$(-6, 1, 9, -4)$ &
+		$(-6, 9, -5, 2)$
+		\\[.2cm]
+		\includegraphics[scale=2]{rectangulation8} &
+		\includegraphics[scale=2]{rectangulation9} &
+		\includegraphics[scale=2]{rectangulation10} &
+		\includegraphics[scale=2]{rectangulation11} &
+		\includegraphics[scale=2]{rectangulation12} &
+		\includegraphics[scale=2]{rectangulation13}
+		\\[-.1cm]
+		$(-2, 5, -9, 6)$ &
+		$(4, -9, -1, 6)$ &
+		$(-6, 9, 1, -4)$ &
+		$(-4, 1, 9, -6)$ &
+		$(2, -9, 9, -2)$ &
+		$(-2, 9, -9, 2)$
+		\\[.2cm]
+		\includegraphics[scale=2]{rectangulation14} &
+		\includegraphics[scale=2]{rectangulation15} &
+		\includegraphics[scale=2]{rectangulation16} &
+		\includegraphics[scale=2]{rectangulation17} &
+		\includegraphics[scale=2]{rectangulation19} &
+		\includegraphics[scale=2]{rectangulation20}
+		\\[-.1cm]
+		$(6, -9, -1, 4)$ &
+		$(4, -1, -9, 6)$ &
+		$(-4, 9, 1, -6)$ &
+		$(2, -5, 9, -6)$ &
+		$(6, -9, 5, -2)$ &
+		$(6, -1, -9, 4)$
+		\\[.2cm]
+		\includegraphics[scale=2]{rectangulation21} &
+		\includegraphics[scale=2]{rectangulation22} &
+		\includegraphics[scale=2]{rectangulation23} &
+		\includegraphics[scale=2]{rectangulation24} &
+		\includegraphics[scale=2]{rectangulation7} &
+		\includegraphics[scale=2]{rectangulation18}
+		\\[-.1cm]
+		$(-1, 9, -2, -6)$ &
+		$(6, -5, 5, -6)$ &
+		$(6, 2, -9, 1)$ &
+		$(6, 2, -2, -6)$ &
+		$(1, -9, 9, -1)$ &
+		$(-1, 9, -9, 1)$
+	\end{tabular}
+	}
+	\caption{The vertices of the strong rectangulotope~$\SRP(4)$ of \cref{fig:strongRectangulotope}.}
+	\label{tab:verticesStrongRectangulotope}
+\end{table}
+
+\begin{example}
+The weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
+The coordinates of the vertices corresponding to all weak rectangulations are gathered in \cref{tab:verticesWeakRectangulotope}.
+The facet inequalities corresponding to each nonempty proper subset of~$[n]$ are gathered in \cref{tab:facetsWeakRectangulotope}.
+
+%\begin{table}[p]
+%	\centerline{
+%	\begin{tabular}{cccccc}
+%		\includegraphics[scale=1.5]{rectangulation1} &
+%		\includegraphics[scale=1.5]{rectangulation2} &
+%		\includegraphics[scale=1.5]{rectangulation3} &
+%		\includegraphics[scale=1.5]{rectangulation4} &
+%		\includegraphics[scale=1.5]{rectangulation5} &
+%		\includegraphics[scale=1.5]{rectangulation6}
+%		\\
+%		$(-3, -1, 1, 3)$ &
+%		$(-3, -1, 5, -1)$ &
+%		$(-3, 3, -3, 3)$ &
+%		$(1, -5, 1, 3)$ &
+%		$(-3, 0, 5, -2)$ &
+%		$(-3, 5, -3, 1)$
+%		\\[.2cm]
+%		\includegraphics[scale=1.5]{rectangulation8} &
+%		\includegraphics[scale=1.5]{rectangulation9} &
+%		\includegraphics[scale=1.5]{rectangulation10} &
+%		\includegraphics[scale=1.5]{rectangulation11} &
+%		\includegraphics[scale=1.5]{rectangulation12} &
+%		\includegraphics[scale=1.5]{rectangulation13}
+%		\\
+%		$(-1, 3, -5, 3)$ &
+%		$(2, -5, 0, 3)$ &
+%		$(-3, 5, 0, -2)$ &
+%		$(-2, 0, 5, -3)$ &
+%		$(1, -5, 5, -1)$ &
+%		$(-1, 5, -5, 1)$
+%		\\[.2cm]
+%		\includegraphics[scale=1.5]{rectangulation14} &
+%		\includegraphics[scale=1.5]{rectangulation15} &
+%		\includegraphics[scale=1.5]{rectangulation16} &
+%		\includegraphics[scale=1.5]{rectangulation17} &
+%		\includegraphics[scale=1.5]{rectangulation19} &
+%		\includegraphics[scale=1.5]{rectangulation20}
+%		\\
+%		$(3, -5, 0, 2)$ &
+%		$(2, 0, -5, 3)$ &
+%		$(-2, 5, 0, -3)$ &
+%		$(1, -3, 5, -3)$ &
+%		$(3, -5, 3, -1)$ &
+%		$(3, 0, -5, 2)$
+%		\\[.2cm]
+%		&		
+%		\includegraphics[scale=1.5]{rectangulation21} &
+%		\includegraphics[scale=1.5]{rectangulation22} &
+%		\includegraphics[scale=1.5]{rectangulation23} &
+%		\includegraphics[scale=1.5]{rectangulation24} &
+%		\\
+%		&
+%		$(-1, 5, -1, -3)$ &
+%		$(3, -3, 3, -3)$ &
+%		$(3, 1, -5, 1)$ &
+%		$(3, 1, -1, -3)$ &
+%	\end{tabular}
+%	}
+%	\caption{The vertices of the weak rectangulotope~$\WRP(4)$ of \cref{fig:weakRectangulotope}.}
+%	\label{tab:verticesWeakRectangulotope}
+%\end{table}
 
@@
 The strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotope}.
-The coordinates of the vertex corresponding to each strong rectangulation are gathered in \cref{tab:verticesStrongRectangulotope}.
+The coordinates of the vertices corresponding to all strong rectangulations are gathered in \cref{tab:verticesStrongRectangulotope}.
 The facet inequalities corresponding to each nonempty proper subset of~$[n]$ are gathered in \cref{tab:facetsStrongRectangulotope}.
 
-\begin{table}
-	\centerline{
-	\begin{tabular}{cccccc}
-		\includegraphics[scale=1.5]{rectangulation1} &
-		\includegraphics[scale=1.5]{rectangulation2} &
-		\includegraphics[scale=1.5]{rectangulation3} &
-		\includegraphics[scale=1.5]{rectangulation4} &
-		\includegraphics[scale=1.5]{rectangulation5} &
-		\includegraphics[scale=1.5]{rectangulation6}
-		\\
-		$(-6, -2, 2, 6)$ &
-		$(-6, -2, 9, -1)$ &
-		$(-6, 5, -5, 6)$ &
-		$(1, -9, 2, 6)$ &
-		$(-6, 1, 9, -4)$ &
-		$(-6, 9, -5, 2)$
-		\\[.2cm]
-		\includegraphics[scale=1.5]{rectangulation8} &
-		\includegraphics[scale=1.5]{rectangulation9} &
-		\includegraphics[scale=1.5]{rectangulation10} &
-		\includegraphics[scale=1.5]{rectangulation11} &
-		\includegraphics[scale=1.5]{rectangulation12} &
-		\includegraphics[scale=1.5]{rectangulation13}
-		\\
-		$(-2, 5, -9, 6)$ &
-		$(4, -9, -1, 6)$ &
-		$(-6, 9, 1, -4)$ &
-		$(-4, 1, 9, -6)$ &
-		$(2, -9, 9, -2)$ &
-		$(-2, 9, -9, 2)$
-		\\[.2cm]
-		\includegraphics[scale=1.5]{rectangulation14} &
-		\includegraphics[scale=1.5]{rectangulation15} &
-		\includegraphics[scale=1.5]{rectangulation16} &
-		\includegraphics[scale=1.5]{rectangulation17} &
-		\includegraphics[scale=1.5]{rectangulation19} &
-		\includegraphics[scale=1.5]{rectangulation20}
-		\\
-		$(6, -9, -1, 4)$ &
-		$(4, -1, -9, 6)$ &
-		$(-4, 9, 1, -6)$ &
-		$(2, -5, 9, -6)$ &
-		$(6, -9, 5, -2)$ &
-		$(6, -1, -9, 4)$
-		\\[.2cm]
-		\includegraphics[scale=1.5]{rectangulation21} &
-		\includegraphics[scale=1.5]{rectangulation22} &
-		\includegraphics[scale=1.5]{rectangulation23} &
-		\includegraphics[scale=1.5]{rectangulation24} &
-		\includegraphics[scale=1.5]{rectangulation7} &
-		\includegraphics[scale=1.5]{rectangulation18}
-		\\
-		$(-1, 9, -2, -6)$ &
-		$(6, -5, 5, -6)$ &
-		$(6, 2, -9, 1)$ &
-		$(6, 2, -2, -6)$ &
-		$(1, -9, 9, -1)$ &
-		$(-1, 9, -9, 1)$
-	\end{tabular}
-	}
-	\caption{The vertices of the strong rectangulotope~$\SRP(4)$ of \cref{fig:strongRectangulotope}.}
-	\label{tab:verticesStrongRectangulotope}
-\end{table}
+%\begin{table}
+%	\centerline{
+%	\begin{tabular}{cccccc}
+%		\includegraphics[scale=1.5]{rectangulation1} &
+%		\includegraphics[scale=1.5]{rectangulation2} &
+%		\includegraphics[scale=1.5]{rectangulation3} &
+%		\includegraphics[scale=1.5]{rectangulation4} &
+%		\includegraphics[scale=1.5]{rectangulation5} &
+%		\includegraphics[scale=1.5]{rectangulation6}
+%		\\
+%		$(-6, -2, 2, 6)$ &
+%		$(-6, -2, 9, -1)$ &
+%		$(-6, 5, -5, 6)$ &
+%		$(1, -9, 2, 6)$ &
+%		$(-6, 1, 9, -4)$ &
+%		$(-6, 9, -5, 2)$
+%		\\[.2cm]
+%		\includegraphics[scale=1.5]{rectangulation8} &
+%		\includegraphics[scale=1.5]{rectangulation9} &
+%		\includegraphics[scale=1.5]{rectangulation10} &
+%		\includegraphics[scale=1.5]{rectangulation11} &
+%		\includegraphics[scale=1.5]{rectangulation12} &
+%		\includegraphics[scale=1.5]{rectangulation13}
+%		\\
+%		$(-2, 5, -9, 6)$ &
+%		$(4, -9, -1, 6)$ &
+%		$(-6, 9, 1, -4)$ &
+%		$(-4, 1, 9, -6)$ &
+%		$(2, -9, 9, -2)$ &
+%		$(-2, 9, -9, 2)$
+%		\\[.2cm]
+%		\includegraphics[scale=1.5]{rectangulation14} &
+%		\includegraphics[scale=1.5]{rectangulation15} &
+%		\includegraphics[scale=1.5]{rectangulation16} &
+%		\includegraphics[scale=1.5]{rectangulation17} &
+%		\includegraphics[scale=1.5]{rectangulation19} &
+%		\includegraphics[scale=1.5]{rectangulation20}
+%		\\
+%		$(6, -9, -1, 4)$ &
+%		$(4, -1, -9, 6)$ &
+%		$(-4, 9, 1, -6)$ &
+%		$(2, -5, 9, -6)$ &
+%		$(6, -9, 5, -2)$ &
+%		$(6, -1, -9, 4)$
+%		\\[.2cm]
+%		\includegraphics[scale=1.5]{rectangulation21} &
+%		\includegraphics[scale=1.5]{rectangulation22} &
+%		\includegraphics[scale=1.5]{rectangulation23} &
+%		\includegraphics[scale=1.5]{rectangulation24} &
+%		\includegraphics[scale=1.5]{rectangulation7} &
+%		\includegraphics[scale=1.5]{rectangulation18}
+%		\\
+%		$(-1, 9, -2, -6)$ &
+%		$(6, -5, 5, -6)$ &
+%		$(6, 2, -9, 1)$ &
+%		$(6, 2, -2, -6)$ &
+%		$(1, -9, 9, -1)$ &
+%		$(-1, 9, -9, 1)$
+%	\end{tabular}
+%	}
+%	\caption{The vertices of the strong rectangulotope~$\SRP(4)$ of \cref{fig:strongRectangulotope}.}
+%	\label{tab:verticesStrongRectangulotope}
+%\end{table}
 
@@
 We now briefly recall some results on the weak Bruhat order and its quotients, both from a lattice and geometric perspectives.
-We refer to \cite{Reading-arcDiagrams, MR3645055, MR3645056, MR4584712} for details.
+We refer to \cite{MR2142177, Reading-arcDiagrams, MR3645055, MR3645056, MR3964495, MR4584712} for details.
 % J: Reading-Chapters: which one is this ?
-
-\begin{figure}
-	\capstart
-	\centerline{\includegraphics[scale=.6]{weakOrderLeft4} \; \includegraphics[scale=.6]{weakOrderCongruence4} \; \includegraphics[scale=.48]{TamariLattice4}}
-	\caption{The weak Bruhat order on~$\f{S}_4$ (left), the sylvester congruence~$\equiv_\textrm{sylv}$~(middle), and the Tamari lattice (right). \cite[Fig.~1 \& 2]{MR3964495}}
-	\label{fig:sylvesterCongruence}
-\end{figure}
 
@@
 
+\afterpage{
+\begin{figure}
+	\capstart
+	\centerline{\includegraphics[scale=.6]{weakOrderLeft4} \; \includegraphics[scale=.6]{weakOrderCongruence4} \; \includegraphics[scale=.48]{TamariLattice4}}
+	\caption{The weak Bruhat order on~$\f{S}_4$ (left), the sylvester congruence~$\equiv_\textrm{sylv}$~(middle), and the Tamari lattice (right). \cite[Fig.~1 \& 2]{MR3964495}}
+	\label{fig:sylvesterCongruence}
+\end{figure}
+}
+
 A \defn{lattice congruence} of the weak Bruhat order is an equivalence relation~$\equiv$ on~$\f{S}_n$ that respects the meet and join operations, \ie such that $x \equiv x'$ and~$y \equiv y'$ implies $x \meet y \, \equiv \, x' \meet y'$ and~$x \join y \, \equiv \, x' \join y'$.
@@
 %Moreover, the sets of arcs of the form~$\c{A}_\equiv$ are easily described.
-\vspace{-.1cm}
+\vspace{-.2cm}
 \parpic(5cm,2.8cm)(10pt, 90pt)[r][b]{\includegraphics[scale=.5]{subarcOrder}}{
@@
 	\capstart
-	\centerline{\includegraphics[scale=.6]{braidFanLeft4} \qquad \raisebox{-.3cm}{\includegraphics[scale=.6]{sylvesterFan4}}}
+	\centerline{\includegraphics[scale=.75]{braidFanLeft4} \quad \raisebox{-.3cm}{\includegraphics[scale=.75]{sylvesterFan4}}}
 	\caption{The braid fan (left) and the sylvester fan (right). \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
@@
 The \defn{shard polytope} of~$\alpha$ is the convex hull~$\SP(\alpha)$ of the characteristic vectors of all $\alpha$-alternating matchings.
-We refer to Figure~\ref{fig:shardPolytopes} for an illustration of this definition.
+We refer to \cref{fig:shardPolytopes} for an illustration of this definition.
 It was shown in~\cite{MR4584712} that any Minkowski sum of positive scalings of the shard polytopes~$\SP(\alpha)$ for all arcs~$\alpha$ in the arc ideal~$\c{A}_\equiv$ is a quotientope for~$\equiv$.
@@
 
+\pagebreak
 \section{Weak rectangulotopes}
@@
 
-%%%%%%%%%%%%%
+This section is devoted to weak rectangulations and the proof of \cref{thm:weakRectangulotope}.
+
+%%%%%%%%%%%%%
+
 \subsection{The weak poset}
@@
 %%%%%%%%%%%%%
+
 \subsection{The weak rectangulation congruence}
@@
 
-We recall that \emph{Baxter permutations} are permutations avoiding the vincular pattern $2\underline{41}3$ and $3\underline{14}2$, where the underlining indicates that the symbols must appear contiguously in the permutation.
+We recall that \emph{Baxter permutations} are permutations avoiding the vincular patterns $2\underline{41}3$ and $3\underline{14}2$, where the underlining indicates that the symbols must appear contiguously in the permutation.
 Similarly, \defn{twisted Baxter permutations} are permutations avoiding the patterns $2\underline{41}3$ and $3\underline{41}2$, and \defn{co-twisted Baxter permutations} are permutations avoiding the patterns $2\underline{14}3$ and $3\underline{14}2$.
@@
   \item the congruence classes are one-to-one with weak rectangulations,  
-  \item the minimal elements of each congruence class are the twisted Baxter permutations,
-  \item the maximal elements are the co-twisted Baxter permutations,
+  \item the minimal elements of the congruence classes are the twisted Baxter permutations,
+  \item the maximal elements of the congruence classes are the co-twisted Baxter permutations,
   \item the congruence classes each contain a single Baxter permutation, and are one-to-one with Baxter permutations.
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Up and down arcs}
@@
 
-\begin{lemma}
+The arc ideal of the weak rectangulation congruence was already considered in~\cite[Exm.~4.10]{Reading-arcDiagrams}.
+
+\begin{lemma}[{\cite[Exm.~4.10]{Reading-arcDiagrams}}]
 The weak rectangulation congruence $\weakeq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that do not cross the horizontal line.
@@
 
+%%%%%%%%%%%%%
+
 \subsection{Shard polytopes of up and down arcs}
@@
 
-From the results of section~\ref{subsec:quotientopes}, the quotientope is realized by a Minkowski sum of the shard polytopes of the up arcs and the down arcs.
-These quotientopes are simplices and anti-simplices (see Figure~\ref{fig:shardPolytopes}, left).
+From the results of \cref{subsec:quotientopes}, the quotientope is realized by a Minkowski sum of the shard polytopes of the up arcs and the down arcs.
+These quotientopes are simplices and anti-simplices.
+%See \cref{fig:shardPolytopes}\,(left).
 
@@
 
-\begin{lemma}
-  \label{lem:weakMinkowski}
+We refer to \cref{fig:shardPolytopes} for examples of up and down shard polytopes.
+
+\begin{proposition}
+  \label{prop:weakMinkowski}
   The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all up and down shard polytopes.
   In particular, it is the Minkowski sum of two opposite associahedra of J.-L.~Loday~\cite{MR2871762}.
-\end{lemma}
+\end{proposition}
 
@@
 
-The description of the weak rectangulotope~$\WRP(n)$ as a Minkowski sum in \cref{lem:weakMinkowski} allows us to easily compute the corresponding submodular function.
-The arc ideal of the weak rectangulation congruence contains exactly all up arcs and all down arcs, hence from the result of~\cite{MR4328906} mentioned in section~\ref{subsec:quotientFans}, all $2^n-2$ rays of the braid fan are preserved in the fan of the weak rectangulotope.
+The description of the weak rectangulotope~$\WRP(n)$ as a Minkowski sum in \cref{prop:weakMinkowski} allows us to easily compute the corresponding submodular function.
+The arc ideal of the weak rectangulation congruence contains exactly all up arcs and all down arcs, hence from the result of~\cite{MR4328906} mentioned in \cref{subsec:quotientFans}, all $2^n-2$ rays of the braid fan are preserved in the fan of the weak rectangulotope.
 This implies that every nonempty proper subset will define a facet of the weak rectangulotope, and the description of the weak rectangulotope by its submodular function is actually an irredundant facet description.
@@
 
-For every weak rectangulation $R$, we wish to give the coordinates of a point $\b{p}(R)\in\R^n$ such that $\WRP(n)$ is the convex hull of the points~$\b{p}(R)$ for all weak rectangulations~$R$.
+For every weak rectangulation $R$, we wish to give the coordinates of a point $\b{p}(R)\in\R^n$ such that $\WRP(n)$ is the convex hull of the points~$\b{p}(R)$ for all weak rectangulations~$R$ of size~$n$.
 Recall that in a Minkowski sum of polytopes, the vertex that is extremal in a generic direction is the sum of the vertices of the summands that are extremal in this direction.
@@
   The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~of the translated down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\pi^{-1}$ is $\b{e}_i$ (resp.~$-\b{e}_i$), where $i$ is the maximum (resp.
-the minimum) of $\pi^{-1}$ in $I$.
+the minimum) of~$\pi^{-1}$ in $I$.
 \end{lemma}
@@
 
-The proof of the following lemma actually follows from Loday's realization in Theorem~\ref{thm:Loday}.
+The proof of the following lemma actually follows from Loday's realization in~\cref{thm:loday}.
 For the sake of completeness, and because we will reuse the same reasoning in the next section, we give an elementary proof.
@@
   \label{lem:weakCoord}
-  Let $R$ be a rectangulation of size~$n$, with target tree~$T$, and let $\b{p}(R)$ the vertex of $\WRP(n)$ associated with $R$.
-  Given $i\in [n]$, the number of up arcs~$\upArc{I}$ such that $\b{e}_i$ is the extremal vertex of $\SP(\upArc{I})$ contributing to~$\b{p}(R)$ is
+  Let $R$ be a rectangulation of size~$n$, with target tree~$T$, and let $\b{p}(R)$ the vertex of~$\WRP(n)$ associated with $R$.
+  Given $i\in [n]$, the number of up arcs~$\upArc{I}$ such that $\b{e}_i$ is the extremal vertex of~$\SP(\upArc{I})$ contributing to~$\b{p}(R)$ is
   \[
@@
   Let $\pi$ be any permutation in $\mathcal{L}(P_w(R))$.
-  From \cref{lem:lodaymax}, for an $i\in [n]$, a shard polytope $\SP(\upArc{I})$ contributes to $\b{e}_i$ only if $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
+  From \cref{lem:lodaymax}, for an $i\in [n]$, a shard polytope~$\SP(\upArc{I})$ contributes to $\b{e}_i$ if and only if $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
   We claim that the number of intervals $I$ satisfying this condition is the given product.
@@
   A symmetric argument shows that there are exactly $v^{T}_i$ choices for the right endpoint of $I$.
+  The combined number of choices is thus the product~$h^{T}_i\cdot v^{T}_i$.
 \end{proof}
 
-A symmetric analogue of Lemma~\ref{lem:weakCoord} holds for down shard polytopes, where the product contributes to $-\b{e}_i$ instead of $\b{e}_i$, and we consider the tree $S(R)$ instead of the tree $T(R)$.
-This proves Theorem~\ref{thm:weakRectangulotope}.
+A symmetric analogue of \cref{lem:weakCoord} holds for down shard polytopes, where the product contributes to $-\b{e}_i$ instead of $\b{e}_i$, and we consider the tree $S(R)$ instead of the tree $T(R)$.
+This proves \cref{thm:weakRectangulotope}.
 
@@
 
-%%%%%%%%%%%%%
+This section is devoted to strong rectangulations and the proof of \cref{thm:strongRectangulotope}.
+
+%%%%%%%%%%%%%
+
 \subsection{The strong poset}
@@
 Given a rectangulation $R$, we define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of $R$ by considering two relations on $[n]$~\cite{ACFF24}:
+\vincent{Add picture weak and strong orders?}
 \begin{enumerate}
 \item The \defn{adjacency poset} $([n],\tri)$ of $R$, defined by considering the adjacency graph of the rectangles of $R$ (identified by their inorder labels in $[n]$), orienting its edges from left to right and from bottom to top, and taking the transitive closure of this directed acyclic graph.
-\item The \defn{special relation} $([n],\btri)$ of $R$, in which $a\btri b$ if the two rectangles with inorder labels $a$ and $b$ are such that either (i) the right side of $b$ and the left side of $a$ lie on the same vertical segment, and the bottom-right corner of $b$ is above the top-left corner of $a$, or (ii) the top side of $b$ and the bottom side of rectangle $a$ lie on the same horizontal segment, and the top-left corner of $b$ lies on the right of the bottom-right corner of  $a$.
+\item The \defn{special relation} $([n],\btri)$ of $R$, in which $a\btri b$ if the two rectangles with inorder labels $a$ and $b$ are such that 
+	\begin{enumerate}[(i)]
+	\item either the right side of $b$ and the left side of $a$ lie on the same vertical segment, and the bottom-right corner of $b$ is above the top-left corner of $a$, 
+	\item or the top side of $b$ and the bottom side of rectangle $a$ lie on the same horizontal segment, and the top-left corner of $b$ lies on the right of the bottom-right corner~of~$a$.
+	\end{enumerate}
 \end{enumerate}
@@
 \begin{proof}
-  Clearly, the strong poset is an extension of the adjacency poset $([n],\tri)$ of $R$, since it is obtained by adding pairs from the special relation.
+  Clearly, the strong poset is an extension of the adjacency poset $([n],\tri)$ of $R$, since it is obtained by adding pairs from the special relation~$([n],\btri)$.
   It remains to observe that the adjacency poset of $R$ is an extension of the weak poset of $R$.
-  For this, we consider a diagonal representative $D$ of the weak equivalence class of $R$, and recall that the weak poset of $R$ is the adjacency poset of $D$.
+  For this, we consider a diagonal representative~$D$ of the weak equivalence class of $R$, and recall that the weak poset of $R$ is the adjacency poset of $D$.
   By definition, $R$ can be obtained from $D$ by a sequence of wall slides.
@@
 %%%%%%%%%%%%%
+
 \subsection{The strong rectangulation congruence}
@@
 The \defn{co-2-clumped permutations} are defined similarly, as permutations avoiding the mirror images of these four patterns.
-He showed that 2-clumped permutations were in bijection with strong rectangulations.
+He showed that 2-clumped permutations are in bijection with strong rectangulations.
 
@@
   \item the congruence classes are one-to-one with strong rectangulations,  
-  \item the minimal elements of each congruence class are the 2-clumped permutations,
-  \item the maximal elements are the co-2-clumped permutations.
+  \item the minimal elements of the congruence classes are the 2-clumped permutations,
+  \item the maximal elements of the congruence classes are the co-2-clumped permutations.
   \end{itemize}
@@
 We call the congruence $\strongeq$ the \defn{strong rectangulation congruence}.
-The \defn{strong rectangulotope} $\SRP(n)$ is the quotientope of the strong rectangulation congruence~${\strongeq}$.
+The \defn{strong rectangulotope}~$\SRP(n)$ is the quotientope of the strong rectangulation congruence~${\strongeq}$.
 
@@
 
-\begin{lemma}
+The arc ideal of the weak rectangulation congruence was already considered in~\cite[Exm.~4.11]{Reading-arcDiagrams}.
+
+\begin{lemma}[{\cite[Exm.~4.11]{Reading-arcDiagrams}}]
   The strong rectangulation congruence $\strongeq$ on $\f{S}_n$ is defined by the arc ideal composed of arcs that cross the horizontal line at most once.
@@
 
-We classify the arcs into two families, depending on whether the arcs cross from above or below.
+We classify the arcs into two families, depending on whether the arcs do not cross the horizontal line from below or from above.
 For two consecutive intervals~$I$ and~$J$ of~$[n]$, the \defn{yin arc}~$\yinArc{I}{J}$ (resp.~\defn{yang arc}~$\yangArc{I}{J}$) is the arc starts at~$\min(I)$ and ends at~$\max(J)$, and goes above (resp.~below) the remaining elements of $I$ and below (resp.~above) the remaining elements of $J$, that is,
@@
 
-The following lemmas are direct applications of the results from~\cite{MR4584712} given in section~\ref{subsec:quotientopes}.
+From the results of \cref{subsec:quotientopes}, the quotientope is realized by a Minkowski sum of the shard polytopes of the yin arcs and the yang arcs.
+%The following lemmas are direct applications of the results from~\cite{MR4584712} given in \cref{subsec:quotientopes}.
+%See \cref{fig:shardPolytopes}.
+These shard polytopes can be described as follows.
 
@@
 
-We refer to Figure~\ref{fig:shardPolytopes} for examples of yin and yang shard polytopes.
+We refer to \cref{fig:shardPolytopes} for examples of yin and yang shard polytopes.
 %Note that yin shard polytopes are translates of the up shard polytopes when $|J|=1$, and of the down shard polytopes when $|I|=1$.
@@
 
-\begin{lemma}
-  \label{lem:strongMinkowski}
+\begin{proposition}
+  \label{prop:strongMinkowski}
   The strong rectangulotope $\SRP(n)$ is realized by the Minkowski sum of all yin and yang shard polytopes.
-\end{lemma}
+\end{proposition}
 
@@
 
-\cref{lem:strongMinkowski} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin congruence is defined by the arc ideal consisting of all yin arcs, and the yang congruence is defined by the yang arcs.
+\cref{prop:strongMinkowski} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin (resp.~yang) congruence is defined by the arc ideal consisting of all yin (resp.~yang) arcs.
 It can be checked that the equivalence classes of these congruences are counted by \defn{semi-Baxter permutations}, defined as permutations avoiding $2\underline{41}3$, hence only one of the two forbidden Baxter patterns.
@@
 
-From the results of section~\ref{subsec:quotientopes}, the submodular function corresponding to our realization of $\SRP(n)$ can be obtained as before by summing the functions defining the yin and yang shard polytopes in \cref{lem:yinsp,lem:yangsp}.
+From the results of \cref{subsec:quotientopes}, the submodular function corresponding to our realization of $\SRP(n)$ can be obtained as before by summing the functions defining the yin and yang shard polytopes in \cref{lem:yinsp,lem:yangsp}.
 Note again that all rays of the braid fan are preserved in the fan of the strong rectangulotope, hence our characterization of the strong rectangulotope is an irredundant facet description.
@@
   Let $R$ be a strong rectangulation of size~$n$, let~$S$ and $T$ be its source and target trees, let~$([n],\prec_s)$ be its strong order, and let~$\b{p}(R)$ be the vertex of $\SRP(n)$ corresponding to $R$.
-  Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $A$ such that $\b{e}_i-\b{e}_j$ is the extremal vertex of $\SP(A)$ contributing to $p(R)$ is
+  Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $\alpha$ such that $\b{e}_i-\b{e}_j$ is the extremal vertex of~$\SP(\alpha)$ contributing to $p(R)$ is
   \[
@@
 \begin{proof}
-  The proof is not much more involved that of Lemma~\ref{lem:weakCoord}.
-  Let $\pi$ be any permutation in $\mathcal{L}(P_s(R))$.
-  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc $A$ whose shard polytope $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ only if $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.
-  Provided this holds, the set of yin arcs $A$ such that $\SP(A)$ contributes to $\b{e}_i-\b{e}_j$ are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that $i$ is the index of the maximum of $\pi^{-1}$ in $I$, and $j$ is the index of the maximum of $\pi^{-1}$ in $J$.
+  The proof follows the same lines as that of \cref{lem:weakCoord}.
+  Let $\pi$ be any permutation in~$\mathcal{L}(P_s(R))$.
+  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc~$\alpha$ whose shard polytope $\SP(\alpha)$ contributes to $\b{e}_i-\b{e}_j$ if and only if $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.
+  Provided this holds, the set of yin arcs $\alpha$ such that $\SP(\alpha)$ contributes to $\b{e}_i-\b{e}_j$ are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that $i$ is the index of the maximum of $\pi^{-1}$ in $I$, and $j$ is the index of the maximum of $\pi^{-1}$ in $J$.
   We claim that the number of choices of the three endpoints defining such a pair of intervals $I$ and $J$ is the given product.
@@
   Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
-  Since from Lemma~\ref{lem:sextw} the strong poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
-  Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
+  Since from \cref{lem:sextw} the strong poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
+  Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of~$\pi^{-1}$ in $I$.
   Now consider the leaf of $T$ that is immediately left of the leftmost leaf of the horizontal subtree of $i$.
@@
 
-This concludes the proof of Theorem~\ref{thm:strongRectangulotope}.
+This concludes the proof of \cref{thm:strongRectangulotope}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{comment}

### 2024-04-21 draft/rectangulotopes.tex

@@
 
-We recall that \emph{Baxter permutations} are permutations avoiding the vincular patterns $2\underline{41}3$ and $3\underline{14}2$, where the underlining indicates that the symbols must appear contiguously in the permutation.
+We recall that \defn{Baxter permutations} are permutations avoiding the vincular patterns $2\underline{41}3$ and $3\underline{14}2$, where the underlining indicates that the symbols must appear contiguously in the permutation.
 Similarly, \defn{twisted Baxter permutations} are permutations avoiding the patterns $2\underline{41}3$ and $3\underline{41}2$, and \defn{co-twisted Baxter permutations} are permutations avoiding the patterns $2\underline{14}3$ and $3\underline{14}2$.
@@
 A symmetric analogue of \cref{lem:weakCoord} holds for down shard polytopes, where the product contributes to $-\b{e}_i$ instead of $\b{e}_i$, and we consider the tree $S(R)$ instead of the tree $T(R)$.
-This proves \cref{thm:weakRectangulotope}.
+This concludes the proof of \cref{thm:weakRectangulotope}.
 

### 2024-04-22 draft/rectangulotopes.tex
remove example environements

@@
 
-\begin{example}
-The weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
+%\begin{example}
+For instance, the weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
 The coordinates of the vertices corresponding to all weak rectangulations are gathered in \cref{tab:verticesWeakRectangulotope}.
@@
 \end{table}
-\end{example}
+%\end{example}
 
@@
 \end{comment}
+\vincent{not sure if we want to reintroduce this example.}
 
@@
 
-\begin{example}
-The strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotope}.
+%\begin{example}
+For instance, the strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotope}.
 The coordinates of the vertices corresponding to all strong rectangulations are gathered in \cref{tab:verticesStrongRectangulotope}.

### 2024-04-22 draft/rectangulotopes.tex
l23

--- 2024-04-21 draft/rectangulotopes.tex
@@
 \end{lemma}
+\begin{proof}
+  Consider the two rectangles $r_i$ and $r_j$ with inorder labels $i$ and $j$ in $R$.
+  We first observe that the source vertex $s(r_j)$ is not contained in the interior of the lower left quadrant of origin $t(r_i)$.
+  Indeed, if we suppose otherwise, then the line segment from $s(r_j)$ to $t(r_i)$ has positive slope, which implies that $i\tri j$, which in turn implies $i\succ_s j$, contradicting our hypothesis.
+  We now proceed to a simple case analysis.
+
+  First suppose that $r_i$ and $r_j$ are disjoint, and that none of their edges lie on the same segment of $R$.
+  Then we claim that $cv^{T,S}_{i,j} = 0$.
+  Otherwise, suppose that $r_i$ and $r_j$ are disjoint, but they share a segment.
+  Without loss of generality (the other case is symmetric), let us suppose that they share a vertical segment.
+
+  Hence it cannot be the case that $r_i$ and $r_j$ are disjoint, and they must be touching.
+  Then, from our first observation, it can only be the case that $r_i$ is left of $r_j$, or below $r_j$, hence that $i\, \backslash \, j$, as claimed.
+  \end{proof}
 

### 2024-04-22 draft/rectangulotopes.tex
fig 2

@@
   \includegraphics[width=.9\textwidth]{rectangulations.pdf}
-	\caption{Three rectangulations. All three rectangulations are weakly equivalent, but only the first two are strongly equivalent. Only the third one is diagonal.}
+	\caption{Four rectangulations.
+The leftmost rectangulation is not weakly equivalent to any other.
+The rightmost three are weakly equivalent, but only the first two are strongly equivalent.
+Only the third one is diagonal.}
 	\label{fig:rectequiv}
 \end{figure}
-\vincent{Add one which is not weakly equivalent?}
 

### 2024-04-22 draft/rectangulotopes.tex
tentative proof of last lemma

@@
 \usepackage{todonotes}
-\newcommand{\vincent}[1]{\todo[size=\scriptsize, color=blue!30]{\rm #1 \\ \hfill --- V.}}
-\newcommand{\Vincent}[1]{\todo[inline, size=\scriptsize, color=blue!30]{\rm #1 \\ \hfill --- V.}}
-\newcommand{\jean}[2][]{\todo[size=\scriptsize, color=orange!30,#1]{\rm #2 \\ \hfill --- J.}}
-\newcommand{\Jean}[2][]{\todo[inline, size=\scriptsize, color=orange!30,#1]{\rm #2 \\ \hfill --- J.}}
+\newcommand{\vincent}[1]{\todo[size=\normalsize, color=blue!30]{\rm #1 \\ \hfill --- V.}}
+\newcommand{\Vincent}[1]{\todo[inline, size=\normalsize, color=blue!30]{\rm #1 \\ \hfill --- V.}}
+\newcommand{\jean}[2][]{\todo[size=\normalsize, color=orange!30,#1]{\rm #2 \\ \hfill --- J.}}
+\newcommand{\Jean}[2][]{\todo[inline, size=\normalsize, color=orange!30,#1]{\rm #2 \\ \hfill --- J.}}
 
@@
 \newcommand{\btri}{\LHD}
+
+% just in case -J
+\begin{comment}
+\tikzstyle{wall}=[line width=0.6pt]
+
+\newcommand{\td}{
+\begin{tikzpicture}[baseline=0.1mm,scale=0.10]
+\protect\draw[wall] (0,1.7)--(2,1.7);
+\protect\draw[wall] (1,0)--(1,1.7);
+\end{tikzpicture}
+}
+
+\newcommand{\tu}{
+\begin{tikzpicture}[baseline=-0.1mm,scale=0.10]
+\protect\draw[wall] (0,0)--(2,0);
+\protect\draw[wall] (1,0)--(1,1.7);
+\end{tikzpicture}
+}
+\end{comment}
 
@@
 Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$.
-We denote by $i \, \backslash \, j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$.
+
+We denote by $i\verticalPattern j$ the situation where 
+and by $i\horizontalPattern j$ the situation where \Jean{Complete with the right definition here.}
+
 We also use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.
@@
   \[
-    \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \neg \, i \, \backslash \, j \rrbracket
+    \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \neg \, i \verticalPattern j \rrbracket
     \qquad\text{and}\qquad
-    \yang{w}^R_{i,j} \eqdef v^S_i \cdot ch^{S,T}_{i,j}\cdot v^T_j \cdot \llbracket \neg \, j \, \backslash \, i \rrbracket,
+    \yang{w}^R_{i,j} \eqdef v^S_i \cdot ch^{S,T}_{i,j}\cdot v^T_j \cdot \llbracket \neg \, j \horizontalPattern i \rrbracket,
   \]
@@
 We refer to \cite{MR2142177, Reading-arcDiagrams, MR3645055, MR3645056, MR3964495, MR4584712} for details.
-% J: Reading-Chapters: which one is this ?
 
@@
 Note that the cover relations in the weak Bruhat order are given by the transpositions of two adjacent letters.
-%The weak Bruhat order is known to be a congruence uniform lattice~\cite{}.
-%In particular, it is join semidistributive, so that any permutation admits a canonical join representation.
 
@@
 
-\pagebreak
+%\pagebreak
 \section{Weak rectangulotopes}
@@
   \label{lem:lodaysp}
-%  Given an up arc~$\alpha$ defined by an interval $I$, the up shard polytope $\SP(\alpha)$ is a translate of the standard simplex~$\conv \set{ \b{e}_i }{ i\in I }$.
   The up shard polytope $\SP(\upArc{I})$ is a translate of the simplex~$\conv \set{ \b{e}_i }{ i\in I }$.
@@
   \]
-  %that maps $S$ to the number of intervals $I$ of $[n]$ such that $I\cap S\not=\emptyset$ and $I\not\subseteq S$. 
 \end{lemma}
@@
 \weakeq\sigma$ if and only if for any interval $I$ of $[n]$,
-%the maximum and the minimum of $\pi^{-1}$ and $\sigma^{-1}$ on $I$ have the same index.
 $\arg\max_{j\in I} \pi^{-1}(j)=\arg\max_{j\in I} \sigma^{-1}(j)$, and
@@
 \item The \defn{adjacency poset} $([n],\tri)$ of $R$, defined by considering the adjacency graph of the rectangles of $R$ (identified by their inorder labels in $[n]$), orienting its edges from left to right and from bottom to top, and taking the transitive closure of this directed acyclic graph.
-\item The \defn{special relation} $([n],\btri)$ of $R$, in which $a\btri b$ if the two rectangles with inorder labels $a$ and $b$ are such that 
+\item The \defn{special relation} $([n],\btri)$ of $R$, in which $a\btri b$ if the two rectangles $r_a$ and $r_b$ with respective inorder labels $a$ and $b$ are such that 
 	\begin{enumerate}[(i)]
-	\item either the right side of $b$ and the left side of $a$ lie on the same vertical segment, and the bottom-right corner of $b$ is above the top-left corner of $a$, 
-	\item or the top side of $b$ and the bottom side of rectangle $a$ lie on the same horizontal segment, and the top-left corner of $b$ lies on the right of the bottom-right corner~of~$a$.
+	\item either the right side of $r_b$ and the left side of $r_a$ lie on the same vertical segment, and the bottom-right corner of $r_b$ is above the top-left corner of $r_a$, 
+	\item or the top side of $r_b$ and the bottom side of rectangle $r_a$ lie on the same horizontal segment, and the top-left corner of $r_b$ lies on the right of the bottom-right corner~of~$r_a$.
 	\end{enumerate}
@@
 From the results of \cref{subsec:quotientopes}, the quotientope is realized by a Minkowski sum of the shard polytopes of the yin arcs and the yang arcs.
-%The following lemmas are direct applications of the results from~\cite{MR4584712} given in \cref{subsec:quotientopes}.
-%See \cref{fig:shardPolytopes}.
 These shard polytopes can be described as follows.
@@
 We refer to \cref{fig:shardPolytopes} for examples of yin and yang shard polytopes.
-%Note that yin shard polytopes are translates of the up shard polytopes when $|J|=1$, and of the down shard polytopes when $|I|=1$.
-%Similarly, up and down shard polytopes are translates of some yang shard polytopes.
 
@@
 \begin{lemma}
+  \label{lem:yincount}
   Let $R$ be a strong rectangulation of size~$n$, let~$S$ and $T$ be its source and target trees, let~$([n],\prec_s)$ be its strong order, and let~$\b{p}(R)$ be the vertex of $\SRP(n)$ corresponding to $R$.
@@
 \begin{proof}
-  The proof follows the same lines as that of \cref{lem:weakCoord}.
   Let $\pi$ be any permutation in~$\mathcal{L}(P_s(R))$.
@@
   We first show that the number of choices for the left endpoint of $I$ is equal to the number $h^T_i$ of leaves in the horizontal subtree of $i$ in $T$.
-  Each such leaf determines a left endpoint for $I$, defined as the node of the tree that follows it in an inorder traversal.
+  The proof of this follows the exact same lines as that of \cref{lem:weakCoord}.
+  Each leaf determines a left endpoint for $I$, defined as the next node in an inorder traversal.
   Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
@@
 
-It remains to observe that the only case where this product is nonzero and $i,j$ is not an inversion in $\pi$ is when $i \, \backslash \, j$.
+\cref{lem:yincount} essentially gives us the coefficient $\yin{w}^R_{i,j}$ in \cref{thm:strongRectangulotope}, except that the condition $\neg i\verticalPattern j$ is replaced by $i\succ_s j$.
+
+Recall that $i\verticalPattern j$ denote the situation where the rectangles $r_i$ and $r_j$ of inorder labels $i$ and $j$ are such that $r_i$ and $r_j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
+
+Our aim is to provide a formula for the coordinates of the vertices of the strong rectangulotope that does not involve the computation of the strong poset or one of its linear extension.
+For that purpose, we prove that for every pair $i<j\in [n]$ such that $i\prec_s j$, hence which do not have any associated yin arc, we have either $cv^{T,S}_{i,j} = 0$ (hence the product in \cref{lem:yincount} is 0 as well and adding it does not harm the result), or $i\verticalPattern j$.
 
 \begin{lemma}
+  \label{lem:whentosum}
   Let $R$ be a rectangulation of size~$n$, let $S$ and $T$ be its source and target trees, and let~$([n],\prec_s)$ be the associated strong order.
@@
   \[
-    cv^{T,S}_{i,j} > 0 \quad \Longrightarrow \quad i \, \backslash \, j.
+    cv^{T,S}_{i,j} > 0 \quad \Longrightarrow \quad i \verticalPattern j.
   \]
@@
   We first observe that the source vertex $s(r_j)$ is not contained in the interior of the lower left quadrant of origin $t(r_i)$.
-  Indeed, if we suppose otherwise, then the line segment from $s(r_j)$ to $t(r_i)$ has positive slope, which implies that $i\tri j$, which in turn implies $i\succ_s j$, contradicting our hypothesis.
-  We now proceed to a simple case analysis.
-
-  First suppose that $r_i$ and $r_j$ are disjoint, and that none of their edges lie on the same segment of $R$.
-  Then we claim that $cv^{T,S}_{i,j} = 0$.
-  Otherwise, suppose that $r_i$ and $r_j$ are disjoint, but they share a segment.
-  Without loss of generality (the other case is symmetric), let us suppose that they share a vertical segment.
-
-  Hence it cannot be the case that $r_i$ and $r_j$ are disjoint, and they must be touching.
-  Then, from our first observation, it can only be the case that $r_i$ is left of $r_j$, or below $r_j$, hence that $i\, \backslash \, j$, as claimed.
+  Indeed, if we suppose otherwise, then the line segment from $s(r_j)$ to $t(r_i)$ has positive slope, which implies that $j\tri i$, which in turn implies $j\prec_s i$, contradicting our hypothesis.
+
+  We now claim that there can only be one segment on which leaves common to the vertical subtree of $i$ in $T$ and the vertical subtree of $j$ in $S$ may lie, and this segment is the vertical segment supporting the right edge of $r_i$ and the left edge of $r_j$.
+  \Jean{This requires a case analysis. Any shortcut?}
+
+  Let us assume that this segment is shared.
+  If the bottom edge of $r_i$ is higher than the top edge of $r_j$, we must have $j\btri i$, which implies $j\prec_s i$, contradicting our hypothesis again.
+  Hence we must have that the bottom edge of $r_i$ is lower than the top edge of $r_j$, hence that $i\verticalPattern j$, as claimed.
 \end{proof}
 
+We omit the proofs of the analogues of \cref{lem:yincount} and \cref{lem:whentosum} for the case of the yang arcs: The arguments are completely symmetric.
 This concludes the proof of \cref{thm:strongRectangulotope}.

### 2024-04-22 draft/rectangulotopes.tex
minor changes

--- 2024-04-22 draft/rectangulotopes.tex
@@
 \begin{figure}
-  \includegraphics[width=.9\textwidth]{rectangulations.pdf}
+  \includegraphics[width=.65\textwidth]{rectangulations.pdf}
 	\caption{Three rectangulations. All three rectangulations are weakly equivalent, but only the first two are strongly equivalent. Only the third one is diagonal.}
@@
 %The weak rectangulotope is realized by the polytope~$\WRP (n) \subset \R^n\!$ defined equivalently~as
-The weak rectangulotope $\WRP (n)$ is realized by: \vincent{We loose the information that the two descriptions define the same polytope...}
+The weak rectangulotope $\WRP (n)$ is realized by the polytope equivalently described as:
   \begin{enumerate}[(i)]
@@
 
-\begin{comment}
-\begin{example}
+%\begin{example}
   The vertex of~$\WRP (16)$ corresponding to the weak rectangulation of \cref{fig:weakRectangulation} is
@@
   For instance, the $4$th coordinate is $h^T_4 \cdot v^T_4 - h^S_4 \cdot v^S_4 = 1 \cdot 5 - 1 \cdot 3 = 2$.
-  The facet of~$\WRP (16)$ corresponding to the subset~$\{3,4,7,8,9,10,14\}$ is given by the inequality
-  \[
-  x_3 + x_4 + x_7 + x_8 + x_9 + x_{10} + x_{14} \le ???
-  \]
-  \vincent{Compute and check...}
-\end{example}
-\end{comment}
-\vincent{not sure if we want to reintroduce this example.}
+%  The facet of~$\WRP (16)$ corresponding to the subset~$\{3,4,7,8,9,10,14\}$ is given by the inequality
+%  \[
+%  x_3 + x_4 + x_7 + x_8 + x_9 + x_{10} + x_{14} \le ???
+%  \]
+%  \vincent{Compute and check...}
+%\end{example}
 
@@
 % The strong rectangulotope is realized by the polytope~$\SRP (n)$ defined equivalently as
-The strong rectangulotope~$\SRP (n)$ is realized by:
+The strong rectangulotope~$\SRP (n)$ is realized by the polytope equivalently described as:
   \begin{enumerate}[(i)]
@@
 
-\begin{comment}
-\begin{example}
+%\begin{comment}
+%\begin{example}
   The vertex of~$\SRP (16)$ corresponding to the strong rectangulation of \cref{fig:strongRectangulation} is
@@
   \]
-  For instance, the contribution of~$\b{e}_5 - \b{e}_7$ is ...
-  \vincent{Compute and check...}
-  The facet of~$\SRP (16)$ corresponding to the subset~$\{3,4,7,8,9,10,14\}$ is given by the inequality
-  \[
-  x_3 + x_4 + x_7 + x_8 + x_9 + x_{10} + x_{14} \le ???
-  \]
-  \vincent{Compute and check...}
-\end{example}
-\end{comment}
+%  For instance, the contribution of~$\b{e}_5 - \b{e}_7$ is ...
+%  \vincent{Compute and check...}
+%  The facet of~$\SRP (16)$ corresponding to the subset~$\{3,4,7,8,9,10,14\}$ is given by the inequality
+%  \[
+%  x_3 + x_4 + x_7 + x_8 + x_9 + x_{10} + x_{14} \le ???
+%  \]
+%  \vincent{Compute and check...}
+%\end{example}
+%\end{comment}
 
@@
 
-\pagebreak
 \section{Weak rectangulotopes}

### 2024-04-22 draft/rectangulotopes.tex
typos

--- 2024-04-22 draft/rectangulotopes.tex
@@
   \label{lem:yincount}
-  Let $R$ be a strong rectangulation of size~$n$, let~$S$ and $T$ be its source and target trees, let~$([n],\prec_s)$ be its strong order, and let~$\b{p}(R)$ be the vertex of $\SRP(n)$ corresponding to $R$.
-  Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $\alpha$ such that $\b{e}_i-\b{e}_j$ is the extremal vertex of~$\SP(\alpha)$ contributing to $p(R)$ is
+  Let $R$ be a strong rectangulation of size~$n$, let~$S$ and $T$ be its source and target trees, let~$([n],\prec_s)$ be its strong poset, and let~$\b{p}(R)$ be the vertex of $\SRP(n)$ corresponding to $R$.
+  Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $\alpha$ such that $\b{e}_i-\b{e}_j$ is the extremal vertex of~$\SP(\alpha)$ contributing to $\b{p}(R)$ is
   \[

### 2024-04-22 draft/rectangulotopes.tex

--- 2024-04-22 draft/rectangulotopes.tex
@@
 \begin{figure}
-  \includegraphics[width=.9\textwidth]{rectangulations.pdf}
+  \includegraphics[width=\textwidth]{rectangulations.pdf}
 	\caption{Four rectangulations.
@@
   \]
-  For instance, the $4$th coordinate is $h^T_4 \cdot v^T_4 - h^S_4 \cdot v^S_4 = 1 \cdot 5 - 1 \cdot 3 = 2$.
-  The facet of~$\WRP (16)$ corresponding to the subset~$\{3,4,7,8,9,10,14\}$ is given by the inequality
-  \[
-  x_3 + x_4 + x_7 + x_8 + x_9 + x_{10} + x_{14} \le ???
-  \]
-  \vincent{Compute and check...}
-\end{example}
-\end{comment}
-\vincent{not sure if we want to reintroduce this example.}
+%  For instance, the $4$th coordinate is $h^T_4 \cdot v^T_4 - h^S_4 \cdot v^S_4 = 1 \cdot 5 - 1 \cdot 3 = 2$.
+%  The facet of~$\WRP (16)$ corresponding to the subset~$\{3,4,7,8,9,10,14\}$ is given by the inequality
+%  \[
+%  x_3 + x_4 + x_7 + x_8 + x_9 + x_{10} + x_{14} \le ???
+%  \]
+%  \vincent{Compute and check...}
+%\end{example}
 
@@
 
-\begin{figure}[b]
+\begin{figure}
 	\capstart
@@
 
-%\pagebreak
 \section{Weak rectangulotopes}
@@
 The weak poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
+See \cref{fig:weakPoset}.
+
+\begin{figure}
+	\centerline{\includegraphics[width=.4\textwidth]{weakPoset1} \qquad \includegraphics[width=.4\textwidth]{weakPoset2}}
+	\caption{A weak poset of the rectangulation of \cref{fig:strongRectangulation}, visualized on the rectangulation of \cref{fig:weakRectangulation} (left) and on its diagonal rectangulation of \cref{fig:weakRectangulation} (right).
+Example from \cite{ACFF24}.}
+	\label{fig:weakPoset}
+\end{figure}
 
@@
 \weakeq$ on the weak Bruhat order on $\f{S}_n$.
+
+\vincent{Add here sentence about geometric interpretation of weak poset. Bijection cover relations weak poset -- facets -- flips...}
 
@@
 Given a rectangulation $R$, we define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of $R$ by considering two relations on $[n]$~\cite{ACFF24}:
-\vincent{Add picture weak and strong orders?}
 \begin{enumerate}
@@
 The strong poset is the transitive closure of the union of the two relations $\tri$ and $\btri$.
+See \cref{fig:strongPoset}.
 The strong poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$. The following statement will be used later.
+
+\begin{figure}
+	\centerline{\includegraphics[width=.4\textwidth]{adjacencyPoset} \qquad \includegraphics[width=.4\textwidth]{strongPoset}}
+	\caption{A adjacency poset (left) and the strong poset (right) of the rectangulation of \cref{fig:strongRectangulation}. Example from \cite{ACFF24}.}
+	\label{fig:strongPoset}
+\end{figure}
 

### 2024-04-22 draft/rectangulotopes.tex
getting there

@@
 If the rectangulation $R$ is diagonal, we retrieve the well-studied pair of twin binary trees~\cite{MR1417289,MR2914637}, see~\cref{fig:weakRectangulation}.
+
 Recall that the \defn{inorder labeling} of a binary tree~$T$ is the labeling of the nodes of~$T$ such that the label of each node~$t$ of~$T$ is larger than all labels in the left subtree of~$t$ and larger than all labels in the right subtree of~$t$.
-%\begin{lemma}
 For any rectangle~$r$ of~$R$, the inorder label of the source~$s(r)$ in~$S(R)$ coincides with the inorder label of the target~$t(r)$ in~$T(R)$.
-%\end{lemma}
+The inorder can also be retrieved from the above-below and left-right relationships defined before, as the transitive closure of the union of the below and right partial orders.
 This enables to unambiguously label the rectangles of~$R$ by the inorder.
@@
 
-We denote by $i\verticalPattern j$ the situation where 
+We denote by $i\verticalPattern j$ the situation where the rectangles $r_i$ and $r_j$ of inorder labels $i$ and $j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
 and by $i\horizontalPattern j$ the situation where \Jean{Complete with the right definition here.}
@@
 \begin{theorem}[\cite{MR2871762}]
+  \label{thm:weakCong}
   The subsets $\mathcal{L}(P_w(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, 
@@
 
-\vincent{Add here sentence about geometric interpretation of weak poset. Bijection cover relations weak poset -- facets -- flips...}
+From \cref{thm:weakCong}, we obtain a bijection between the cover relations in the weak poset and the facets of the dual cone corresponding to the rectangulation, which are in turn one-to-one with feasible flips in the rectangulation.
 
@@
 \begin{theorem}[\cite{MR2864445,ACFF24}]
+  \label{thm:strongCong}
   The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, are equivalence classes of a congruence relation $\strongeq$ on the weak Bruhat order on $\f{S}_n$.
@@
 The \defn{strong rectangulotope}~$\SRP(n)$ is the quotientope of the strong rectangulation congruence~${\strongeq}$.
+
+Note that, similarly to the weak case, \cref{thm:strongCong} yields a bijection between the cover relations in the strong poset and the facets of the dual cone corresponding to the rectangulation, which are in turn one-to-one with feasible flips in the rectangulation.
 
@@
 \cref{lem:yincount} essentially gives us the coefficient $\yin{w}^R_{i,j}$ in \cref{thm:strongRectangulotope}, except that the condition $\neg i\verticalPattern j$ is replaced by $i\succ_s j$.
-
-Recall that $i\verticalPattern j$ denote the situation where the rectangles $r_i$ and $r_j$ of inorder labels $i$ and $j$ are such that $r_i$ and $r_j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
-
+Recall that $i\verticalPattern j$ denote the situation where the rectangles $r_i$ and $r_j$ of inorder labels $i$ and $j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
 Our aim is to provide a formula for the coordinates of the vertices of the strong rectangulotope that does not involve the computation of the strong poset or one of its linear extension.
-For that purpose, we prove that for every pair $i<j\in [n]$ such that $i\prec_s j$, hence which do not have any associated yin arc, we have either $cv^{T,S}_{i,j} = 0$ (hence the product in \cref{lem:yincount} is 0 as well and adding it does not harm the result), or $i\verticalPattern j$.
+For that purpose, we prove that for every pair $i<j\in [n]$ such that $i\prec_s j$, which do not have any associated yin arc, we have either $cv^{T,S}_{i,j} = 0$ (hence the product in \cref{lem:yincount} is 0 as well and adding it does not harm the result), or $i\verticalPattern j$.
 
@@
   Indeed, if we suppose otherwise, then the line segment from $s(r_j)$ to $t(r_i)$ has positive slope, which implies that $j\tri i$, which in turn implies $j\prec_s i$, contradicting our hypothesis.
-
-  We now claim that there can only be one segment on which leaves common to the vertical subtree of $i$ in $T$ and the vertical subtree of $j$ in $S$ may lie, and this segment is the vertical segment supporting the right edge of $r_i$ and the left edge of $r_j$.
-  \Jean{This requires a case analysis. Any shortcut?}
-
-  Let us assume that this segment is shared.
+  We also observe that since $i < j$, by definition of the inorder, the rectangle $r_j$ must be below or on the right of $r_i$.
+  Combining those two observations, we conclude that the left edge of $r_j$ cannot be on the left of the right edge of $r_i$.
+
+  Therefore, there can only be one segment on which a leaf common to the vertical subtree of $i$ in $T$ and the vertical subtree of $j$ in $S$ may lie, which is the vertical segment supporting the right edge of $r_i$ and the left edge of $r_j$.
+  Let us assume that this segment is shared by those two edges.
   If the bottom edge of $r_i$ is higher than the top edge of $r_j$, we must have $j\btri i$, which implies $j\prec_s i$, contradicting our hypothesis again.
-  Hence we must have that the bottom edge of $r_i$ is lower than the top edge of $r_j$, hence that $i\verticalPattern j$, as claimed.
+  Thus the bottom edge of $r_i$ is lower than the top edge of $r_j$, and $i\verticalPattern j$, as claimed.
 \end{proof}
 
-We omit the proofs of the analogues of \cref{lem:yincount} and \cref{lem:whentosum} for the case of the yang arcs: The arguments are completely symmetric.
+We omit the proofs of the analogues of \cref{lem:yincount} and \cref{lem:whentosum} for the case of the yang arcs, which use completely symmetric arguments.
 This concludes the proof of \cref{thm:strongRectangulotope}.

### 2024-04-23 draft/rectangulotopes.tex
minor changes

--- 2024-04-22 draft/rectangulotopes.tex
@@
 \newcommand{\weakeq}{\asymp}
-\newcommand{\strongeq}{\mathbin{\smash{\begin{smallmatrix} \backsim \\[-.3cm] \sim \end{smallmatrix}}}}% % J: I would like to stack \backsim and \sim instead
+\newcommand{\strongeq}{\mathbin{\smash{\begin{smallmatrix} \backsim \\[-.3cm] \sim \end{smallmatrix}}}}
 
@@
 This defines two partial orders on $[n]$.
-The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of these two partial orders defined by~$S(R)$ and~$T(R)$.
+The \defn{weak poset}~$P_w(R)=([n],\preccurlyeq_w)$ of the rectangulation~$R$ is the transitive closure of the union of these two partial orders defined by~$S(R)$ and~$T(R)$.
 Equivalently, the weak poset can be defined by considering the adjacency graph of the rectangles in a diagonal representative of the weak equivalence class of $R$, orienting its edges from left to right and from bottom to top, and taking the transitive closure of this directed acyclic graph.
@@
 We call this congruence the \defn{weak rectangulation congruence} (it is sometimes referred to as the \defn{Baxter congruence}).
-The \defn{weak rectangulotope} $\WRP(n)$ is the quotientope of the weak rectangulation congruence $
-\weakeq$ on the weak Bruhat order on $\f{S}_n$.
-
-\vincent{Add here sentence about geometric interpretation of weak poset. Bijection cover relations weak poset -- facets -- flips...}
+It is the intersection of the sylvester congruence with the antisylvester congruence.
+The \defn{weak rectangulotope} $\WRP(n)$ is the quotientope of the weak rectangulation congruence $\weakeq$ on the weak Bruhat order on $\f{S}_n$.
+
+By construction, the quotient fan~$\c{F}_\weakeq$ has a maximal cone~$C_R$ for each weak rectangulation~$R$ of size~$n$.
+\vincent{Notations? $\preccurlyeq_R^w$ and~$C_R^w$?}
+This cone is given by~$C_R = \set{\b{x} \in \R^n}{x_i \le x_j \text{ if } i \preccurlyeq_w j}$, where~$\preccurlyeq_w$ is the weak poset of~$R$.
+Note that there is a bijection between the facets of~$C_R$, the cover relations of~$\preccurlyeq_w$, and the flips from~$R$.
+The definition of the weak poset just translates the fact that the weak rectangulation fan~$\c{F}_\weakeq$ is the common refinement of the sylvester fan and the anti-sylvester fan.
+Namely, as~$C_R$ is the intersection of the cone of the sylvester fan corresponding to the target tree~$T(R)$ with the cone of the anti-sylvester fan corresponding to the source tree~$S(R)$, the weak poset~$\preccurlyeq_w$ is the transitive closure of the union of the posets associated to~$S(R)$ and~$T(R)$.
+
+Note that weak rectangulotopes were first constructed by S.~Law and N.~Reading in~\cite{MR2871762} as Minkowski sums of two opposite associahedra.
+Our vertex and facet descriptions of \cref{thm:weakRectangulotope} can be directly derived from this construction.
+Here, we provide a more pedestrian approach based on Minkowski sums of shard polytopes, as it will serve as a template for the more involved proof of \cref{thm:strongRectangulotope} in \cref{sec:strongRectangulotopes}.
 
@@
 Note that the basic arc~$(i, i+1, \varnothing, \varnothing)$ is both an up and down arc.
+The up arcs correspond to the sylvester congruence, while the down arcs correspond to the anti-sylvester congruence.
 
@@
 
-Given a rectangulation $R$, we define the \defn{strong poset} $P_s(R)=([n],\prec_s)$ of $R$ by considering two relations on $[n]$~\cite{ACFF24}:
+Given a rectangulation $R$, we define the \defn{strong poset} $P_s(R)=([n],\preccurlyeq_s)$ of $R$ by considering two relations on $[n]$~\cite{ACFF24}:
 \begin{enumerate}
@@
 
+By construction, the quotient fan~$\c{F}_{\strongeq}$ has a maximal cone~$C_R$ for each strong rectangulation~$R$ of size~$n$.
+This cone is given by~$C_R = \set{\b{x} \in \R^n}{x_i \le x_j \text{ if } i \preccurlyeq_s j}$, where~$\preccurlyeq_s$ is the strong poset of~$R$.
+Note that there is a bijection between the facets of~$C_R$, the cover relations of~$\preccurlyeq_s$, and the flips from~$R$.
+
 %%%%%%%%%%%%%
@@
 These equivalence classes are in turn in bijection with rectangulations that avoid one of the two patterns defining diagonal rectangulations, see~\cite{ACFF24}.
+\vincent{Move this}
 
@@
   \label{lem:yincount}
-  Let $R$ be a strong rectangulation of size~$n$, let~$S$ and $T$ be its source and target trees, let~$([n],\prec_s)$ be its strong poset, and let~$\b{p}(R)$ be the vertex of $\SRP(n)$ corresponding to $R$.
+  Let $R$ be a strong rectangulation of size~$n$, let~$S$ and $T$ be its source and target trees, let~$([n],\preccurlyeq_s)$ be its strong poset, and let~$\b{p}(R)$ be the vertex of $\SRP(n)$ corresponding to $R$.
   Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $\alpha$ such that $\b{e}_i-\b{e}_j$ is the extremal vertex of~$\SP(\alpha)$ contributing to $\b{p}(R)$ is
@@
   \label{lem:whentosum}
-  Let $R$ be a rectangulation of size~$n$, let $S$ and $T$ be its source and target trees, and let~$([n],\prec_s)$ be the associated strong order.
+  Let $R$ be a rectangulation of size~$n$, let $S$ and $T$ be its source and target trees, and let~$([n],\preccurlyeq_s)$ be the associated strong order.
   Given a pair $i,j\in [n]$ such that $i<j$ and $i\prec_s j$, we have

### 2024-04-23 draft/rectangulotopes.tex

@@
 Here, we provide a more pedestrian approach based on Minkowski sums of shard polytopes, as it will serve as a template for the more involved proof of \cref{thm:strongRectangulotope} in \cref{sec:strongRectangulotopes}.
+\jean{From \cref{thm:weakCong}, we obtain a bijection between the cover relations in the weak poset and the facets of the dual cone corresponding to the rectangulation, which are in turn one-to-one with feasible flips in the rectangulation.}
 
@@
 Note that there is a bijection between the facets of~$C_R$, the cover relations of~$\preccurlyeq_s$, and the flips from~$R$.
+\jean{Note that, similarly to the weak case, \cref{thm:strongCong} yields a bijection between the cover relations in the strong poset and the facets of the dual cone corresponding to the rectangulation, which are in turn one-to-one with feasible flips in the rectangulation.}
 

### 2024-04-23 draft/rectangulotopes.tex
add rectangulotopes already constructed

@@
 \author{Vincent Pilaud}
-\address{Universitat de Barcelona}
+\address{Universitat de Barcelona (UB)}
 \email{vincent.pilaud@ub.edu}
@@
 The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
+We note that weak rectangulotopes were already constructed as Minkowski sums of two opposite associahedra in~\cite{MR2871762}, before the general constructions of quotientopes of~\cite{MR3964495,MR4584712}.
 For~$n = 4$, the weak rectangulation lattice is represented in \cref{fig:weakRectangulationLattice} and the weak rectangulotope is represented in \cref{fig:weakRectangulotope}.
@@
 The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by~$\SRP(n)$.
+Strong rectangulotopes can be obtained as quotientopes~\cite{MR3964495,MR4584712} for the strong rectangulation congruence.
 For~$n = 4$, the strong rectangulation lattice is represented in \cref{fig:strongRectangulationLattice} and the strong rectangulotope is represented in \cref{fig:strongRectangulotope}.
@@
   \]
-  (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and 0 otherwise.)
+  (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and~$0$ otherwise.)
   Similarly, the shard polytope~$\SP(\downArc{I})$ of a down arc~$\downArc{I}$ is given by the submodular function
@@
   \]
-  (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and 0 otherwise.)
+  (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and~$0$ otherwise.)
   Similarly, the shard polytope~$\SP(\yangArc{I}{J})$ of a yang arc~$\yangArc{I}{J}$ is given by the submodular function
@@
 \cref{lem:yincount} essentially gives us the coefficient $\yin{w}^R_{i,j}$ in \cref{thm:strongRectangulotope}, except that the condition $\neg i\verticalPattern j$ is replaced by $i\succ_s j$.
-Recall that $i\verticalPattern j$ denote the situation where the rectangles $r_i$ and $r_j$ of inorder labels $i$ and $j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
+Recall that $i\verticalPattern j$ denote the situation where the rectangles $r_i$ and~$r_j$ of inorder labels $i$ and $j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
 Our aim is to provide a formula for the coordinates of the vertices of the strong rectangulotope that does not involve the computation of the strong poset or one of its linear extension.
-For that purpose, we prove that for every pair $i<j\in [n]$ such that $i\prec_s j$, which do not have any associated yin arc, we have either $cv^{T,S}_{i,j} = 0$ (hence the product in \cref{lem:yincount} is 0 as well and adding it does not harm the result), or $i\verticalPattern j$.
+For that purpose, we prove that for every pair $i<j\in [n]$ such that~$i\prec_s j$, which do not have any associated yin arc, we have either $cv^{T,S}_{i,j} = 0$ (hence the product in \cref{lem:yincount} is~$0$ as well and adding it does not harm the result), or $i\verticalPattern j$.
 
@@
   Let $R$ be a rectangulation of size~$n$, let $S$ and $T$ be its source and target trees, and let~$([n],\preccurlyeq_s)$ be the associated strong order.
-  Given a pair $i,j\in [n]$ such that $i<j$ and $i\prec_s j$, we have
+  Given a pair $i,j\in [n]$ such that $i<j$ and $i\prec_s j$, we~have
   \[

### 2024-04-23 draft/rectangulotopes.tex
minor change feasible flips

@@
 This cone is given by~$C_R = \set{\b{x} \in \R^n}{x_i \le x_j \text{ if } i \preccurlyeq_w j}$, where~$\preccurlyeq_w$ is the weak poset of~$R$.
-Note that there is a bijection between the facets of~$C_R$, the cover relations of~$\preccurlyeq_w$, and the flips from~$R$.
+Note that there is a bijection between the facets of~$C_R$, the cover relations of~$\preccurlyeq_w$, and the feasible flips in~$R$.
 The definition of the weak poset just translates the fact that the weak rectangulation fan~$\c{F}_\weakeq$ is the common refinement of the sylvester fan and the anti-sylvester fan.
@@
 This cone is given by~$C_R = \set{\b{x} \in \R^n}{x_i \le x_j \text{ if } i \preccurlyeq_s j}$, where~$\preccurlyeq_s$ is the strong poset of~$R$.
-Note that there is a bijection between the facets of~$C_R$, the cover relations of~$\preccurlyeq_s$, and the flips from~$R$.
+Note that there is a bijection between the facets of~$C_R$, the cover relations of~$\preccurlyeq_s$, and the feasible flips in~$R$.
 \jean{Note that, similarly to the weak case, \cref{thm:strongCong} yields a bijection between the cover relations in the strong poset and the facets of the dual cone corresponding to the rectangulation, which are in turn one-to-one with feasible flips in the rectangulation.}

### 2024-04-23 draft/rectangulotopes.tex
last important changes

--- 2024-04-23 draft/rectangulotopes.tex
@@
 	\capstart
-	\centerline{\includegraphics[scale=.75]{permutahedronLeft4} \quad \raisebox{-.4cm}{\includegraphics[scale=.75]{associahedron4}}}
+	\centerline{\includegraphics[scale=.75]{permutahedronLeft4} \quad \raisebox{-.5cm}{\includegraphics[scale=.75]{associahedron4}}}
 	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$ (right). \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
@@
 \begin{itemize}
-\item \defn{Simple flips} replace a vertical segment incident to exactly two rectangles by a horizontal segment, and vice-versa,
-\item \defn{Pivoting flips} transform a pair of adjacent rectangles whose union is not a rectangle, changing a left-right pair into an above-below pair, and vice-versa,
-\item \defn{Wall slide flips} are wall slides that modify the adjacency of the rectangles, removing exactly one adjacency and adding exactly one.
-These flips are only relevant in strong rectangulations.
+\item \defn{simple flips} replace a vertical segment incident to exactly two rectangles by a horizontal segment, and vice-versa,
+\item \defn{pivoting flips} transform a pair of adjacent rectangles whose union is not a rectangle, changing a left-right pair into an above-below pair, and vice-versa,
+\item \defn{wall slide flips} are wall slides that modify the adjacency of the rectangles, removing exactly one adjacency and adding exactly one. These are only relevant in strong rectangulations.
 \end{itemize}
@@
 In order to exhibit a similar realization for the strong rectangulotopes, we need to count subsets of leaves that lie in some subtrees of both the source and target trees of a rectangulation.
+%
 Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$.
-
-We denote by $i\verticalPattern j$ the situation where the rectangles $r_i$ and $r_j$ of inorder labels $i$ and $j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
-and by $i\horizontalPattern j$ the situation where \Jean{Complete with the right definition here.}
-
+%
+We denote by ${}^i \!\!\!
+\raisebox{.03cm}{\verticalPattern} \!\!\!
+{}_j$ the situation where the rectangles $r_i$ and $r_j$ of inorder labels $i$ and $j$ share a vertical segment containing the right edge of~$r_i$ and the left edge of~$r_j$, and the bottom edge of $r_i$ is below the top edge of~$r_j$.
+Similarly, we denote by~$\; {}^i \!\!\!\!\!
+\horizontalPattern \!\!\!\!\!
+\raisebox{-.06cm}{$_j$} \,$ the situation where the rectangles $r_i$ and $r_j$ share a horizontal segment containing the bottom edge of~$r_i$ and the top edge of~$r_j$, and the right edge of~$r_i$ is on the right of the left edge of~$r_j$.
 We also use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.
@@
   \[
-    \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \neg \, i \verticalPattern j \rrbracket
+    \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \; \neg \; {}^i \!\!\! \raisebox{.03cm}{\verticalPattern} \!\!\! {}_j \; \rrbracket
     \qquad\text{and}\qquad
-    \yang{w}^R_{i,j} \eqdef v^S_i \cdot ch^{S,T}_{i,j}\cdot v^T_j \cdot \llbracket \neg \, j \horizontalPattern i \rrbracket,
+    \yang{w}^R_{i,j} \eqdef v^S_i \cdot ch^{S,T}_{i,j}\cdot v^T_j \cdot \llbracket \; \neg \;\; {}^i \!\!\!\!\! \horizontalPattern \!\!\!\!\! \raisebox{-.06cm}{$_j$} \;\; \rrbracket,
   \]
@@
 	\capstart
-	\centerline{\includegraphics[scale=.75]{braidFanLeft4} \quad \raisebox{-.3cm}{\includegraphics[scale=.75]{sylvesterFan4}}}
+	\centerline{\includegraphics[scale=.75]{braidFanLeft4} \quad \raisebox{-.35cm}{\includegraphics[scale=.75]{sylvesterFan4}}}
 	\caption{The braid fan (left) and the sylvester fan (right). \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
@@
 
+\pagebreak
 A set function $f:2^{[n]}\mapsto \R$ is said to be \defn{submodular} when $f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$.
@@
 This defines two partial orders on $[n]$.
-The \defn{weak poset}~$P_w(R)=([n],\preccurlyeq_w)$ of the rectangulation~$R$ is the transitive closure of the union of these two partial orders defined by~$S(R)$ and~$T(R)$.
+The \defn{weak poset}~$([n],\prec_w^R)$ of the rectangulation~$R$ is the transitive closure of the union of these two partial orders defined by~$S(R)$ and~$T(R)$.
 Equivalently, the weak poset can be defined by considering the adjacency graph of the rectangles in a diagonal representative of the weak equivalence class of $R$, orienting its edges from left to right and from bottom to top, and taking the transitive closure of this directed acyclic graph.
@@
 
-We define the set~$\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
-\[
-\mathcal{L}(P_w(R)) \eqdef \bigset{\pi\in\f{S}_n }{ i\prec_w j\implies \pi^{-1}(i) < \pi^{-1}(j)}.
+We define the set~$\mathcal{L}(\prec_w^R)$ as the linear extensions of the weak poset:
+\[
+\mathcal{L}(\prec_w^R) \eqdef \bigset{\pi\in\f{S}_n }{ i\prec_w^R j\implies \pi^{-1}(i) < \pi^{-1}(j)}.
 \]
@@
 
+\pagebreak
+
 \begin{theorem}[\cite{MR2871762}]
   \label{thm:weakCong}
-  The subsets $\mathcal{L}(P_w(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, 
+  The subsets $\mathcal{L}(\prec_w^R)$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, 
   are equivalence classes of a congruence relation $\weakeq$ on the weak Bruhat order on $\f{S}_n$.
@@
 We call this congruence the \defn{weak rectangulation congruence} (it is sometimes referred to as the \defn{Baxter congruence}).
-It is the intersection of the sylvester congruence with the antisylvester congruence.
+It is the intersection of the sylvester congruence with the anti-sylvester congruence.
 The \defn{weak rectangulotope} $\WRP(n)$ is the quotientope of the weak rectangulation congruence $\weakeq$ on the weak Bruhat order on $\f{S}_n$.
 
-By construction, the quotient fan~$\c{F}_\weakeq$ has a maximal cone~$C_R$ for each weak rectangulation~$R$ of size~$n$.
-\vincent{Notations? $\preccurlyeq_R^w$ and~$C_R^w$?}
-This cone is given by~$C_R = \set{\b{x} \in \R^n}{x_i \le x_j \text{ if } i \preccurlyeq_w j}$, where~$\preccurlyeq_w$ is the weak poset of~$R$.
-Note that there is a bijection between the facets of~$C_R$, the cover relations of~$\preccurlyeq_w$, and the flips from~$R$.
+By construction, the quotient fan~$\c{F}_\weakeq$ has a maximal cone~$\polytope{C}_w^R$ for each weak rectangulation~$R$ of size~$n$.
+This cone is given by~$\polytope{C}_w^R = \set{\b{x} \in \R^n}{x_i \le x_j \text{ if } i \prec_w^R j}$. %, where~$\prec_w$ is the weak poset of~$R$.
+Note that there is a bijection between the facets of~$\polytope{C}_w^R$, the cover relations of~$\prec_w^R$, and the feasible flips in~$R$.
 The definition of the weak poset just translates the fact that the weak rectangulation fan~$\c{F}_\weakeq$ is the common refinement of the sylvester fan and the anti-sylvester fan.
-Namely, as~$C_R$ is the intersection of the cone of the sylvester fan corresponding to the target tree~$T(R)$ with the cone of the anti-sylvester fan corresponding to the source tree~$S(R)$, the weak poset~$\preccurlyeq_w$ is the transitive closure of the union of the posets associated to~$S(R)$ and~$T(R)$.
+Namely, as~$\polytope{C}_w^R$ is the intersection of the cone of the sylvester fan corresponding to the target tree~$T(R)$ with the cone of the anti-sylvester fan corresponding to the source tree~$S(R)$, the weak poset~$\prec_w^R$ is the transitive closure of the union of the posets associated to~$S(R)$ and~$T(R)$.
 
@@
 Here, we provide a more pedestrian approach based on Minkowski sums of shard polytopes, as it will serve as a template for the more involved proof of \cref{thm:strongRectangulotope} in \cref{sec:strongRectangulotopes}.
-\jean{From \cref{thm:weakCong}, we obtain a bijection between the cover relations in the weak poset and the facets of the dual cone corresponding to the rectangulation, which are in turn one-to-one with feasible flips in the rectangulation.}
 
@@
 \begin{proof}
-  Let $\pi$ be any permutation in $\mathcal{L}(P_w(R))$.
+  Let $\pi$ be any permutation in $\mathcal{L}(\prec_w^R)$.
   From \cref{lem:lodaymax}, for an $i\in [n]$, a shard polytope~$\SP(\upArc{I})$ contributes to $\b{e}_i$ if and only if $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
@@
   Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
-  Since the weak poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
+  Since the weak poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s^R i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
   Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
@@
 
-Given a rectangulation $R$, we define the \defn{strong poset} $P_s(R)=([n],\preccurlyeq_s)$ of $R$ by considering two relations on $[n]$~\cite{ACFF24}:
+Given a rectangulation $R$, we define the \defn{strong poset} $([n],\prec_s^R)$ of $R$ by considering two relations on $[n]$~\cite{ACFF24}:
 \begin{enumerate}
 \item The \defn{adjacency poset} $([n],\tri)$ of $R$, defined by considering the adjacency graph of the rectangles of $R$ (identified by their inorder labels in $[n]$), orienting its edges from left to right and from bottom to top, and taking the transitive closure of this directed acyclic graph.
-\item The \defn{special relation} $([n],\btri)$ of $R$, in which $a\btri b$ if the two rectangles $r_a$ and $r_b$ with respective inorder labels $a$ and $b$ are such that 
+\item The \defn{special relation} $([n],\btri)$ of $R$, in which $i\btri j$ if the two rectangles $r_i$ and $r_j$ with respective inorder labels $i$ and $j$ are such that 
 	\begin{enumerate}[(i)]
-	\item either the right side of $r_b$ and the left side of $r_a$ lie on the same vertical segment, and the bottom-right corner of $r_b$ is above the top-left corner of $r_a$, 
-	\item or the top side of $r_b$ and the bottom side of rectangle $r_a$ lie on the same horizontal segment, and the top-left corner of $r_b$ lies on the right of the bottom-right corner~of~$r_a$.
+	\item either the right side of $r_j$ and the left side of $r_i$ lie on the same vertical segment, and the bottom-right corner of $r_j$ is above the top-left corner of $r_i$, 
+	\item or the top side of $r_j$ and the bottom side of rectangle $r_i$ lie on the same horizontal segment, and the top-left corner of $r_j$ lies on the right of the bottom-right corner~of~$r_i$.
 	\end{enumerate}
@@
   \label{lem:sextw}
-  The strong poset $P_s(R)$ of a rectangulation $R$ is an extension of its weak poset $P_w(R)$.
+  The strong poset $\prec_s^R$ of a rectangulation $R$ is an extension of its weak poset $\prec_w^R$.
 \end{lemma}
@@
   Clearly, the strong poset is an extension of the adjacency poset $([n],\tri)$ of $R$, since it is obtained by adding pairs from the special relation~$([n],\btri)$.
-  It remains to observe that the adjacency poset of $R$ is an extension of the weak poset of $R$.
+  It remains to observe that the adjacency poset of $R$ is an extension of the weak poset of $R$ (compare for instance \cref{fig:weakPoset}\,(left) and \cref{fig:strongPoset}\,(left)).
   For this, we consider a diagonal representative~$D$ of the weak equivalence class of $R$, and recall that the weak poset of $R$ is the adjacency poset of $D$.
@@
 
-Let $\mathcal{L}(P_s)$ be the set of permutations of $\f{S}_n$ corresponding to linear extensions of the strong poset $P_s$:
-\[
-\mathcal{L}(P_s(R)) \eqdef \bigset{\pi\in\f{S}_n }{ i\prec_s j\implies \pi^{-1}(i) < \pi^{-1}(j)}.
+Let $\mathcal{L}(\prec_s^R)$ be the set of permutations of $\f{S}_n$ corresponding to linear extensions of the strong poset $\prec_s^R$:
+\[
+\mathcal{L}(\prec_s^R) \eqdef \bigset{\pi\in\f{S}_n }{ i\prec_s^R j\implies \pi^{-1}(i) < \pi^{-1}(j)}.
 \]
@@
 
+\pagebreak
 \begin{theorem}[\cite{MR2864445,ACFF24}]
   \label{thm:strongCong}
-  The subsets $\mathcal{L}(P_s(R))$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, are equivalence classes of a congruence relation $\strongeq$ on the weak Bruhat order on $\f{S}_n$.
+  The subsets $\mathcal{L}(\prec_s^R)$ of $\f{S}_n$, for all rectangulations $R$ of size~$n$, are equivalence classes of a congruence relation $\strongeq$ on the weak Bruhat order on $\f{S}_n$.
     Furthermore:
@@
 
-By construction, the quotient fan~$\c{F}_{\strongeq}$ has a maximal cone~$C_R$ for each strong rectangulation~$R$ of size~$n$.
-This cone is given by~$C_R = \set{\b{x} \in \R^n}{x_i \le x_j \text{ if } i \preccurlyeq_s j}$, where~$\preccurlyeq_s$ is the strong poset of~$R$.
-Note that there is a bijection between the facets of~$C_R$, the cover relations of~$\preccurlyeq_s$, and the flips from~$R$.
-\jean{Note that, similarly to the weak case, \cref{thm:strongCong} yields a bijection between the cover relations in the strong poset and the facets of the dual cone corresponding to the rectangulation, which are in turn one-to-one with feasible flips in the rectangulation.}
+Similarly to the weak case, the quotient fan~$\c{F}_{\strongeq}$ has a maximal cone~$\polytope{C}_s^R$ for each strong rectangulation~$R$ of size~$n$.
+This cone is given by~$\polytope{C}_s^R = \set{\b{x} \in \R^n}{x_i \le x_j \text{ if } i \prec_s^R j}$.
+Note that there is a bijection between the facets of~$\polytope{C}_s^R$, the cover relations of~$\prec_s^R$, and the feasible flips in~$R$.
 
@@
 
+Note that the yin (resp.~yang) arcs alone form an arc ideal, and we call \defn{yin} (resp.~\defn{yang}) \defn{congruence} the corresponding congruence of the weak order.
+It can be checked that the equivalence classes of these congruences are counted by \defn{semi-Baxter permutations}, defined as permutations avoiding $2\underline{41}3$, hence only one of the two forbidden Baxter patterns.
+These equivalence classes are in turn in bijection with rectangulations that avoid one of the two patterns defining diagonal rectangulations, see~\cite{ACFF24}.
+The strong rectangulation congruence is the intersection of the yin and yang congruences.
+
 %%%%%%%%%%%%%
@@
 
-\cref{prop:strongMinkowski} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes defined by the \defn{yin} and \defn{yang congruences} respectively, where the yin (resp.~yang) congruence is defined by the arc ideal consisting of all yin (resp.~yang) arcs.
-It can be checked that the equivalence classes of these congruences are counted by \defn{semi-Baxter permutations}, defined as permutations avoiding $2\underline{41}3$, hence only one of the two forbidden Baxter patterns.
-These equivalence classes are in turn in bijection with rectangulations that avoid one of the two patterns defining diagonal rectangulations, see~\cite{ACFF24}.
-\vincent{Move this}
+\cref{prop:strongMinkowski} gives us an alternative description of $\SRP(n)$ as the sum of two opposite quotientopes corresponding to the yin and yang congruences.
 
@@
   \label{lem:yincount}
-  Let $R$ be a strong rectangulation of size~$n$, let~$S$ and $T$ be its source and target trees, let~$([n],\preccurlyeq_s)$ be its strong poset, and let~$\b{p}(R)$ be the vertex of $\SRP(n)$ corresponding to $R$.
+  Let $R$ be a strong rectangulation of size~$n$, let~$S$ and $T$ be its source and target trees, let~$([n],\prec_s^R)$ be its strong poset, and let~$\b{p}(R)$ be the vertex of $\SRP(n)$ corresponding to $R$.
   Given a pair $i,j\in [n]$ with $i<j$, the number of yin arcs $\alpha$ such that $\b{e}_i-\b{e}_j$ is the extremal vertex of~$\SP(\alpha)$ contributing to $\b{p}(R)$ is
@@
 \begin{proof}
-  Let $\pi$ be any permutation in~$\mathcal{L}(P_s(R))$.
+  Let $\pi$ be any permutation in~$\mathcal{L}(\prec_s^R)$.
   From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc~$\alpha$ whose shard polytope $\SP(\alpha)$ contributes to $\b{e}_i-\b{e}_j$ if and only if $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.
@@
   Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
-  Since from \cref{lem:sextw} the strong poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
+  Since from \cref{lem:sextw} the strong poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s^R i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
   Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of~$\pi^{-1}$ in $I$.
@@
 
-\cref{lem:yincount} essentially gives us the coefficient $\yin{w}^R_{i,j}$ in \cref{thm:strongRectangulotope}, except that the condition $\neg i\verticalPattern j$ is replaced by $i\succ_s j$.
-Recall that $i\verticalPattern j$ denote the situation where the rectangles $r_i$ and~$r_j$ of inorder labels $i$ and $j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
+\cref{lem:yincount} essentially gives us the coefficient $\yin{w}^R_{i,j}$ in \cref{thm:strongRectangulotope}, except that the condition $\neg \; {}^i \!\!\!
+\raisebox{.03cm}{\verticalPattern} \!\!\!
+{}_j$ is replaced by $i\succ_s j$.
+Recall that ${}^i \!\!\!
+\raisebox{.03cm}{\verticalPattern} \!\!\!
+{}_j$ denote the situation where the rectangles $r_i$ and~$r_j$ of inorder labels $i$ and $j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
 Our aim is to provide a formula for the coordinates of the vertices of the strong rectangulotope that does not involve the computation of the strong poset or one of its linear extension.
-For that purpose, we prove that for every pair $i<j\in [n]$ such that~$i\prec_s j$, which do not have any associated yin arc, we have either $cv^{T,S}_{i,j} = 0$ (hence the product in \cref{lem:yincount} is~$0$ as well and adding it does not harm the result), or $i\verticalPattern j$.
+For that purpose, we prove that for every pair $i<j\in [n]$ such that~$i\prec_s^R j$, which do not have any associated yin arc, we have either ${}^i \!\!\!
+\raisebox{.03cm}{\verticalPattern} \!\!\!
+{}_j$ or $cv^{T,S}_{i,j} = 0$ (hence the product in \cref{lem:yincount} is~$0$ as well and adding it does not harm the result).
 
@@
   \label{lem:whentosum}
-  Let $R$ be a rectangulation of size~$n$, let $S$ and $T$ be its source and target trees, and let~$([n],\preccurlyeq_s)$ be the associated strong order.
-  Given a pair $i,j\in [n]$ such that $i<j$ and $i\prec_s j$, we~have
-  \[
-    cv^{T,S}_{i,j} > 0 \quad \Longrightarrow \quad i \verticalPattern j.
+  Let $R$ be a rectangulation of size~$n$, with source and target trees~$S$ and~$T$, and strong order~$\prec_s^R$.
+  Given a pair $i,j\in [n]$ such that $i<j$ and $i\prec_s^R j$, we~have
+  \[
+    cv^{T,S}_{i,j} > 0 \quad \Longrightarrow \quad {}^i \!\!\! \raisebox{.03cm}{\verticalPattern} \!\!\! {}_j.
   \]
@@
   Consider the two rectangles $r_i$ and $r_j$ with inorder labels $i$ and $j$ in $R$.
-  We first observe that the source vertex $s(r_j)$ is not contained in the interior of the lower left quadrant of origin $t(r_i)$.
-  Indeed, if we suppose otherwise, then the line segment from $s(r_j)$ to $t(r_i)$ has positive slope, which implies that $j\tri i$, which in turn implies $j\prec_s i$, contradicting our hypothesis.
-  We also observe that since $i < j$, by definition of the inorder, the rectangle $r_j$ must be below or on the right of $r_i$.
-  Combining those two observations, we conclude that the left edge of $r_j$ cannot be on the left of the right edge of $r_i$.
+  We first observe that the source vertex $s(r_j)$ is not contained in the interior of the lower left quadrant of origin~$t(r_i)$.
+  Indeed, if we assume otherwise, then the line segment from $s(r_j)$ to $t(r_i)$ has positive slope, which implies that $j\tri i$, which in turn implies $j\prec_s^R i$, contradicting our hypothesis.
+  We also observe that since $i < j$, by definition of the inorder on~$S$ and~$T$, the rectangle $r_j$ must be below or on the right of~$r_i$.
+  Combining those two observations, we conclude that the right edge of $r_i$ cannot be on the right of the left edge of~$r_j$.
 
@@
   Let us assume that this segment is shared by those two edges.
-  If the bottom edge of $r_i$ is higher than the top edge of $r_j$, we must have $j\btri i$, which implies $j\prec_s i$, contradicting our hypothesis again.
-  Thus the bottom edge of $r_i$ is lower than the top edge of $r_j$, and $i\verticalPattern j$, as claimed.
+  If the bottom edge of $r_i$ is higher than the top edge of $r_j$, we must have $j\btri i$, which implies $j\prec_s^R i$, contradicting our hypothesis again.
+  Thus the bottom edge of $r_i$ is lower than the top edge of $r_j$, and ${}^i \!\!\! \raisebox{.03cm}{\verticalPattern} \!\!\! {}_j$, as claimed.
 \end{proof}

### 2024-04-23 draft/rectangulotopes.tex
added references

@@
 
-The combinatorics of rectangulations has been studied for nearly two decades~\cite{MR2233287,MR2763051,MR2871762,MR2864445,MR3084577,MR3192492,MR3878132,MR4598046}.
-Interestingly, several combinatorial results were initially motivated by applications to the design of \defn{floorplans} for very large scale integrated circuits, and published in electrical engineering journals, see for instance R.
+The combinatorics of rectangulations has been studied for nearly two decades~\cite{MR2233287,MR2763051,MR2871762,MR2864445,MR3084577,MR3192492,MR3878132,MR4598046,AB24,ACFF24}\footnote{In fact, several combinatorial ideas can even be traced back to the problem of ``squaring the square'' studied since the 1930s~\cite{A30,Abe32,MR0000470,MR0003040}.
+Also, some results were initially motivated by applications to the design of \defn{floorplans} for very large scale integrated circuits, and published in electrical engineering journals, see for instance R.
 Fujimaki, Y.
 Inoue, and T.
-Takahashi~\cite{FT07,TF08,ITF09,FIT09}.
+Takahashi~\cite{FT07,TF08,ITF09,FIT09}.}
 
@@
 
-\pagebreak
+%\pagebreak
 A set function $f:2^{[n]}\mapsto \R$ is said to be \defn{submodular} when $f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$.
@@
 
-\pagebreak
+%\pagebreak
 
@@
 	\centerline{\includegraphics[width=.4\textwidth]{adjacencyPoset} \qquad \includegraphics[width=.4\textwidth]{strongPoset}}
-	\caption{A adjacency poset (left) and the strong poset (right) of the rectangulation of \cref{fig:strongRectangulation}. Example from \cite{ACFF24}.}
+	\caption{The adjacency poset (left) and the strong poset (right) of the rectangulation of \cref{fig:strongRectangulation}. Example from \cite{ACFF24}.}
 	\label{fig:strongPoset}
@@
   \label{lem:sextw}
-  The strong poset $\prec_s^R$ of a rectangulation $R$ is an extension of its weak poset $\prec_w^R$.
+  The strong poset $([n], \prec_s^R)$ of a rectangulation $R$ is an extension of its weak poset $([n], \prec_w^R)$.
 \end{lemma}
@@
 
-\pagebreak
+%\pagebreak
 \begin{theorem}[\cite{MR2864445,ACFF24}]

### 2024-04-23 draft/rectangulotopes.tex
added references

@@
 The combinatorics of rectangulations has been studied for nearly two decades~\cite{MR2233287,MR2763051,MR2871762,MR2864445,MR3084577,MR3192492,MR3878132,MR4598046,AB24,ACFF24}\footnote{In fact, several combinatorial ideas can even be traced back to the problem of ``squaring the square'' studied since the 1930s~\cite{A30,Abe32,MR0000470,MR0003040}.
-Also, some results were initially motivated by applications to the design of \defn{floorplans} for very large scale integrated circuits, and published in electrical engineering journals, see for instance R.
+Also, some results were initially motivated by applications to the design of \defn{floorplans} for very large scale integrated circuits, and published in electrical engineering journals, see for instance Z.
+C.
+Shen and C.
+C.
+N.
+Chu~\cite{SC03}, and R.
 Fujimaki, Y.

### 2024-04-23 draft/rectangulotopes.tex
minor changes

@@
 
-The combinatorics of rectangulations has been studied for nearly two decades~\cite{MR2233287,MR2763051,MR2871762,MR2864445,MR3084577,MR3192492,MR3878132,MR4598046,AB24,ACFF24}\footnote{In fact, several combinatorial ideas can even be traced back to the problem of ``squaring the square'' studied since the 1930s~\cite{A30,Abe32,MR0000470,MR0003040}.
+The combinatorics of rectangulations has been studied for nearly two decades~\cite{MR2233287,MR2763051,MR2871762,MR2864445,MR3084577,MR3192492,MR3878132,MR4598046,AB24,ACFF24}.
+In fact, several combinatorial ideas can even be traced back to the problem of ``squaring the square'' studied since the 1930s~\cite{A30,Abe32,MR0000470,MR0003040}.
 Also, some results were initially motivated by applications to the design of \defn{floorplans} for very large scale integrated circuits, and published in electrical engineering journals, see for instance Z.
@@
 Inoue, and T.
-Takahashi~\cite{FT07,TF08,ITF09,FIT09}.}
+Takahashi~\cite{FT07,TF08,ITF09,FIT09}.
 
@@
 	}
+	\vspace{-.2cm}
 	\caption{The facets of the weak rectangulotope~$\WRP(4)$ of \cref{fig:weakRectangulotope}.}
@@
 	}
+	\vspace{-.2cm}
 	\caption{The facets of the strong rectangulotope~$\SRP(4)$ of \cref{fig:strongRectangulotope}.}
@@
 
+\enlargethispage{.2cm}
 We think of the formula of \cref{thm:strongRectangulotope}\,(i) as the analogue of J.-L.-Loday's product formula of \cref{thm:loday}\,(i) for the coordinates of the vertices of the associahedron.
@@
 
-%\pagebreak
+\pagebreak
 A set function $f:2^{[n]}\mapsto \R$ is said to be \defn{submodular} when $f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$.
@@
 
-%\pagebreak
+\pagebreak
 
@@
 
-%\pagebreak
+\pagebreak
 \begin{theorem}[\cite{MR2864445,ACFF24}]

### 2024-04-24 draft/rectangulotopes.tex
minor

@@
   Rectangulations are decompositions of a square into finitely many axis-aligned rectangles.
-  We describe realizations of $(n-1)$-dimensional polytopes associated with two families of rectangulations composed of $n$ rectangles.
+  We describe realizations of $(n-1)$-dimensional polytopes associated with two combinatorial families of rectangulations composed of $n$ rectangles.
   They are defined as quotientopes of natural lattice congruence on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
@@
 
-Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined combinatorially by D.~Tamari~\cite{T51} and by J.~Stasheff~\cite{S63} in a topological context.
+Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined by D.~Tamari~\cite{T51} and by J.~Stasheff~\cite{S63} in a topological context.
 It has since then been identified as a fundamental object in many other areas of mathematics, including operads, cluster algebras, combinatorial Hopf algebras, and physics (see for instance the recent survey from V.
@@
 \item \defn{simple flips} replace a vertical segment incident to exactly two rectangles by a horizontal segment, and vice-versa,
-\item \defn{pivoting flips} transform a pair of adjacent rectangles whose union is not a rectangle, changing a left-right pair into an above-below pair, and vice-versa,
+\item \defn{pivoting flips} transform a pair of adjacent rectangles whose union is L-shaped, changing a left-right pair into an above-below pair, and vice-versa,
 \item \defn{wall slide flips} are wall slides that modify the adjacency of the rectangles, removing exactly one adjacency and adding exactly one. These are only relevant in strong rectangulations.
 \end{itemize}
-
 The various types of flips are illustrated in \cref{fig:flips}.
@@
 For vertices, we describe the coordinates of the vertex corresponding to a weak or strong rectangulation with simple product formulas in the spirit of J.-L.~Loday's realization in~\cref{thm:loday}\,(i).
-As mentioned in~\cite{PilaudSantosZiegler}, there is no such formula for arbitrary quotientopes, which makes rectangulotopes join the restricted club of Loday type quotientopes, jointly with associahedra and permutreehedra~\cite{MR3856522}.
+As mentioned in~\cite{PilaudSantosZiegler}, there is no such formula for arbitrary quotientopes, which makes rectangulotopes join the restricted club of Loday type quotientopes, together with associahedra and permutreehedra~\cite{MR3856522}.
 For facets, we show that both weak and strong rectangulotopes have $2^n-2$ facets (the maximal possible number of facets of a deformed permutahedron), and provide simple formulas for the right hand side of the inequality corresponding to a given proper nonempty subset of~$[n]$.
@@
 
-\subsection{Weak rectangulotopes}
+\subsection{Realizations of weak rectangulotopes}
 \label{subsec:weakRectangulotopes}
@@
 
-\subsection{Strong rectangulotopes}
+\subsection{Realizations of strong rectangulotopes}
 \label{subsec:strongRectangulotopes}
@@
 
+\jean{We use the notation $\pi^{-1}(i)$ later. Should we switch to $\pi^{-1}_ii$, or use $\sigma (i)$ in what follows?}
+
 Denote by~$\f{S}_n$ the set of permutations of~$[n]$.
@@
 Similarly, \defn{twisted Baxter permutations} are permutations avoiding the patterns $2\underline{41}3$ and $3\underline{41}2$, and \defn{co-twisted Baxter permutations} are permutations avoiding the patterns $2\underline{14}3$ and $3\underline{14}2$.
-These permutation families are central in the theory of rectangulations and appear in many other contexts~\cite{MR2763051}.
+These permutation families are equinumerous and their sizes are the \defn{Baxter numbers} (see sequence A001181 in the Online Encyclopedia of Integer Sequences).
+They are central in the theory of rectangulations and appear in many other contexts~\cite{MR2763051}.
 
@@
   \item the maximal elements of the congruence classes are the co-twisted Baxter permutations,
-  \item the congruence classes each contain a single Baxter permutation, and are one-to-one with Baxter permutations.
+  \item the congruence classes each contain a single Baxter permutation. %and are one-to-one with Baxter permutations. % J: redundant
   \end{itemize}
@@
 The \defn{co-2-clumped permutations} are defined similarly, as permutations avoiding the mirror images of these four patterns.
-He showed that 2-clumped permutations are in bijection with strong rectangulations.
+These two permutation families are equinumerous (see sequence A342141 in the Online Encyclopedia of Integer Sequences).
+Reading showed that 2-clumped permutations are in bijection with strong rectangulations.
 

### 2024-04-24 draft/rectangulotopes.tex
pagebreaks issues

@@
 For facets, we show that both weak and strong rectangulotopes have $2^n-2$ facets (the maximal possible number of facets of a deformed permutahedron), and provide simple formulas for the right hand side of the inequality corresponding to a given proper nonempty subset of~$[n]$.
-
 We now give the necessary definitions for precisely stating these vertex and facet descriptions of the weak and strong rectangulotopes.
@@
 
-\pagebreak
+%\pagebreak 
 A set function $f:2^{[n]}\mapsto \R$ is said to be \defn{submodular} when $f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$.
@@
 %%%%%%%%%%%%%
+%\pagebreak
 
@@
 They are central in the theory of rectangulations and appear in many other contexts~\cite{MR2763051}.
-
-\pagebreak
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Strong rectangulotopes}

### 2024-04-26 draft/rectangulotopes.tex
minor changes

--- 2024-04-24 draft/rectangulotopes.tex
@@
 An elementary yet striking example of this idea is the \defn{permutahedron} $\Perm (n)$.
-It is the $(n-1)$-dimensional polytope defined either as the convex hull of the points $\sum_{i\in [n]} \sigma^{-1}(i)\cdot \b{e}_i$ for all permutations $\sigma \in \f{S}_n$, or as the intersection of the hyperplane of~$\R^n$ given by~$\sum_{i \in [n]} x_i = \smash{\binom{n+1}{2}}$ with the halfspaces given by~$\sum_{i \in I} x_i \ge \smash{\binom{|I|+1}{2}}$ for all nonempty proper subsets~$\varnothing \ne I \subsetneq [n]$.
+It is the $(n-1)$-dimensional polytope defined either as the convex hull of the points $\sum_{i\in [n]} \sigma^{-1}_i\cdot \b{e}_i$ for all permutations $\sigma \in \f{S}_n$, or as the intersection of the hyperplane of~$\R^n$ given by~$\sum_{i \in [n]} x_i = \smash{\binom{n+1}{2}}$ with the halfspaces given by~$\sum_{i \in I} x_i \ge \smash{\binom{|I|+1}{2}}$ for all nonempty proper subsets~$\varnothing \ne I \subsetneq [n]$.
 Its skeleton is the Cayley graph of the symmetric group for the generators consisting of adjacent transpositions.
@@
 
-Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined by D.~Tamari~\cite{T51} and by J.~Stasheff~\cite{S63} in a topological context.
+Among those, the \defn{associahedron} is a classical and ubiquitous polytope, first defined by D.~Tama\-ri~\cite{T51} and by J.~Stasheff~\cite{S63} in a topological context.
 It has since then been identified as a fundamental object in many other areas of mathematics, including operads, cluster algebras, combinatorial Hopf algebras, and physics (see for instance the recent survey from V.
@@
 
-\subsection{Realizations of weak rectangulotopes}
+\subsection{Weak rectangulotopes}
 \label{subsec:weakRectangulotopes}
@@
 
-\subsection{Realizations of strong rectangulotopes}
+\subsection{Strong rectangulotopes}
 \label{subsec:strongRectangulotopes}
@@
 
-\jean{We use the notation $\pi^{-1}(i)$ later. Should we switch to $\pi^{-1}_ii$, or use $\sigma (i)$ in what follows?}
-
 Denote by~$\f{S}_n$ the set of permutations of~$[n]$.
@@
 \[
-\mathcal{L}(\prec_w^R) \eqdef \bigset{\pi\in\f{S}_n }{ i\prec_w^R j\implies \pi^{-1}(i) < \pi^{-1}(j)}.
+\mathcal{L}(\prec_w^R) \eqdef \bigset{\sigma\in\f{S}_n }{ i\prec_w^R j\implies \sigma^{-1}_i < \sigma^{-1}_j}.
 \]
@@
   \label{lem:lodaymax}
-  Let $\pi \in \f{S}_n$ and $I$ be an interval of~$[n]$ of size at least~$2$.
-  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~of the translated down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\pi^{-1}$ is $\b{e}_i$ (resp.~$-\b{e}_i$), where $i$ is the maximum (resp.
-the minimum) of~$\pi^{-1}$ in $I$.
+  Let $\sigma \in \f{S}_n$ and $I$ be an interval of~$[n]$ of size at least~$2$.
+  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~of the translated down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\sigma^{-1}$ is $\b{e}_i$ (resp.~$-\b{e}_i$), where $i$ is the maximum (resp.
+the minimum) of~$\sigma^{-1}$ in $I$.
 \end{lemma}
 
-Note that this extremal vertex should only depend on the weak rectangulation congruence class of $\pi$.
+Note that this extremal vertex should only depend on the weak rectangulation congruence class of $\sigma$.
 This, in passing, gives us an alternative characterization of the weak rectangulation congruence:
-$\pi
-\weakeq\sigma$ if and only if for any interval $I$ of $[n]$,
-$\arg\max_{j\in I} \pi^{-1}(j)=\arg\max_{j\in I} \sigma^{-1}(j)$, and
-$\arg\min_{j\in I} \pi^{-1}(j)=\arg\min_{j\in I} \sigma^{-1}(j)$.
+$\sigma \weakeq \tau$ if and only if for any interval $I$ of $[n]$,
+$\arg\max_{j\in I} \sigma^{-1}_j=\arg\max_{j\in I} \tau^{-1}_j$, and
+$\arg\min_{j\in I} \sigma^{-1}_j=\arg\min_{j\in I} \tau^{-1}_j$.
 
@@
 \begin{proof}
-  Let $\pi$ be any permutation in $\mathcal{L}(\prec_w^R)$.
-  From \cref{lem:lodaymax}, for an $i\in [n]$, a shard polytope~$\SP(\upArc{I})$ contributes to $\b{e}_i$ if and only if $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
+  Let $\sigma$ be any permutation in $\mathcal{L}(\prec_w^R)$.
+  From \cref{lem:lodaymax}, for an $i\in [n]$, a shard polytope~$\SP(\upArc{I})$ contributes to $\b{e}_i$ if and only if $i$ is the index of the maximum of $\sigma^{-1}$ in $I$.
   We claim that the number of intervals $I$ satisfying this condition is the given product.
@@
   Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
-  Since the weak poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s^R i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
-  Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of $\pi^{-1}$ in $I$.
+  Since the weak poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s^R i$, hence $\sigma^{-1}_\ell \leq \sigma^{-1}_i$.
+  Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of $\sigma^{-1}$ in $I$.
   Now consider the leaf of $T$ that is immediately left of the leftmost leaf of the horizontal subtree of $i$.
   This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T$.
-  Therefore, we must have $\ell\succ_s i$, and $\pi^{-1}(\ell) > \pi^{-1}(i)$, and we cannot extend $I$ past this index on the left.
+  Therefore, we must have $\ell\succ_s i$, and $\sigma^{-1}_\ell > \sigma^{-1}_i$, and we cannot extend $I$ past this index on the left.
   This shows that there are exactly $h^T_i$ choices, and not more, for the choice of the left endpoint of $I$.
@@
 \[
-\mathcal{L}(\prec_s^R) \eqdef \bigset{\pi\in\f{S}_n }{ i\prec_s^R j\implies \pi^{-1}(i) < \pi^{-1}(j)}.
+\mathcal{L}(\prec_s^R) \eqdef \bigset{\sigma\in\f{S}_n }{ i\prec_s^R j\implies \sigma^{-1}_i < \sigma^{-1}_j}.
 \]
@@
   \label{lem:yinminmax}
-  Let $\pi\in\f{S}_n$, let~$I$ and~$J$ be two consecutive intervals of~$[n]$, and let $i \eqdef \arg\max_{k\in I} \pi^{-1}(k)$ and $j \eqdef \arg\min_{k\in J} \pi^{-1}(k)$.
-  The unique vertex of the yin shard polytope~$\SP(\yinArc{I}{J})$ that is extreme with respect to the direction $\pi^{-1}$ is $\b{e}_i-\b{e}_j$ if $\pi^{-1}(i)>\pi^{-1}(j)$, and $\b{0}$ otherwise.
+  Let $\sigma\in\f{S}_n$, let~$I$ and~$J$ be two consecutive intervals of~$[n]$, and let $i \eqdef \arg\max_{k\in I} \sigma^{-1}_k$ and $j \eqdef \arg\min_{k\in J} \sigma^{-1}_k$.
+  The unique vertex of the yin shard polytope~$\SP(\yinArc{I}{J})$ that is extreme with respect to the direction $\sigma^{-1}$ is $\b{e}_i-\b{e}_j$ if $\sigma^{-1}_i > \sigma^{-1}_j$, and $\b{0}$ otherwise.
 \end{lemma}
 
-Note again that by definition, this extremal vertex should only depend on the strong congruence class of $\pi$.
+Note again that by definition, this extremal vertex should only depend on the strong congruence class of $\sigma$.
 
@@
 \begin{proof}
-  Let $\pi$ be any permutation in~$\mathcal{L}(\prec_s^R)$.
-  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc~$\alpha$ whose shard polytope $\SP(\alpha)$ contributes to $\b{e}_i-\b{e}_j$ if and only if $\pi^{-1}(i)>\pi^{-1}(j)$, hence if $i\succ_s j$.
-  Provided this holds, the set of yin arcs $\alpha$ such that $\SP(\alpha)$ contributes to $\b{e}_i-\b{e}_j$ are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that $i$ is the index of the maximum of $\pi^{-1}$ in $I$, and $j$ is the index of the maximum of $\pi^{-1}$ in $J$.
+  Let $\sigma$ be any permutation in~$\mathcal{L}(\prec_s^R)$.
+  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc~$\alpha$ whose shard polytope $\SP(\alpha)$ contributes to $\b{e}_i-\b{e}_j$ if and only if $\sigma^{-1}_i > \sigma^{-1}_j$, hence if $i\succ_s j$.
+  Provided this holds, the set of yin arcs $\alpha$ such that $\SP(\alpha)$ contributes to $\b{e}_i-\b{e}_j$ are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that $i$ is the index of the maximum of $\sigma^{-1}$ in $I$, and $j$ is the index of the maximum of $\sigma^{-1}$ in $J$.
   We claim that the number of choices of the three endpoints defining such a pair of intervals $I$ and $J$ is the given product.
@@
   Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
-  Since from \cref{lem:sextw} the strong poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s^R i$, hence $\pi^{-1}(\ell)\leq \pi^{-1}(i)$.
-  Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of~$\pi^{-1}$ in $I$.
+  Since from \cref{lem:sextw} the strong poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s^R i$, hence $\sigma^{-1}_\ell \leq \sigma^{-1}_i$.
+  Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of~$\sigma^{-1}$ in $I$.
   Now consider the leaf of $T$ that is immediately left of the leftmost leaf of the horizontal subtree of $i$.
   This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T$.
-  Therefore, we must have $\ell\succ_s i$, and $\pi^{-1}(\ell) > \pi^{-1}(i)$, and we cannot extend $I$ past this index on the left.
+  Therefore, we must have $\ell\succ_s i$, and $\sigma^{-1}_\ell > \sigma^{-1}_i$, and we cannot extend $I$ past this index on the left.
   This shows that there are exactly $h^T_i$ choices, and not more, for the choice of the left endpoint of $I$.
@@
 \begin{align*}
-w\ell(i, \sigma) & \eqdef \max \; \{0\} \cup \set{j}{j < i \text{ and } \sigma^{-1}(j) < \sigma^{-1}(i)} \\
-wr(i, \sigma) & \eqdef \min \; \{n+1\} \cup \set{j}{i < j \text{ and } \sigma^{-1}(i) < \sigma^{-1}(j)} \\
-wd(i, \sigma) & \eqdef \min \; \{n+1\} \cup \set{j}{i < j \text{ and } \sigma^{-1}(j) < \sigma^{-1}(i)} \\
+w\ell(i, \sigma) & \eqdef \max \; \{0\} \cup \set{j}{j < i \text{ and } \sigma^{-1}_j < \sigma^{-1}_i} \\
+wr(i, \sigma) & \eqdef \min \; \{n+1\} \cup \set{j}{i < j \text{ and } \sigma^{-1}_i < \sigma^{-1}_j} \\
+wd(i, \sigma) & \eqdef \min \; \{n+1\} \cup \set{j}{i < j \text{ and } \sigma^{-1}_j < \sigma^{-1}_i} \\
 \text{and} \qquad
-wu(i, \sigma) & \eqdef \max \; \{0\} \cup \set{j}{j < i \text{ and } \sigma^{-1}(i) < \sigma^{-1}(j)}.
+wu(i, \sigma) & \eqdef \max \; \{0\} \cup \set{j}{j < i \text{ and } \sigma^{-1}_i < \sigma^{-1}_j}.
 \end{align*}

### 2024-04-26 draft/rectangulotopes.tex
proofreading Vincent

@@
 %\renewcommand\labelitemi{$\diamond$} % redefine itemize default symbol
+\newcommand{\OEIS}[1]{{\rm \href{http://oeis.org/#1}{\cite[\texttt{#1}]{OEIS}}}}
 
@@
   We describe realizations of $(n-1)$-dimensional polytopes associated with two combinatorial families of rectangulations composed of $n$ rectangles.
-  They are defined as quotientopes of natural lattice congruence on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
+  They are defined as quotientopes of natural lattice congruences on the weak Bruhat order on permutations in $\f{S}_n$, and their skeleta are flip graphs on rectangulations.
   We give simple vertex and facet descriptions of these polytopes, in particular elementary formulas for computing the coordinates of the vertex corresponding to each rectangulation, in the spirit of J.-L.~Loday's realization of the associahedron.
@@
 See \cref{fig:quotientopes}\,(left) for an illustration when~$n = 4$.
-The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra}~\cite{MR2520477,Postnikov,MR4064768,MR4651496} (also called \defn{generalized permutahedra}, or \defn{polymatroids}).
+The many generalizations and extensions of the permutahedra gave rise to a flourishing theory of \defn{deformed permutahedra}~\cite{Postnikov,MR2520477,MR4064768,MR4651496} (also called \defn{generalized permutahedra}, or \defn{polymatroids}).
 
@@
 Santos and G.
-Ziegler~\cite{PilaudSantosZiegler} and references therein).
+Ziegler~\cite{PilaudSantosZiegler} and the references therein).
 It was first thought of as a purely combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894,MR2321739}.
-The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
+The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of non-intersecting diagonals of a convex $(n+3)$-gon.
 In particular, its skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.
-In~2004, J.-L.~Loday published the following elegant description of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex given the corresponding binary tree.
+In~2004, J.-L.~Loday published the following elegant description of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex corresponding to a given binary tree.
 We note that the inequality description of the same polytope was actually provided in~1993 by S.~Shnider and S.~Sternberg~\cite{ShniderSternberg}, but Loday's vertex description largely popularized this realization.
@@
 Those polytopes are deformed permutahedra that they called \defn{quotientopes}.
-Associahedra are therefore the quotientopes of the sylvester congruence of the weak Bruhat order whose classes are in bijection with triangulations.
+The associahedron is therefore the quotientope of the sylvester congruence of the weak Bruhat order whose classes are in bijection with triangulations.
 
@@
 The \defn{segments} of a rectangulation are the inclusionwise maximal line segments composed of edges of the rectangles, excluding the edges of the decomposed square.
-We suppose throughout that no four rectangles have a common vertex, hence that the rectangulations are \defn{generic} in that sense.
+We suppose throughout that the rectangulations are \defn{generic}, in the sense that no four rectangles have a common vertex.
 
@@
 We define \defn{on the left of} and \defn{on the right of} similarly.
-Two rectangulations are said to be \defn{weakly equivalent} if there exists a bijection between their rectangles that preserves the above-below and left-right relations between the rectangles.
+Two rectangulations are said to be \defn{weakly equivalent} if there exists a bijection between their rectangles that preserves the \mbox{above--below} and left--right relations between the rectangles.
 Weak equivalence classes of rectangulations will be referred to as \defn{weak rectangulations}.
 (Note that this is a slight, yet convenient abuse of terminology, as weak rectangulations are really sets of rectangulations.)
-On the other hand, two rectangulations are said to be \defn{strongly equivalent} when there exists a bijection between their rectangles that not only preserves the above-below and left-right relations, but also the adjacency relation between rectangles.
+On the other hand, two rectangulations are said to be \defn{strongly equivalent} when there exists a bijection between their rectangles that not only preserves the above--below and left--right relations, but also the adjacency relation between rectangles.
 We will naturally refer to strong equivalence classes of rectangulations as \defn{strong rectangulations}.
@@
 However, if the wall slide changes the adjacency relation between the rectangles, then the resulting rectangulation is not strongly equivalent to $R$ anymore.
-A \defn{diagonal rectangulation} is a rectangulation in which every rectangle intersects the top-left to bottom-right diagonal of the square.
+A \defn{diagonal rectangulation} is a rectangulation in which every rectangle intersects the top--left to bottom--right diagonal of the square.
 Examples of rectangulations are given in \cref{fig:rectequiv}.
@@
 
-The \defn{weak rectangulation congruence} (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762} and revisited in~\cite[Thm.~1.1 \& Exm.~4.10]{Reading-arcDiagrams} in terms of arc diagrams.
+The \defn{weak rectangulation congruence}~$\weakeq$ (or \defn{Baxter congruence}) was first explicitly studied by S.~Law and N.~Reading~\cite{MR2871762} and revisited in~\cite[Thm.~1.1 \& Exm.~4.10]{Reading-arcDiagrams} in terms of arc diagrams.
 Its classes are in bijection with weak rectangulations, but also with \defn{Baxter permutations}~\cite{MR0491652,MR0555815}, with \defn{twin binary trees}~\cite{MR1417289,MR2914637}, and with many other combinatorial families~\cite{MR2763051}.
 The corresponding quotientopes will be referred to as the \defn{weak rectangulotopes} and denoted by $\WRP(n)$.
-We note that weak rectangulotopes were already constructed as Minkowski sums of two opposite associahedra in~\cite{MR2871762}, before the general constructions of quotientopes of~\cite{MR3964495,MR4584712}.
+We note that weak rectangulotopes were already constructed as Minkowski sums of two opposite associahedra in~\cite{MR2871762}, even before the general constructions of quotientopes of~\cite{MR3964495,MR4584712}.
 For~$n = 4$, the weak rectangulation lattice is represented in \cref{fig:weakRectangulationLattice} and the weak rectangulotope is represented in \cref{fig:weakRectangulotope}.
 
-The \defn{strong rectangulation congruence} was introduced by N.~Reading~\cite{MR2864445}, revisited in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams} in terms of arc diagrams, and studied more recently by E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
+The \defn{strong rectangulation congruence}~$\strongeq$ was studied by N.~Reading~\cite{MR2864445}, revisited~in~\cite[Thm.~1.2 \& Exm.~4.11]{Reading-arcDiagrams} in terms of arc diagrams, and studied more recently by E.~Meehan~\cite{MR3697823}, and A.~Asinowski, J.~Cardinal, S.~Felsner, and É.~Fusy~\cite{ACFF24}.
 Its classes are in bijection with strong rectangulations.
@@
 Strong rectangulotopes can be obtained as quotientopes~\cite{MR3964495,MR4584712} for the strong rectangulation congruence.
-For~$n = 4$, the strong rectangulation lattice is represented in \cref{fig:strongRectangulationLattice} and the strong rectangulotope is represented in \cref{fig:strongRectangulotope}.
+For~$n = 4$, the strong rectangulation lattice is represented in \cref{fig:strongRectangulationLattice} and the strong rectangulotope is represented in~\cref{fig:strongRectangulotope}.
 
@@
 An important motivation for studying these quotientopes is the informative structure of their skeleta, which are \defn{flip graphs} on the rectangulations.
-A flip is a local, reversible change in the structure of the rectangulation.
+A \defn{flip} is a local, reversible change in the structure of the rectangulation.
 Flips in rectangulations come in three flavors:
@@
 \item \defn{simple flips} replace a vertical segment incident to exactly two rectangles by a horizontal segment, and vice-versa,
-\item \defn{pivoting flips} transform a pair of adjacent rectangles whose union is L-shaped, changing a left-right pair into an above-below pair, and vice-versa,
+\item \defn{pivoting flips} transform a pair of adjacent rectangles whose union is L-shaped, changing a left--right pair into an above--below pair, and vice-versa,
 \item \defn{wall slide flips} are wall slides that modify the adjacency of the rectangles, removing exactly one adjacency and adding exactly one. These are only relevant in strong rectangulations.
@@
     \end{center}
-	\caption{The three types of flips between rectangulations. The flips are possible only if the grey areas do not intersect any segment of the rectangulation.}
+	\caption{The three types of flips between rectangulations. The flips are possible only if the grey areas do not intersect any segment of the rectangulation. \cite[Fig.~19]{ACFF24}}
 	\label{fig:flips}
@@
 
-Fix a rectangulation~$R$ of size~$n$, and consider the directed graph~$D(R)$ whose vertex set is the set of all vertices of~$R$, and whose edges are obtained as follows.
+Fix a rectangulation~$R$ of size~$n$, and consider the directed graph~$D(R)$ whose vertices are all vertices of~$R$, and whose edges are obtained as follows.
 For each rectangle~$r$ of~$R$, we include in~$D(R)$ four edges joining the vertices of~$r$, where the horizontal edges are oriented from left to right, and the vertical edges are oriented from bottom to top.
-Hence, the bottom left corner of~$r$ is its source~$s(r)$, while its top right corner is its target~$t(r)$.
+Hence, in the rectangle~$r$, its bottom--left corner is its source~$s(r)$, while its top--right corner is its target~$t(r)$.
 Note that since~$R$ is generic, the sources~$\set{s(r)}{r \in R}$ and the targets~$\set{t(r)}{r \in R}$ are disjoint.
 The \defn{source tree}~$S(R)$ (resp.~\defn{target tree}~$T(R)$) is the subgraph of~$D(R)$ induced by the sources~$\set{s(r)}{r \in R}$ (resp.~by the targets~$\set{t(r)}{r \in R}$).
-It is not difficult to check that the source tree~$S(R)$ (resp.~target tree~$T(R)$) is a binary tree, rooted at the bottom left (resp.~top right) corner of~$R$, and oriented from (resp.~towards) its root.
+It is not difficult to check that the source tree~$S(R)$ (resp.~target tree~$T(R)$) is a binary tree, rooted at the bottom--left (resp.~top--right) corner of~$R$, and oriented from (resp.~towards) its root.
 
@@
 By symmetry, we only prove the statement for the source tree~$S(R)$.
-As the rectangulation is generic, each source~$s(r)$ distinct from the bottom left corner of~$R$ is either on the top edge of a rectangle~$r'$ or on the right edge of a rectangle~$r'$ (but not both).
-This shows that~$r$ distinct from the bottom left rectangle of~$R$ has a unique parent~$r'$ in~$S(R)$, so that~$S(R)$ is indeed a tree on~$[n]$.
+As the rectangulation is generic, each source~$s(r)$ distinct from the bottom--left corner of~$R$ is either on the top edge of a rectangle~$r'$ or on the right edge of a rectangle~$r'$ (but not both).
+This shows that~$r$ distinct from the bottom--left rectangle of~$R$ has a unique parent~$r'$ in~$S(R)$, so that~$S(R)$ is indeed a tree on~$[n]$.
 This tree is binary as each node has at most one vertical child and one horizontal child, and no other children.
@@
 For any rectangle~$r$ of~$R$, the inorder label of the source~$s(r)$ in~$S(R)$ coincides with the inorder label of the target~$t(r)$ in~$T(R)$.
-The inorder can also be retrieved from the above-below and left-right relationships defined before, as the transitive closure of the union of the below and right partial orders.
+The inorder can also be retrieved from the above--below and left--right relationships defined before, as the transitive closure of the union of the below and right partial orders.
 This enables to unambiguously label the rectangles of~$R$ by the inorder.
 The resulting labeling coincides with the NW--SE labeling of~\cite{ACFF24}.
-From now on, the labels of the rectangles in a rectangulation are the inorder labels, and the two trees $T(R)$ and $S(R)$ are defined on the vertex set $[n]$ accordingly.
+From now on, the labels of the rectangles in a rectangulation are the inorder labels, and the two trees~$T(R)$ and $S(R)$ are defined on the vertex set $[n]$ accordingly.
 For each $i\in [n]$, we refer to its two subtrees in $T(R)$ as the \defn{horizontal} and \defn{vertical subtrees}, and similarly for $S(R)$.
@@
 For instance, the weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
-The coordinates of the vertices corresponding to all weak rectangulations are gathered in \cref{tab:verticesWeakRectangulotope}.
-The facet inequalities corresponding to each nonempty proper subset of~$[n]$ are gathered in \cref{tab:facetsWeakRectangulotope}.
+The coordinates of the vertices of~$\WRP(4)$ corresponding to all weak rectangulations are gathered in \cref{tab:verticesWeakRectangulotope}.
+The facet inequalities of~$\WRP(4)$ corresponding to each nonempty proper subset of~$[4]$ are gathered~in~\cref{tab:facetsWeakRectangulotope}.
 
@@
 \raisebox{-.06cm}{$_j$} \,$ the situation where the rectangles $r_i$ and $r_j$ share a horizontal segment containing the bottom edge of~$r_i$ and the top edge of~$r_j$, and the right edge of~$r_i$ is on the right of the left edge of~$r_j$.
-We also use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.
+We use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.
 
@@
 For instance, the strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotope}.
-The coordinates of the vertices corresponding to all strong rectangulations are gathered in \cref{tab:verticesStrongRectangulotope}.
-The facet inequalities corresponding to each nonempty proper subset of~$[n]$ are gathered in \cref{tab:facetsStrongRectangulotope}.
+The coordinates of the vertices of~$\SRP(4)$ corresponding to all strong rectangulations are gathered in \cref{tab:verticesStrongRectangulotope}.
+The facet inequalities of~$\SRP(4)$ corresponding to each nonempty proper subset of~$[4]$ are gathered in \cref{tab:facetsStrongRectangulotope}.
 
@@
 
-\subsection*{Acknowledgments}
-
-This work was initiated at the Workshop on Combinatorics, Algorithms, and Geometry held on March 4-8, 2024 in Dresden, Germany.
+\subsection{Acknowledgments}
+
+This work was initiated at the Workshop on Combinatorics, Algorithms, and Geometry held on March 4--8, 2024 in Dresden, Germany.
 The authors thank Namrata and Torsten M\"utze for the organization and the other participants of the workshop for the stimulating interactions, notably on other combinatorial aspects of rectangulations.
@@
 We now briefly recall some results on the weak Bruhat order and its quotients, both from a lattice and geometric perspectives.
-We refer to \cite{MR2142177, Reading-arcDiagrams, MR3645055, MR3645056, MR3964495, MR4584712} for details.
+We refer to \cite{MR2142177, Reading-arcDiagrams, MR3645056, MR3645055, MR3964495, MR4584712} for details.
 
@@
 
-For instance, the join irreducible permutations minimal in their sylvester congruence class are precisely the \defn{up arcs}, \ie the arcs of the form~$(a, b, {]a,b[}, \varnothing)$ (similarly, we call \defn{down arcs} the arcs of the form~$(a, b, \varnothing, {]a,b[})$
-).
+For instance, the join irreducible permutations minimal in their sylvester congruence class are precisely the \defn{up arcs}, \ie the arcs of the form~$(a, b, {]a,b[}, \varnothing)$ (similarly, we call \defn{down arcs} the arcs of the form~$(a, b, \varnothing, {]a,b[})$).
 The corresponding noncrossing arc diagrams are famously known as noncrossing partitions.
@@
 For instance, \cref{fig:fans}\,(right) represents the \defn{sylvester fan} (the quotient fan of the sylvester congruence of \cref{fig:sylvesterCongruence}\,(middle)) for~$n = 4$, whose dual graph is the cover graph of the Tamari lattice of \cref{fig:sylvesterCongruence}\,(right).
+The maximal cone corresponding to a binary tree~$T$ is given by~$\polytope{C}^T = \set{\b{x} \in \R^n}{x_i \le x_i \text{ for each edge } i \to j \text{ in } T}$.
 
@@
 For instance, the associahedron~$\Asso(n)$ (\cref{fig:quotientopes}\,(right) when~$n = 4$) is a quotientope for the sylvester congruence (\cref{fig:sylvesterCongruence}\,(middle) when~$n = 4$), its normal fan is the sylvester fan (\cref{fig:fans}\,(right) when~$n = 4$), and its skeleton is the cover graph of the Tamari lattice (\cref{fig:sylvesterCongruence}\,(right) when~$n = 4$).
-The existence of quotientopes was first proved in \cite{MR3964495} and later better understood in~\cite{MR4584712} using Minkowski sums of shard polytopes.
+The existence of quotientopes was first proved in \cite{MR3964495} and later better understood in~\cite{MR4584712} using Minkowski sums of shard polytopes, the approach we will use throughout the paper.
 
@@
 Recall that for any vector~$\b{v} \ne \b{0}$ of~$\R^n$, the face of~$\polytope{P} + \polytope{Q}$ maximizing the scalar product with~$\b{v}$ is the Minkowski sum of the faces of~$\polytope{P}$ and~$\polytope{Q}$ maximizing the scalar product with~$\b{v}$.
-In particular, if~$\b{v}$ is a generic direction, the vertex of~$\polytope{P} + \polytope{Q}$ extremal in direction~$\b{v}$ is just the sum of the vertices of~$\polytope{P}$ and~$\polytope{Q}$ that are extremal in direction~$\b{v}$.
+In particular, if~$\b{v}$ is a generic direction, the vertex of~$\polytope{P} + \polytope{Q}$ that is extremal in direction~$\b{v}$ is just the sum of the vertices of~$\polytope{P}$ and~$\polytope{Q}$ that are extremal in direction~$\b{v}$.
 Moreover, the normal fan of the Minkowski sum~$\polytope{P} + \polytope{Q}$ is the common refinement of the normal fans of~$\polytope{P}$ and~$\polytope{Q}$.
@@
 An \defn{$\alpha$-alternating matching} is a sequence~$a \le i_1 < j_1 < i_2 < j_2 < \dots < i_q < j_q \le b$ such that~$i_p \in \{a\} \cup A$ and~$j_p \in \{b\} \cup B$ for all~$p \in [q]$.
-Its \defn{characteristic vector} is~$\sum_{p \in [q]} \b{e}_{i_p} - \b{e}_{j_p}$.
+Its \defn{characteristic vector} is~$\sum_{p \in [q]} (\b{e}_{i_p} - \b{e}_{j_p})$.
 The \defn{shard polytope} of~$\alpha$ is the convex hull~$\SP(\alpha)$ of the characteristic vectors of all $\alpha$-alternating matchings.
@@
 	\centerline{\includegraphics[scale=1]{shardPolytopes}}
-	\caption{The shard polytopes of the up arc~$(1, 4, \{2, 3\}, \varnothing)$, down arc~$(1, 4, \varnothing, \{2, 3\})$, yin arc~$(1, 4, \{2\}, \{3\})$, and yang arc~$(1, 4, \{3\}, \{2\})$.
+	\caption{The shard polytopes of the up arc~$(1, 4, \{2, 3\}, \varnothing)$, the down arc~$(1, 4, \varnothing, \{2, 3\})$, the yin arc~$(1, 4, \{2\}, \{3\})$, and the yang arc~$(1, 4, \{3\}, \{2\})$.
 The vertices of the shard polytope of an arc~$\alpha \eqdef (a, b, A, B)$ are labeled by the corresponding $\alpha$-alternating matchings, where we use solid dots~$\bullet$ for elements in~$\{a\} \cup A$ and hollow dots~$\circ$ for elements in~$B \cup \{b\}$.
@@
 
-Given a rectangulation~$R$, let us consider the source and target trees~$S(R)$ and~$T(R)$, and orient the horizontal edges from left to right, and the vertical edges from bottom to top.
+Given a rectangulation~$R$, let us consider its source and target trees~$S(R)$ and~$T(R)$, and orient the horizontal edges from left to right, and the vertical edges from bottom to top.
 This defines two partial orders on $[n]$.
@@
 	\centerline{\includegraphics[width=.4\textwidth]{weakPoset1} \qquad \includegraphics[width=.4\textwidth]{weakPoset2}}
-	\caption{A weak poset of the rectangulation of \cref{fig:strongRectangulation}, visualized on the rectangulation of \cref{fig:weakRectangulation} (left) and on its diagonal rectangulation of \cref{fig:weakRectangulation} (right).
+	\caption{The weak poset of the rectangulation of \cref{fig:strongRectangulation}, visualized on the rectangulation of \cref{fig:strongRectangulation} (left) and on its diagonal rectangulation of \cref{fig:weakRectangulation} (right).
 Example from \cite{ACFF24}.}
@@
 
-We define the set~$\mathcal{L}(\prec_w^R)$ as the linear extensions of the weak poset:
+We consider the set~$\mathcal{L}(\prec_w^R)$ of linear extensions of the weak poset~$\prec_w^R$:
 \[
@@
 
-We recall that \defn{Baxter permutations} are permutations avoiding the vincular patterns $2\underline{41}3$ and $3\underline{14}2$, where the underlining indicates that the symbols must appear contiguously in the permutation.
+We recall that \defn{Baxter permutations}~\cite{MR0491652,MR0555815} are permutations avoiding the vincular patterns $2\underline{41}3$ and $3\underline{14}2$, where the underlining indicates that the symbols must appear contiguously in the permutation.
 Similarly, \defn{twisted Baxter permutations} are permutations avoiding the patterns $2\underline{41}3$ and $3\underline{41}2$, and \defn{co-twisted Baxter permutations} are permutations avoiding the patterns $2\underline{14}3$ and $3\underline{14}2$.
-These permutation families are equinumerous and their sizes are the \defn{Baxter numbers} (see sequence A001181 in the Online Encyclopedia of Integer Sequences).
+These permutation families are equinumerous and their sizes are the \defn{Baxter numbers}, see \OEIS{A001181}.
 They are central in the theory of rectangulations and appear in many other contexts~\cite{MR2763051}.
@@
 
-These arcs can be classified into two families of arcs: the up arcs and the down arcs.
-For an interval $I$ of $[n]$ of size at least~$2$, the \defn{up arc}~$\upArc{I}$ (resp.~\defn{down arc}~$\downArc{I}$) is the arc that starts at~$\min(I)$ and ends at~$\max(I)$, and passes above (resp.~below) all remaining elements of~$I$, that is
+We classified these arcs into two families: the up arcs and the down arcs.
+For an interval~$I$ of~$[n]$ of size at least~$2$, the \defn{up arc}~$\upArc{I}$ (resp.~\defn{down arc}~$\downArc{I}$) is the arc that starts at~$\min(I)$ and ends at~$\max(I)$, and passes above (resp.~below) all remaining elements of~$I$, that is,
 \begin{align*}
@@
   The weak rectangulotope $\WRP(n)$ is realized by the Minkowski sum of all up and down shard polytopes.
-  In particular, it is the Minkowski sum of two opposite associahedra of J.-L.~Loday~\cite{MR2871762}.
+%  In particular, it is the Minkowski sum of two opposite associahedra of J.-L.~Loday~\cite{MR2871762}.
 \end{proposition}
+
+Since the combinatorial type of a Minkowski sum is invariant to translation of the summands, we can use the simplices defined in \cref{lem:lodaysp,lem:antilodaysp} as a definition of the up and down shard polytopes.
+Also note that since the basic arcs~$(i, i+1, \varnothing, \varnothing)$ are both up and down, each basic shard polytope is summed twice in the proposed realization.
+This, again, does not change the combinatorial type of the polytope.
+
+\cref{prop:weakMinkowski} in particular recovers the result of~\cite{MR2871762} that $\WRP(n)$ is the sum of two opposite associahedra of J.-L.~Loday~\cite{MR2871762}, corresponding to the sylvester and anti-sylvester congruences.
 
@@
   Let $\sigma \in \f{S}_n$ and $I$ be an interval of~$[n]$ of size at least~$2$.
-  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~of the translated down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\sigma^{-1}$ is $\b{e}_i$ (resp.~$-\b{e}_i$), where $i$ is the maximum (resp.
+  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\sigma^{-1}$ is $\b{e}_i$ (resp.~$-\b{e}_i$), where $i$ is the maximum (resp.
 the minimum) of~$\sigma^{-1}$ in $I$.
@@
 The proof of the following lemma actually follows from Loday's realization in~\cref{thm:loday}.
-For the sake of completeness, and because we will reuse the same reasoning in the next section, we give an elementary proof.
+For the sake of completeness, and because we will reuse the same reasoning in \cref{sec:strongRectangulotopes}, we give an elementary proof.
 
@@
   Therefore, we must have $\ell\succ_s i$, and $\sigma^{-1}_\ell > \sigma^{-1}_i$, and we cannot extend $I$ past this index on the left.
-  This shows that there are exactly $h^T_i$ choices, and not more, for the choice of the left endpoint of $I$.
+  This shows that there are exactly $h^T_i$ choices for the choice of the left endpoint of $I$.
   A symmetric argument shows that there are exactly $v^{T}_i$ choices for the right endpoint of $I$.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Strong rectangulotopes}
@@
 	\begin{enumerate}[(i)]
-	\item either the right side of $r_j$ and the left side of $r_i$ lie on the same vertical segment, and the bottom-right corner of $r_j$ is above the top-left corner of $r_i$, 
-	\item or the top side of $r_j$ and the bottom side of rectangle $r_i$ lie on the same horizontal segment, and the top-left corner of $r_j$ lies on the right of the bottom-right corner~of~$r_i$.
+	\item either the right side of $r_j$ and the left side of $r_i$ lie on the same vertical segment, and the bottom--right corner of $r_j$ is above the top--left corner of $r_i$, 
+	\item or the top side of $r_j$ and the bottom side of rectangle $r_i$ lie on the same horizontal segment, and the top--left corner of $r_j$ lies on the right of the bottom--right corner~of~$r_i$.
 	\end{enumerate}
@@
   \label{lem:sextw}
-  The strong poset $([n], \prec_s^R)$ of a rectangulation $R$ is an extension of its weak poset $([n], \prec_w^R)$.
+  For any rectangulation $R$, its strong poset $([n], \prec_s^R)$ is an extension of its weak poset~$([n], \prec_w^R)$.
 \end{lemma}
@@
   (These are the flips described in~\cref{fig:wallSlides}, performed in bottom to top order on the figure.) 
-  Note that each of these we can only add new pairs in the adjacency poset, hence the adjacency poset of $R$ is always an extension of that of $D$. 
+  Note that each of these wall slides can only add new pairs in the adjacency poset, hence the adjacency poset of $R$ is always an extension of that of $D$. 
 \end{proof}
@@
 
-Let $\mathcal{L}(\prec_s^R)$ be the set of permutations of $\f{S}_n$ corresponding to linear extensions of the strong poset $\prec_s^R$:
+We consider the set~$\mathcal{L}(\prec_s^R)$ of linear extensions of the strong poset~$\prec_s^R$:
 \[
@@
 N.
-Reading~\cite{MR2864445} introduced \defn{2-clumped permutations}, defined as the permutations avoiding the vincular patterns $24\underline{51}3$, $42\underline{51}3$, $3\underline{51}24$, and $3\underline{51}42$.
+Reading introduced \defn{2-clumped permutations}~\cite{MR2864445}, defined as the permutations avoiding the vincular patterns $24\underline{51}3$, $42\underline{51}3$, $3\underline{51}24$, and $3\underline{51}42$.
 The \defn{co-2-clumped permutations} are defined similarly, as permutations avoiding the mirror images of these four patterns.
-These two permutation families are equinumerous (see sequence A342141 in the Online Encyclopedia of Integer Sequences).
+These two permutation families are equinumerous, see \OEIS{A342141}.
 Reading showed that 2-clumped permutations are in bijection with strong rectangulations.
@@
 
-We classify the arcs into two families, depending on whether the arcs do not cross the horizontal line from below or from above.
+We classify these arcs into two families, depending on whether the arcs do not cross the horizontal line from below or from above.
 For two consecutive intervals~$I$ and~$J$ of~$[n]$, the \defn{yin arc}~$\yinArc{I}{J}$ (resp.~\defn{yang arc}~$\yangArc{I}{J}$) is the arc starts at~$\min(I)$ and ends at~$\max(J)$, and goes above (resp.~below) the remaining elements of $I$ and below (resp.~above) the remaining elements of $J$, that is,
@@
 Since the combinatorial type of a Minkowski sum is invariant to translation of the summands, we can use the polytopes defined in \cref{lem:yinsp,lem:yangsp} as a definition of the yin and yang shard polytopes.
-Also note that since the up and down shard polytopes are, up to a translation, both yin and yang, each up and down shard polytope is summed twice in the proposed realization.
+Also note that since the up and down shard polytopes are, up to a translation, both yin and yang, each up and down shard polytope is summed twice (hence all basic shard polytopes are summed four times) in the proposed realization.
 This, again, does not change the combinatorial type of the polytope.
@@
 \raisebox{.03cm}{\verticalPattern} \!\!\!
-{}_j$ denote the situation where the rectangles $r_i$ and~$r_j$ of inorder labels $i$ and $j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
-Our aim is to provide a formula for the coordinates of the vertices of the strong rectangulotope that does not involve the computation of the strong poset or one of its linear extension.
+{}_j$ denotes the situation where the rectangles $r_i$ and~$r_j$ of inorder labels $i$ and $j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
+Our aim is to provide a formula for the coordinates of the vertices of the strong rectangulotope that does not involve the computation of the strong poset or one of its linear extensions.
 For that purpose, we prove that for every pair $i<j\in [n]$ such that~$i\prec_s^R j$, which do not have any associated yin arc, we have either ${}^i \!\!\!
@@
 
-  Therefore, there can only be one segment on which a leaf common to the vertical subtree of $i$ in $T$ and the vertical subtree of $j$ in $S$ may lie, which is the vertical segment supporting the right edge of $r_i$ and the left edge of $r_j$.
+  Therefore, there can only be one segment which may contain a leaf to the vertical subtree of $i$ in $T$ and a leaf of the vertical subtree of $j$ in $S$, which is the vertical segment supporting the right edge of $r_i$ and the left edge of $r_j$.
   Let us assume that this segment is shared by those two edges.

### 2024-04-26 draft/rectangulotopes.tex
update caption figure 2

@@
 	\caption{Four rectangulations.
-The leftmost rectangulation is not weakly equivalent to any other.
-The rightmost three are weakly equivalent, but only the first two are strongly equivalent.
+The rightmost rectangulation is not weakly equivalent to any other.
+The leftmost three are weakly equivalent, but only the first two are strongly equivalent.
 Only the third one is diagonal.}

### 2024-04-26 draft/rectangulotopes.tex
min max inversion characterization of strong eq

--- 2024-04-26 draft/rectangulotopes.tex
@@
 Note that this extremal vertex should only depend on the weak rectangulation congruence class of $\sigma$.
-This, in passing, gives us an alternative characterization of the weak rectangulation congruence:
-$\sigma \weakeq \tau$ if and only if for any interval $I$ of $[n]$,
+This, in passing, gives us alternative characterizations of the sylvester congruence $\frown$ and the weak rectangulation congruence $\weakeq$:
+$\sigma \frown \tau$ if and only if for any interval $I$ of $[n]$,
 $\arg\max_{j\in I} \sigma^{-1}_j=\arg\max_{j\in I} \tau^{-1}_j$, and
+$\sigma \weakeq \tau$ if the previous property holds, and in addition
 $\arg\min_{j\in I} \sigma^{-1}_j=\arg\min_{j\in I} \tau^{-1}_j$.
@@
   \label{lem:yinminmax}
-  Let $\sigma\in\f{S}_n$, let~$I$ and~$J$ be two consecutive intervals of~$[n]$, and let $i \eqdef \arg\max_{k\in I} \sigma^{-1}_k$ and $j \eqdef \arg\min_{k\in J} \sigma^{-1}_k$.
-  The unique vertex of the yin shard polytope~$\SP(\yinArc{I}{J})$ that is extreme with respect to the direction $\sigma^{-1}$ is $\b{e}_i-\b{e}_j$ if $\sigma^{-1}_i > \sigma^{-1}_j$, and $\b{0}$ otherwise.
+  Let $\sigma\in\f{S}_n$, let~$I$ and~$J$ be two consecutive intervals of~$[n]$, and let $i\eqdef \arg\max_{k\in I} \sigma^{-1}_k$ and $j \eqdef \arg\min_{k\in J} \sigma^{-1}_k$ (respectively, for yang arcs, $i\eqdef \arg\min_{k\in I} \sigma^{-1}_k$ and $j \eqdef \arg\max_{k\in J} \sigma^{-1}_k$).
+  The unique vertex of the yin (resp.
+yang) shard polytope~$\SP(\yinArc{I}{J})$ (resp.
+$\SP(\yangArc{I}{J})$) that is extreme with respect to the direction $\sigma^{-1}$ is $\b{e}_i-\b{e}_j$ (resp.
+$\b{e}_j-\b{e}_i$) if $\sigma^{-1}_i > \sigma^{-1}_j$ (resp.
+$\sigma^{-1}_i < \sigma^{-1}_j$), and $\b{0}$ otherwise.
 \end{lemma}
@@
 Note again that by definition, this extremal vertex should only depend on the strong congruence class of $\sigma$.
+This, again, gives us alternative characterizations of the yin congruence $\sim$, the yang congruence $\backsim$, and the strong rectangulation congruence $\strongeq$.
+We have $\sigma \sim \tau$ if and only if for any pair of consecutive intervals $I$ and $J$ of $[n]$, 
+$\arg\max_{k\in I} \sigma^{-1}_k = \arg\max_{k\in I} \tau^{-1}_k$, 
+$\arg\min_{k\in J} \sigma^{-1}_k = \arg\min_{k\in J} \tau^{-1}_k$,
+and, if we let $i$ and $j$ be defined as in Lemma~\ref{lem:yinminmax} as the indices of the maximum in $I$ and the minimum in $J$,
+$\sigma^{-1}_i > \sigma^{-1}_j \Leftrightarrow \tau^{-1}_i > \tau^{-1}_j$.
+A similar statement holds for the yang congruence $\backsim$, by exchanging $\min$ and $\max$, and flipping the sign of the inequality.
+Finally, $\sigma\strongeq\tau$ if both sets of conditions hold.
 

### 2024-04-26 draft/rectangulotopes.tex
minor change

--- 2024-04-26 draft/rectangulotopes.tex
@@
 From the results of \cref{subsec:quotientopes}, the submodular function corresponding to our realization of $\SRP(n)$ can be obtained as before by summing the functions defining the yin and yang shard polytopes in \cref{lem:yinsp,lem:yangsp}.
-Note again that all rays of the braid fan are preserved in the fan of the strong rectangulotope, hence our characterization of the strong rectangulotope is an irredundant facet description.
+Note again that all rays of the braid fan are preserved in the fan of the strong rectangulotope, hence our submodular description of the strong rectangulotope is an irredundant facet description.
 

### 2024-04-26 draft/rectangulotopes.tex
add ref

--- 2024-04-26 draft/rectangulotopes.tex
@@
 
-The combinatorics of rectangulations has been studied for nearly two decades~\cite{MR2233287,MR2763051,MR2871762,MR2864445,MR3084577,MR3192492,MR3878132,MR4598046,AB24,ACFF24}.
+The combinatorics of rectangulations has been studied for nearly two decades~\cite{MR2233287,MR2763051,MR2871762,MR2864445,MR3084577,MR3192492,MR3878132,MR4598046,MR4667583,AB24,ACFF24}.
 In fact, several combinatorial ideas can even be traced back to the problem of ``squaring the square'' studied since the 1930s~\cite{A30,Abe32,MR0000470,MR0003040}.
@@
   Let $\sigma \in \f{S}_n$ and $I$ be an interval of~$[n]$ of size at least~$2$.
-  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\sigma^{-1}$ is $\b{e}_i$ (resp.~$-\b{e}_i$), where $i$ is the maximum (resp.
-the minimum) of~$\sigma^{-1}$ in $I$.
+  The unique vertex of the up shard polytope~$\SP(\upArc{I})$ (resp.~down shard polytope~$\SP(\downArc{I})$) that is extreme with respect to the direction $\sigma^{-1}$ is $\b{e}_i$ (resp.~$-\b{e}_i$), where $i$ is the maximum (resp.~the minimum) of~$\sigma^{-1}$ in $I$.
 \end{lemma}
@@
 Note that this extremal vertex should only depend on the weak rectangulation congruence class of $\sigma$.
-This, in passing, gives us an alternative characterization of the weak rectangulation congruence:
-$\sigma \weakeq \tau$ if and only if for any interval $I$ of $[n]$,
-$\arg\max_{j\in I} \sigma^{-1}_j=\arg\max_{j\in I} \tau^{-1}_j$, and
-$\arg\min_{j\in I} \sigma^{-1}_j=\arg\min_{j\in I} \tau^{-1}_j$.
+This, in passing, gives us alternative characterizations of the sylvester congruence~$\frown$, the anti-sylvester congruence~$\smile$, and the weak rectangulation congruence $\weakeq$.
+We have~$\sigma \frown \tau$ if and only if $\arg\max_{j\in I} \sigma^{-1}_j=\arg\max_{j\in I} \tau^{-1}_j$ for any interval $I$ of $[n]$.
+Similarly, $\sigma \smile \tau$ if and only if $\arg\min_{j\in I} \sigma^{-1}_j=\arg\min_{j\in I} \tau^{-1}_j$ for any interval $I$ of $[n]$.
+Finally, $\sigma\weakeq\tau$ if both conditions hold.
 
@@
   \label{lem:yinminmax}
-  Let $\sigma\in\f{S}_n$, let~$I$ and~$J$ be two consecutive intervals of~$[n]$, and let $i \eqdef \arg\max_{k\in I} \sigma^{-1}_k$ and $j \eqdef \arg\min_{k\in J} \sigma^{-1}_k$.
-  The unique vertex of the yin shard polytope~$\SP(\yinArc{I}{J})$ that is extreme with respect to the direction $\sigma^{-1}$ is $\b{e}_i-\b{e}_j$ if $\sigma^{-1}_i > \sigma^{-1}_j$, and $\b{0}$ otherwise.
+  Let $\sigma\in\f{S}_n$, let~$I$ and~$J$ be two consecutive intervals of~$[n]$, and let $i\eqdef \arg\max_{k\in I} \sigma^{-1}_k$ and $j \eqdef \arg\min_{k\in J} \sigma^{-1}_k$ (resp.~$i\eqdef \arg\min_{k\in I} \sigma^{-1}_k$ and $j \eqdef \arg\max_{k\in J} \sigma^{-1}_k$).
+  The unique vertex of the yin (resp.~yang) shard polytope~$\SP(\yinArc{I}{J})$ (resp.~$\SP(\yangArc{I}{J})$) that is extreme with respect to the direction $\sigma^{-1}$ is $\b{e}_i-\b{e}_j$ (resp.~$\b{e}_j-\b{e}_i$) if $\sigma^{-1}_i > \sigma^{-1}_j$ (resp.~$\sigma^{-1}_i < \sigma^{-1}_j$), and $\b{0}$ otherwise.
 \end{lemma}
@@
 Note again that by definition, this extremal vertex should only depend on the strong congruence class of $\sigma$.
+This, again, gives us alternative characterizations of the yin congruence $\sim$, the yang congruence $\backsim$, and the strong rectangulation congruence $\strongeq$.
+We have $\sigma \sim \tau$ if and only if for any pair of consecutive intervals $I$ and $J$ of $[n]$, 
+$\arg\max_{k\in I} \sigma^{-1}_k = \arg\max_{k\in I} \tau^{-1}_k$ and 
+$\arg\min_{k\in J} \sigma^{-1}_k = \arg\min_{k\in J} \tau^{-1}_k$ (that is, if $\sigma \weakeq \tau$),
+and, if we let $i$ and $j$ be defined as in Lemma~\ref{lem:yinminmax} as the indices of the maximum in $I$ and the minimum in $J$,
+$\sigma^{-1}_i > \sigma^{-1}_j \Leftrightarrow \tau^{-1}_i > \tau^{-1}_j$.
+A similar statement holds for the yang congruence $\backsim$, by exchanging $\min$ and $\max$.
+Finally, $\sigma\strongeq\tau$ if both sets of conditions hold.
 

### 2024-05-15 draft/rectangulotopes.tex
minor bug
pagebreak
typo thanks

@@
 
-\thanks{VP was partially supported by the Spanish grant PID2022-137283NB-C21 of MCIN/AEI/10.13039/501100011033 / FEDER, UE, by Departament de Recerca i Universitats de la Generalitat de Catalunya (2021 SGR 00697), by the French grant CHARMS (ANR-19-CE40-0017), and by the French--Austrian projects PAGCAP (ANR-21-CE48-0020 \& FWF I 5788).}
+\thanks{VP was partially supported by the Spanish grant PID2022-137283NB-C21 of MCIN/AEI/10.13039/501100011033 / FEDER, UE, by Departament de Recerca i Universitats de la Generalitat de Catalunya (2021 SGR 00697), by the French grant CHARMS (ANR-19-CE40-0017), and by the French--Austrian project PAGCAP (ANR-21-CE48-0020 \& FWF I 5788).}
 
@@
 %Moreover, the sets of arcs of the form~$\c{A}_\equiv$ are easily described.
-\vspace{-.2cm}
+\vspace{-.1cm}
 \parpic(5cm,2.8cm)(10pt, 90pt)[r][b]{\includegraphics[scale=.5]{subarcOrder}}{
@@
 
+\enlargethispage{.3cm}
 Note again that by definition, this extremal vertex should only depend on the strong congruence class of $\sigma$.

### 2024-10-08 draft/rectangulotopes.tex
some corrections from report X

@@
     \]
-    for all binary trees~$T$ with $n$ internal nodes, where $\ell^T_i$ and $r^T_i$ denote the number of leaves in the left and right subtrees of $i$ in $T$, see~\cite{MR2108555},
+    for all binary trees~$T$ with $n$ internal nodes, where $\ell^T_i$ and $r^T_i$ denote the number of leaves in the left and right subtrees of~$i$ in $T$ (labeled in inorder), see~\cite{MR2108555},
   \item the intersection of the hyperplane defined by $\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by
@@
 For~$n = 4$, the strong rectangulation lattice is represented in \cref{fig:strongRectangulationLattice} and the strong rectangulotope is represented in~\cref{fig:strongRectangulotope}.
+Note that, contrarily to what could suggest this $n = 4$ case, the strong rectangulation lattice is not isomorphic to the weak order, and the strong rectangulotope is not combinatorially equivalent to the permutahedron as soon as~$n > 4$.
 
@@
 For vertices, we describe the coordinates of the vertex corresponding to a weak or strong rectangulation with simple product formulas in the spirit of J.-L.~Loday's realization in~\cref{thm:loday}\,(i).
-As mentioned in~\cite{PilaudSantosZiegler}, there is no such formula for arbitrary quotientopes, which makes rectangulotopes join the restricted club of Loday type quotientopes, together with associahedra and permutreehedra~\cite{MR3856522}.
+As mentioned in~\cite{PilaudSantosZiegler}, there is no known such formula for arbitrary quotientopes, which makes rectangulotopes join the restricted club of Loday type quotientopes, together with associahedra and permutreehedra~\cite{MR3856522}.
 For facets, we show that both weak and strong rectangulotopes have $2^n-2$ facets (the maximal possible number of facets of a deformed permutahedron), and provide simple formulas for the right hand side of the inequality corresponding to a given proper nonempty subset of~$[n]$.
@@
 It is not difficult to check that the source tree~$S(R)$ (resp.~target tree~$T(R)$) is a binary tree, rooted at the bottom--left (resp.~top--right) corner of~$R$, and oriented from (resp.~towards) its root.
+Observe also that $S(R)$ and~$T(R)$ only depend on the weak class of the rectangulation~$R$.
 
@@
   \[
-  \sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{I \not\subseteq S \text{ and } I \not\subseteq [n] \ssm S}
-  \]
-  for all subsets~$\varnothing \ne S \subsetneq [n]$.
+  \sum_{i \in X} x_i \le \#\set{I \text{ interval of } [n]}{I \not\subseteq X \text{ and } I \not\subseteq [n] \ssm X}
+  \]
+  for all subsets~$\varnothing \ne X \subsetneq [n]$.
   \end{enumerate}
@@
 \raisebox{.03cm}{\verticalPattern} \!\!\!
-{}_j$ the situation where the rectangles $r_i$ and $r_j$ of inorder labels $i$ and $j$ share a vertical segment containing the right edge of~$r_i$ and the left edge of~$r_j$, and the bottom edge of $r_i$ is below the top edge of~$r_j$.
+{}_j$ the situation where the rectangles $r_i$ and $r_j$ of inorder labels $i$ and $j$ share a vertical segment containing the right edge of~$r_i$ and the left edge of~$r_j$, and the bottom edge of $r_i$ is strictly below the top edge of~$r_j$.
 Similarly, we denote by~$\; {}^i \!\!\!\!\!
 \horizontalPattern \!\!\!\!\!
-\raisebox{-.06cm}{$_j$} \,$ the situation where the rectangles $r_i$ and $r_j$ share a horizontal segment containing the bottom edge of~$r_i$ and the top edge of~$r_j$, and the right edge of~$r_i$ is on the right of the left edge of~$r_j$.
+\raisebox{-.06cm}{$_j$} \,$ the situation where the rectangles $r_i$ and $r_j$ share a horizontal segment containing the bottom edge of~$r_i$ and the top edge of~$r_j$, and the right edge of~$r_i$ is strictly on the right of the left edge of~$r_j$.
 We use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.
@@
   \[
-  \sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}
-  \]
-  for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
+  \sum_{i \in X} x_i \le \#\set{I,J}{I \not\subseteq X \text{ and } J \not\subseteq [n] \ssm X} + \#\set{I,J}{I \not\subseteq [n] \ssm X \text{ and } J \not\subseteq X}
+  \]
+  for all subsets~$\varnothing \ne X \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
   \end{enumerate}
@@
 The authors thank Namrata and Torsten M\"utze for the organization and the other participants of the workshop for the stimulating interactions, notably on other combinatorial aspects of rectangulations.
+We are grateful to two anonymous referees for many suggestions on this paper.
 
@@
 \parpic(5cm,2.8cm)(10pt, 90pt)[r][b]{\includegraphics[scale=.5]{subarcOrder}}{
-An arc~$(a, b, A, B)$ is a \defn{subarc} of an arc~$(a', b', A', B')$ if \linebreak $a' \le a < b \le b$ and~$A \subseteq A'$ while~$B \subseteq B'$.
+An arc~$(a, b, A, B)$ is a \defn{subarc} of an arc~$(a', b', A', B')$ if \linebreak $a' \le a < b \le b'$ and~$A \subseteq A'$ while~$B \subseteq B'$.
 An \defn{arc ideal} is a subset of arcs closed by subarcs.
@@
 For instance, \cref{fig:fans}\,(right) represents the \defn{sylvester fan} (the quotient fan of the sylvester congruence of \cref{fig:sylvesterCongruence}\,(middle)) for~$n = 4$, whose dual graph is the cover graph of the Tamari lattice of \cref{fig:sylvesterCongruence}\,(right).
-The maximal cone corresponding to a binary tree~$T$ is given by~$\polytope{C}^T = \set{\b{x} \in \R^n}{x_i \le x_i \text{ for each edge } i \to j \text{ in } T}$.
+The maximal cone corresponding to a binary tree~$T$ is given by~$\polytope{C}^T = \set{\b{x} \in \R^n}{x_i \le x_j \text{ for each edge } i \to j \text{ in } T}$.
 
@@
   \[
-  \Bigset{x \in \R^n}{ \sum_{i \in [n]} x_i = f([n]) \text{ and } \sum_{i \in S} x_i \le f(S) \text{ for all } \varnothing \ne S \subseteq [n]},
+  \Bigset{x \in \R^n}{ \sum_{i \in [n]} x_i = f([n]) \text{ and } \sum_{i \in X} x_i \le f(X) \text{ for all } \varnothing \ne X \subseteq [n]},
   \]
@@
 
+\begin{proof}
+$\SP(\upArc{I}) = \conv \{\b{0}\} \cup \set{ \b{e}_i - \b{e}_{\max(I)} }{ i \in I \ssm \{\max(I)\} } = \conv \set{ \b{e}_i }{ i\in I } - \b{e}_{\max(I)}$.
+\end{proof}
+
 \begin{lemma}
@@
 \end{lemma}
+
+\begin{proof}
+$\SP(\downArc{I}) = \conv \{\b{0}\} \cup \set{ \b{e}_{\min(I)} - \b{e}_i }{ i \in I \ssm \{\min(I)\} } = \conv \set{ -\b{e}_i }{ i\in I } + \b{e}_{\min(I)}$.
+\end{proof}
 
@@
   \[
-  f_{\weakeq}(S) \eqdef \#\bigset{ I\text{ interval of } [n] }{ I \not\subseteq S \text{ and } I \not\subseteq [n]\ssm S }.
+  f_{\weakeq}(X) \eqdef \#\bigset{ I\text{ interval of } [n] }{ I \not\subseteq X \text{ and } I \not\subseteq [n]\ssm X }.
   \]
@@
   \[
-  f_{\upArc{I}}(S) \eqdef \llbracket S\cap I\not=\emptyset \rrbracket .
+  f_{\upArc{I}}(X) \eqdef \llbracket X\cap I\not=\emptyset \rrbracket .
   \]
@@
   \[
-  f_{\downArc{I}}(S) \eqdef - \llbracket I\subseteq S \rrbracket .
+  f_{\downArc{I}}(X) \eqdef - \llbracket I\subseteq X \rrbracket .
   \]
@@
   \[
-  f_{\weakeq}(S) = \sum_{I\text{ interval of }[n]} f_{\upArc{I}}(S) + f_{\downArc{I}}(S)
-  \]
-  is therefore exactly the number of intervals of $I$ that intersect $S$ but are not contained in $S$, hence the number of intervals contained neither in~$S$ nor in its complement.
+  f_{\weakeq}(X) = \sum_{I\text{ interval of }[n]} f_{\upArc{I}}(X) + f_{\downArc{I}}(X)
+  \]
+  is therefore exactly the number of intervals of $I$ that intersect $X$ but are not contained in $X$, hence the number of intervals contained neither in~$X$ nor in its complement.
 \end{proof}
@@
 \end{lemma}
+
 \begin{proof}
@@
   Any such node $\ell$ either lies in the horizontal subtree of $i$, or is $i$ itself.
-  Since the weak poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_s^R i$, hence $\sigma^{-1}_\ell \leq \sigma^{-1}_i$.
+  Since the weak poset extends the order induced by $T$, we have that either $\ell=i$ or $\ell\prec_w^R i$, hence $\sigma^{-1}_\ell \leq \sigma^{-1}_i$.
   Therefore, each such choice of left endpoint for $I$ does not contradict the fact that $i$ is the index of the maximum of $\sigma^{-1}$ in $I$.
@@
   This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T$.
-  Therefore, we must have $\ell\succ_s i$, and $\sigma^{-1}_\ell > \sigma^{-1}_i$, and we cannot extend $I$ past this index on the left.
+  Therefore, we must have $\ell\succ_w i$, and $\sigma^{-1}_\ell > \sigma^{-1}_i$, and we cannot extend $I$ past this index on the left.
   This shows that there are exactly $h^T_i$ choices for the choice of the left endpoint of $I$.
@@
 
-The arc ideal of the weak rectangulation congruence was already considered in~\cite[Exm.~4.11]{Reading-arcDiagrams}.
+The arc ideal of the strong rectangulation congruence was already considered in~\cite[Exm.~4.11]{Reading-arcDiagrams}.
 
@@
   We consider the forbidden vincular patterns $24\underline{51}3$, $42\underline{51}3$, $3\underline{51}24$, and $3\underline{51}42$ that define the 2-clumped permutations, the minimal elements of each congruence class of $\strongeq$.
-  We observe that every occurrence of a pattern in a permutation corresponds to an occurrence of an arc that cross the horizontal line at least twice in the arc diagram of the permutation.
+  We observe that every occurrence of a pattern in a permutation corresponds to an occurrence of an arc that crosses the horizontal line at least twice in the arc diagram of the permutation.
   The 2-clumped permutations are therefore the permutations whose arc diagrams are composed only of arcs that cross the horizontal line at most once.
@@
 We classify these arcs into two families, depending on whether the arcs do not cross the horizontal line from below or from above.
-For two consecutive intervals~$I$ and~$J$ of~$[n]$, the \defn{yin arc}~$\yinArc{I}{J}$ (resp.~\defn{yang arc}~$\yangArc{I}{J}$) is the arc starts at~$\min(I)$ and ends at~$\max(J)$, and goes above (resp.~below) the remaining elements of $I$ and below (resp.~above) the remaining elements of $J$, that is,
+For two consecutive intervals~$I$ and~$J$ of~$[n]$, the \defn{yin arc}~$\yinArc{I}{J}$ (resp.~\defn{yang arc}~$\yangArc{I}{J}$) is the arc that starts at~$\min(I)$ and ends at~$\max(J)$, and goes above (resp.~below) the remaining elements of $I$ and below (resp.~above) the remaining elements of $J$, that is,
 \begin{align*}
@@
 
+\begin{proof}
+$\SP(\yangArc{I}{J}) = \conv \{\b{0}\} \cup \{\b{e}_{\min(I)} - \b{e}_{\max(J)}\} \cup \set{ \b{e}_{\min(I)} - \b{e}_i  + \b{e}_j - \b{e}_{\max(J)}}{ i \in I \ssm \{\min(I)\}, j \in J \ssm \{\max(I)\} } = \conv \{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{i \in I,\; j \in J} + \b{e}_{\min(I)} - \b{e}_{\max(J)}$.
+\vincent{horrible...}
+\end{proof}
+
 We refer to \cref{fig:shardPolytopes} for examples of yin and yang shard polytopes.
@@
   \[
-  f_{\strongeq}(S) \eqdef \# \bigset{ I, J }{ I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S } +
-  \#\bigset{ I, J }{ I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S },
+  f_{\strongeq}(X) \eqdef \# \bigset{ I, J }{ I \not\subseteq [n] \ssm X \text{ and } J \not\subseteq X } +
+  \#\bigset{ I, J }{ I \not\subseteq X \text{ and } J \not\subseteq [n] \ssm X },
   \]
@@
   \[
-  f_{\yinArc{I}{J}}(S) \eqdef \llbracket I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S \rrbracket .
+  f_{\yinArc{I}{J}}(X) \eqdef \llbracket I \not\subseteq [n] \ssm X \text{ and } J \not\subseteq X \rrbracket .
   \]
@@
   \[
-  f_{\yangArc{I}{J}}(S) \eqdef \llbracket I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S \rrbracket .
+  f_{\yangArc{I}{J}}(X) \eqdef \llbracket I \not\subseteq X \text{ and } J \not\subseteq [n] \ssm X \rrbracket .
   \]
@@
   \[
-  f_{\strongeq}(S) \eqdef \sum_{I,J} f_{\yinArc{I}{J}}(S) + f_{\yangArc{I}{J}}(S)
+  f_{\strongeq}(X) \eqdef \sum_{I,J} f_{\yinArc{I}{J}}(X) + f_{\yangArc{I}{J}}(X)
   \]
@@
   \begin{itemize}
-  \item $h^T_i$ denotes the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
+  \item $h^T_i$ denotes the number of leaves in the horizontal subtree of $i$ in~$T$,
   \item $cv^{T,S}_{i,j}$ denotes the number of common leaves of the vertical subtree of $i$ in $T$ and the vertical subtree of $j$ in $S$.

### 2024-10-14 draft/rectangulotopes.tex
some corrections from report Y

@@
 	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$ (right). \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
+	\vspace{-.4cm}
 	\label{fig:quotientopes}
@@
 Only the third one is diagonal.}
+	\vspace{-.5cm}
 	\label{fig:rectequiv}
@@
 
+\enlargethispage{.2cm}
 Weak and strong rectangulations of size $n$ have been shown to define congruences of the weak Bruhat order on $\f{S}_n$.
@@
 
-Recall that the \defn{inorder labeling} of a binary tree~$T$ is the labeling of the nodes of~$T$ such that the label of each node~$t$ of~$T$ is larger than all labels in the left subtree of~$t$ and larger than all labels in the right subtree of~$t$.
+Recall that the \defn{inorder labeling} of a binary tree~$T$ is the labeling of the nodes of~$T$ such that the label of each node~$t$ of~$T$ is larger than all labels in the left subtree of~$t$ and smaller than all labels in the right subtree of~$t$.
 For any rectangle~$r$ of~$R$, the inorder label of the source~$s(r)$ in~$S(R)$ coincides with the inorder label of the target~$t(r)$ in~$T(R)$.
@@
 From now on, the labels of the rectangles in a rectangulation are the inorder labels, and the two trees~$T(R)$ and $S(R)$ are defined on the vertex set $[n]$ accordingly.
-For each $i\in [n]$, we refer to its two subtrees in $T(R)$ as the \defn{horizontal} and \defn{vertical subtrees}, and similarly for $S(R)$.
+For each vertex in $T(R)$ (or in~$S(R)$), we refer to its subtree reached by a horizontal (resp.~vertical) edge as its \defn{horizontal} (resp.~\defn{vertical})~\defn{subtree}.
 
@@
 \raisebox{-.06cm}{$_j$} \,$ the situation where the rectangles $r_i$ and $r_j$ share a horizontal segment containing the bottom edge of~$r_i$ and the top edge of~$r_j$, and the right edge of~$r_i$ is strictly on the right of the left edge of~$r_j$.
-We use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.
+We use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise, and~$\neg \, \varphi$ for the logical negation of~$\varphi$.
 
@@
   \]
-  for all subsets~$\varnothing \ne X \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
+  for all subsets~$\varnothing \ne X \subsetneq [n]$, where~$I,J$ are nonempty consecutive intervals of~$[n]$, meaning that~$\max(I) = \min(J)-1$.
   \end{enumerate}
@@
 
-\enlargethispage{.2cm}
 We think of the formula of \cref{thm:strongRectangulotope}\,(i) as the analogue of J.-L.-Loday's product formula of \cref{thm:loday}\,(i) for the coordinates of the vertices of the associahedron.
@@
 
-We now briefly recall some results on the weak Bruhat order and its quotients, both from a lattice and geometric perspectives.
+We now briefly recall some results on the weak Bruhat order and its quotients, both from a lattice and geometric perspective.
 We refer to \cite{MR2142177, Reading-arcDiagrams, MR3645056, MR3645055, MR3964495, MR4584712} for details.
@@
 
-\pagebreak 
 A set function $f:2^{[n]}\mapsto \R$ is said to be \defn{submodular} when $f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$.
@@
 
+\pagebreak
 \subsection{The weak rectangulation congruence}
@@
 
-\pagebreak
 \begin{theorem}[\cite{MR2864445,ACFF24}]
@@
   \]
-  where $I, J$ are pairs of nonempty consecutive intervals of~$[n]$.
+  where $I, J$ are nonempty consecutive intervals of~$[n]$.
 \end{lemma}
@@
 
-\enlargethispage{.3cm}
 Note again that by definition, this extremal vertex should only depend on the strong congruence class of $\sigma$.
 This, again, gives us alternative characterizations of the yin congruence $\sim$, the yang congruence $\backsim$, and the strong rectangulation congruence $\strongeq$.
-We have $\sigma \sim \tau$ if and only if for any pair of consecutive intervals $I$ and $J$ of $[n]$, 
+We have $\sigma \sim \tau$ if and only if for any two consecutive intervals $I$ and $J$ of $[n]$, 
 $\arg\max_{k\in I} \sigma^{-1}_k = \arg\max_{k\in I} \tau^{-1}_k$ and 

### 2024-10-17 response/response.tex
move response

\documentclass{article}
\usepackage{verbatim,a4wide}
\title{Eur.~J.~Comb. 240426: Response to referees}
\author{Jean Cardinal \and Vincent Pilaud}
\begin{document}
\maketitle

The authors warmly thank the referees for their careful reading and numerous insightful comments. We consider each of them below.

\section{Report X}

\begin{verbatim}
a question out of curiosity: is there an explicit formula (in each case, and 
also for the associahedron) for a direction under which the poset is increasing, 
for the given geometric realization? (if there is an answer it could be 
mentioned)
\end{verbatim}

\begin{verbatim}
p2 in Theorem 1 it should be mentioned that "node i" means the ith node in 
infix order
\end{verbatim}

Done.

\begin{verbatim}
bottom p2 "describe the geometry", should it not be "provide an explicit 
geometric realization" ? (since the geometric realization is not unique)
\end{verbatim}

?

\begin{verbatim}
p3 the recursive definition of "above" is not ideal. Maybe cut it into 
defining "directly above" and then "above" (recursively using "directly 
above")
\end{verbatim}

?

\begin{verbatim}
p4 when citing Figure 6 it could be mentioned (e.g. in a footnote) that 
even if for n=4 the strong rectangulotope is the permutohedron 
(combinatorially), the geometric realization is not the same (I am not 
very familiar with quotientopes but am a bit amazed by this fact, is it 
that if you remove deletable hyperplanes from the geometric realization 
of the permutohedron you would have a geometric realization of the strong 
rectangulotope with more involved vertex coordinates ?)
\end{verbatim}

Added a remark?

\begin{verbatim}
p7 Figure 7 caption, you should mention if it is for weak of strong 
(or both), and what the gray areas mean
\end{verbatim}

J: These are the flips in strong rectangulations.

\begin{verbatim}
p8 l2 "there is no such formula" -> "there is no known such formula" 
(or is it sure no such formula can exist?)
\end{verbatim}

Done.

\begin{verbatim}
p8 it could be mentioned when defining S(R) and T(R) that they only 
depend (?) on the weak class of R
\end{verbatim}

Done.

\begin{verbatim}
p8 Theorem 2, it is a bit confusing to have notation S with two meanings, 
maybe it could be changed for the subset notation in (ii) (and also other 
instances)
\end{verbatim}

Done.

\begin{verbatim}
p10 first paragraph of 1.6: definition of the small inline symbol, the text 
seems to allow for the top edge of r_i to be aligned with the top edge of 
r_j, and bottom edge of r_i to be aligned with bottom edge of r_j, which 
is not what the drawing suggests
\end{verbatim}

J: I think this is OK as is.

\begin{verbatim}
For theorem 3 you could comment on what quantities in the vertex coordinates 
formula depends on the strong type and not just on the weak type (it seems 
to be only the Iverson bracket?)
\end{verbatim}

J: Good point. The number of common leaves can also change, right?

\begin{verbatim}
p10 l.-2 when you say "quadratic in n" do you mean that the ith coordinate 
is O(n^2) or just that the sum is over a quadratic set of indices ?
\end{verbatim}

J: at least the second.

\begin{verbatim}
p12 second line of paragraph starting with "An arc...", b should be b' in 
"a'\leq a<b\leq b" 
\end{verbatim}

Done.

\begin{verbatim}
p12 "It has a ray" is a bit vague, maybe give more visual intuition on what 
is a ray (is it a codimension 1 cone? or a union of some?)
\end{verbatim}

We added a remark.

\begin{verbatim}
p13 l4 x_i\leq x_i should be x_i\leq x_j
\end{verbatim}

Done.

\begin{verbatim}
p13 l11 the remark "(in other words, if it is above..." may be too early 
(not sure it follows from something explained before)
\end{verbatim}

?

\begin{verbatim}
p15 "bijection between the facets" -> "bijection between the facets of 
dimension 1" (?) also at the end of Section 4.2
\end{verbatim}

J: No, it is the facets (eg faces of codimension 1) of the cone, as written.

\begin{verbatim}
p16 it could be good to give a short proof for Lemma 6 (e.g. explain which 
translate)
\end{verbatim}

Todo...

\begin{verbatim}
p17 again it would be good to give a short proof of Lemma 10 (or a citation)
\end{verbatim}

?

\begin{verbatim}
p17 in proof of Lemma 11, \prec_s^R whould be \prec_w^R,   and \succ_s 
should be \succ_w^R
\end{verbatim}

Done.

\begin{verbatim}
p19 l1 of 4.3 "of the weak rectangulation congruence" -> "of the strong 
rectangulation congruence"
\end{verbatim}

Done.

\begin{verbatim}
p19 l4 of proof of Lemma 14: "cross" -> "crosses"
\end{verbatim}

Done.

\begin{verbatim}
p19 middle "the arc starts" -> "the arc that starts"
\end{verbatim}

Done.

\begin{verbatim}
p19 bottom "the up and down shard polytopes" -> "the up shard polytopes 
and the down shard polytopes" (same for "each up and down...")
\end{verbatim}

?

\begin{verbatim}
p20 Lemma 20 in the first bullet: "and vertical" should be deleted
\end{verbatim}

Done.

\begin{verbatim}
p21 two lines before Lemma 21 (announcing Lemma 21) you write "which do 
not have any associated yin arc", but this does not appear anymore in the 
statement of Lemma 21 (I do not see well how it relates)
\end{verbatim}

Todo.

\begin{verbatim}
p21 in the proof of Lemma 21 "the right edge of r_i can not be on the 
right of the left edge of r_j" I wonder if it is well defined (it may 
depend on the geometric realization of the strong rectangulation)
\end{verbatim}

Todo.

\begin{verbatim}
after Lemma 21, I am not sure Lemmas 20 and 21 are enough to give the 
first part of Theorem 3, this would be the case if \prec_s was a total order. 
The case of pairs i,j not comparable in the strong order may also have to be 
covered in Lemma 21
\end{verbatim}

?

\begin{verbatim}
p21 after "this concludes the proof of Theorem 3" you could add a remark on 
whether there is an analogue of Theorem 3 for semi-Baxter permutations and 
their corresponding strong rectangulations, this could be a nice corollary
\end{verbatim}

Todo.

\section{Report Y}

\begin{verbatim}
Page 4:

"The strong rectangulation congruence... Its classes are in bijection with 
strong rectangulations.  The corresponding quotientopes will be referred 
to as the strong rectangulotopes and denoted by SR(n). Strong rectangulotopes 
can be obtained as quotientopes [PS19, PPR23] for the strong rectangulation 
congruence."

The last sentence seems to just repeat what has already been 
said earlier in the paragraph (except that it adds important references).  
Perhaps this paragraph can be revised.
\end{verbatim}

Todo.

\begin{verbatim}
Page 8:

In the definition of "inorder", the word "larger" appears twice and one of 
the instances should be "smaller".  ("...smaller than all labels in the right 
subtree...")
\end{verbatim}

Done.

\begin{verbatim}
The definition of horizontal and vertical subtrees could be made clearer, 
especially since it will be applied to two different trees that appear in 
two orientations.  I assume that the horizontal subtree for i is the subtree 
that is reached from i by a horizontal edge, but it would be good to say 
that explicitly (or correct my misunderstanding).
\end{verbatim}

Done.

\begin{verbatim}
Page 10:

Some readers may appreciate a sentence defining the logical not symbol 
that you use (perhaps in the sentence after defining the Iverson bracket.
\end{verbatim}

Done. (J:?)

\begin{verbatim}
In Theorem 3, there needs to be more clarification on "consecutive intervals".  
At first glance, I wasn't even sure whether that was a description of each 
interval separately or a description of the pair of intervals.  And then it 
wasn't clear whether they could be consecutive in either order or if I is always 
smaller.  After looking deeper into the paper, I conclude that you mean that 
the smallest element of J is one more than the largest element of I.  But you 
should say that explicitly somewhere (or correct my misunderstanding), 
probably right before the statement of Theorem 3.
\end{verbatim}

Done.

\begin{verbatim}
Page 11:

Typo in "...both from a lattice and geometric perspectives".  
Either an "a" or an "s" should be deleted.
\end{verbatim}

Done.

\begin{verbatim}
Page 12:

"It has a chamber for each permutation sigma of [n], given by the set of vectors 
whose coordinates are ordered as sigma."  It's a little unclear what you mean 
here, but as a "public service", it might be worth being very clear and pointing 
out to make these regions to match the vertices of the permutohedron in such 
a way that the edges are *right* weak order cover relations.  (As you know, 
it's *not* the obvious thing, "make the entries of sigma into a vector and 
let sigma label the region that contains that vector", but everyone assumes 
it is.)  Probably you could say something useful in far fewer words than I 
just used.
\end{verbatim}

Todo.

\begin{verbatim}
Page 13:

"An immediate consequence is that all 2n - 2 rays of the braid arrangement 
are preserved in the quotient fan F if and only if the arc ideal A contains 
all up arcs and all down arcs (in other words, if it is above the weak 
rectangulation congruence in the lattice of congruences of the weak Bruhat order)."  
This is a very interesting property of the weak rectangulation congruence.  
I wonder if you should make more of it (perhaps in the introduction and maybe 
also by making it a numbered result). (But also, maybe it belongs later 
in the paper than page 13, after you have said more about this congruence?  
On page 13, you don't yet know the connection between this congruence and up/down arcs.  
Maybe all of this belongs at the end of Section 3.3.)
\end{verbatim}

Todo.

\begin{verbatim}
Page 15:

There is a missing head on the arrow to 12 on the right drawing of Figure 13.
\end{verbatim}

Done.

\end{document}

### 2024-10-17 response/response.tex
many corrections from referees

@@
 \documentclass{article}
-\usepackage{verbatim,a4wide}
+\usepackage{verbatim,a4wide,bm,amssymb}
 \title{Eur.~J.~Comb. 240426: Response to referees}
@@
 
+Done. We have added the explicit direction in the first paragraph of Section 2.4.
+
 \begin{verbatim}
@@
 
-Done.
+Done. Thank you for noticing it.
 
@@
 
-?
+Done. It was indeed unclear.
 
@@
 
-?
+Done.
+We unrolled the recursion with the sentence ``there is a sequence~${r = r_0, \dots, r_k = s}$ of rectangles such that the bottom edge of~$r_{i-1}$ lies in the same segment as the top edge of~$r_i$ for all~$i \in [k]$''.
 
@@
 
-Added a remark?
+Done.
+We have added the following sentence: ``Note that, contrarily to what could suggest this $n = 4$ case, the strong rectangulation lattice is not isomorphic to the weak order, and the strong rectangulotope is not combinatorially equivalent to the permutahedron as soon as~$n > 4$.''
 
@@
 
-J: These are the flips in strong rectangulations.
+Done.
+We have added ``strong'' in the caption of Figure 7.
+We believe that the second sentence of the caption of Figure 7 explains the meaning of the grey areas.
+We also slightly modified the last paragraph of page 4 to be more precise on the definitions of the weak and strong rectangulation flip graphs.
 
@@
 
-J: I think this is OK as is.
+Done. We slightly modified the symbols to avoid this problem.
 
@@
 
-J: Good point. The number of common leaves can also change, right?
+Done. We added a the sentence ``Note that the common leaves of~$T(R)$ and~$S(R)$ only depend on the weak class of the rectangulation~$R$'' after the definition of common leaves.
 
@@
 
-J: at least the second.
+Done. We meant a quadratic set of indices.
 
@@
 
-We added a remark.
+Done. A ray is a dimension~$1$ cone. We have modified the sentence just after to explain it properly.
 
@@
 
-?
+Done. We have delayed this observation to Remark 10.
 
@@
 
-J: No, it is the facets (eg faces of codimension 1) of the cone, as written.
+Our sentence is correct as the facets of a cone are its codimension~$1$ faces.
 
@@
 
-Todo...
+Done. We added the explicit translation vector in Lemmas 6, 7, and 17.
 
@@
 
-?
+Done.
 
@@
 
-?
+We have decided to keep our formulation for brievety.
 
@@
 
-Todo.
+Done. We have reformulated the last two sentences of the paragraph to make the argument clearer.
 
@@
 
-?
+Done. You are right, it was incorrect. We replaced $i \prec_s^R j$ by $i \not\succ_s^R j$ in the hypothesis of Lemma 22. This is actually what the proof shows.
 
@@
 
-Todo.
+Done. We reformulated the last sentence to ``They can be obtained from the constructions of [PS19, PPR23]''.
 
@@
 
-Done.
+Done. Thank you for noticing this.
 
@@
 
-Done.
+Done. This is indeed clearer.
 
@@
 
-Done. (J:?)
+Done. Just before Theorem~3.
 
@@
 
-Done.
+Done. We have decided to do it directly inside Theorem~3. The meaning is really that $\max(I) = \min(J)-1$.
 
@@
 
-Todo.
+Done.
+You are right.
+We replaced our unclear sentence by a formula: ``It has a chamber $\{\boldsymbol{x} \in \mathbb{R}^n \,|\, x_{\sigma_1} \le \dots \le x_{\sigma_n}\}$ for each permutation~$\sigma$ of~$[n]$''.
 
@@
 
-Todo.
+Done. We have delayed this observation to Remark 10.
 
@@
 
-Done.
+Done. Good point!
 

### 2024-10-17 draft/rectangulotopes.tex
many corrections from referees

@@
 
-The goal of this paper is to describe the geometry of the quotientopes for two particular congruences of the weak Bruhat order whose classes are in bijection with equivalence classes of rectangulations, defined as decompositions of a square into rectangles.
+The goal of this paper is to provide explicit geometric realizations of the quotientopes for two particular congruences of the weak Bruhat order whose classes are in bijection with equivalence classes of rectangulations, defined as decompositions of a square into rectangles.
 
@@
 	\centerline{\includegraphics[scale=.75]{permutahedronLeft4} \quad \raisebox{-.5cm}{\includegraphics[scale=.75]{associahedron4}}}
+	\vspace{-.2cm}
 	\caption{The permutahedron~$\Perm(4)$ (left) and the associahedron~$\Asso(4)$ (right). \cite[Fig.~1]{MR3964495} \&~\cite[Fig.~5]{MR4584712}}
+	\label{fig:quotientopes}
 	\vspace{-.4cm}
-	\label{fig:quotientopes}
 \end{figure}
@@
 
+\enlargethispage{.3cm}
 The combinatorics of rectangulations has been studied for nearly two decades~\cite{MR2233287,MR2763051,MR2871762,MR2864445,MR3084577,MR3192492,MR3878132,MR4598046,MR4667583,AB24,ACFF24}.
@@
 In order to focus on the combinatorial structure of a rectangulation, we recall two equivalence relations between rectangulations (see for instance~\cite{ACFF24} and references therein).
-We say that a rectangle~$r$ is \defn{above} another rectangle~$s$ (and~$s$ is \defn{below}~$r$) if either the bottom edge of~$r$ lies in the same segment as the top edge of~$s$, or if~$r$ is above another rectangle~$t$ that is above~$s$.
+We say that a rectangle~$r$ is \defn{above} another rectangle~$s$ (and~$s$ is \defn{below}~$r$) if there is a sequence~${r = r_0, \dots, r_k = s}$ of rectangles such that the bottom edge of~$r_{i-1}$ lies in the same segment as the top edge of~$r_i$ for all~$i \in [k]$.
+%either the bottom edge of~$r$ lies in the same segment as the top edge of~$s$, or, recursively, if~$r$ is above another rectangle~$t$ that is above~$s$.
 We define \defn{on the left of} and \defn{on the right of} similarly.
@@
 Only the third one is diagonal.}
-	\vspace{-.5cm}
 	\label{fig:rectequiv}
@@
 
-\enlargethispage{.2cm}
 Weak and strong rectangulations of size $n$ have been shown to define congruences of the weak Bruhat order on $\f{S}_n$.
@@
 The corresponding quotientopes will be referred to as the \defn{strong rectangulotopes} and denoted by~$\SRP(n)$.
-Strong rectangulotopes can be obtained as quotientopes~\cite{MR3964495,MR4584712} for the strong rectangulation congruence.
+They can be obtained from the constructions of~\cite{MR3964495,MR4584712}.
 For~$n = 4$, the strong rectangulation lattice is represented in \cref{fig:strongRectangulationLattice} and the strong rectangulotope is represented in~\cref{fig:strongRectangulotope}.
@@
 The various types of flips are illustrated in \cref{fig:flips}.
-We refer to \cite{MR2871762,MR3878132,MR3697823,ACFF24} for detailed descriptions of the flip graphs on both weak and strong rectangulations.
+Those flips define the flip graph on strong rectangulations.
+We refer to \cite{MR2871762,MR3878132,MR3697823} for a detailed description of the flip graphs on weak rectangulations.
 Just like associahedra encode the rotation graphs on binary trees, the rectangulotopes directly yield flip graphs on rectangulations.
@@
     \end{center}
-	\caption{The three types of flips between rectangulations. The flips are possible only if the grey areas do not intersect any segment of the rectangulation. \cite[Fig.~19]{ACFF24}}
+	\caption{The three types of flips between strong rectangulations. The flips are possible only if the grey areas do not intersect any segment of the rectangulation. \cite[Fig.~19]{ACFF24}}
 	\label{fig:flips}
@@
 Our main results are elementary vertex and facet descriptions of both families of quotientopes.
-For vertices, we describe the coordinates of the vertex corresponding to a weak or strong rectangulation with simple product formulas in the spirit of J.-L.~Loday's realization in~\cref{thm:loday}\,(i).
+For vertices, we describe the coordinates of the vertex corresponding to a weak or strong rectangulation with simple formulas in the spirit of J.-L.~Loday's realization in~\cref{thm:loday}\,(i).
 As mentioned in~\cite{PilaudSantosZiegler}, there is no known such formula for arbitrary quotientopes, which makes rectangulotopes join the restricted club of Loday type quotientopes, together with associahedra and permutreehedra~\cite{MR3856522}.
@@
 Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$.
-%
-We denote by ${}^i \!\!\!
+Note that the common leaves of~$T(R)$ and~$S(R)$ only depend on the weak class of the rectangulation~$R$.
+
+We denote by ${}_i \!\!\!
 \raisebox{.03cm}{\verticalPattern} \!\!\!
-{}_j$ the situation where the rectangles $r_i$ and $r_j$ of inorder labels $i$ and $j$ share a vertical segment containing the right edge of~$r_i$ and the left edge of~$r_j$, and the bottom edge of $r_i$ is strictly below the top edge of~$r_j$.
-Similarly, we denote by~$\; {}^i \!\!\!\!\!
-\horizontalPattern \!\!\!\!\!
-\raisebox{-.06cm}{$_j$} \,$ the situation where the rectangles $r_i$ and $r_j$ share a horizontal segment containing the bottom edge of~$r_i$ and the top edge of~$r_j$, and the right edge of~$r_i$ is strictly on the right of the left edge of~$r_j$.
+{}^j$ the situation where the rectangles $r_i$ and $r_j$ of inorder labels $i$ and $j$ share a vertical segment containing the right edge of~$r_i$ and the left edge of~$r_j$, and the bottom edge of $r_i$ is below the top edge of~$r_j$.
+Similarly, we denote by~$\;\;\; {}^i \!\!\!\!\!\!\!\!
+\horizontalPattern \!\!\!\!\!\!\!\!
+\raisebox{-.06cm}{$_j$} \;\;\;$ the situation where the rectangles $r_i$ and $r_j$ share a horizontal segment containing the bottom edge of~$r_i$ and the top edge of~$r_j$, and the right edge of~$r_i$ is on the right of the left edge of~$r_j$.
 We use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise, and~$\neg \, \varphi$ for the logical negation of~$\varphi$.
@@
   \[
-    \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \; \neg \; {}^i \!\!\! \raisebox{.03cm}{\verticalPattern} \!\!\! {}_j \; \rrbracket
+    \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \; \neg \; {}_i \!\!\! \raisebox{.03cm}{\verticalPattern} \!\!\! {}^j \; \rrbracket
     \qquad\text{and}\qquad
-    \yang{w}^R_{i,j} \eqdef v^S_i \cdot ch^{S,T}_{i,j}\cdot v^T_j \cdot \llbracket \; \neg \;\; {}^i \!\!\!\!\! \horizontalPattern \!\!\!\!\! \raisebox{-.06cm}{$_j$} \;\; \rrbracket,
+    \yang{w}^R_{i,j} \eqdef v^S_i \cdot ch^{S,T}_{i,j}\cdot v^T_j \cdot \llbracket \; \neg \;\;\;\; {}^i \!\!\!\!\!\!\!\!
+\horizontalPattern \!\!\!\!\!\!\!\!
+\raisebox{-.06cm}{$_j$} \;\; \;\; \rrbracket,
   \]
@@
 We think of the formula of \cref{thm:strongRectangulotope}\,(i) as the analogue of J.-L.-Loday's product formula of \cref{thm:loday}\,(i) for the coordinates of the vertices of the associahedron.
-We note however that our formula is quadratic in~$n$.
+We note however that our formula has a quadratic number of terms.
 It remains unclear to us if there is a simple linear formula to compute the coordinates of the vertex of~$\SRP(n)$ corresponding to a given strong rectangulation.
@@
 
-The \defn{braid arrangement} is the hyperplane arrangement consisting of the hyperplanes~$\set{\b{x} \in \R^n}{x_a = x_b}$ for all~$1 \le a < b \le n$.
-It has a chamber for each permutation~$\sigma$ of~$[n]$, given by the set of vectors whose coordinates are ordered as~$\sigma$.
+The \defn{braid arrangement} is the arrangement consisting of the hyper\-planes~$\set{\b{x} \in \R^n}{x_a = x_b}$ for all~$1 \le a < b \le n$.
+It has a chamber~$\set{\b{x} \in \R^n}{x_{\sigma_1} \le \dots \le x_{\sigma_n}}$ for each permutation~$\sigma$ of~$[n]$. %, given by the set of vectors whose coordinates are ordered as~$\sigma$.
 It has a ray for each nonempty proper subset~$\varnothing \ne J \subsetneq [n]$.
-(To be more precise when speaking about rays, we should actually consider the essentialization of the braid arrangement obtained by slicing it with a hyperplane normal to its lineality space, which is generated by the vector~$(1, \dots, 1)$.)
 See \cref{fig:fans}\,(left).
+(A ray is a dimension~$1$ cone in the essentialization of the braid arrangement obtained by slicing it with a hyperplane normal to its lineality space, which is generated by the vector~$(1, \dots, 1)$.)
 
@@
 It was shown in~\cite[Lem.~23]{MR4328906} that the ray of the braid arrangement corresponding to a nonempty proper subset~$\varnothing \ne J \subsetneq [n]$ is preserved in the quotient fan~$\c{F}_\equiv$ if and only if the arc ideal~$\c{A}_\equiv$ contains the $|J|-1$ down arcs joining two consecutive elements of~$J$ and the~$n-|J|-1$ up arcs joining two consecutive elements of~$[n] \ssm J$.
-An immediate consequence is that all $2^n-2$ rays of the braid arrangement are preserved in the quotient fan~$\c{F}_\equiv$ if and only if the arc ideal~$\c{A}_\equiv$ contains all up arcs and all down arcs (in other words, if it is above the weak rectangulation congruence in the lattice of congruences of the weak Bruhat order).
+An immediate consequence is that all $2^n-2$ rays of the braid arrangement are preserved in the quotient fan~$\c{F}_\equiv$ if and only if the arc ideal~$\c{A}_\equiv$ contains all up arcs and all down arcs.
 
@@
 A \defn{quotientope} for a lattice congruence~$\equiv$ of the weak Bruhat order is a polytope whose normal fan is the quotient fan~$\c{F}_\equiv$.
-In particular, the skeleton of a quotientope is isomorphic to the cover graph of the quotient~$\f{S}_n/{\equiv}$.
-For instance, the associahedron~$\Asso(n)$ (\cref{fig:quotientopes}\,(right) when~$n = 4$) is a quotientope for the sylvester congruence (\cref{fig:sylvesterCongruence}\,(middle) when~$n = 4$), its normal fan is the sylvester fan (\cref{fig:fans}\,(right) when~$n = 4$), and its skeleton is the cover graph of the Tamari lattice (\cref{fig:sylvesterCongruence}\,(right) when~$n = 4$).
+In particular, the skeleton of a quotientope, oriented in the direction~$(n,n-1, \dots, 1) - (1, 2, \dots, n) = (n-1, n-3, \dots, 3-n, 1-n)$, is isomorphic to the Hasse diagram of the quotient~$\f{S}_n/{\equiv}$.
+For instance, the associahedron~$\Asso(n)$ (\cref{fig:quotientopes}\,(right) when~$n = 4$) is a quotientope for the sylvester congruence (\cref{fig:sylvesterCongruence}\,(middle) when~$n = 4$), its normal fan is the sylvester fan (\cref{fig:fans}\,(right) when~$n = 4$), and its oriented skeleton is the Hasse diagram of the Tamari lattice (\cref{fig:sylvesterCongruence}\,(right) when~$n = 4$).
 The existence of quotientopes was first proved in \cite{MR3964495} and later better understood in~\cite{MR4584712} using Minkowski sums of shard polytopes, the approach we will use throughout the paper.
@@
 From the results of \cref{subsec:quotientopes}, the quotientope is realized by a Minkowski sum of the shard polytopes of the up arcs and the down arcs.
-These quotientopes are simplices and anti-simplices.
+It immediately follows from the definition that these quotientopes are simplices and anti-simplices.
 %See \cref{fig:shardPolytopes}\,(left).
@@
   \label{lem:lodaysp}
-  The up shard polytope $\SP(\upArc{I})$ is a translate of the simplex~$\conv \set{ \b{e}_i }{ i\in I }$.
+  The up shard polytope $\SP(\upArc{I})$ is the translate of the simplex~$\conv \set{ \b{e}_i }{ i\in I }$ by the vector~$- \b{e}_{\max(I)}$.
 \end{lemma}
 
-\begin{proof}
-$\SP(\upArc{I}) = \conv \{\b{0}\} \cup \set{ \b{e}_i - \b{e}_{\max(I)} }{ i \in I \ssm \{\max(I)\} } = \conv \set{ \b{e}_i }{ i\in I } - \b{e}_{\max(I)}$.
-\end{proof}
+%\begin{proof}
+%$\SP(\upArc{I}) = \conv \{\b{0}\} \cup \set{ \b{e}_i - \b{e}_{\max(I)} }{ i \in I \ssm \{\max(I)\} } = \conv \set{ \b{e}_i }{ i\in I } - \b{e}_{\max(I)}$.
+%\end{proof}
 
@@
   \label{lem:antilodaysp}
-  The down shard polytope $\SP(\downArc{I})$ is a translate of the negative simplex~$\conv \set{ - \b{e}_i }{ i\in I }$.
+  The down shard polytope $\SP(\downArc{I})$ is the translate of the negative simplex~$\conv \set{ - \b{e}_i }{ i\in I }$ by the vector~$\b{e}_{\min(I)}$.
 \end{lemma}
 
-\begin{proof}
-$\SP(\downArc{I}) = \conv \{\b{0}\} \cup \set{ \b{e}_{\min(I)} - \b{e}_i }{ i \in I \ssm \{\min(I)\} } = \conv \set{ -\b{e}_i }{ i\in I } + \b{e}_{\min(I)}$.
-\end{proof}
+%\begin{proof}
+%$\SP(\downArc{I}) = \conv \{\b{0}\} \cup \set{ \b{e}_{\min(I)} - \b{e}_i }{ i \in I \ssm \{\min(I)\} } = \conv \set{ -\b{e}_i }{ i\in I } + \b{e}_{\min(I)}$.
+%\end{proof}
 
@@
 
+\begin{remark}
+Note that the result of~\cite{MR4328906} actually implies that all $2^n-2$ rays of the braid arrangement are preserved in the quotient fan~$\c{F}_\equiv$ if and only if~$\equiv$ refines the weak rectangulation congruence~$\weakeq$.
+\end{remark}
+
 %%%%%%%%%%%%%
@@
 \end{lemma}
+
+\begin{proof}
+Recall from \cref{lem:lodaysp,lem:antilodaysp} that the up shard polytope~$\SP(\upArc{I})$ (resp.~down shard polytope~$\SP(\downArc{I})$) is the simplex~$\conv \set{ \b{e}_i }{ i\in I }$ (resp.~the negative simplex~$\conv \set{ - \b{e}_i }{ i\in I }$).
+Hence, the optimal vertex of~$\SP(\upArc{I})$ (resp.~$\SP(\downArc{I})$) in direction~$\sigma^{-1}$ is $\b{e}_i$ (resp.~$-\b{e}_i$), where $i$ is the maximum (resp.~the minimum) of~$\sigma^{-1}$ in $I$.
+\end{proof}
 
@@
   This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T$.
-  Therefore, we must have $\ell\succ_w i$, and $\sigma^{-1}_\ell > \sigma^{-1}_i$, and we cannot extend $I$ past this index on the left.
+  Therefore, we must have $\ell\succ_w^R i$, and $\sigma^{-1}_\ell > \sigma^{-1}_i$, and we cannot extend $I$ past this index on the left.
   This shows that there are exactly $h^T_i$ choices for the choice of the left endpoint of $I$.
@@
   \label{lem:yangsp}
-  The yang shard polytope $\SP(\yangArc{I}{J})$ is a translate of~$\conv \{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{i \in I,\; j \in J}$.
+  The yang shard polytope $\SP(\yangArc{I}{J})$ is the translate of~$\conv \{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{i \in I,\; j \in J}$ by the vector~$\b{e}_{\min(I)} - \b{e}_{\max(J)}$.
 \end{lemma}
 
-\begin{proof}
-$\SP(\yangArc{I}{J}) = \conv \{\b{0}\} \cup \{\b{e}_{\min(I)} - \b{e}_{\max(J)}\} \cup \set{ \b{e}_{\min(I)} - \b{e}_i  + \b{e}_j - \b{e}_{\max(J)}}{ i \in I \ssm \{\min(I)\}, j \in J \ssm \{\max(I)\} } = \conv \{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{i \in I,\; j \in J} + \b{e}_{\min(I)} - \b{e}_{\max(J)}$.
-\vincent{horrible...}
-\end{proof}
+%\begin{proof}
+%%$\SP(\yangArc{I}{J}) = \conv \{\b{0}\} \cup \{\b{e}_{\min(I)} - \b{e}_{\max(J)}\} \cup \set{ \b{e}_{\min(I)} - \b{e}_i  + \b{e}_j - \b{e}_{\max(J)}}{ i \in I \ssm \{\min(I)\}, j \in J \ssm \{\max(I)\} } = \conv \{\b{0}\} \cup \set{\b{e}_j - \b{e}_i}{i \in I,\; j \in J} + \b{e}_{\min(I)} - \b{e}_{\max(J)}$.
+%\end{proof}
 
@@
   \]
-  if $i\succ_s j$, and $0$ otherwise, where
+  if $i\succ_s^R j$, and $0$ otherwise, where
   \begin{itemize}
@@
   Let $\sigma$ be any permutation in~$\mathcal{L}(\prec_s^R)$.
-  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc~$\alpha$ whose shard polytope $\SP(\alpha)$ contributes to $\b{e}_i-\b{e}_j$ if and only if $\sigma^{-1}_i > \sigma^{-1}_j$, hence if $i\succ_s j$.
+  From \cref{lem:yinminmax}, for a pair $i<j \in [n]$, there exists a yin arc~$\alpha$ whose shard polytope $\SP(\alpha)$ contributes to $\b{e}_i-\b{e}_j$ if and only if $\sigma^{-1}_i > \sigma^{-1}_j$, hence if $i\succ_s^R j$.
   Provided this holds, the set of yin arcs $\alpha$ such that $\SP(\alpha)$ contributes to $\b{e}_i-\b{e}_j$ are defined by pairs of contiguous intervals $I,J$ of $[n]$ with $i\in I$, $j\in J$, such that $i$ is the index of the maximum of $\sigma^{-1}$ in $I$, and $j$ is the index of the maximum of $\sigma^{-1}$ in $J$.
@@
   This leaf is associated with the node $\ell$ that comes next in inorder traversal, which must be an ancestor of $i$ in $T$.
-  Therefore, we must have $\ell\succ_s i$, and $\sigma^{-1}_\ell > \sigma^{-1}_i$, and we cannot extend $I$ past this index on the left.
+  Therefore, we must have $\ell\succ_s^R i$, and $\sigma^{-1}_\ell > \sigma^{-1}_i$, and we cannot extend $I$ past this index on the left.
   This shows that there are exactly $h^T_i$ choices, and not more, for the choice of the left endpoint of $I$.
@@
 
-\cref{lem:yincount} essentially gives us the coefficient $\yin{w}^R_{i,j}$ in \cref{thm:strongRectangulotope}, except that the condition $\neg \; {}^i \!\!\!
+\cref{lem:yincount} essentially gives us the coefficient $\yin{w}^R_{i,j}$ in \cref{thm:strongRectangulotope}, except that the condition $\neg \; {}_i \!\!\!
 \raisebox{.03cm}{\verticalPattern} \!\!\!
-{}_j$ is replaced by $i\succ_s j$.
-Recall that ${}^i \!\!\!
+{}^j$ is replaced by $i\succ_s^R j$.
+Recall that ${}_i \!\!\!
 \raisebox{.03cm}{\verticalPattern} \!\!\!
-{}_j$ denotes the situation where the rectangles $r_i$ and~$r_j$ of inorder labels $i$ and $j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
+{}^j$ denotes the situation where the rectangles $r_i$ and~$r_j$ of inorder labels $i$ and $j$ share a vertical segment, $r_i$ is on the left of $r_j$, and the bottom edge of $r_i$ is lower than the top edge of $r_j$.
 Our aim is to provide a formula for the coordinates of the vertices of the strong rectangulotope that does not involve the computation of the strong poset or one of its linear extensions.
-For that purpose, we prove that for every pair $i<j\in [n]$ such that~$i\prec_s^R j$, which do not have any associated yin arc, we have either ${}^i \!\!\!
+For that purpose, we prove that for every pair $i<j\in [n]$ such that~$i \not\succ_s^R j$, we have either ${}_i \!\!\! \raisebox{.03cm}{\verticalPattern} \!\!\! {}^j$ or $cv^{T,S}_{i,j} = 0$.
+Hence, we obtain that the product~$h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \; \neg \; {}_i \!\!\!
 \raisebox{.03cm}{\verticalPattern} \!\!\!
-{}_j$ or $cv^{T,S}_{i,j} = 0$ (hence the product in \cref{lem:yincount} is~$0$ as well and adding it does not harm the result).
+{}^j \; \rrbracket$ is always~$0$ when~$i \not\succ_s^R j$, so that adding it in \cref{thm:strongRectangulotope} does not harm the result.
 
@@
   Let $R$ be a rectangulation of size~$n$, with source and target trees~$S$ and~$T$, and strong order~$\prec_s^R$.
-  Given a pair $i,j\in [n]$ such that $i<j$ and $i\prec_s^R j$, we~have
-  \[
-    cv^{T,S}_{i,j} > 0 \quad \Longrightarrow \quad {}^i \!\!\! \raisebox{.03cm}{\verticalPattern} \!\!\! {}_j.
+  Given a pair $i,j\in [n]$ such that $i<j$ and $i\not\succ_s^R j$, we~have
+  \[
+    cv^{T,S}_{i,j} > 0 \quad \Longrightarrow \quad {}_i \!\!\! \raisebox{.03cm}{\verticalPattern} \!\!\! {}^j.
   \]
@@
   If the bottom edge of $r_i$ is higher than the top edge of $r_j$, we must have $j\btri i$, which implies $j\prec_s^R i$, contradicting our hypothesis again.
-  Thus the bottom edge of $r_i$ is lower than the top edge of $r_j$, and ${}^i \!\!\! \raisebox{.03cm}{\verticalPattern} \!\!\! {}_j$, as claimed.
+  Thus the bottom edge of $r_i$ is lower than the top edge of $r_j$, and ${}_i \!\!\! \raisebox{.03cm}{\verticalPattern} \!\!\! {}^j$, as claimed.
 \end{proof}

### 2024-10-25 draft/rectangulotopes.tex
last two comments of ref X

@@
 
+\cref{lem:yincount} and \cref{lem:whentosum} together already provide a realization of the quotientopes of the yin congruence, the vertices of which are in bijection with rectangulations that avoid one of the two patterns defining diagonal rectangulations~\cite{ACFF24}, and with $2\underline{41}3$-avoiding permutations, as mentioned in \cref{subsec:yinYangArcs}.
+
 We omit the proofs of the analogues of \cref{lem:yincount} and \cref{lem:whentosum} for the case of the yang arcs, which use completely symmetric arguments.

### 2024-10-28 response/response.tex
last two comments of ref X
minor changes letter

@@
 
-Done.
 We have added the following sentence: ``Note that, contrarily to what could suggest this $n = 4$ case, the strong rectangulation lattice is not isomorphic to the weak order, and the strong rectangulotope is not combinatorially equivalent to the permutahedron as soon as~$n > 4$.''
+
+To answer your question more specifically, when $n=4$, the strong rectangulation congruence is the trivial congruence (all classes are singletons), so the quotientope is a permutahedron (but indeed not the standard one).
+Note that both the weak and strong rectangulotopes have has many facets as the permutahedron, and are thus obtained by moving facets, not deleting facets.
 
@@
 
-Done.
+Done. We have replaced everywhere $S \subseteq [n]$ by $X \subseteq [n]$.
 
@@
 
-Done.
+Done. Thank you!
 
@@
 
-Todo.
+We believe this is correct as is.
 
@@
 
-Todo.
+There is indeed an analogue, we added a remark.
 
@@
 
-Done. We have decided to do it directly inside Theorem~3. The meaning is really that $\max(I) = \min(J)-1$.
+Done. It was indeed not precise enough. We have decided to do it directly inside Theorem~3. The meaning is really that $\max(I) = \min(J)-1$.
 
